### 2020-05-09 article.tex
Update to get template working locally.

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, animals, antibiotics, antibodies, antioxidants, applmech, applnano, applsci, arts, asc, asi, atmosphere, atoms, automation, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, bloods, brainsci, breath, buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejbc, ejihpe, electricity, electrochem, electronicmat, electronics, endocrines, energies, entropy, environments, epidemiologia, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, fuels, futureinternet, futurephys, galaxies, games, gardens, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, journalmedia, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, obesities, oceans, ohbm, osteology, optics, organics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, suschem, sustainability, world, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2020}
\copyrightyear{2020}
%\externaleditor{Academic Editor: name}
\history{Received: date; Accepted: date; Published: date}
%\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%------------------------------------------------------------------
% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
%\pdfoutput=1

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}

% Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3}
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: Place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: Describe briefly the main methods or treatments applied; (3) Results: Summarize the article's main findings; and (4) Conclusion: Indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences:
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data:
%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately. If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal. In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}

%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%\setcounter{secnumdepth}{4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact latex@mdpi.com.
%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

\section{Introduction}
The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}. Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\begin{quote}
This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection}
\unskip
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}[leftmargin=*,labelsep=5.8mm]
\item	First bullet
\item	Second bullet
\item	Third bullet
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
\item	First item
\item	Second item
\item	Third item
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=2 cm]{Definitions/logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

Text

Text

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

Text

Text

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\textwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\textwidth}{1pt}
%\end{listing}

\subsection{Formatting of Mathematical Components}

This is an example of an equation:

\begin{equation}
a + b = c
\end{equation}
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}
This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\abbreviations{The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & linear dichroism
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendix
\section{}
\unskip
\subsection{}
The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.

\section{}
All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\externalbibliography{yes}
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book}
Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
\end{thebibliography}

% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences. For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\sampleavailability{Samples of the compounds ...... are available from the authors.}

%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2020-05-09 article.tex
check

@@
 %=================================================================
-\documentclass[journal,article,submit,moreauthors,pdftex]{Definitions/mdpi}
-
-% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
-The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
-This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
-For preprints.org, the editorial staff will make this change immediately prior to posting.
-
-%--------------------
-% Class Options:
-%--------------------
-%----------
-% journal
-%----------
-% Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, animals, antibiotics, antibodies, antioxidants, applmech, applnano, applsci, arts, asc, asi, atmosphere, atoms, automation, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, bloods, brainsci, breath, buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, civileng, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, ejbc, ejihpe, electricity, electrochem, electronicmat, electronics, endocrines, energies, entropy, environments, epidemiologia, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, fuels, futureinternet, futurephys, galaxies, games, gardens, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, hearts, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jce, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jne, jnt, jof, joitmc, journalmedia, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, ijns, neurosci, neuroglia, nitrogen, notspecified, nutrients, obesities, oceans, ohbm, osteology, optics, organics, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, pollutants, polymers, polysaccharides, preprints , proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reprodmed, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, suschem, sustainability, world, symmetry, systems, technologies, telecom, test, tourismhosp, toxics, toxins, transplantology, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
-
-%---------
-% article
-%---------
-% The default type of manuscript is "article", but can be replaced by:
-% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
-% supfile = supplementary materials
-
-%----------
-% submit
-%----------
-% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
-This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
-Also, line numbering will be removed.
-Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
-
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
-
-%---------
-% pdftex
-%---------
-% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 
-%------------------------------------------------------------------
-% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
-%\pdfoutput=1
-
 %=================================================================
@@
 % Full title of the paper (Capitalized)
-\Title{Title}
-
-% Author Orchid ID: enter ID or remove command
-%\newcommand{\orcidauthorA}{0000-0000-000-000X} % Add \orcidA{} behind the author's name
-%\newcommand{\orcidauthorB}{0000-0000-000-000X} % Add \orcidB{} behind the author's name
+\Title{Design, Implementation, and Evaluation of Open Power Quality}
 
 % Authors, for the paper (add full first names)
-\Author{Firstname Lastname $^{1,\dagger,\ddagger}$, Firstname Lastname $^{1,\ddagger}$ and Firstname Lastname $^{2,}$*}
+\Author{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}
 
 % Authors, for metadata in PDF
-\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+\AuthorNames{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}
 
@@
 \address{%
-$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
-$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
-
-% Contact information of the corresponding author
-\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
-
-% Current address and/or shared authorship
-\firstnote{Current address: Affiliation 3}
-\secondnote{These authors contributed equally to this work.}
-% The commands \thirdnote{} till \eighthnote{} are available for further notes
-
-%\simplesumm{} % Simple summary
-
-%\conference{} % An extended version of a conference paper
+Department of Information and Computer Sciences\\
+University of Hawaii at Manoa\\
+Honolulu, HI 96822\\
+johnson@hawaii.edu}
 
@@
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline.)}
-
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
+\keyword{Power Quality, Open Source, Renewable Energy, Grid Stability}
 

### 2020-05-10 article.tex
checkpoint

@@
 
-% Abstract (Do not insert blank lines, i.e. \\)
-\abstract{A single paragraph of about 200 words maximum.
-For research articles, abstracts should give a pertinent overview of the work.
-We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: Place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: Describe briefly the main methods or treatments applied; (3) Results: Summarize the article's main findings; and (4) Conclusion: Indicate the main conclusions or interpretations.
-The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
+% Abstract (single paragraph, 200 words max, background, methods, results, conclusions. Currently 193 words.)
+
+\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture which must accomodate distributed, intermittent generation.
+This transition also means that the formerly sharp distinction between energy producers (i.e.
+utility companies) and consumers (residences, businesses, etc) are blurring: residences can now both produce and consume energy, making energy policy more complex.
+The Open Power Quality (OPQ) project began in 2013 with the goal of designing and implementing a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.
+Since then, we have designed low cost hardware devices that monitor power quality  and low-cost cloud-based software services that can economically analyze the data and detect a variety of anomalies.
+In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.
+Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions.
+We conclude that OPQ provides a new and useful approach to power quality monitoring.}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
+\begin{document}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences:
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data:
-%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
-If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
-In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
-
-%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%\setcounter{secnumdepth}{4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{section}{-1} %% Remove this when starting to work on the template.
-\section{How to Use this Template}
-The template details the sections that can be used in a manuscript.
-Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section).
-Please check the instructions for authors page of the journal to verify the correct order and names.
-For any questions, please contact the editorial office of the journal or support@mdpi.com.
-For LaTeX related questions please contact latex@mdpi.com.
 %The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision

### 2020-05-10 dissertation-christe/ci.py

@@ -1,0 +1,46 @@
+#!/usr/bin/env python3
+
+import glob
+import subprocess
+import sys
+
+if __name__ == "__main__":
+    tex_files = glob.glob("src/*.tex")
+    print("Checking spelling of {} tex source files...".format(len(tex_files)))
+    errors = []
+
+    for tex_file in tex_files:
+        if "appendix" in tex_file:
+            continue
+
+        result = subprocess.run(["hunspell",
+                                 "-u3",
+                                 "-d",
+                                 "en_US",
+                                 "-p",
+                                 "dict.txt",
+                                 "-t",
+                                 tex_file],
+                                stdout=subprocess.PIPE)
+        stdout = result.stdout
+        if len(stdout) > 0:
+            errs = list(map(lambda line: line.strip(), stdout.strip().split(b"\n")))
+            errs = list(filter(lambda err: "Ph".encode() not in err, errs))
+            errors.extend(errs)
+
+        warnings = ["todo", " table ", "(table", " figure ", "(figure", " id ", " ids "]
+        with open(tex_file, "r") as fin:
+            for line_no, line in enumerate(fin.readlines()):
+                for warning in warnings:
+                    if warning in line:
+                        print("WARN: {} {}:{} -- {}".format(warning, tex_file, line_no, line))
+
+    if len(errors) > 0:
+        print("Errors found...")
+        for error in errors:
+            print(error)
+        sys.exit(1)
+    else:
+        print("No errors found...")
+        sys.exit(0)
+

### 2020-05-10 dissertation-christe/clean.sh

@@ -1,0 +1,31 @@
+#!/usr/bin/env bash
+
+rm -f *.aux
+rm -f *.bbl
+rm -f *.blg
+rm -f *.lof
+rm -f *.log
+rm -f *.lot
+rm -f *.out
+rm -f *.pdf
+rm -f *.gz
+rm -f *.toc
+rm -f *.make
+rm -f *.d
+rm -f *.fls
+rm -f *.txss
+rm -f src/*.aux
+rm -f src/*.bbl
+rm -f src/*.blg
+rm -f src/*.lof
+rm -f src/*.log
+rm -f src/*.lot
+rm -f src/*.out
+rm -f src/*.pdf
+rm -f src/*.gz
+rm -f src/*.toc
+rm -f src/*.make
+rm -f src/*.d
+rm -f src/*.fls
+rm -f src/*.txss
+rm -f out/*

### 2020-05-10 dissertation-christe/convert_table.py

@@ -1,0 +1,99 @@
+import typing
+
+table = r"""
+\begin{table}[H]
+	\centering
+	\caption{OPQ Deployment}
+	\begin{tabular}{|c|c|c|c|}
+		\hline 
+		Box & Location & Coordinates & Description \\ 
+		\hline
+		1000 & POST (CSDL) & -157.816237,21.297438 & Same line as 1002, 1009, 1010, 1013. Sensitive computer equipment. \\
+		\hline
+		1001 & Hamilton (Basement 3F) & -157.816173,21.300332 & Same line as 1008. Ease of access. \\
+		\hline
+		1002 & POST (ICSpace) & -157.826327,21.30046 & Same line as 1000, 1009, 1010, 1013. Sensitive computer equipment.   \\
+		\hline
+		1003 & IT Building & -157.826327,21.30046 & Same line as 1004. Sensitive computer equipment. \\
+		\hline
+		1004 & Building 37 & -157.817821,21.298732 & Same line as 1003. LAVA Lab. \\
+		\hline
+		1005 & Parking Structure (Phase II) & -157.819234, 21.296042 & Same line as 1011, 1012. Proximity to PV. \\
+		\hline
+		1006 & Frog I & -157.823122,21.298050 & Same line as 1007, 1014. Proximity to PV. PV ground truth access. \\
+		\hline
+		1007 & Frog II & -157.822819,21.298046 & Same line as 1006, 1014. Proximity to PV. PV ground truth access.  \\
+		\hline
+		1008 & Ag. Engineering & -157.816107,21.301564 & Same line as 1001. Known PQ issues. \\
+		\hline
+		1009 & Watanabe & -157.815817,21.298351 & Same line as 1000, 1002, 1010, 1013. Sensitive physics instruments. \\
+		\hline
+		1010 & Holmes & -157.816104,21.297011 & Same line as 1000, 1002, 1009, 1013. Machining shops and electronic equipment. \\
+		\hline
+		1011 & Parking Structure (Phase I) & -157.817430,21.295338 & Same line as 1005, 1012. Proximity to PV. \\
+		\hline
+		1012 & Law Library & -157.817361,21.296328 & Same line as 1005, 1011. \\
+		\hline
+		1013  & Kennedy Theater & -157.815225,21.299282 & Same line as 1000, 1002, 1009, 1010. Demanding electronic equipment. \\
+		\hline
+		1014 & Sinclair Library & -157.820499,21.298475 & Same line as 1006, 1007. \\
+		\hline
[…]

### 2020-05-10 dissertation-christe/make-clean.sh

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+cd src && make clean

### 2020-05-10 dissertation-christe/make.sh

@@ -1,0 +1,3 @@
+#!/usr/bin/env bash
+
+cd src && make

### 2020-05-10 dissertation-christe/plotting/data_grabber.py

@@ -1,0 +1,60 @@
+import numpy as np
+import pymongo
+import pymongo.database
+import sys
+
+if __name__ == "__main__":
+    mongo_client: pymongo.MongoClient = pymongo.MongoClient()
+    db: pymongo.database.Database = mongo_client.opq
+
+    # Measurements
+    stats = db.command("collstats", "measurements")
+    print(f"measurements mean size={stats['avgObjSize']}")
+
+    # Trends
+    stats = db.command("collstats", "trends")
+    print(f"trends mean size={stats['avgObjSize']}")
+
+    # Events
+    events = db.events.find({"target_event_start_timestamp_ms": {"$gt": 0},
+                             "target_event_end_timestamp_ms": {"$gt": 0}},
+                            projection={"target_event_start_timestamp_ms": True,
+                                        "target_event_end_timestamp_ms": True,
+                                        "boxes_received": True})
+    events_bytes = []
+    timestamps = []
+
+    for event in events:
+        timestamps.append(event["target_event_start_timestamp_ms"])
+        timestamps.append(event["target_event_end_timestamp_ms"])
+        range_s = (event["target_event_end_timestamp_ms"] - event["target_event_start_timestamp_ms"]) / 1000.0
+        event_bytes = range_s * 12_000 * 2 * len(event["boxes_received"])
+        events_bytes.append(event_bytes)
+
+    total_bytes = sum(events_bytes)
+    total_s = (max(timestamps) - min(timestamps)) / 1000.0
+    mean_dr = total_bytes / total_s
+    sigma_dr = np.sqrt((1.0/total_s) * sum([(x - mean_dr)**2 for x in events_bytes]))
+
+    print("event total_bytes", total_bytes)
+    print("event total_s", total_s)
+    print("event mean_dr", mean_dr)
+    print("event sigma_dr", sigma_dr)
+
+    incidents = db.incidents.find(projection={"start_timestamp_ms": True,
+                                              "end_timestamp_ms": True,
+                                              "classifications": True})
+    incidents_bytes = []
+    for incident in incidents:
+        le = (incident["end_timestamp_ms"] - incident["start_timestamp_ms"]) / 1000.0
+        if le < 0 or "OUTAGE" in incident["classifications"]:
+            continue
+        incidents_bytes.append(le * 12_000 * 2)
+
+    total_bytes = sum(incidents_bytes)
[…]

### 2020-05-10 dissertation-christe/plotting/laha/dl.py

@@ -1,0 +1,71 @@
+import numpy as np
+
+import laha.errors as errors
+
+
+def mu_s_sd(samps: float,
+            detections: float,
+            mu_s_samp: float,
+            sigma_s_samp: float,
+            mu_sr: float,
+            sigma_sr: float,
+            mu_t_sd: float,
+            sigma_t_sd: float) -> (float, float):
+    # Result
+    _mu_s_sd = mu_s_samp * mu_sr * mu_t_sd
+
+    # Errors
+    delta_s_samp = errors.sem(sigma_s_samp, samps)
+    delta_sr = errors.sem(sigma_sr, samps)
+    delta_t_sd = errors.sem(sigma_t_sd, detections)
+
+    delta_s_sd = errors.propagate_multiplication(_mu_s_sd,
+                                                 (mu_s_samp, delta_s_samp),
+                                                 (mu_sr, delta_sr),
+                                                 (mu_t_sd, delta_t_sd))
+
+    return _mu_s_sd, delta_s_sd
+
+
+def mu_s_d(samples: float,
+           detections: float,
+           mu_s_samp: float,
+           mu_sr: float,
+           mu_t_sd: float,
+           sigma_t_sd: float,
+           mu_sd: float,
+           sigma_sd: float) -> (float, float):
+    (_mu_s_sd, delta_s_sd) = mu_s_sd(samples, detections, mu_s_samp, mu_sr, mu_t_sd, sigma_t_sd)
+
+    _mu_s_d = _mu_s_sd * mu_sd
+
+    delta_sd = errors.sem(sigma_sd, detections)
+    delta_s_d = errors.propagate_multiplication(_mu_s_d,
+                                                (_mu_s_sd, delta_s_sd),
+                                                (mu_sd, delta_sd))
+
+    return _mu_s_d, delta_s_d
+
+
+def mu_s_dl(samples: float,
+            detections: float,
+            mu_s_samp: float,
+            mu_sr: float,
+            mu_t_sd: float,
+            sigma_t_sd: float,
+            mu_sd: float,
+            sigma_sd: float,
+            mu_dr: float,
+            sigma_dr: float,
+            t: float) -> (float, float):
+    _mu_s_d, delta_s_d = mu_s_d(samples, detections, mu_s_samp, mu_sr, mu_t_sd, sigma_t_sd, mu_sd, sigma_sd)
+
+    _mu_s_dl = _mu_s_d * mu_dr * t
+
[…]

### 2020-05-10 dissertation-christe/plotting/laha/errors.py

@@ -1,0 +1,22 @@
+import functools
+
+import numpy as np
+
+
+def sem(sigma: float, n: float) -> float:
+    return sigma / np.sqrt(n)
+
+
+def propagate_constant(delta: float, *c: float) -> float:
+    _c = functools.reduce(lambda x, y: x * y, c)
+    return delta * np.abs(_c)
+
+
+def propagate_sum(*deltas: float) -> float:
+    summed_squares = sum([delta**2 for delta in deltas])
+    return np.sqrt(summed_squares)
+
+
+def propagate_multiplication(mu: float, *mu_deltas: (float, float)) -> float:
+    summed_squares = sum([(d / m)**2 for (m, d) in mu_deltas])
+    return np.abs(mu) * np.sqrt(summed_squares)

### 2020-05-10 dissertation-christe/plotting/laha/example_stats.py

@@ -1,0 +1,74 @@
+
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def example_1():
+    sample_size_bytes = 2
+    sample_rate_hz = 12_000
+    sensors_sending_data = np.array([10, 10, 10, 10, 10, 10, 10, 10, 8, 11])  # 80% 10, 10% 8, 10% 11
+    mean_sensors_sending_data = sensors_sending_data.mean()
+    sigma_sensors_sending_data = sensors_sending_data.std()
+
+    x_values = np.arange(1, 31_540_000, step=86_400)  # seconds in year by seconds in day
+    y_values = sample_size_bytes * sample_rate_hz * mean_sensors_sending_data * x_values
+
+    e_values = []
+    for i in range(len(x_values)):
+        delta_sensors_sending_data = sigma_sensors_sending_data / np.sqrt(x_values[i])
+        delta_mean_size = delta_sensors_sending_data * np.abs(sample_size_bytes * sample_rate_hz * x_values[i])
+        e_values.append(delta_mean_size)
+    e_values = np.array(e_values)
+
+    plt.plot(x_values, y_values)
+    plt.plot(x_values, y_values + e_values)
+    plt.plot(x_values, y_values - e_values)
+
+    plt.title("Example 1")
+    plt.xlabel("Time (S)")
+    plt.ylabel("Bytes")
+
+    plt.show()
+
+
+def example_2():
+    sensors_sending_data = np.array([10, 10, 10, 10, 10, 10, 10, 10, 8, 11])  # 80% 10, 10% 8, 10% 11
+    mean_sample_size_bytes = 2
+    sigma_sample_size_bytes = 0.0
+    mean_sample_rate_hz = 12_000
+    sigma_sample_rate_hz = 0.0
+    mean_sensors_sending_data = sensors_sending_data.mean()
+    sigma_sensors_sending_data = sensors_sending_data.std()
+
+    x_values = np.arange(1, 31_540_000, step=86_400)  # seconds in year by seconds in day
+    y_values = mean_sample_size_bytes * mean_sample_rate_hz * mean_sensors_sending_data * x_values
+
+    e_values = []
+    for i in range(len(x_values)):
+        num_samples = x_values[i] * mean_sample_rate_hz
+        delta_sample_size_bytes = sigma_sample_size_bytes / np.sqrt(num_samples)
+        delta_sample_rate_hz = sigma_sample_rate_hz / np.sqrt(num_samples)
[…]

### 2020-05-10 dissertation-christe/plotting/laha/iml.py

@@ -1,0 +1,53 @@
+import numpy as np
+
+import laha.errors as errors
+
+
+def s_sen(s_samp: float, sr: float, t: float) -> float:
+    return s_samp * sr * t
+
+
+def s_iml(s_sens: np.ndarray) -> float:
+    return s_sens.sum()
+
+
+def mu_s_sen(samp: int,
+             mu_s_samp: float,
+             sigma_s_samp: float,
+             mu_sr: float,
+             sigma_sr: float,
+             t: float) -> (float, float):
+    _mu_s_sen = mu_s_samp * mu_sr * t
+
+    delta_s_samp = errors.sem(sigma_s_samp, samp)
+    delta_sr = errors.sem(sigma_sr, samp)
+
+    e = errors.propagate_multiplication(_mu_s_sen,
+                                        (mu_s_samp, delta_s_samp),
+                                        (mu_sr, delta_sr))
+    delta_s_sen = errors.propagate_constant(e, t)
+
+    return _mu_s_sen, delta_s_sen
+
+
+def mu_s_iml(samp: int,
+             mu_s_samp: float,
+             sigma_s_samp: float,
+             mu_sr: float,
+             sigma_sr: float,
+             mu_b: float,
+             sigma_b: float,
+             t: float) -> (float, float):
+    _mu_s_iml = mu_s_samp * mu_sr * mu_b * t
+
+    delta_s_samp = errors.sem(sigma_s_samp, samp)
+    delta_sr = errors.sem(sigma_sr, samp)
+    delta_b = errors.sem(sigma_b, samp)
+
+    e = errors.propagate_multiplication(_mu_s_iml,
+                                        (mu_s_samp, delta_s_samp),
+                                        (mu_sr, delta_sr),
+                                        (mu_b, delta_b))
+    delta_s_iml = errors.propagate_constant(e, t)
+
+    return _mu_s_iml, delta_s_iml

### 2020-05-10 dissertation-christe/plotting/lokahi/actual_data.py

@@ -1,0 +1,1525 @@
+import collections
+import datetime
+import functools
+import os
+import os.path
+from dataclasses import dataclass
+from typing import Dict, Set, List, Callable, Tuple, Union
+import urllib.parse
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+import scipy.stats as stats
+
+seconds_in_day = 86400
+seconds_in_two_weeks = seconds_in_day * 14
+seconds_in_month = seconds_in_day * 30.4167
+seconds_in_year = seconds_in_month * 12
+seconds_in_2_years = seconds_in_year * 2
+
+
+@dataclass
+class SeriesSpec:
+    values: List
+    dt_func: Callable
+    v_func: Callable
+
+
+def bin_dt_by_min(dt: datetime.datetime) -> datetime.datetime:
+    return datetime.datetime(dt.year, dt.month, dt.day, dt.hour, dt.minute, 0, 0, tzinfo=datetime.timezone.utc)
+
+
+def bin_dt_by_day(dt: datetime.datetime) -> datetime.datetime:
+    return datetime.datetime(dt.year, dt.month, dt.day, 0, 0, 0, 0)
+
+
+def intersect_lists(lists: List[List[datetime.datetime]]) -> Set[datetime.datetime]:
+    sets: List[Set] = list(map(set, lists))
+    return set.intersection(*sets)
+
+
+def align_data_multi(series: List[SeriesSpec]) -> List[Tuple[np.ndarray, np.ndarray]]:
+    all_dts: List[List[datetime.datetime]] = []
+    all_binned_dts: List[List[datetime.datetime]] = []
+    all_vals: List[List[float]] = []
+    already_seens: List[Set[datetime.datetime]] = []
+
+    for _ in range(len(series)):
+        already_seens.append(set())
+
+    for serie in series:
+        dts: List[datetime.datetime] = list(map(serie.dt_func, serie.values))
+        binned_dts: List[datetime.datetime] = list(map(bin_dt_by_day, dts))
+        vals: List = list(map(serie.v_func, serie.values))
+        all_dts.append(dts)
+        all_binned_dts.append(binned_dts)
+        all_vals.append(vals)
+
+    intersecting_dts: Set[datetime.datetime] = intersect_lists(all_binned_dts)
+
+    res: List[Tuple[np.ndarray, np.ndarray]] = []
+
+    for i, binned_dts in enumerate(all_binned_dts):
+        dts = []
[…]

### 2020-05-10 dissertation-christe/plotting/lokahi/historical_device_locations.py

@@ -1,0 +1,121 @@
+import datetime
+import os
+from dataclasses import dataclass
+from typing import Dict, List
+import urllib.parse
+
+import fastkml
+import pymongo
+import pymongo.database
+import shapely
+import shapely.geometry
+
+
+@dataclass
+class HistoricalDevice:
+    last_active_ts: int
+    id_uuid: str
+    lat: float
+    lng: float
+
+    @staticmethod
+    def from_line(line: str) -> 'HistoricalDevice':
+        split_line = line.strip().split(" ")
+        last_active_ts: int = int(split_line[0])
+        id_uuid: str = split_line[1]
+        lat: float = float(split_line[2])
+        lng: float = float(split_line[3])
+
+        return HistoricalDevice(last_active_ts, id_uuid, lat, lng)
+
+
+def parse_historical_devices(path: str) -> List[HistoricalDevice]:
+    with open(path, "r") as fin:
+        lines: List[str] = fin.readlines()
+        historical_devices: List[HistoricalDevice] = list(map(HistoricalDevice.from_line, lines))
+        return sorted(historical_devices, key=lambda historical_device: historical_device.last_active_ts)
+
+
+def format_mongo_auth_uri_ssl(user: str,
+                              password: str,
+                              host: str,
+                              port: int,
+                              cert: str) -> str:
+    escaped_pass = urllib.parse.quote_plus(password)
+    return f"mongodb://{user}:{escaped_pass}@{host}:{port}/?ssl=true&ssl_ca_certs={cert}&replicaSet=rs0"
+
+
+def get_client(host: str, port: int, user: str, password: str, cert: str) -> pymongo.MongoClient:
+    return pymongo.MongoClient(format_mongo_auth_uri_ssl(user, password, host, port, cert))
+
+
+def get_historical_devices(mongo_client: pymongo.MongoClient) -> None:
+    db: pymongo.database.Database = mongo_client["redvox"]
+    coll: pymongo.collection.Collection = db["HistoricalDevice"]
+    dt_start_window: datetime.datetime = datetime.datetime(2019, 10, 1, 0, 0, 0, 0)
+
+    query: Dict = {
+        "lastActive": {"$gte": dt_start_window}
+    }
+
+    projection: Dict = {
+        "_id": False,
+        "deviceId": True,
+        "uuid": True,
+        "lastActive": True,
+        "lastLocation.coordinates": True
[…]

### 2020-05-10 dissertation-christe/plotting/lokahi/pl.py

@@ -1,0 +1,412 @@
+from dataclasses import dataclass
+import datetime
+import os
+from typing import Callable, Dict, List, Tuple, TypeVar, Optional, Set
+import urllib.parse
+
+import bson
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+from scipy import stats
+
+
+@dataclass
+class SeriesSpec:
+    values: List
+    dt_func: Callable
+    v_func: Callable
+
+
+def bin_dt_by_min(dt: datetime.datetime) -> datetime.datetime:
+    return datetime.datetime(dt.year, dt.month, dt.day, dt.hour, dt.minute, 0, 0)
+
+
+# def bin_dt_by_day(dt: datetime.datetime) -> datetime.datetime:
+#     return datetime.datetime(dt.year, dt.month, dt.day, 0, 0, 0, 0)
+
+
+def intersect_lists(lists: List[List[datetime.datetime]]) -> Set[datetime.datetime]:
+    sets: List[Set] = list(map(set, lists))
+    return set.intersection(*sets)
+
+def align_data_multi(series: List[SeriesSpec]) -> List[Tuple[np.ndarray, np.ndarray]]:
+    all_dts: List[List[datetime.datetime]] = []
+    all_binned_dts: List[List[datetime.datetime]] = []
+    all_vals: List[List[float]] = []
+    already_seens: List[Set[datetime.datetime]] = []
+
+    for _ in range(len(series)):
+        already_seens.append(set())
+
+    for serie in series:
+        dts: List[datetime.datetime] = list(map(serie.dt_func, serie.values))
+        binned_dts: List[datetime.datetime] = list(map(bin_dt_by_min, dts))
+        vals: List = list(map(serie.v_func, serie.values))
+        all_dts.append(dts)
+        all_binned_dts.append(binned_dts)
+        all_vals.append(vals)
+
+    intersecting_dts: Set[datetime.datetime] = intersect_lists(all_binned_dts)
+
+    res: List[Tuple[np.ndarray, np.ndarray]] = []
+
+    for i, binned_dts in enumerate(all_binned_dts):
+        dts = []
+        vs = []
+        for j, binned_dt in enumerate(binned_dts):
+            if binned_dt in intersecting_dts and binned_dt not in already_seens[i]:
+                dts.append(binned_dt)
+                vs.append(all_vals[i][j])
+                already_seens[i].add(binned_dt)
[…]

### 2020-05-10 dissertation-christe/plotting/lokahi/plot_laha.py

@@ -1,0 +1,582 @@
+from typing import *
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+seconds_in_day = 86400
+seconds_in_two_weeks = seconds_in_day * 14
+seconds_in_month = seconds_in_day * 30.4167
+seconds_in_year = seconds_in_month * 12
+seconds_in_2_years = seconds_in_year * 2
+
+
+class Data:
+    def __init__(self,
+                 time: int,
+                 total_samples: int,
+                 total_samples_b: int,
+                 total_measurements: int,
+                 total_measurements_b: int,
+                 total_orphaned_measurements: int,
+                 total_orphaned_measurements_b: int,
+                 total_event_measurements: int,
+                 total_event_measurements_b: int,
+                 total_incident_measurements: int,
+                 total_incident_measurements_b: int,
+                 total_phenomena_measurements: int,
+                 total_phenomena_measurements_b: int,
+                 total_trends: int,
+                 total_trends_b: int,
+                 total_orphaned_trends: int,
+                 total_orphaned_trends_b: int,
+                 total_event_trends: int,
+                 total_event_trends_b: int,
+                 total_incident_trends: int,
+                 total_incident_trends_b: int,
+                 total_phenomena_trends: int,
+                 total_phenomena_trends_b: int,
+                 total_events: int,
+                 total_events_b: int,
+                 total_orphaned_events: int,
+                 total_orphaned_events_b: int,
+                 total_incident_events: int,
+                 total_incident_events_b: int,
+                 total_phenomena_events: int,
+                 total_phenomena_events_b: int,
+                 total_incidents: int,
+                 total_incidents_b: int,
+                 total_phenomena_incidents: int,
+                 total_phenomena_incidents_b: int,
+                 total_phenomena: int,
+                 total_phenomena_b: int,
+                 total_laha_b: int,
+                 total_iml_b: int,
+                 total_aml_b: int,
+                 total_dl_b: int,
[…]

### 2020-05-10 dissertation-christe/plotting/lokahi/results.py

@@ -1,0 +1,160 @@
+import collections
+import datetime
+import os
+import os.path
+from typing import Dict, Set
+import urllib.parse
+
+import pymongo
+import pymongo.database
+
+
+class DailyMetrics:
+    def __init__(self,
+                 devices_80hz: Set[str],
+                 devices_800hz: Set[str],
+                 devices_8000hz: Set[str],
+                 total_80hz_packets: int,
+                 total_800hz_packets: int,
+                 total_8000hz_packets: int):
+
+        self.devices_80hz: Set[str] = devices_80hz
+        self.devices_800hz: Set[str] = devices_800hz
+        self.devices_8000hz: Set[str] = devices_8000hz
+        self.total_80hz_packets: int = total_80hz_packets
+        self.total_800hz_packets: int = total_800hz_packets
+        self.total_8000hz_packets: int = total_8000hz_packets
+
+    @staticmethod
+    def new() -> 'DailyMetrics':
+        return DailyMetrics(set(), set(), set(), 0, 0, 0)
+
+    def total_devices_80hz(self) -> int:
+        return len(self.devices_80hz)
+
+    def total_devices_800hz(self) -> int:
+        return len(self.devices_800hz)
+
+    def total_devices_8000hz(self) -> int:
+        return len(self.devices_8000hz)
+
+    def total_devices(self) -> int:
+        return self.total_devices_80hz() + self.total_devices_800hz() + self.total_devices_8000hz()
+
+    def total_packets(self) -> int:
+        return self.total_80hz_packets + self.total_800hz_packets + self.total_8000hz_packets
+
+    def total_data_bytes_80hz(self) -> float:
+        return self.total_80hz_packets * 4096 * 4
+
+    def total_data_bytes_800hz(self) -> float:
+        return self.total_800hz_packets * 32768 * 4
+
+    def total_data_bytes_8000hz(self) -> float:
+        return self.total_8000hz_packets * 262144 * 4
+
+    def total_data_bytes(self) -> float:
+        return self.total_data_bytes_80hz() + self.total_devices_800hz() + self.total_devices_8000hz()
+
+
+def format_mongo_auth_uri_ssl(user: str,
+                              password: str,
+                              host: str,
+                              port: int,
+                              cert: str) -> str:
[…]

### 2020-05-10 dissertation-christe/plotting/lokahi/stats.py

@@ -1,0 +1,233 @@
+import os
+from typing import Dict, List, Optional, TypeVar
+import urllib.parse
+
+import boto3
+import botocore
+import numpy as np
+import pymongo
+import pymongo.database
+import bson
+
+T = TypeVar("T")
+
+
+def doc_get(doc: Dict, key: str, default: Optional[T] = None) -> Optional[T]:
+    if key in doc:
+        return doc[key]
+
+    return default
+
+
+def get_sample_rate(mongo_client: pymongo.MongoClient, id_uuid: str, start_ts_s: int, end_ts_s: int) -> float:
+    db: pymongo.database.Database = mongo_client["redvox"]
+    coll: pymongo.collection.Collection = db["RedvoxPacketApi900"]
+    s = id_uuid.split(":")
+    redvox_id = s[0]
+    redvox_uuid = s[1]
+    query = {"appFileStartTimestampMachine": {"$gte": start_ts_s * 1_000_000.0,
+                                              "$lte": end_ts_s * 1_000_000.0},
+             "redvoxId": redvox_id,
+             "redvoxUuid": redvox_uuid}
+    projection = {"_id": False,
+                  "appFileStartTimestampMachine": True,
+                  "redvoxId": True,
+                  "redvoxUuid": True,
+                  "evenlySampledChannels": True}
+
+    packet = coll.find_one(query, projection=projection)
+    return packet["evenlySampledChannels"][0]["sampleRateHz"] if packet is not None else 0
+
+class RedvoxReport:
+    def __init__(self,
+                 start_timestamp_s: int,
+                 end_timestamp_s: int,
+                 web_based_products: List[str],
+                 report_id: str,
+                 is_public: bool,
+                 is_private: bool,
+                 receivers: int,
+                 devices: List[str],
+                 mongo_client: pymongo.MongoClient,
+                 meta_size_bytes: int,
+                 s3_client):
+        self.start_timestamp_s = start_timestamp_s
+        self.end_timestamp_s = end_timestamp_s
+        self.web_based_products = web_based_products
+        self.report_id = report_id
+        self.is_public = is_public
+        self.is_private = is_private
+        self.receivers = receivers
+        self.devices = devices
+        self.estimated_data_bytes = self.__data_bytes(mongo_client, s3_client)
+        self.product_bytes = self.__product_bytes(s3_client)
[…]

### 2020-05-10 dissertation-christe/plotting/lokahi/stats_resultz.py

@@ -1,0 +1,221 @@
+import collections
+import datetime
+import os
+from typing import Dict, List, Optional, TypeVar
+import urllib.parse
+
+import boto3
+import botocore
+import numpy as np
+import pymongo
+import pymongo.database
+
+T = TypeVar("T")
+
+
+def doc_get(doc: Dict, key: str, default: Optional[T] = None) -> Optional[T]:
+    if key in doc:
+        return doc[key]
+
+    return default
+
+
+def get_sample_rate(mongo_client: pymongo.MongoClient, id_uuid: str, start_ts_s: int, end_ts_s: int) -> float:
+    db: pymongo.database.Database = mongo_client["redvox"]
+    coll: pymongo.collection.Collection = db["RedvoxPacketApi900"]
+    s = id_uuid.split(":")
+    redvox_id = s[0]
+    redvox_uuid = s[1]
+    query = {"appFileStartTimestampMachine": {"$gte": start_ts_s * 1_000_000.0,
+                                              "$lte": end_ts_s * 1_000_000.0},
+             "redvoxId": redvox_id,
+             "redvoxUuid": redvox_uuid}
+    projection = {"_id": False,
+                  "appFileStartTimestampMachine": True,
+                  "redvoxId": True,
+                  "redvoxUuid": True,
+                  "evenlySampledChannels": True}
+
+    packet = coll.find_one(query, projection=projection)
+    return packet["evenlySampledChannels"][0]["sampleRateHz"] if packet is not None else 0
+
+
+class RedvoxReport:
+    def __init__(self,
+                 start_timestamp_s: int,
+                 end_timestamp_s: int,
+                 web_based_products: List[str],
+                 report_id: str,
+                 is_public: bool,
+                 is_private: bool,
+                 receivers: int,
+                 devices: List[str],
+                 mongo_client: pymongo.MongoClient,
+                 s3_client):
+        self.start_timestamp_s = start_timestamp_s
+        self.end_timestamp_s = end_timestamp_s
+        self.web_based_products = web_based_products
+        self.report_id = report_id
+        self.is_public = is_public
+        self.is_private = is_private
+        self.receivers = receivers
+        self.devices = devices
+        self.estimated_data_bytes = self.__data_bytes(mongo_client, s3_client)
+        self.product_bytes = self.__product_bytes(s3_client)
+        self.total_bytes = self.estimated_data_bytes + self.product_bytes
+        self.is_estimated = False
+
+    def duration(self) -> int:
[…]

### 2020-05-10 dissertation-christe/plotting/opq/actual_data.py

@@ -1,0 +1,1703 @@
+from dataclasses import dataclass
+import datetime
+from typing import Any, Dict, List, Optional, Tuple, Union, TypeVar, Callable, Set
+import pickle
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+import scipy.stats as stats
+
+DB: str = "opq"
+COLL: str = "laha_stats"
+
+S_IN_DAY = 86_400
+S_IN_YEAR = 31_540_000
+
+seconds_in_day = 86400
+seconds_in_two_weeks = seconds_in_day * 14
+seconds_in_month = seconds_in_day * 30.4167
+seconds_in_year = seconds_in_month * 12
+seconds_in_2_years = seconds_in_year * 2
+
+A = TypeVar("A")
+B = TypeVar("B")
+C = TypeVar("C")
+D = TypeVar("D")
+
+
+def bin_dt_by_min(dt: datetime.datetime) -> datetime.datetime:
+    return datetime.datetime(dt.year, dt.month, dt.day, dt.hour, dt.minute, 0, 0)
+
+
+def align_data(series_a: List,
+               series_b: List,
+               dt_func_a: Callable,
+               dt_func_b: Callable,
+               val_func_a: Callable,
+               val_func_b: Callable) -> Tuple[np.ndarray, np.ndarray, np.ndarray, np.ndarray]:
+    a_dts: List[datetime.datetime] = list(map(dt_func_a, series_a))
+    b_dts: List[datetime.datetime] = list(map(dt_func_b, series_b))
+    a_vals: List = list(map(val_func_a, series_a))
+    b_vals: List = list(map(val_func_b, series_b))
+    a_binned_dts: List[datetime.datetime] = list(map(bin_dt_by_min, a_dts))
+    b_binned_dts: List[datetime.datetime] = list(map(bin_dt_by_min, b_dts))
+
+    intersecting_dts: Set[datetime.datetime] = set(a_binned_dts).intersection(set(b_binned_dts))
+
+    resulting_dts_a: List[datetime.datetime] = []
+    resulting_dts_b: List[datetime.datetime] = []
+    resulting_a_vals: List = []
+    resulting_b_vals: List = []
+
+    already_seen_a_dts: Set[datetime.datetime] = set()
+    already_seen_b_dts: Set[datetime.datetime] = set()
+
+    for i in range(len(series_a)):
+        dt = a_binned_dts[i]
[…]

### 2020-05-10 dissertation-christe/plotting/opq/aml_stats.py

@@ -1,0 +1,110 @@
+from typing import *
+
+import numpy as np
+import pymongo
+import pymongo.database
+
+if __name__ == "__main__":
+    mongo_client = pymongo.MongoClient()
+    db: pymongo.database.Database = mongo_client["opq"]
+    events_coll: pymongo.collection.Collection = db["events"]
+    box_events_coll: pymongo.collection.Collection = db["box_events"]
+    incidents_coll: pymongo.collection.Collection = db["incidents"]
+
+    # Statistics for events
+    events_query = {}
+    events_projection = {"_id": False,
+                         "event_id": True}
+
+    events: pymongo.cursor.Cursor = events_coll.find(events_query, projection=events_projection)
+    events: List[Dict] = list(events)
+    event_ids: List[int] = list(map(lambda event: event["event_id"], events))
+
+    box_events_query = {"event_id": {"$in": event_ids}}
+    box_events_projection: Dict[str, bool] = {"_id": False,
+                                              "event_start_timestamp_ms": True,
+                                              "event_end_timestamp_ms": True,
+                                              "event_id": True}
+
+    box_events: pymongo.cursor.Cursor = box_events_coll.find(box_events_query, projection=box_events_projection)
+    box_events: List[Dict] = list(box_events)
+
+    event_min_ts_ms: int = min(list(map(lambda doc: doc["event_start_timestamp_ms"], box_events)))
+    event_max_ts_ms: int = max(list(map(lambda doc: doc["event_end_timestamp_ms"], box_events)))
+    event_min_ts_s: float = event_min_ts_ms / 1000.0
+    event_max_ts_s: float = event_max_ts_ms / 1000.0
+
+    event_durations_ms: np.ndarray = np.array(
+            list(map(lambda doc: doc["event_end_timestamp_ms"] - doc["event_start_timestamp_ms"], box_events)))
+    event_durations_s: np.ndarray = event_durations_ms / 1000.0
+
+    event_total_duration_ms: int = event_max_ts_ms - event_min_ts_ms
+    event_total_duration_s: float = event_total_duration_ms / 1000.0
+
+    event_data_duration_s: float = event_durations_s.sum()
+    event_mean_data_duration_s: float = event_durations_s.mean()
+    events_per_second = len(box_events) / event_total_duration_s
+
+    event_data_stored: np.ndarray = 12_000.0 * 2 * event_durations_s
+    event_total_data = event_data_stored.sum()
+    event_mean_data = event_data_stored.mean()
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/ground_truth.py

@@ -1,0 +1,58 @@
+import datetime
+import glob
+import os.path
+from typing import List
+
+import numpy as np
+import pymongo
+
+import util.frequency as frequency
+import util.io as io
+import util.thd as thd
+import util.voltage as voltage
+import util.events as events
+
+
+def find_ground_truth_data_range(ground_truth_root: str):
+    all_files: List[str] = glob.glob(os.path.join(ground_truth_root, "*", "*"), recursive=True)
+    files: List[str] = list(filter(os.path.isfile, all_files))
+
+    data_points: List[io.DataPoint] = []
+
+    for file in files:
+        file_data_points: List[io.DataPoint] = io.parse_file(file)
+        data_points.extend(file_data_points)
+
+    timestamps = np.array(list(map(lambda data_point: data_point.ts_s, data_points)))
+    min_timestamp = timestamps.min()
+    max_timestamp = timestamps.max()
+
+    print(f"min_ts {min_timestamp} {datetime.datetime.utcfromtimestamp(min_timestamp)}")
+    print(f"min_ts {max_timestamp} {datetime.datetime.utcfromtimestamp(max_timestamp)}")
+
+
+def main():
+    mongo_client: pymongo.MongoClient = pymongo.MongoClient()
+    start_ts_s: int = 1574977260
+    end_ts_s: int = 1577664000
+    gt_root: str = "/Users/anthony/scrap/ground_truth_data2"
+    # gt_root: str = "/Users/anthony/Desktop/forAnthony/data"
+    out_dir: str = "/Users/anthony/Development/dissertation/src/figures"
+
+    # find_ground_truth_data_range(gt_root)
+
+    # frequency.compare_frequencies(start_ts_s, end_ts_s, gt_root, mongo_client, out_dir)
+    # thd.compare_thds(start_ts_s, end_ts_s, gt_root, mongo_client, out_dir)
+    # voltage.compare_vrms(start_ts_s, end_ts_s, gt_root, mongo_client, out_dir)
+
+    frequency.compare_frequency_incidents(start_ts_s, end_ts_s, gt_root, mongo_client, out_dir)
+    # voltage.compare_vrms_incidents(start_ts_s, end_ts_s, gt_root, mongo_client, out_dir)
+    # thd.compare_thd_incidents(start_ts_s, end_ts_s, gt_root, mongo_client, out_dir)
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/process.py

@@ -1,0 +1,54 @@
+import matplotlib.pyplot as plt
+import glob
+import math
+keys = {}
+
+for dir in glob.glob("data/*"):
+	try:
+		file1  = open(dir + "/VBC")
+		file2  = open(dir + "/VAB")
+		file3  = open(dir + "/VCA")
+	except:
+		print("Dir " + dir + " broked")
+		continue
+	if dir == "data/KELLER_HALL_MAIN_MTR":
+		continue
+	data_min = []
+	data_max = []
+	t_min = []
+	t_max = []
+	print(dir)
+	for line1 in file1:
+		line2 = file2.readline();
+		line3 = file3.readline();
+		min1 = float(line1.split()[2])
+		min2 = float(line2.split()[2])
+		min3 = float(line3.split()[2])
+		min_rms = math.sqrt(min1**2 + min2**2 + min3**2)/4/math.sqrt(3)
+		if min_rms > 60:
+			t_min.append(int(line1.split()[0]))
+			data_min.append(min_rms)
+		
+		max1 = float(line1.split()[3])
+		max2 = float(line2.split()[3])
+		max3 = float(line3.split()[3])
+		max_rms = math.sqrt(max1**2 + max2**2 + max3**2)/4/math.sqrt(3)
+		if max_rms > 60:
+			t_max.append(int(line1.split()[0]))
+			data_max.append(max_rms)
+		if data_max[-1] > 125 or data_min[-1] <115:
+			if not line1.split()[0] in keys:
+				keys[line1.split()[0]] = []
+			keys[line1.split()[0]].append(dir.split("/")[1])		
+	plt.xlim((1574889300 - 10*60, 1574889300 + 10*60))
+	plt.title("Device: " + line1.split()[0])
+	plt.plot(t_min, data_min)
+	plt.plot(t_max, data_max)
+	plt.show()
+	#break;
+
+for key in keys:
+	if len(keys[key]) >1:
+		print(key)
+		for meter in keys[key]:
+			print("\t" + meter)

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/__init__.py

@@ -1,0 +1,96 @@
+from typing import Dict, List
+import os.path
+
+opq_box_to_uhm_meters: Dict[str, List[str]] = {
+    "1000": ["POST_MAIN_1",
+             "POST_MAIN_2"],
+    "1001": ["HAMILTON_LIB_PH_III_CH_1_MTR",
+             "HAMILTON_LIB_PH_III_CH_2_MTR",
+             "HAMILTON_LIB_PH_III_CH_3_MTR",
+             "HAMILTON_LIB_PH_III_MAIN_1_MTR",
+             "HAMILTON_LIB_PH_III_MAIN_2_MTR",
+             "HAMILTON_LIB_PH_III_MCC_AC1_MTR",
+             "HAMILTON_LIB_PH_III_MCC_AC2_MTR"],
+    "1002": ["POST_MAIN_1",
+             "POST_MAIN_2"],
+    "1003": ["KELLER_HALL_MAIN_MTR"],
+    "1005": [],
+    "1006": [],
+    "1007": [],
+    "1008": [],
+    "1009": [],
+    "1010": [],
+    "1021": ["MARINE_SCIENCE_MAIN_A_MTR",
+             "MARINE_SCIENCE_MAIN_B_MTR",
+             "MARINE_SCIENCE_MCC_MTR"],
+    "1022": ["AG_ENGINEERING_MAIN_MTR",
+             "AG_ENGINEERING_MCC_MTR"],
+    "1023": ["LAW_LIB_MAIN_MTR"],
+    "1024": [],
+    "1025": ["KENNEDY_THEATRE_MAIN_MTR"]
+}
+
+uhm_meter_to_opq_box: Dict[str, str] = {}
+for opq_box, uhm_meters in opq_box_to_uhm_meters.items():
+    for uhm_meter in uhm_meters:
+        uhm_meter_to_opq_box[uhm_meter] = opq_box
+
+
+def latex_figure_source(path: str) -> str:
+    file_name: str = os.path.split(path)[-1]
+    file_label: str = file_name.split(".")[0]
+    return """
+Figure~\\ref{fig:FILE_LABEL}    
+    
+\\begin{figure}[H]
+    \\centering
+    \\includegraphics[width=\\linewidth]{figures/FILE_NAME}
+    \\caption{}
+    \\label{fig:FILE_LABEL}
+\\end{figure}
+""".replace("FILE_NAME", file_name).replace("FILE_LABEL", file_label)
+
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/align_data.py

@@ -1,0 +1,101 @@
+import datetime
+from dataclasses import dataclass
+import functools
+from typing import Callable, List, Tuple, Set, Iterable
+
+import numpy as np
+
+
+@dataclass
+class SeriesSpec:
+    values: List
+    dt_func: Callable
+    v_func: Callable
+
+
+def bin_dt_by_min(dt: datetime.datetime) -> datetime.datetime:
+    return datetime.datetime(dt.year, dt.month, dt.day, dt.hour, dt.minute, 0, 0, tzinfo=datetime.timezone.utc)
+
+
+def intersect_lists(lists: List[List[datetime.datetime]]) -> Set[datetime.datetime]:
+    sets: List[Set] = list(map(set, lists))
+    return set.intersection(*sets)
+
+
+def align_data_multi(series: List[SeriesSpec]) -> List[Tuple[np.ndarray, np.ndarray]]:
+    all_dts: List[List[datetime.datetime]] = []
+    all_binned_dts: List[List[datetime.datetime]] = []
+    all_vals: List[List[float]] = []
+    already_seens: List[Set[datetime.datetime]] = []
+
+    for _ in range(len(series)):
+        already_seens.append(set())
+
+    for serie in series:
+        dts: List[datetime.datetime] = list(map(serie.dt_func, serie.values))
+        binned_dts: List[datetime.datetime] = list(map(bin_dt_by_min, dts))
+        vals: List = list(map(serie.v_func, serie.values))
+        all_dts.append(dts)
+        all_binned_dts.append(binned_dts)
+        all_vals.append(vals)
+
+    intersecting_dts: Set[datetime.datetime] = intersect_lists(all_binned_dts)
+
+    res: List[Tuple[np.ndarray, np.ndarray]] = []
+
+    for i, binned_dts in enumerate(all_binned_dts):
+        dts = []
+        vs = []
+        for j, binned_dt in enumerate(binned_dts):
+            if binned_dt in intersecting_dts and binned_dt not in already_seens[i]:
+                dts.append(binned_dt)
+                vs.append(all_vals[i][j])
+                already_seens[i].add(binned_dt)
+        res.append((np.array(dts), np.array(vs)))
+
+    return res
+
+
+
+
+def align_data_by_min(series_a: List,
+                      series_b: List,
+                      dt_func_a: Callable,
+                      dt_func_b: Callable,
+                      v_func_a: Callable,
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/events.py

@@ -1,0 +1,256 @@
+import datetime
+from typing import Dict, List, Tuple
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+import scipy.stats as stats
+
+import util
+import util.align_data as align
+import util.io as io
+
+
+def events(opq_start_ts_s: int,
+           opq_end_ts_s: int,
+           opq_box_id: str,
+           ground_truth_root: str,
+           uhm_sensor: str,
+           mongo_client: pymongo.MongoClient,
+           out_dir: str) -> None:
+    # if opq_box_id != "1021" and uhm_sensor != "MARINE_SCIENCE_MCC_MTR":
+    #     return
+
+
+    ground_truth_path_f: str = f"{ground_truth_root}/{uhm_sensor}/Frequency"
+    uhm_data_points_f: List[io.DataPoint] = io.parse_file(ground_truth_path_f)
+
+    ground_truth_path_thd: str = f"{ground_truth_root}/{uhm_sensor}/AVG_VOLTAGE_THD"
+    uhm_data_points_thd: List[io.DataPoint] = io.parse_file(ground_truth_path_thd)
+
+    ground_truth_path_vab: str = f"{ground_truth_root}/{uhm_sensor}/VAB"
+    uhm_data_points_vab: List[io.DataPoint] = io.parse_file(ground_truth_path_vab)
+
+    ground_truth_path_vbc: str = f"{ground_truth_root}/{uhm_sensor}/VBC"
+    uhm_data_points_vbc: List[io.DataPoint] = io.parse_file(ground_truth_path_vbc)
+
+    ground_truth_path_vca: str = f"{ground_truth_root}/{uhm_sensor}/VCA"
+    uhm_data_points_vca: List[io.DataPoint] = io.parse_file(ground_truth_path_vca)
+
+    series: List[align.SeriesSpec] = [
+        align.SeriesSpec(uhm_data_points_f,
+                         lambda data_point: datetime.datetime.utcfromtimestamp(data_point.ts_s),
+                         lambda data_point: data_point.min_v),
+        align.SeriesSpec(uhm_data_points_f,
+                         lambda data_point: datetime.datetime.utcfromtimestamp(data_point.ts_s),
+                         lambda data_point: data_point.max_v),
+        align.SeriesSpec(uhm_data_points_thd,
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/frequency.py

@@ -1,0 +1,214 @@
+import datetime
+import os.path
+from typing import Dict, List, Tuple
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+import scipy.stats as stats
+
+import util
+import util.align_data as align
+import util.io as io
+
+
+def plot_frequency(opq_start_ts_s: int,
+                   opq_end_ts_s: int,
+                   opq_box_id: str,
+                   ground_truth_root: str,
+                   uhm_sensor: str,
+                   mongo_client: pymongo.MongoClient,
+                   out_dir: str) -> str:
+    ground_truth_path: str = f"{ground_truth_root}/{uhm_sensor}/Frequency"
+    uhm_data_points: List[io.DataPoint] = io.parse_file(ground_truth_path)
+
+    db: pymongo.database.Database = mongo_client["opq"]
+    coll: pymongo.collection.Collection = db["trends"]
+    query: Dict = {
+        "box_id": opq_box_id,
+        "timestamp_ms": {"$gte": opq_start_ts_s * 1000,
+                         "$lte": opq_end_ts_s * 1000},
+        "frequency": {"$exists": True}
+    }
+
+    projection: Dict[str, bool] = {
+        "_id": False,
+        "box_id": True,
+        "timestamp_ms": True,
+        "frequency": True,
+    }
+
+    cursor: pymongo.cursor.Cursor = coll.find(query, projection=projection)
+    opq_trend_docs: List[Dict] = list(cursor)
+    opq_trends: List[io.Trend] = list(map(io.Trend.from_doc, opq_trend_docs))
+
+    aligned_data: Tuple[np.ndarray, np.ndarray, np.ndarray, np.ndarray] = align.align_data_by_min(
+            opq_trends,
+            uhm_data_points,
+            lambda trend: datetime.datetime.utcfromtimestamp(trend.timestamp_ms / 1000.0),
+            lambda data_point: datetime.datetime.utcfromtimestamp(data_point.ts_s),
+            lambda trend: trend.frequency.average,
+            lambda data_point: data_point.avg_v
+    )
+
+    aligned_opq_dts: np.ndarray = aligned_data[0]
+    aligned_opq_vs: np.ndarray = aligned_data[1]
+    aligned_uhm_dts: np.ndarray = aligned_data[2]
+    aligned_uhm_vs: np.ndarray = aligned_data[3]
+
+    diffs: np.ndarray = aligned_uhm_vs - aligned_opq_vs
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/io.py

@@ -1,0 +1,101 @@
+import dataclasses
+from dataclasses import dataclass
+from typing import Dict, List, Optional, Union
+
+@dataclass
+class Incident:
+    incident_id: int
+    event_id: int
+    box_id: str
+    start_timestamp_ms: int
+    end_timestamp_ms: int
+    measurement_type: str
+    deviation_from_nominal: float
+    classifications: List[str]
+
+    @staticmethod
+    def field_names() -> List[str]:
+        return list(map(lambda field: field.name, dataclasses.fields(Incident)))
+
+    @staticmethod
+    def from_doc(doc: Dict) -> 'Incident':
+        field_vals: List = list(map(lambda field_name: doc[field_name], Incident.field_names()))
+        return Incident(*field_vals)
+
+    @staticmethod
+    def projection() -> Dict[str, bool]:
+        doc: Dict[str, bool] = {
+            "_id": False
+        }
+
+        for field_name in Incident.field_names():
+            doc[field_name] = True
+
+        return doc
+
+
+@dataclass
+class TrendMetric:
+    min_v: float
+    max_v: float
+    average: float
+
+    @staticmethod
+    def from_doc(doc: Dict[str, Union[float, int]]) -> 'TrendMetric':
+        return TrendMetric(doc["min"],
+                           doc["max"],
+                           doc["average"])
+
+
+@dataclass
+class Trend:
+    box_id: str
+    timestamp_ms: int
+    thd: Optional[TrendMetric]
+    voltage: Optional[TrendMetric]
+    frequency: Optional[TrendMetric]
+
+    @staticmethod
+    def from_doc(doc: Dict) -> 'Trend':
+        return Trend(doc["box_id"],
+                     doc["timestamp_ms"],
+                     TrendMetric.from_doc(doc["thd"]) if "thd" in doc else None,
+                     TrendMetric.from_doc(doc["voltage"]) if "voltage" in doc else None,
+                     TrendMetric.from_doc(doc["frequency"]) if "frequency" in doc else None)
+
+
+class DataPoint:
+    def __init__(self,
+                 ts_s: int,
+                 actual_v: float,
+                 min_v: float,
+                 max_v: float,
+                 avg_v: float,
+                 stddev_v: float) -> None:
+        self.ts_s: int = ts_s
+        self.actual_v: float = actual_v
+        self.min_v: float = min_v
+        self.max_v: float = max_v
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/thd.py

@@ -1,0 +1,373 @@
+import datetime
+from typing import Dict, List, Tuple
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+import scipy.stats as stats
+
+import util
+import util.align_data as align
+import util.io as io
+
+THD_TYPES = ["AVG_VOLTAGE_THD"]
+
+
+def normal(mu: float, sigma: float, bins: np.ndarray, percent_density: float) -> np.ndarray:
+    return ((1 / (np.sqrt(2 * np.pi) * sigma)) *
+            np.exp(-0.5 * (1 / sigma * (bins - mu)) ** 2)) * percent_density
+
+def print_tex_table(opq_box_id: str,
+                    uhm_meter: str,
+                    mus: List[float],
+                    sigmas: List[float]) -> None:
+    mu_strs = list(map(lambda mu: f"{mu:.4f}", mus))
+    sigma_strs = list(map(lambda sigma: f"{sigma:.4f}", sigmas))
+
+    print(f"{opq_box_id} & {uhm_meter} & {' '.join(mu_strs)} & {' '.join(sigma_strs)} \\\\")
+
+def plot_thd(opq_start_ts_s: int,
+             opq_end_ts_s: int,
+             opq_box_id: str,
+             ground_truth_root: str,
+             uhm_sensor: str,
+             mongo_client: pymongo.MongoClient,
+             out_dir: str) -> str:
+    ground_truth_path: str = f"{ground_truth_root}/{uhm_sensor}/AVG_VOLTAGE_THD"
+    uhm_data_points: List[io.DataPoint] = io.parse_file(ground_truth_path)
+
+    db: pymongo.database.Database = mongo_client["opq"]
+    coll: pymongo.collection.Collection = db["trends"]
+    query: Dict = {
+        "box_id": opq_box_id,
+        "timestamp_ms": {"$gte": opq_start_ts_s * 1000,
+                         "$lte": opq_end_ts_s * 1000},
+        "thd": {"$exists": True}
+    }
+
+    projection: Dict[str, bool] = {
+        "_id": False,
+        "box_id": True,
+        "timestamp_ms": True,
+        "thd": True,
+    }
+
+    cursor: pymongo.cursor.Cursor = coll.find(query, projection=projection)
+    opq_trend_docs: List[Dict] = list(cursor)
[…]

### 2020-05-10 dissertation-christe/plotting/opq/ground_truth/util/voltage.py

@@ -1,0 +1,656 @@
+import datetime
+from typing import Dict, List, Tuple
+
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+import scipy.stats as stats
+
+import util
+import util.align_data as align
+import util.io as io
+
+
+def normal(mu: float, sigma: float, bins: np.ndarray, percent_density: float) -> np.ndarray:
+    return ((1 / (np.sqrt(2 * np.pi) * sigma)) *
+            np.exp(-0.5 * (1 / sigma * (bins - mu)) ** 2)) * percent_density
+
+def print_tex_table(opq_box_id: str,
+                    uhm_meter: str,
+                    mus: List[float],
+                    sigmas: List[float]) -> None:
+    mu_strs = list(map(lambda mu: f"{mu:.4f}", mus))
+    sigma_strs = list(map(lambda sigma: f"{sigma:.4f}", sigmas))
+
+    print(f"{opq_box_id} & {uhm_meter} & {' '.join(mu_strs)} & {' '.join(sigma_strs)}")
+
+
+def plot_voltage(opq_start_ts_s: int,
+                 opq_end_ts_s: int,
+                 opq_box_id: str,
+                 ground_truth_root: str,
+                 uhm_sensor: str,
+                 mongo_client: pymongo.MongoClient,
+                 out_dir: str) -> None:
+    ground_truth_path_vab: str = f"{ground_truth_root}/{uhm_sensor}/VAB"
+    uhm_data_points_vab: List[io.DataPoint] = io.parse_file(ground_truth_path_vab)
+
+    ground_truth_path_vbc: str = f"{ground_truth_root}/{uhm_sensor}/VBC"
+    uhm_data_points_vbc: List[io.DataPoint] = io.parse_file(ground_truth_path_vbc)
+
+    ground_truth_path_vca: str = f"{ground_truth_root}/{uhm_sensor}/VCA"
+    uhm_data_points_vca: List[io.DataPoint] = io.parse_file(ground_truth_path_vca)
+
+    db: pymongo.database.Database = mongo_client["opq"]
+    coll: pymongo.collection.Collection = db["trends"]
+    query: Dict = {
+        "box_id": opq_box_id,
+        "timestamp_ms": {"$gte": opq_start_ts_s * 1000,
+                         "$lte": opq_end_ts_s * 1000},
[…]

### 2020-05-10 dissertation-christe/plotting/opq/kml_parser.py

@@ -1,0 +1,50 @@
+from typing import List, Dict
+
+import fastkml
+
+
+box_to_location: Dict[str, str] = {
+    "1000": "POST 1",
+    "1001": "Hamilton",
+    "1002": "POST 2",
+    "1003": "LAVA Lab",
+    "1005": "Parking Structure Ph II",
+    "1006": "Frog 1",
+    "1007": "Frog 2",
+    "1008": "Mile's Office",
+    "1009": "Watanabe",
+    "1010": "Holmes",
+    "1021": "Marine Science Building",
+    "1022": "Ag. Engineering",
+    "1023": "Law Library",
+    "1024": "IT Building",
+    "1025": "Kennedy Theater"
+}
+
+def main():
+    with open("doc.kml", "rb") as fin:
+        kml: fastkml.kml.KML = fastkml.kml.KML()
+        kml.from_string(fin.read())
+
+
+        features = list(kml.features())
+        document: fastkml.kml.Document = features[0]
+        folder: fastkml.kml.Folder = list(document.features())[0]
+        placemarks: List[fastkml.kml.Placemark] = list(folder.features())
+
+        for placemark in placemarks:
+            name: str = placemark.name
+            box_id: str = name.split("-")[1].strip()
+            geometry = placemark.geometry
+            coords = geometry.coords[0]
+            lat: float = coords[0]
+            lng: float = coords[1]
+            loc: str = box_to_location[box_id]
+            # print(f"{box_id} & {loc} & {lat} & {lng} \\\\")
+            print(f"({lat}, {lng}),")
+
+
+
+
+if __name__ == "__main__":
+    main()

### 2020-05-10 dissertation-christe/plotting/opq/make_kml.py

@@ -1,0 +1,49 @@
+from typing import List, Tuple
+
+import lxml
+import fastkml
+import shapely
+import shapely.geometry
+
+
+def main():
+    box_locations: List[Tuple[float, float]] = [
+        (-157.816237, 21.297438),
+        (-157.816173, 21.300332),
+        (-157.816305, 21.297663),
+        (-157.816034748669, 21.29974948387028),
+        (-157.819234, 21.296042),
+        (-157.823122, 21.29805),
+        (-157.822819, 21.298046),
+        (-157.8137681609306, 21.30386147625208),
+        (-157.815817, 21.298351),
+        (-157.816104, 21.297011),
+        (-157.8156900462205, 21.29789461271471),
+        (-157.8154874938278, 21.30163608338939),
+        (-157.817361, 21.296328),
+        (-157.816451, 21.29886),
+        (-157.815225, 21.299282)
+    ]
+
+    kml: fastkml.kml.KML = fastkml.kml.KML()
+    ns: str = '{http://www.opengis.net/kml/2.2}'
+
+    document: fastkml.kml.Document = fastkml.kml.Document(ns)
+    kml.append(document)
+
+    folder: fastkml.kml.Folder = fastkml.kml.Folder(ns)
+
+    document.append(folder)
+
+    for lat_lng in box_locations:
+        placemark: fastkml.kml.Placemark = fastkml.kml.Placemark(ns)
+        placemark.geometry = shapely.geometry.Point(lat_lng[0], lat_lng[1])
+        # placemark.name = historical_device.id_uuid
+        folder.append(placemark)
+
+    with open("opq_devices.kml", "w") as fout:
+        fout.write(kml.to_string(prettyprint=True))
+
+
+if __name__ == "__main__":
+    main()

### 2020-05-10 dissertation-christe/plotting/opq/pl.py

@@ -1,0 +1,337 @@
+import datetime
+from dataclasses import dataclass
+from typing import Dict, List, Callable, Set, Tuple
+
+import bson
+import matplotlib.pyplot as plt
+import numpy as np
+import pymongo
+import pymongo.database
+from scipy import stats
+
+@dataclass
+class SeriesSpec:
+    values: List
+    dt_func: Callable
+    v_func: Callable
+
+
+def bin_dt_by_min(dt: datetime.datetime) -> datetime.datetime:
+    return datetime.datetime(dt.year, dt.month, dt.day, dt.hour, dt.minute, 0, 0, tzinfo=datetime.timezone.utc)
+
+
+# def bin_dt_by_day(dt: datetime.datetime) -> datetime.datetime:
+#     return datetime.datetime(dt.year, dt.month, dt.day, 0, 0, 0, 0)
+
+
+def intersect_lists(lists: List[List[datetime.datetime]]) -> Set[datetime.datetime]:
+    sets: List[Set] = list(map(set, lists))
+    return set.intersection(*sets)
+
+def align_data_multi(series: List[SeriesSpec]) -> List[Tuple[np.ndarray, np.ndarray]]:
+    all_dts: List[List[datetime.datetime]] = []
+    all_binned_dts: List[List[datetime.datetime]] = []
+    all_vals: List[List[float]] = []
+    already_seens: List[Set[datetime.datetime]] = []
+
+    for _ in range(len(series)):
+        already_seens.append(set())
+
+    for serie in series:
+        dts: List[datetime.datetime] = list(map(serie.dt_func, serie.values))
+        binned_dts: List[datetime.datetime] = list(map(bin_dt_by_min, dts))
+        vals: List = list(map(serie.v_func, serie.values))
+        all_dts.append(dts)
+        all_binned_dts.append(binned_dts)
+        all_vals.append(vals)
+
+    intersecting_dts: Set[datetime.datetime] = intersect_lists(all_binned_dts)
+
+    res: List[Tuple[np.ndarray, np.ndarray]] = []
+
+    for i, binned_dts in enumerate(all_binned_dts):
+        dts = []
+        vs = []
+        for j, binned_dt in enumerate(binned_dts):
+            if binned_dt in intersecting_dts and binned_dt not in already_seens[i]:
+                dts.append(binned_dt)
+                vs.append(all_vals[i][j])
+                already_seens[i].add(binned_dt)
+        res.append((np.array(dts), np.array(vs)))
+
+    return res
+
+
[…]

### 2020-05-10 dissertation-christe/plotting/opq/plot_laha.py

@@ -1,0 +1,388 @@
+from typing import *
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+seconds_in_day = 86400
+seconds_in_two_weeks = seconds_in_day * 14
+seconds_in_month = seconds_in_day * 30.4167
+seconds_in_year = seconds_in_month * 12
+seconds_in_2_years = seconds_in_year * 2
+
+
+class Data:
+    def __init__(self,
+                 time: int,
+                 total_samples: int,
+                 total_samples_b: int,
+                 total_measurements: int,
+                 total_measurements_b: int,
+                 total_orphaned_measurements: int,
+                 total_orphaned_measurements_b: int,
+                 total_event_measurements: int,
+                 total_event_measurements_b: int,
+                 total_incident_measurements: int,
+                 total_incident_measurements_b: int,
+                 total_phenomena_measurements: int,
+                 total_phenomena_measurements_b: int,
+                 total_trends: int,
+                 total_trends_b: int,
+                 total_orphaned_trends: int,
+                 total_orphaned_trends_b: int,
+                 total_event_trends: int,
+                 total_event_trends_b: int,
+                 total_incident_trends: int,
+                 total_incident_trends_b: int,
+                 total_phenomena_trends: int,
+                 total_phenomena_trends_b: int,
+                 total_events: int,
+                 total_events_b: int,
+                 total_orphaned_events: int,
+                 total_orphaned_events_b: int,
+                 total_incident_events: int,
+                 total_incident_events_b: int,
+                 total_phenomena_events: int,
+                 total_phenomena_events_b: int,
+                 total_incidents: int,
+                 total_incidents_b: int,
+                 total_phenomena_incidents: int,
+                 total_phenomena_incidents_b: int,
+                 total_phenomena: int,
+                 total_phenomena_b: int,
+                 total_laha_b: int,
+                 total_iml_b: int,
+                 total_aml_b: int,
+                 total_dl_b: int,
+                 total_il_b: int,
+                 total_pl_b: int):
+        self.time: int = time
[…]

### 2020-05-10 dissertation-christe/plotting/opq/results.py

@@ -1,0 +1,213 @@
+from collections import defaultdict
+import datetime
+from typing import Dict, List
+
+import pymongo
+import pymongo.database
+
+import matplotlib.pyplot as plt
+
+
+def global_events_summary():
+    mongo_client: pymongo.MongoClient = pymongo.MongoClient()
+    db: pymongo.database.Database = mongo_client["opq"]
+    events_coll: pymongo.collection.Collection = db["events"]
+
+    query: Dict = {
+        "target_event_start_timestamp_ms": {"$gt": 1569888000000}
+    }
+
+    projection: Dict[str, bool] = {"_id": False,
+                                   "target_event_start_timestamp_ms": True,
+                                   "boxes_received": True}
+
+    events_cursor: pymongo.cursor.Cursor = events_coll.find(query, projection=projection)
+    event_docs: List[Dict] = list(events_cursor)
+
+    num_boxes_received_to_num_events: Dict[int, int] = defaultdict(lambda: 0)
+    for event_doc in event_docs:
+        num_boxes_received: int = len(event_doc["boxes_received"])
+        num_boxes_received_to_num_events[num_boxes_received] += 1
+
+    total_incidents = []
+    for k, v in num_boxes_received_to_num_events.items():
+        total_incidents.append(v)
+        print(f"{v} & {k} \\")
+
+    print(sum(total_incidents))
+
+
+def incidents_summary():
+    mongo_client: pymongo.MongoClient = pymongo.MongoClient()
+    db: pymongo.database.Database = mongo_client["opq"]
+    incidents_coll: pymongo.collection.Collection = db["incidents"]
+
+    query: Dict = {
+        "start_timestamp_ms": {"$gt": 1569888000000}
+    }
+
+    projection: Dict[str, bool] = {"_id": False,
+                                   "start_timestamp_ms": True,
+                                   "classifications": True,
+                                   "incident_id": True}
+
+    incidents_cursor: pymongo.cursor.Cursor = incidents_coll.find(query, projection=projection)
+    incident_docs: List[Dict] = list(incidents_cursor)
+
+    incident_classification_to_cnt: Dict[str, int] = defaultdict(lambda: 0)
+
+    for incident_doc in incident_docs:
+        classification: str = incident_doc["classifications"][0]
[…]

### 2020-05-10 dissertation-christe/plotting/opq/similarity/io.py

@@ -1,0 +1,8 @@
+from typing import List
+import os
+
+
+def find_incident_directories(base_dir: str) -> List[str]:
+    return [f.path for f in os.scandir(base_dir) if f.is_dir()]
+
+

### 2020-05-10 dissertation-christe/plotting/opq/similarity/results.py

@@ -1,0 +1,20 @@
+def res(incident_type: str,
+        cluster_id: int,
+        num_incidents: int,
+        mean_duration_ms: float,
+        sr_hz: float,
+        sample_size_bytes: int,
+        mean_incident_size_bytes: int) -> None:
+    size_bytes: float = num_incidents * ((mean_duration_ms / 1_000.0) * sr_hz * sample_size_bytes + mean_incident_size_bytes)
+    size_gb: float = size_bytes / 1_000_000_000.0
+    print(f"{incident_type} {cluster_id} & {size_gb:.2f}")
+
+def main():
+    res("FSag", 1, 116769, 98.91, 12000, 2, 365)
+    res("FSag", 6, 35957, 809.18, 12000, 2, 365)
+    res("FSwell", 7, 78577, 138.63, 12000, 2, 365)
+    res("FSwell", 0, 95488, 158.29, 12000, 2, 365)
+
+
+if __name__ == "__main__":
+    main()

### 2020-05-10 dissertation-christe/plotting/opq/similarity/similarity.py

@@ -1,0 +1,12 @@
+from typing import List
+
+import io
+
+
+def main(base_dir: str):
+    incident_dirs: List[str] = io.find_incident_directories(base_dir)
+    print(len(incident_dirs))
+
+
+if __name__ == "__main__":
+    main("/home/opq/scrap/incident_data")

### 2020-05-10 dissertation-christe/plotting/opq/stats.py

@@ -1,0 +1,191 @@
+from typing import Dict, List, Set
+
+import numpy as np
+import pymongo
+import pymongo.database
+
+
+def event_stats(mongo_client: pymongo.MongoClient) -> List[Dict]:
+    db: pymongo.database.Database = mongo_client["opq"]
+    box_events_coll: pymongo.collection.Collection = db["box_events"]
+    events_coll: pymongo.collection.Collection = db["events"]
+
+    events_query = {"target_event_start_timestamp_ms": {"$gt": 0},
+                    "target_event_end_timestamp_ms": {"$gt": 0}}
+
+    events_projection = {"_id": False,
+                         "target_event_start_timestamp_ms": True,
+                         "target_event_end_timestamp_ms": True}
+
+    events: List[Dict] = events_coll.find(events_query, projection=events_projection)
+    durations_events_ms: np.ndarray = np.array(
+        list(map(lambda doc: doc["target_event_end_timestamp_ms"] - doc["target_event_start_timestamp_ms"], events)))
+    durations_events_s: np.ndarray = durations_events_ms / 1_000.0
+
+    data_duration_events_s = durations_events_s.sum()
+
+    box_events_query = {"event_start_timestamp_ms": {"$gt": 0},
+                        "event_end_timestamp_ms": {"$gt": 0}}
+    box_events_projection = {"_id": False,
+                             "event_id": True,
+                             "event_start_timestamp_ms": True,
+                             "event_end_timestamp_ms": True}
+
+    box_events: pymongo.cursor.Cursor = box_events_coll.find(box_events_query, projection=box_events_projection)
+    box_events: List[Dict] = list(box_events)
+
+    min_ts_ms: int = min(list(map(lambda doc: doc["event_start_timestamp_ms"], box_events)))
+    max_ts_ms: int = max(list(map(lambda doc: doc["event_end_timestamp_ms"], box_events)))
+
+    durations_box_events_ms: np.ndarray = np.array(
+        list(map(lambda doc: doc["event_end_timestamp_ms"] - doc["event_start_timestamp_ms"], box_events)))
+    durations_box_events_s: np.ndarray = durations_box_events_ms / 1000.0
+    data_stored: np.ndarray = 12_000.0 * 2 * durations_box_events_s
+
+    total_duration_ms: int = max_ts_ms - min_ts_ms
[…]

### 2020-05-10 dissertation-christe/plotting/plots.py

@@ -1,0 +1,752 @@
+import typing
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+import laha.iml as iml
+import laha.dl as dl
+
+S_IN_DAY = 86_400
+S_IN_YEAR = 31_540_000
+
+
+def plot_iml_level_opq():
+    plt.figure(figsize=(12, 5))
+    sample_size_bytes = 2
+    sample_rate_hz = 12_000
+    x_values = np.arange(S_IN_YEAR, step=S_IN_DAY)
+    y_values = x_values * sample_size_bytes * sample_rate_hz
+
+    plt.plot(x_values, y_values)
+
+    plt.title("IML Size (Lokahi) Sample Size=4, SR=12000, Len=1yr")
+    plt.xlabel("Time (S)")
+    plt.ylabel("Bytes")
+
+    plt.savefig("../src/figures/plot_iml_level_opq.png")
+    plt.show()
+
+
+def plot_iml_level_lokahi():
+    plt.figure(figsize=(12, 5))
+    sample_size_bytes = 4
+    sample_rates_hz = [80, 800, 8000]
+    x_values = np.arange(S_IN_YEAR, step=S_IN_DAY)
+    for sample_rate_hz in sample_rates_hz:
+        y_values = x_values * sample_size_bytes * sample_rate_hz
+        plt.plot(x_values, y_values, label="%d Hz" % sample_rate_hz)
+
+    plt.title("IML Size (Lokahi) Sample Size=4, SR=[80,800,8000], Len=1yr")
+    plt.xlabel("Time (S)")
+    plt.ylabel("Bytes")
+
+    plt.legend()
+    plt.savefig("../src/figures/plot_iml_level_lokahi.png")
+    plt.show()
+
+
+def plot_iml_level_no_opt_var_sample_size(sample_sizes_bytes: typing.List[int],
+                                          sample_rate_hz: int,
+                                          window_length_s: int):
+    x_values = np.arange(window_length_s, step=S_IN_DAY)
+
+    plt.figure(figsize=(12, 5))
+
+    for sample_size_bytes in sample_sizes_bytes:
+        y_values = x_values * sample_size_bytes * sample_rate_hz
+        plt.plot(x_values, y_values, label="%d bytes" % sample_size_bytes)
+
+    plt.title("IML No Opt: Sample Size Bytes=%s, Sample Rate Hz=%d, Window Length S=%d" % (
+        str(sample_sizes_bytes),
+        sample_rate_hz,
+        window_length_s
+    ))
[…]

### 2020-05-10 dissertation-christe/plotting/scrap.py

@@ -1,0 +1,19 @@
+from typing import List
+
+import numpy as np
+
+def sum_series(series: np.ndarray) -> np.ndarray:
+    result: List[int] = [series[0]]
+    for i in range(1, len(series)):
+        result.append(result[i - 1] + series[i])
+
+    return np.array(result)
+
+
+def main():
+    a = np.array([1, 2, 3, 4, 5])
+    print(sum_series(a))
+
+
+if __name__ == "__main__":
+    main()

### 2020-05-10 dissertation-christe/sim/src/config.rs

@@ -1,0 +1,22 @@
+pub struct Config {
+    pub ticks: usize,
+    pub percent_event_duration: f64,
+    pub percent_event_to_incident: f64,
+    pub mean_event_len: usize,
+    pub samples_ttl: usize,
+    pub measurements_ttl: usize,
+    pub trends_ttl: usize,
+    pub events_ttl: usize,
+    pub incidents_ttl: usize,
+    pub phenomena_ttl: usize,
+    pub num_sensors: usize,
+    pub bytes_per_sample: usize,
+    pub bytes_per_measurement: usize,
+    pub bytes_per_trend: usize,
+    pub bytes_per_event: usize,
+    pub bytes_per_incident: usize,
+    pub bytes_per_phenomena: usize,
+    pub print_info_every_n_ticks: usize,
+    pub write_info_every_n_ticks: usize,
+    pub out_file: String,
+}

### 2020-05-10 dissertation-christe/sim/src/constants.rs

@@ -1,0 +1,46 @@
+pub const SAMPLES_PER_SECOND: usize = 12_000;
+pub const BYTES_PER_SAMPLE: usize = 2;
+
+pub const SECONDS_PER_MINUTE: usize = 60;
+pub const MINUTES_PER_HOUR: usize = 60;
+pub const HOURS_PER_DAY: usize = 24;
+pub const DAYS_PER_WEEK: usize = 7;
+pub const DAYS_PER_MONTH: usize = 30;
+pub const MONTHS_PER_YEAR: usize = 12;
+
+pub const SECONDS_PER_FIFTEEN_MINUTES: usize = SECONDS_PER_MINUTE * 15;
+pub const SECONDS_PER_HOUR: usize = SECONDS_PER_MINUTE * MINUTES_PER_HOUR;
+pub const SECONDS_PER_DAY: usize = SECONDS_PER_HOUR * HOURS_PER_DAY;
+pub const SECONDS_PER_WEEK: usize = SECONDS_PER_DAY * DAYS_PER_WEEK;
+pub const SECONDS_PER_TWO_WEEKS: usize = SECONDS_PER_WEEK * 2;
+pub const SECONDS_PER_MONTH: usize = SECONDS_PER_DAY * DAYS_PER_MONTH;
+pub const SECONDS_PER_YEAR: usize = SECONDS_PER_MONTH * MONTHS_PER_YEAR;
+
+pub const DEFAULT_MEASUREMENT_TTL: usize = SECONDS_PER_DAY;
+pub const DEFAULT_TRENDS_TTL: usize = SECONDS_PER_TWO_WEEKS;
+pub const DEFAULT_EVENTS_TTL: usize = SECONDS_PER_MONTH;
+pub const DEFAULT_INCIDENTS_TTL: usize = SECONDS_PER_YEAR;
+pub const DEFAULT_PHENOMENA_TTL: usize = SECONDS_PER_YEAR * 2;
+
+pub const ESTIMATED_EVENTS_PER_SECOND: f64 = 0.009610162129845402;
+pub const ESTIMATED_INCIDENTS_PER_SECOND: f64 = 0.018677002968930212;
+
+// New
+pub const ESTIMATED_PHENOMENA_PER_SECOND: f64 = 0.00034763162434612893;
+
+pub const ESTIMATED_PERCENT_DATA_DURATION: f64 = 0.09301198295969353;
[…]

### 2020-05-10 dissertation-christe/sim/src/main.rs

@@ -1,0 +1,32 @@
+pub mod config;
+pub mod constants;
+pub mod sim;
+pub mod storage;
+
+fn main() {
+    let conf = config::Config {
+        ticks: constants::SECONDS_PER_YEAR * 3,
+        percent_event_duration: constants::ESTIMATED_PERCENT_DATA_DURATION,
+        percent_event_to_incident: constants::ESTIMATED_PERCENT_EVENT_TO_INCIDENT,
+        mean_event_len: constants::ESTIMATED_EVENT_LEN_S,
+        samples_ttl: constants::SECONDS_PER_FIFTEEN_MINUTES,
+        measurements_ttl: constants::DEFAULT_MEASUREMENT_TTL,
+        trends_ttl: constants::DEFAULT_TRENDS_TTL,
+        events_ttl: constants::DEFAULT_EVENTS_TTL,
+        incidents_ttl: constants::DEFAULT_INCIDENTS_TTL,
+        phenomena_ttl: constants::DEFAULT_PHENOMENA_TTL,
+        num_sensors: 1,
+        bytes_per_sample: constants::ESTIMATED_BYTES_PER_META_SAMPLE,
+        bytes_per_measurement: constants::ESTIMATED_BYTES_PER_MEASUREMENT,
+        bytes_per_trend: constants::ESTIMATED_BYTES_PER_TREND,
+        bytes_per_event: constants::ESTIMATED_BYTES_PER_EVENT,
+        bytes_per_incident: constants::ESTIMATED_BYTES_PER_INCIDENT,
+        bytes_per_phenomena: constants::ESTIMATED_BYTES_PER_PHENOMENA,
+        print_info_every_n_ticks: 500_000,
+        write_info_every_n_ticks: 3600,
+        out_file: "/home/opq/scrap/sim_data_opq.txt".to_string(),
+        //        out_file: "/home/opq/Documents/anthony/dissertation/plotting/opq/sim_data.txt".to_string(),
+    };
+    let mut simulation = sim::Simulation::new(conf);
+    simulation.run_simulation();
+}

### 2020-05-10 dissertation-christe/sim/src/sim.rs

@@ -1,0 +1,882 @@
+use crate::config::Config;
+use crate::constants;
+use crate::storage::{Storage, StorageItem};
+use crate::{config, storage};
+use rand;
+use rand::prelude::ThreadRng;
+use rand::Rng;
+use std::io::{BufWriter, Write};
+
+#[inline]
+fn percent_chance(chance: f64, rng: &mut ThreadRng) -> bool {
+    rng.gen_range(0.0, 1.0) < chance
+}
+
+#[inline]
+pub fn fmt_size_mb(total_bytes: usize) -> String {
+    format!("{:.*}MB", 2, total_bytes as f64 / 1_000_000.0)
+}
+
+#[inline]
+pub fn fmt_percent(percent: f64) -> String {
+    format!("{:.*}%", 2, percent * 100.0)
+}
+
+pub struct Simulation {
+    rng: ThreadRng,
+    storage: Storage,
+    buf_writer: BufWriter<std::fs::File>,
+    conf: Config,
+
+    total_storage_items: usize,
+    total_samples: usize,
+    total_orphaned_samples: usize,
+    total_event_samples: usize,
+    total_incident_samples: usize,
+    total_phenomena_samples: usize,
+    total_measurements: usize,
+    total_orphaned_measurements: usize,
+    total_event_measurements: usize,
+    total_phenomena_measurements: usize,
+    total_incident_measurements: usize,
+    total_trends: usize,
+    total_orphaned_trends: usize,
+    total_event_trends: usize,
+    total_incident_trends: usize,
+    total_phenomena_trends: usize,
+    total_events: usize,
+    total_orphaned_events: usize,
+    total_incident_events: usize,
+    total_phenomena_events: usize,
+    total_incidents: usize,
+    total_orphaned_incidents: usize,
+    total_phenomena_incidents: usize,
+    total_phenomena: usize,
+}
+
+impl Simulation {
+    pub fn new(conf: Config) -> Simulation {
+        let file = std::fs::File::create(conf.out_file.clone()).unwrap();
+        let buf_writer = BufWriter::new(file);
+        Simulation {
+            rng: rand::thread_rng(),
+            storage: Storage::new(),
+            buf_writer,
+            conf,
+            total_storage_items: 0,
+            total_samples: 0,
+            total_orphaned_samples: 0,
+            total_event_samples: 0,
+            total_incident_samples: 0,
+            total_phenomena_samples: 0,
+            total_measurements: 0,
+            total_orphaned_measurements: 0,
[…]

### 2020-05-10 dissertation-christe/sim/src/storage.rs

@@ -1,0 +1,298 @@
+use crate::constants;
+use std::cmp::Ordering;
+
+#[inline]
+fn find_index<T: Ord>(v: &Vec<T>, val: &T) -> usize {
+    match v.binary_search(val) {
+        Ok(idx) => idx + 1,
+        Err(idx) => idx,
+    }
+}
+
+#[derive(PartialEq, Clone)]
+pub enum StorageType {
+    Sample(usize),
+    Measurement(usize),
+    Trend(usize),
+    Event(usize),
+    Incident(usize),
+    Phenomena(usize),
+}
+
+pub struct StorageItemStatistic {
+    pub items: usize,
+    pub percent_total: f64,
+    pub total_bytes: usize,
+}
+
+impl StorageItemStatistic {
+    pub fn fmt_percent(&self) -> String {
+        format!("{:.*}%", 2, self.percent_total)
+    }
+
+    pub fn fmt_size_mb(&self) -> String {
+        format!("{:.*}MB", 2, self.total_bytes as f64 / 1_000_000.0)
+    }
+}
+
+pub struct StorageItem {
+    pub storage_type: StorageType,
+    pub ts: usize,
+    pub ttl: usize,
+    pub is_event: bool,
+    pub is_incident: bool,
+    pub is_phenomena: bool,
+}
+
+impl StorageItem {
+    pub fn new(
+        storage_type: StorageType,
+        ts: usize,
+        ttl: usize,
+        is_event: Option<bool>,
+        is_incident: Option<bool>,
+        is_phenomena: Option<bool>,
+    ) -> StorageItem {
+        StorageItem {
+            storage_type,
+            ts,
+            ttl,
+            is_event: is_event.unwrap_or(false),
+            is_incident: is_incident.unwrap_or(false),
+            is_phenomena: is_phenomena.unwrap_or(false),
+        }
+    }
+
+    pub fn from_ttl(ttl: usize) -> StorageItem {
+        StorageItem::new_measurement(0, ttl, None, None, None)
+    }
+
+    pub fn new_sample(
+        ts: usize,
+        ttl: usize,
+        is_event: Option<bool>,
+        is_incident: Option<bool>,
+        is_phenomena: Option<bool>,
+    ) -> StorageItem {
+        StorageItem::new(
+            StorageType::Sample(constants::ESTIMATED_BYTES_PER_META_SAMPLE),
+            ts,
+            ttl,
+            is_event,
+            is_incident,
[…]

### 2020-05-10 dissertation-christe/sim_lokahi/src/config.rs

@@ -1,0 +1,21 @@
+pub struct Config {
+    pub ticks: usize,
+    pub percent_event_duration: f64,
+    pub percent_event_to_incident: f64,
+    pub mean_event_len: usize,
+    pub samples_ttl: usize,
+    pub measurements_ttl: usize,
+    pub trends_ttl: usize,
+    pub events_ttl: usize,
+    pub incidents_ttl: usize,
+    pub phenomena_ttl: usize,
+    pub num_sensors: usize,
+    pub bytes_per_sample: usize,
+    pub bytes_per_measurement: usize,
+    pub bytes_per_trend: usize,
+    pub bytes_per_event: usize,
+    pub bytes_per_incident: usize,
+    pub print_info_every_n_ticks: usize,
+    pub write_info_every_n_ticks: usize,
+    pub out_file: String,
+}

### 2020-05-10 dissertation-christe/sim_lokahi/src/constants.rs

@@ -1,0 +1,62 @@
+pub const SAMPLES_PER_SECOND_80: usize = 80;
+pub const SAMPLES_PER_SECOND_800: usize = 800;
+pub const SAMPLES_PER_SECOND_8000: usize = 8000;
+pub const BYTES_PER_SAMPLE: usize = 4;
+pub const META_SAMPLE_80_LEN: usize = 51;
+pub const META_SAMPLE_800_LEN: usize = 41;
+pub const META_SAMPLE_8000_LEN: usize = 33;
+
+pub const SECONDS_PER_MINUTE: usize = 60;
+pub const MINUTES_PER_HOUR: usize = 60;
+pub const HOURS_PER_DAY: usize = 24;
+pub const DAYS_PER_WEEK: usize = 7;
+pub const DAYS_PER_MONTH: usize = 30;
+pub const MONTHS_PER_YEAR: usize = 12;
+
+pub const SECONDS_PER_FIFTEEN_MINUTES: usize = SECONDS_PER_MINUTE * 15;
+pub const SECONDS_PER_HOUR: usize = SECONDS_PER_MINUTE * MINUTES_PER_HOUR;
+pub const SECONDS_PER_DAY: usize = SECONDS_PER_HOUR * HOURS_PER_DAY;
+pub const SECONDS_PER_WEEK: usize = SECONDS_PER_DAY * DAYS_PER_WEEK;
+pub const SECONDS_PER_TWO_WEEKS: usize = SECONDS_PER_WEEK * 2;
+pub const SECONDS_PER_MONTH: usize = SECONDS_PER_DAY * DAYS_PER_MONTH;
+pub const SECONDS_PER_YEAR: usize = SECONDS_PER_MONTH * MONTHS_PER_YEAR;
+
+pub const DEFAULT_MEASUREMENT_TTL: usize = SECONDS_PER_DAY;
+pub const DEFAULT_TRENDS_TTL: usize = SECONDS_PER_TWO_WEEKS;
+pub const DEFAULT_EVENTS_TTL: usize = SECONDS_PER_MONTH;
+pub const DEFAULT_INCIDENTS_TTL: usize = SECONDS_PER_YEAR;
+pub const DEFAULT_PHENOMENA_TTL: usize = SECONDS_PER_YEAR * 2;
+
+pub const ESTIMATED_EVENTS_PER_SECOND: f64 = 0.00003116867161164198;
[…]

### 2020-05-10 dissertation-christe/sim_lokahi/src/main.rs

@@ -1,0 +1,32 @@
+pub mod config;
+pub mod constants;
+pub mod sim;
+pub mod storage;
+
+fn main() {
+    let conf = config::Config {
+        ticks: constants::SECONDS_PER_YEAR * 3,
+        //        ticks: constants::SECONDS_PER_DAY,
+        percent_event_duration: constants::ESTIMATED_PERCENT_EVENT_DATA_DURATION,
+        percent_event_to_incident: 0.0,
+        mean_event_len: constants::ESTIMATED_EVENT_LEN_S,
+        samples_ttl: constants::SECONDS_PER_FIFTEEN_MINUTES,
+        measurements_ttl: constants::DEFAULT_MEASUREMENT_TTL,
+        trends_ttl: constants::DEFAULT_TRENDS_TTL,
+        events_ttl: constants::DEFAULT_EVENTS_TTL,
+        incidents_ttl: constants::DEFAULT_INCIDENTS_TTL,
+        phenomena_ttl: constants::DEFAULT_PHENOMENA_TTL,
+        num_sensors: 1,
+        bytes_per_sample: constants::ESTIMATED_BYTES_PER_META_SAMPLE_8000,
+        bytes_per_measurement: constants::ESTIMATED_BYTES_PER_MEASUREMENT,
+        bytes_per_trend: constants::ESTIMATED_BYTES_PER_TREND,
+        bytes_per_event: constants::ESTIMATED_BYTES_PER_EVENT_8000,
+        bytes_per_incident: constants::ESTIMATED_BYTES_PER_INCIDENT_8000,
+        print_info_every_n_ticks: 500_000,
+        write_info_every_n_ticks: 3600,
+        //        write_info_every_n_ticks: 300,
+        out_file: "/home/opq/scrap/sim_data_8000_lokahi.txt".to_string(),
+    };
+    let mut simulation = sim::Simulation::new(conf);
+    simulation.run_simulation();
+}

### 2020-05-10 dissertation-christe/sim_lokahi/src/sim.rs

@@ -1,0 +1,832 @@
+use crate::config::Config;
+use crate::constants;
+use crate::storage::{Storage, StorageItem};
+use crate::{config, storage};
+use rand;
+use rand::prelude::ThreadRng;
+use rand::Rng;
+use std::io::{BufWriter, Write};
+
+#[inline]
+fn percent_chance(chance: f64, rng: &mut ThreadRng) -> bool {
+    rng.gen_range(0.0, 1.0) < chance
+}
+
+#[inline]
+pub fn fmt_size_mb(total_bytes: usize) -> String {
+    format!("{:.*}MB", 2, total_bytes as f64 / 1_000_000.0)
+}
+
+#[inline]
+pub fn fmt_percent(percent: f64) -> String {
+    format!("{:.*}%", 2, percent * 100.0)
+}
+
+pub struct Simulation {
+    rng: ThreadRng,
+    storage: Storage,
+    buf_writer: BufWriter<std::fs::File>,
+    conf: Config,
+
+    total_storage_items: usize,
+    total_samples: usize,
+    total_orphaned_samples: usize,
+    total_event_samples: usize,
+    total_incident_samples: usize,
+    total_phenomena_samples: usize,
+    total_measurements: usize,
+    total_orphaned_measurements: usize,
+    total_event_measurements: usize,
+    total_phenomena_measurements: usize,
+    total_incident_measurements: usize,
+    total_trends: usize,
+    total_orphaned_trends: usize,
+    total_event_trends: usize,
+    total_incident_trends: usize,
+    total_phenomena_trends: usize,
+    total_events: usize,
+    total_orphaned_events: usize,
+    total_incident_events: usize,
+    total_phenomena_events: usize,
+    total_incidents: usize,
+    total_orphaned_incidents: usize,
+    total_phenomena_incidents: usize,
+    total_phenomena: usize,
+}
+
+impl Simulation {
+    pub fn new(conf: Config) -> Simulation {
+        let file = std::fs::File::create(conf.out_file.clone()).unwrap();
+        let buf_writer = BufWriter::new(file);
+        Simulation {
+            rng: rand::thread_rng(),
+            storage: Storage::new(),
+            buf_writer,
+            conf,
+            total_storage_items: 0,
+            total_samples: 0,
+            total_orphaned_samples: 0,
+            total_event_samples: 0,
+            total_incident_samples: 0,
+            total_phenomena_samples: 0,
+            total_measurements: 0,
+            total_orphaned_measurements: 0,
[…]

### 2020-05-10 dissertation-christe/sim_lokahi/src/storage.rs

@@ -1,0 +1,298 @@
+use crate::constants;
+use std::cmp::Ordering;
+
+#[inline]
+fn find_index<T: Ord>(v: &Vec<T>, val: &T) -> usize {
+    match v.binary_search(val) {
+        Ok(idx) => idx + 1,
+        Err(idx) => idx,
+    }
+}
+
+#[derive(PartialEq, Clone)]
+pub enum StorageType {
+    MetaSample(usize),
+    Measurement(usize),
+    Trend(usize),
+    Event(usize),
+    Incident(usize),
+    Phenomena(usize),
+}
+
+pub struct StorageItemStatistic {
+    pub items: usize,
+    pub percent_total: f64,
+    pub total_bytes: usize,
+}
+
+impl StorageItemStatistic {
+    pub fn fmt_percent(&self) -> String {
+        format!("{:.*}%", 2, self.percent_total)
+    }
+
+    pub fn fmt_size_mb(&self) -> String {
+        format!("{:.*}MB", 2, self.total_bytes as f64 / 1_000_000.0)
+    }
+}
+
+pub struct StorageItem {
+    pub storage_type: StorageType,
+    pub ts: usize,
+    pub ttl: usize,
+    pub is_event: bool,
+    pub is_incident: bool,
+    pub is_phenomena: bool,
+}
+
+impl StorageItem {
+    pub fn new(
+        storage_type: StorageType,
+        ts: usize,
+        ttl: usize,
+        is_event: Option<bool>,
+        is_incident: Option<bool>,
+        is_phenomena: Option<bool>,
+    ) -> StorageItem {
+        StorageItem {
+            storage_type,
+            ts,
+            ttl,
+            is_event: is_event.unwrap_or(false),
+            is_incident: is_incident.unwrap_or(false),
+            is_phenomena: is_phenomena.unwrap_or(false),
+        }
+    }
+
+    pub fn from_ttl(ttl: usize) -> StorageItem {
+        StorageItem::new_measurement(0, ttl, None, None, None)
+    }
+
+    pub fn new_sample(
+        ts: usize,
+        ttl: usize,
+        is_event: Option<bool>,
+        is_incident: Option<bool>,
+        is_phenomena: Option<bool>,
+    ) -> StorageItem {
+        StorageItem::new(
+            StorageType::MetaSample(constants::ESTIMATED_BYTES_PER_META_SAMPLE_8000),
+            ts,
+            ttl,
+            is_event,
+            is_incident,
[…]

### 2020-05-10 dissertation-christe/src/figures/make_dots.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+for dot_file in *.dot 
+do
+	png_file="${dot_file%.dot}.png"
+	dot -Tpng ${dot_file} > ${png_file}
+	echo "Converted ${dot_file} => ${png_file}"
+done

### 2020-05-10 dissertation-negrashov/data/box_evaluation/trans/plot_waveform.py

@@ -1,0 +1,36 @@
+import sys
+import matplotlib.pyplot as plt
+import numpy as np
+if(len(sys.argv) < 2):
+    print("need an argument")
+    sys.exit()
+with open(sys.argv[1]) as f:
+    content = f.readlines()
+data = [];
+
+i = 0
+
+for line in content:
+    try:
+        data.append(float(line))
+    except ValueError:
+        print("not a float " + line)
+        continue
+    if len(data) == 200:
+        break
+print(len(data))
+
+timestep = 1/60.0/200.0
+
+x = []
+for i in range(0,len(data)):
+	x.append(i*timestep*1000)
+
+fig, axs = plt.subplots(1, 1, sharey=True, tight_layout=True)
+
+# We can set the number of bins with the `bins` kwarg
+plt.xlabel("Time(ms)")
+plt.ylabel("Voltage")
+
+axs.plot(x,data);
+plt.show()

### 2020-05-12 article.tex
checkpoint

@@
 %=================================================================
-\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+% \documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[preprints,article,accept,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 \section{Introduction}
-The introduction should briefly place the study in a broad context and highlight why it is important.
+
+%{\em The introduction should briefly place the study in a broad context and highlight why it is important.
 It should define the purpose of the work and its significance.
@@
 Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-And now citing a book reference \cite{ref-book}.
-Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.
+Finally, briefly mention the main aim of the work and highlight the principal conclusions.}
+
+Power quality is not currently a concern for most people in developed nations.
+Just like most people in developed nations assume that their tap water is of adequate quality to drink, most also assume that their electricity is of adequate quality to power their homes and appliances without causing harm.
+And, in both cases, most people assume that public utilities will monitor and correct any quality problems if they occur.
+
+Successfully maintaining adequate power quality and providing sufficient amounts of it to meet the rising needs of consumers has been a triumph of electrical utilities for over 100 years.
+In recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing public concern and interest.
+First, there is a global need to shift to renewable energy sources.
+Second, modern consumer electronics place more stringent demands on power quality.
+Third, effective policy making in this modern context requires better public access to power quality data.
+
+{\em We need more renewable energy.} There is now a global movement away from centralized, fossil-fuel based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar.
+But the economic, environmental, and political advantages of renewable energy comes with significant new technical challenges.
+Wind and solar are intermittent (for example, solar energy cannot be harvested at night) and unpredictable (for example, wind and solar energy fluctuate based upon cloud cover and wind speed).
+In addition, renewable energy generation can be distributed throughout the grid (such as in the case of residential rooftop photovoltaic (PV) systems).
+
+One impact of adding renewable energy generation to an electrical grid is that maintaining adequate power quality is much more challenging.
+This problem is inversely proportional to the size of the electrical grid.
+For example, in Hawaii on the island of Oahu, the public demand for rooftop solar exceeds Hawaiian Electric Company's ability to utilize it and maintain adequate power quality.
+As a result, in many neighborhoods, new rooftop solar installation is prohibited because Hawaiian Electric fears its impact on not only local power quality but global grid stability \cite{TBD}.
+
+{\em Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality.
+Only a few decades ago, computers were a rare presence outside of labs and large institutions.
+Today, computers are everywhere: embedded in phones, washers, refrigerators, thermostats, and so forth.
+These electronic devices not only have higher power quality requirements, but some of them actually introduce power quality problems in the form of harmonic distortion.
+Poor power quality can result in electronic devices failing unpredictably, and/or decreasing their lifespan.
+
+{\em Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers.
+For example, in Hawaii, utilities are required to make a "best effort" to provide non-harmful voltage and frequency values, but are only required to publicly report on power outages of more than 3 minutes.
+There is no requirement for utilities to report potentially harmful forms of voltage, frequency, THD, or transients.
+Indeed, in Hawaii, there is not yet infrastructure in place that would enable utilities to collect that information, even if they were asked to report on it.
+For more details, see Hawaiian Electric Tariff Rule No.
+2 Character of Service and Hawaiian Electric Tariff Rule No.
+16 Interruption of Service.
+
+OPQ intends to provide an unbiased, independent, third party source of accurate power quality data.
+This can be used by consumers to better understand the performance of their public utilities, by researchers to devise improvements to grid control, and by public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
+The OPQ Project has produced a suite of innovations that can help address each of these issues.
+
+{\em OPQ Box.} First, our power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75.
+This is from 10 to 100 times less expensive than commercial power quality monitors.
+This means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building.
+Similarly, instead of monitoring power quality at the substation level of the grid, OPQ Boxes make it feasible to deploy dozens or hundreds into a community to obtain fine-grained data about the impact of solar or other renewables as directly perceived by the customer.
+
+{\em Cloud native.} Second, our architecture is "cloud native", which means that OPQ Boxes are not designed for "stand alone", autonomous use.
+Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and analyzed.
+This creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes.
+It means our cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.
+This enables OPQ installations to be simultaneously responsive and scalable.
+
+{\em Extensibility.} Third, all OPQ software subsystems are designed for extensibility and interoperation.
+Our middleware components (Mauka and Makai) provide a plugin architecture.
+Our visualization component (View) is built from modular UI elements using React.
+We are also developing API endpoints for interoperation with other software systems.
+This means that OPQ is not a monolithic, closed system with a frozen feature set, but rather a design environment for experimentation with advanced classification and analysis of power quality data.
+
+{\em Open Source.} Fourth, our software and hardware designs and implementations are made available using open source licenses.
+This means that organizations choosing OPQ do not face the business risk of committing resources to a single vendor with proprietary hardware and software.
+Our goal is to facilitate the creation of a community of researchers and industry practitioners to replicate, extend, and apply the insights gained from the OPQ system.
+
+Finally, while we believe OPQ offers a compelling combination of capabilities, we want to be clear that all designs involve trade-offs.
+As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs.
+In some cases, OPQ may not be the appropriate choice.
+In other cases, a hybrid solution consisting of OPQ in combination with other technologies for power monitoring may be most appropriate.
+Our goal in this paper is to describe OPQ well enough for organizations to make a thoughtful decision about the use of our technology.
 

### 2020-05-13 01-abstract.tex
add architecture section

--- 2020-05-10 article.tex
@@
-%  LaTeX support: latex@mdpi.com
-%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
-
-%=================================================================
-\documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
-
-%=================================================================
-\firstpage{1}
-\makeatletter
-\setcounter{page}{\@firstpage}
-\makeatother
-\pubvolume{xx}
-\issuenum{1}
-\articlenumber{5}
-\pubyear{2020}
-\copyrightyear{2020}
-%\externaleditor{Academic Editor: name}
-\history{Received: date; Accepted: date; Published: date}
-%\updates{yes} % If there is an update available, un-comment this line
-
-%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
-%\continuouspages{yes}
-
-%=================================================================
-% Add packages and commands here.
-The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
-
-%=================================================================
-%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
-%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
-
-%=================================================================
-% Full title of the paper (Capitalized)
-\Title{Design, Implementation, and Evaluation of Open Power Quality}
-
-% Authors, for the paper (add full first names)
-\Author{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}
-
-% Authors, for metadata in PDF
-\AuthorNames{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}
-
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-Department of Information and Computer Sciences\\
-University of Hawaii at Manoa\\
-Honolulu, HI 96822\\
-johnson@hawaii.edu}
-
 % Abstract (single paragraph, 200 words max, background, methods, results, conclusions. Currently 193 words.)
@@
 \keyword{Power Quality, Open Source, Renewable Energy, Grid Stability}
-
-\begin{document}
-
-%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
-
-\section{Introduction}
-The introduction should briefly place the study in a broad context and highlight why it is important.
-It should define the purpose of the work and its significance.
-The current state of the research field should be reviewed carefully and key publications cited.
-Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-And now citing a book reference \cite{ref-book}.
-Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\begin{quote}
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection}
-\unskip
-\subsubsection{Subsubsection}
-
-Bulleted lists look like this:
-\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-\item	First bullet
-\item	Second bullet
-\item	Third bullet
-\end{itemize}
-
-Numbered lists can be added as follows:
-\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item	First item
-\item	Second item
-\item	Third item
-\end{enumerate}
-
-The text continues here.
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2 cm]{Definitions/logo-mdpi}
-\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
-\end{figure}
-
-Text
-
-Text
-
-\begin{table}[H]
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-entry 1		& data			& data\\
-entry 2		& data			& data\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Text
-
-Text
-
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\textwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\textwidth}{1pt}
-%\end{listing}
-
-\subsection{Formatting of Mathematical Components}
-
-This is an example of an equation:
-
-\begin{equation}
-a + b = c
-\end{equation}
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work reported.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
-
-\noindent
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & linear dichroism
-\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendix
-\section{}
-\unskip
-\subsection{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
-For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: external bibliography
-%=====================================
-%\externalbibliography{yes}
-%\bibliography{your_external_BibTeX_file}
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-\begin{thebibliography}{999}
-% Reference 1
-\bibitem[Author1(year)]{ref-journal}
-Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% Reference 2
-\bibitem[Author2(year)]{ref-book}
-Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-\end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}

### 2020-05-13 02-intro.tex
add architecture section

--- 2020-05-12 article.tex
@@
-%  LaTeX support: latex@mdpi.com
-%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
-
-%=================================================================
-% \documentclass[energies,article,submit,moreauthors,pdftex]{Definitions/mdpi}
-\documentclass[preprints,article,accept,moreauthors,pdftex]{Definitions/mdpi}
-
-%=================================================================
-\firstpage{1}
-\makeatletter
-\setcounter{page}{\@firstpage}
-\makeatother
-\pubvolume{xx}
-\issuenum{1}
-\articlenumber{5}
-\pubyear{2020}
-\copyrightyear{2020}
-%\externaleditor{Academic Editor: name}
-\history{Received: date; Accepted: date; Published: date}
-%\updates{yes} % If there is an update available, un-comment this line
-
-%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
-%\continuouspages{yes}
-
-%=================================================================
-% Add packages and commands here.
-The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
-
-%=================================================================
-%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
-%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
-
-%=================================================================
-% Full title of the paper (Capitalized)
-\Title{Design, Implementation, and Evaluation of Open Power Quality}
-
-% Authors, for the paper (add full first names)
-\Author{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}
-
-% Authors, for metadata in PDF
-\AuthorNames{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}
-
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-Department of Information and Computer Sciences\\
-University of Hawaii at Manoa\\
-Honolulu, HI 96822\\
-johnson@hawaii.edu}
-
-% Abstract (single paragraph, 200 words max, background, methods, results, conclusions. Currently 193 words.)
-
-\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture which must accomodate distributed, intermittent generation.
-This transition also means that the formerly sharp distinction between energy producers (i.e.
-utility companies) and consumers (residences, businesses, etc) are blurring: residences can now both produce and consume energy, making energy policy more complex.
-The Open Power Quality (OPQ) project began in 2013 with the goal of designing and implementing a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.
-Since then, we have designed low cost hardware devices that monitor power quality  and low-cost cloud-based software services that can economically analyze the data and detect a variety of anomalies.
-In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.
-Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions.
-We conclude that OPQ provides a new and useful approach to power quality monitoring.}
-
-% Keywords
-\keyword{Power Quality, Open Source, Renewable Energy, Grid Stability}
-
-\begin{document}
-
-%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
-
 \section{Introduction}
@@
 Our goal in this paper is to describe OPQ well enough for organizations to make a thoughtful decision about the use of our technology.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\begin{quote}
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection}
-\unskip
-\subsubsection{Subsubsection}
-
-Bulleted lists look like this:
-\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-\item	First bullet
-\item	Second bullet
-\item	Third bullet
-\end{itemize}
-
-Numbered lists can be added as follows:
-\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item	First item
-\item	Second item
-\item	Third item
-\end{enumerate}
-
-The text continues here.
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2 cm]{Definitions/logo-mdpi}
-\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
-\end{figure}
-
-Text
-
-Text
-
-\begin{table}[H]
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-entry 1		& data			& data\\
-entry 2		& data			& data\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Text
-
-Text
-
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\textwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\textwidth}{1pt}
-%\end{listing}
-
-\subsection{Formatting of Mathematical Components}
-
-This is an example of an equation:
-
-\begin{equation}
-a + b = c
-\end{equation}
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work reported.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
-
-\noindent
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & linear dichroism
-\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendix
-\section{}
-\unskip
-\subsection{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
-For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: external bibliography
-%=====================================
-%\externalbibliography{yes}
-%\bibliography{your_external_BibTeX_file}
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-\begin{thebibliography}{999}
-% Reference 1
-\bibitem[Author1(year)]{ref-journal}
-Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% Reference 2
-\bibitem[Author2(year)]{ref-book}
-Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-\end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}

### 2020-05-13 03-architecture.tex
add architecture section

\section{System Architecture}

At its most basic, the OPQ system architecture is extremely simple: OPQ Boxes are plugged into wall outlets, they monitor the quality of power at their wall outlets, and the results are communicated via the Internet to a software system called OPQ Cloud. To see the results, users login to the system using a browser.

A slightly less basic description is this: the OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events.

\begin{enumerate}

\item OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components.

\item OPQ Makai monitors incoming low fidelity data from OPQ Boxes, requests high fidelity data when necessary, and stores the results in a MongoDB database.

\item OPQ Mauka analyzes data, creates "events" when it detects anomolies, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis.

\item OPQ View is a visualization platform for displaying the results for data capture and analysis.

\end{enumerate}

\begin{figure}
\center \includegraphics[width=5in]{images/system-diagram.png}
\caption{High level architecture of the OPQ System}
\label{fig:architecture}
\end{figure}

Figure \ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side).  First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQMakai. OPQ Makai analyzes low fidelity measurements, and requests high fidelity waveforms when desirable. Both measurements and waveforms are saved in a MongoDB database. OPQ Mauka analyzes low and high fidelity data, and creates "events" to represent anomalies. OPQ View notifies users of events and allows them to drill down into low and high fidelity data.

OPQ Makai, OPQ Mauka, and OPQ View are all cloud-based software services that collectively form a single "instance" with respect to data transmission, storage, analysis, and visualization. We refer to this collection of software-side components as OPQ Cloud. Every OPQ Box connects to a single instance of an OPQ Cloud. It is possible to have multiple OPQ Cloud instances. For example, a company might install an OPQ Cloud instance behind their firewall along with OPQ Boxes to provide a private mechanism for collecting and analyzing power quality data.

This architecture has a number of benefits. The combination of low and high fidelity data reduces both network overhead and storage requirements, which increases the scalability of the system in terms of the number of OPQ Boxes that can be tied to a single OPQ Cloud instance. OPQ Makai and OPQ Mauka have a plugin architecture, making it easier to extend their functionality to incorporate new triggers for high quality data (in the case of OPQ Makai) and new events and analyses (in the case of OPQ Mauka). Finally, the open source licensing of both hardware and software makes it possible to incorporate new ideas, bug fixes, and enhancements from technologists across the power quality community.

\section{Information Architecture}

OPQ's "information architecture" is designed to facilitate the process of generating actionable, useful insights into the nature of the electrical grid starting with the data collected from a wall outlet. It is a five layer architecture in which data in the form of Measurements collected by an OPQ Box can eventually yield insights in the form of Phenomena. (The Incident and Phenomena levels are currently in development.)

Table \ref{fig:information-architecture} provides an overview of the five layers from lowest level (Box) to highest level (Phenomena). "TTL", to Time To Live, indicates how long entities at each level are stored before they are automatically deleted.

\begin{table}[H]
\caption{OPQ Information Architecture Levels}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{llp{4in}}
\toprule
\textbf{Layer}	& \textbf{TTL}	& \textbf{Purpose}\\
\midrule
Box		& 1 hour			& Collects and holds a rolling one hour window of "high fidelity" wave form data for a single location.\\
Measurement		& 1 day			& Approximately once a second, each box sends a Measurement to OPQ Cloud. Measurements provides "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients). Measurements give OPQ Cloud basic situational awareness of the grid without the overhead of transmitting and storing wave form data. Measurements exist for one day, though a daily roll-up of Measurement summary statistics called "Trends" persists much longer.\\
Event & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request wave form data from one or more boxes for a specific time interval (typically just a second or so). Note that Makai only has one hour to request high fidelity wave form data before it is written over in the Box.\\
Incident & 6 months & Not every non-nominal power data value is significant, i.e. actually meaningful for gaining insight into the grid. Over the years, electrical engineers have developed a variety of standards (IEEE 1159, ITIC, SEMA, etc.) for characterizing significant power quality events. OPQ Cloud's Mauka service provides classification algorithms to analyze each Event to see if it satisfies any of the standards for significance. If so, an Incident is created, indicating that a significant PQ event has occurred at a specific time in a specific location. Each Incident can also be annotated with context, which is additional information about the environment or other physical factors present at the time and location of this incident. Context can be manually provided by users or automatically associated with Incidents through APIs to online services.\\
Phenomena & 1 year & All of the prior levels represent behaviors of a individual Boxes in a single location at a given point in time. This final level of OPQ's information architecture is premised on the idea that insightful, actionable information results from the ability to either explain or else predict multiple Incidents. First, a Phenomena can be created in order to collect data from multiple Incidents with a common cause. An example of this form of phenomena is "the voltage drop experienced by all boxes in Kailua on 5/22/18 at 1:03pm was caused by a transformer trip at the Kailua substation." Second, a Phenomena can collect data from multiple Incidents in order to predict future Incidents. For example, "when winds exceed 40 knots in Kahuku, data from prior Incidents leads to the prediction that there will be a voltage surge of approximately 5V for approximately 3 seconds in all boxes located within 1 mile of Kahuku."\\
\bottomrule
\end{tabular}
\label{fig:information-architecture}
\end{table}

### 2020-05-13 article.tex
add architecture section

--- 2020-05-10 article.tex
@@
 
-% Abstract (single paragraph, 200 words max, background, methods, results, conclusions. Currently 193 words.)
-
-\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture which must accomodate distributed, intermittent generation.
-This transition also means that the formerly sharp distinction between energy producers (i.e.
-utility companies) and consumers (residences, businesses, etc) are blurring: residences can now both produce and consume energy, making energy policy more complex.
-The Open Power Quality (OPQ) project began in 2013 with the goal of designing and implementing a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.
-Since then, we have designed low cost hardware devices that monitor power quality  and low-cost cloud-based software services that can economically analyze the data and detect a variety of anomalies.
-In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.
-Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions.
-We conclude that OPQ provides a new and useful approach to power quality monitoring.}
-
-% Keywords
-\keyword{Power Quality, Open Source, Renewable Energy, Grid Stability}
+\input{01-abstract}
 
@@
 
-%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
+\input{02-intro}
+\input{03-architecture}
 
-\section{Introduction}
-The introduction should briefly place the study in a broad context and highlight why it is important.
-It should define the purpose of the work and its significance.
-The current state of the research field should be reviewed carefully and key publications cited.
-Please highlight controversial and diverging hypotheses when necessary.
-Finally, briefly mention the main aim of the work and highlight the principal conclusions.
-As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
-Citing a journal paper \cite{ref-journal}.
-And now citing a book reference \cite{ref-book}.
-Please use the command \citep{ref-journal} for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Languages, Laws, Religions, Risks, Social Sciences.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\begin{quote}
-This section may be divided by subheadings.
-It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
-\end{quote}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Subsection}
-\unskip
-\subsubsection{Subsubsection}
-
-Bulleted lists look like this:
-\begin{itemize}[leftmargin=*,labelsep=5.8mm]
-\item	First bullet
-\item	Second bullet
-\item	Third bullet
-\end{itemize}
-
-Numbered lists can be added as follows:
-\begin{enumerate}[leftmargin=*,labelsep=4.9mm]
-\item	First item
-\item	Second item
-\item	Third item
-\end{enumerate}
-
-The text continues here.
-
-\subsection{Figures, Tables and Schemes}
-
-All figures and tables should be cited in the main text as Figure 1, Table 1, etc.
-
-\begin{figure}[H]
-\centering
-\includegraphics[width=2 cm]{Definitions/logo-mdpi}
-\caption{This is a figure, Schemes follow the same formatting.
-If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
-(\textbf{b}) Description of what is contained in the second panel.
-Figures should be placed in the main text near to the first time they are cited.
-A caption on a single line should be centered.}
-\end{figure}
-
-Text
-
-Text
-
-\begin{table}[H]
-\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
-\centering
-%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
-\begin{tabular}{ccc}
-\toprule
-\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
-\midrule
-entry 1		& data			& data\\
-entry 2		& data			& data\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Text
-
-Text
-
-%\begin{listing}[H]
-%\caption{Title of the listing}
-%\rule{\textwidth}{1pt}
-%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
-%\rule{\textwidth}{1pt}
-%\end{listing}
-
-\subsection{Formatting of Mathematical Components}
-
-This is an example of an equation:
-
-\begin{equation}
-a + b = c
-\end{equation}
-%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-
-Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
-%% Example of a theorem:
-\begin{Theorem}
-Example text of a theorem.
-\end{Theorem}
-
-The text continues here. Proofs must be formatted as follows:
-
-%% Example of a proof:
-\begin{proof}[Proof of Theorem 1]
-Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
-\end{proof}
-The text continues here.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses.
-The findings and their implications should be discussed in the broadest context possible.
-Future research directions may also be highlighted.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
-Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
-Please disclose at the submission stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
-
-Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers.
-If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
-They must be provided prior to publication.
-
-Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
-This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Patents}
-This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{6pt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work reported.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
-This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\abbreviations{The following abbreviations are used in this manuscript:\\
-
-\noindent
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute\\
-DOAJ & Directory of open access journals\\
-TLA & Three letter acronym\\
-LD & linear dichroism
-\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendix
-\section{}
-\unskip
-\subsection{}
-The appendix is an optional section that can contain details and data supplemental to the main text.
-For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
-Mathematical proofs of results not central to the paper can be added as an appendix.
-
-\section{}
-All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: external bibliography
-%=====================================
-%\externalbibliography{yes}
-%\bibliography{your_external_BibTeX_file}
-
-%=====================================
-% References, variant B: internal bibliography
-%=====================================
-\begin{thebibliography}{999}
-% Reference 1
-\bibitem[Author1(year)]{ref-journal}
-Author1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
-% Reference 2
-\bibitem[Author2(year)]{ref-book}
-Author2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
-\end{thebibliography}
-
-% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
-For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\sampleavailability{Samples of the compounds ...... are available from the authors.}
-
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2020-05-15 01-abstract.tex
update

@@
 
-\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture which must accomodate distributed, intermittent generation.
-This transition also means that the formerly sharp distinction between energy producers (i.e.
-utility companies) and consumers (residences, businesses, etc) are blurring: residences can now both produce and consume energy, making energy policy more complex.
-The Open Power Quality (OPQ) project began in 2013 with the goal of designing and implementing a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.
-Since then, we have designed low cost hardware devices that monitor power quality  and low-cost cloud-based software services that can economically analyze the data and detect a variety of anomalies.
+\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture with significant distributed, intermittent generation.
+This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) will blur: customers will both produce and consume energy, making energy management and public policy more complex.
+The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.
 In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.

### 2020-05-15 02-intro.tex
update

@@
 Just like most people in developed nations assume that their tap water is of adequate quality to drink, most also assume that their electricity is of adequate quality to power their homes and appliances without causing harm.
-And, in both cases, most people assume that public utilities will monitor and correct any quality problems if they occur.
+And, in both cases, most people assume that public utilities will appropriately monitor and correct any quality problems if they occur.
 
 Successfully maintaining adequate power quality and providing sufficient amounts of it to meet the rising needs of consumers has been a triumph of electrical utilities for over 100 years.
-In recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing public concern and interest.
+In recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing concern and interest.
 First, there is a global need to shift to renewable energy sources.
-Second, modern consumer electronics place more stringent demands on power quality.
-Third, effective policy making in this modern context requires better public access to power quality data.
+Second, traditional top-down grid power quality monitoring may not be well suited to bottom-up, distributed generation associated with renewable energy sources such as roof-top solar.
+Third, modern consumer electronics place more stringent demands on power quality.
+Finally, effective public policy making in this modern context can be aided by public access to power quality data.
+Let's look at each of these in a bit more detail.
 
-{\em We need more renewable energy.} There is now a global movement away from centralized, fossil-fuel based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar.
+{\em 1.
+We need more renewable energy.} Concerns including pollution, environmental degradation, and climate change have produced a global movement away from centralized, fossil-fuel based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar.
 But the economic, environmental, and political advantages of renewable energy comes with significant new technical challenges.
 Wind and solar are intermittent (for example, solar energy cannot be harvested at night) and unpredictable (for example, wind and solar energy fluctuate based upon cloud cover and wind speed).
-In addition, renewable energy generation can be distributed throughout the grid (such as in the case of residential rooftop photovoltaic (PV) systems).
+In addition, renewable energy generation is often distributed throughout the grid (such as in the case of residential rooftop photovoltaic (PV) systems).
 
@@
 This problem is inversely proportional to the size of the electrical grid.
-For example, in Hawaii on the island of Oahu, the public demand for rooftop solar exceeds Hawaiian Electric Company's ability to utilize it and maintain adequate power quality.
-As a result, in many neighborhoods, new rooftop solar installation is prohibited because Hawaiian Electric fears its impact on not only local power quality but global grid stability \cite{TBD}.
+For example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai.
+As a result, in many neighborhoods, permits for new rooftop solar installation require complex and expensive "interconnection studies" because Hawaiian Electric fears the impact on grid stability \cite{trabish_solar_2014,anastasi_energy_2009}.
 
-{\em Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality.
+{\em 2. Traditional top-down monitoring is not well-suited to bottom-up energy generation.}
+
+\cite{nakafuji_back--basics_2011}
+
+{\em 2.
+Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality.
 Only a few decades ago, computers were a rare presence outside of labs and large institutions.

### 2020-05-17 02-intro.tex
checkpoint

@@
 Finally, briefly mention the main aim of the work and highlight the principal conclusions.}
+
+\subsection{Motivation}
 
@@
 For example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai.
-As a result, in many neighborhoods, permits for new rooftop solar installation require complex and expensive "interconnection studies" because Hawaiian Electric fears the impact on grid stability \cite{trabish_solar_2014,anastasi_energy_2009}.
+(In contrast, the European electrical grid is approximately 600 GW and the U.S.
+continental grid is over 1000 GW.) For small grids, the unpredictable nature of power generation for distributed, intermittent renewables can quickly create problems for grid stability.
+In the case of Hawaii, consumer demand for grid-tied rooftop solar exceeded the ability of Hawaiian Electric to manage, resulting first in complex and expensive "interconnection studies" \cite{trabish_solar_2014,anastasi_energy_2009}, and later in the requirement for new installations to include batteries and grid-disconnected operation.
 
-{\em 2. Traditional top-down monitoring is not well-suited to bottom-up energy generation.}
+{\em 2.
+Traditional top-down monitoring is not well-suited to bottom-up energy generation.} For traditional grid architectures where generation is centralized and under the complete control of the utility, it is common to monitor power quality only to the substation level, because it can be assumed that the power quality experienced at the substation is a reasonably accurate proxy for the power quality experienced by the 1,000 or so end-users serviced by the substation.
+Furthermore, if an end-user experiences a power quality problem not experienced by the substation, then it is mostly likely due to equipment or electrical issues local to that end-user and not a grid-level problem.
 
+These assumptions might not hold in grids with distributed, intermittent generation by end-users, such as is the case with roof top solar.
+In these cases, the unpredictable nature of generation can lead to local power quality problems that are not experienced by the substation, and that are not due to any single individual, but rather the bottom-up power generation architecture of the grid.
 \cite{nakafuji_back--basics_2011}
 
-{\em 2.
+{\em 3.
 Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality.
@@
 
-{\em Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers.
+{\em 4.
+Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers.
 For example, in Hawaii, utilities are required to make a "best effort" to provide non-harmful voltage and frequency values, but are only required to publicly report on power outages of more than 3 minutes.
-There is no requirement for utilities to report potentially harmful forms of voltage, frequency, THD, or transients.
-Indeed, in Hawaii, there is not yet infrastructure in place that would enable utilities to collect that information, even if they were asked to report on it.
+There is no requirement for utilities to report potentially harmful levels of voltage, frequency, THD, or transients as long as they do not lead to outages.
+Indeed, in places like Hawaii, there is not even infrastructure in place to enable utilities to collect that information, even if they were asked to report on it.
 For more details, see Hawaiian Electric Tariff Rule No.
@@
 
-OPQ intends to provide an unbiased, independent, third party source of accurate power quality data.
-This can be used by consumers to better understand the performance of their public utilities, by researchers to devise improvements to grid control, and by public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
-The OPQ Project has produced a suite of innovations that can help address each of these issues.
+Providing power quality data to address the above issues requires solving several difficult technical problems.
+First, monitoring below the substation level increases the number of required monitoring devices by two to three orders of magnitude.
+This has significant implications for the cost of monitoring in terms of the devices, and the sheer amount of low-level power quality data that will be produced.
+Second, this low-level data must be processed in a manner that does not require inordinate amounts of processing, network bandwidth, or storage.
+Finally, the data must lead to actionable information.
+
+\subsection{Goals of the Open Power Quality Project}
+
+The goal of the OPQ Project is to provide an unbiased, independent, third party source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
+OPQ accomplishes this through a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services.
+An OPQ sensor network is intended to be useful to consumers who want to better understand the performance of their public utilities, to researchers who wish to design improvements to grid management, and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
 
@@
 
-{\em Cloud native.} Second, our architecture is "cloud native", which means that OPQ Boxes are not designed for "stand alone", autonomous use.
+{\em Cloud native.} Second, our architecture is "cloud native", which means that OPQ Boxes are not designed for "stand alone", autonomous use, unlike current commercially available power quality monitors.
 Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and analyzed.
-This creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes.
+This creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
 It means our cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.

### 2020-05-21 literature.tex
updates du joiur

\section{Misc. Literature Quotes}

In February 2014 on the Big Island, the utility reports that 10\% of its circuits hit unstable levels \cite{trabish_solar_2014}.

Photovoltaics affect power quality by introducing harmonics \cite{anurangi_effects_2017}. The power feed-in of PV generation in rural low-voltage grids can influence power quality (PQ) as well as facility operation and reliability \cite{rita_pinto_impact_2016}.

Models, not actual installations \cite{anurangi_effects_2017,bayindir_effects_2016,farhoodnea_power_2012,shafiullah_experimental_2014}. Or data taken for only one week \cite{kucuk_assessment_2013}.

``across all business sectors, the U.S. economy is losing between \$104 billion and \$164 billion a year to outages and another \$15 billion to \$24 billion to PQ phenomena'' \cite{elphick_summary_2015} India lost more than \$9.6 billion in 2008 due to power quality, and Europe lost \$150 billion per year \cite{laskar_power_2012}.

### 2020-05-21 02-intro.tex
checckpoint

@@
 For example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai.
-(In contrast, the European electrical grid is approximately 600 GW and the U.S.
-continental grid is over 1000 GW.) For small grids, the unpredictable nature of power generation for distributed, intermittent renewables can quickly create problems for grid stability.
+(In contrast, the size of the European electrical grid is approximately 600 GW and the size of the U.S.
+continental grid is over 1000 GW.) For small grids, the unpredictable nature of power generation by distributed, intermittent renewables can quickly create problems for grid stability.
 In the case of Hawaii, consumer demand for grid-tied rooftop solar exceeded the ability of Hawaiian Electric to manage, resulting first in complex and expensive "interconnection studies" \cite{trabish_solar_2014,anastasi_energy_2009}, and later in the requirement for new installations to include batteries and grid-disconnected operation.
@@
 In these cases, the unpredictable nature of generation can lead to local power quality problems that are not experienced by the substation, and that are not due to any single individual, but rather the bottom-up power generation architecture of the grid.
-\cite{nakafuji_back--basics_2011}
+Nakafuji notes that Hawaiian Electric is contending with PV penetrations in excess of 60\% on certain distribution circuits, for which traditional rules of thumb for design of protection and distribution systems might not hold \cite{nakafuji_back--basics_2011}.
 
@@
 There is no requirement for utilities to report potentially harmful levels of voltage, frequency, THD, or transients as long as they do not lead to outages.
-Indeed, in places like Hawaii, there is not even infrastructure in place to enable utilities to collect that information, even if they were asked to report on it.
-For more details, see Hawaiian Electric Tariff Rule No.
-2 Character of Service and Hawaiian Electric Tariff Rule No.
-16 Interruption of Service.
+Indeed, in places like Hawaii, there is not yet infrastructure in place to enable utilities to collect that information, even if they were asked to report on it.
 
@@
 Second, this low-level data must be processed in a manner that does not require inordinate amounts of processing, network bandwidth, or storage.
-Finally, the data must lead to actionable information.
+Finally, collecting the data is not useful if it cannot be converted into actionable information in a reasonable amount of time.
 
@@
 
-The goal of the OPQ Project is to provide an unbiased, independent, third party source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
-OPQ accomplishes this through a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services.
-An OPQ sensor network is intended to be useful to consumers who want to better understand the performance of their public utilities, to researchers who wish to design improvements to grid management, and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
+The goal of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
+OPQ accomplishes this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services.
+An OPQ sensor network is designed to be useful to consumers who want to better understand the performance of their public utilities, to researchers who wish to design improvements to grid management, and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
+There are several key features of our project emerge from these goals.
 
-{\em OPQ Box.} First, our power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75.
+{\em OPQ Box.} First, our custom power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75.
 This is from 10 to 100 times less expensive than commercial power quality monitors.
-This means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building.
+This cost differential means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building.
 Similarly, instead of monitoring power quality at the substation level of the grid, OPQ Boxes make it feasible to deploy dozens or hundreds into a community to obtain fine-grained data about the impact of solar or other renewables as directly perceived by the customer.
 
-{\em Cloud native.} Second, our architecture is "cloud native", which means that OPQ Boxes are not designed for "stand alone", autonomous use, unlike current commercially available power quality monitors.
-Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and analyzed.
-This creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
-It means our cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.
+{\em Cloud-native information architecture.} Second, our information architecture is "cloud native", which means that OPQ Boxes are not designed for stand-alone, autonomous use, unlike current commercially available power quality monitors.
+Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and with what fidelity in order to most efficiently produce actionable information.
+Two-way communication between the OPQ Boxes and could services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
+As a result, OPQ cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.
 This enables OPQ installations to be simultaneously responsive and scalable.
 
-{\em Extensibility.} Third, all OPQ software subsystems are designed for extensibility and interoperation.
+{\em Plug-in analysis architecture.} Third, all OPQ software services are designed for extensibility and interoperation.
 Our middleware components (Mauka and Makai) provide a plugin architecture.
 Our visualization component (View) is built from modular UI elements using React.
-We are also developing API endpoints for interoperation with other software systems.
 This means that OPQ is not a monolithic, closed system with a frozen feature set, but rather a design environment for experimentation with advanced classification and analysis of power quality data.
@@
 
-Finally, while we believe OPQ offers a compelling combination of capabilities, we want to be clear that all designs involve trade-offs.
+While we believe OPQ offers a compelling combination of capabilities, we want to be clear that all network designs involve trade-offs.
 As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs.
@@
 In other cases, a hybrid solution consisting of OPQ in combination with other technologies for power monitoring may be most appropriate.
-Our goal in this paper is to describe OPQ well enough for organizations to make a thoughtful decision about the use of our technology.
+
+\subsection{Structure of this paper}
+
+The next sections present an overview of the OPQ system architecture and information architecture.
+The following section presents related research and technology.
+Section \ref{sec:makai} provides more detail on the OPQ Box, and Section \ref{sec:makai} provides more detail on our cloud services.
+Section \ref{sec:pilot} discusses the pilot deployment of a sensor network and its results.
+We conclude in Section \ref{sec:conclusion} with some proposals for future research.

### 2020-05-22 02-intro.tex
update structure.

@@
 
-The next sections present an overview of the OPQ system architecture and information architecture.
-The following section presents related research and technology.
-Section \ref{sec:makai} provides more detail on the OPQ Box, and Section \ref{sec:makai} provides more detail on our cloud services.
+Section \ref{sec:architecture} presents an overview of the OPQ system architecture and information architecture.
+Section \ref{sec:related-work} presents related research and technology.
+Section \ref{sec:opq-makai} discusses OPQ Box and its triggering system, and Section \ref{sec:opq-makai} discusses our cloud-based analysis services.
 Section \ref{sec:pilot} discusses the pilot deployment of a sensor network and its results.
-We conclude in Section \ref{sec:conclusion} with some proposals for future research.
+Section \ref{sec:conclusions} concludes with some proposals for future research.

### 2020-05-22 04-related.tex
update structure.

\section{Related Work}
\label{sec:related-work}

Related work.

### 2020-05-22 01-abstract.tex
finished first draft of architecture section. Progress!

@@
 \abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture with significant distributed, intermittent generation.
-This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) will blur: customers will both produce and consume energy, making energy management and public policy more complex.
-The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to provide make useful information about electrical grids available to producers, consumers, and policy makers.
+This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) will blur: end-users will both produce and consume energy, making energy management and public policy more complex.
+The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to make useful information about electrical grids available to producers, consumers, and policy makers.
 In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.

### 2020-05-22 02-intro.tex
finished first draft of architecture section. Progress!

@@
 First, there is a global need to shift to renewable energy sources.
-Second, traditional top-down grid power quality monitoring may not be well suited to bottom-up, distributed generation associated with renewable energy sources such as roof-top solar.
+Second, traditional approaches to top-down grid power quality monitoring may not be well suited to bottom-up, distributed generation associated with renewable energy sources such as rooftop photovoltaic panels (PV).
 Third, modern consumer electronics place more stringent demands on power quality.
@@
 
-Section \ref{sec:architecture} presents an overview of the OPQ system architecture and information architecture.
+The remainder of this paper is organized as follows.
+Sections \ref{sec:system-architecture} and \ref{sec:information-architecture} present an overview of the OPQ system architecture and information architecture.
 Section \ref{sec:related-work} presents related research and technology.

### 2020-05-22 03-architecture.tex
update structure.
finished first draft of architecture section. Progress!

@@
 \section{System Architecture}
+\label{sec:system-architecture}
 
@@
 \section{Information Architecture}
+\label{sec:information-architecture}
 
 OPQ's "information architecture" is designed to facilitate the process of generating actionable, useful insights into the nature of the electrical grid starting with the data collected from a wall outlet.
-It is a five layer architecture in which data in the form of Measurements collected by an OPQ Box can eventually yield insights in the form of Phenomena.
-(The Incident and Phenomena levels are currently in development.)
+The information architecture must address several requirements.
+First, it must support analyses that require access to high fidelity data, which in the case of power quality is wave form data produced by sampling the wave form 256 times per cycle, which means there are approximately 1.3 billion data points generated per day per OPQ Box.
+Second, for typical grids that are generally stable, only an extremely small fraction of these data points are actually interesting.
+Third, as will be discussed in more detail later, what constitutes ``interesting'' data in one box might be dependent upon the state of data in other nearby boxes.
+Finally, network and storage limitations preclude a solution in which all data is sent from OPQ Boxes to the cloud, and so the information architecture must provide a ``smart'' mechanism for identifying data of interest on the boxes, as well as discarding data in the cloud if it is found to be not useful.
 
-Table \ref{fig:information-architecture} provides an overview of the five layers from lowest level (Box) to highest level (Phenomena).
-"TTL", to Time To Live, indicates how long entities at each level are stored before they are automatically deleted.
+To address these requirements, the OPQ information architecture is organized as a five layer pyramid as summarized in Table \ref{fig:information-architecture}.
+This table illustrates an important conceptual feature of the OPQ information architecture called ``TTL'', or Time To Live.
+The OPQ information architecture implements an active, ``use it or lose it'' approach to information management, in which a data item must either be associated with a higher level of the system within its Time To Live, or else it will be automatically deleted.
+So, for example, the high fidelity waveform data on an OPQ Box have a TTL of one hour: if that data is not requested from a box by a higher layer in the information architecture within an hour of it being generated, it will be lost.
+Similarly, data points in the Measurement layer will be deleted after one day if they are not associated with an Event.
 
@@
 Box		& 1 hour			& Collects and holds a rolling one hour window of "high fidelity" wave form data for a single location.\\
-Measurement		& 1 day			& Approximately once a second, each box sends a Measurement to OPQ Cloud.
-Measurements provides "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients).
-Measurements give OPQ Cloud basic situational awareness of the grid without the overhead of transmitting and storing wave form data.
-Measurements exist for one day, though a daily roll-up of Measurement summary statistics called "Trends" persists much longer.\\
-Event & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request wave form data from one or more boxes for a specific time interval (typically just a second or so).
-Note that Makai only has one hour to request high fidelity wave form data before it is written over in the Box.\\
-Incident & 6 months & Not every non-nominal power data value is significant, i.e.
-actually meaningful for gaining insight into the grid.
-Over the years, electrical engineers have developed a variety of standards (IEEE 1159, ITIC, SEMA, etc.) for characterizing significant power quality events.
-OPQ Cloud's Mauka service provides classification algorithms to analyze each Event to see if it satisfies any of the standards for significance.
-If so, an Incident is created, indicating that a significant PQ event has occurred at a specific time in a specific location.
-Each Incident can also be annotated with context, which is additional information about the environment or other physical factors present at the time and location of this incident.
-Context can be manually provided by users or automatically associated with Incidents through APIs to online services.\\
-Phenomena & 1 year & All of the prior levels represent behaviors of a individual Boxes in a single location at a given point in time.
-This final level of OPQ's information architecture is premised on the idea that insightful, actionable information results from the ability to either explain or else predict multiple Incidents.
-First, a Phenomena can be created in order to collect data from multiple Incidents with a common cause.
-An example of this form of phenomena is "the voltage drop experienced by all boxes in Kailua on 5/22/18 at 1:03pm was caused by a transformer trip at the Kailua substation." Second, a Phenomena can collect data from multiple Incidents in order to predict future Incidents.
-For example, "when winds exceed 40 knots in Kahuku, data from prior Incidents leads to the prediction that there will be a voltage surge of approximately 5V for approximately 3 seconds in all boxes located within 1 mile of Kahuku."\\
+Measurement		& 1 day			& Measurements provide "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients). \\
+Event & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': wave form data from one or more boxes for a specific time interval (typically just a second or so).
+\\
+Incident & 6 months & An Incident is created when an Event satisfies a standard (such as IEEE 1159, ITIC, or SEMA) for a significant power quality problem.\\
+Phenomena & 1 year & Phenomena classify Incidents when they are predictable and/or have a known causal explanation. \\
 \bottomrule
@@
 \end{table}
+
+Let's now discuss how information moves through these levels.
+
+The bottom layer of the information architecture is called Box, and represents all of the OPQ Box hardware devices in an OPQ sensor network.
+Each box maintains onboard storage of the latest hour of high quality waveform data in a circular buffer.
+
+Approximately once a second, each Box sends a Measurement data object to the sensor network's cloud services, which we refer to in aggregate as OPQ Cloud.
+Measurements are the second layer of the information architecture, and provide "low fidelity" summary statistics regarding four power quality measurements (Frequency, Voltage, THD, and Transients).
+Measurements give OPQ Cloud basic situational awareness of the grid without the overhead of transmitting, storing, and analyzing high fidelity wave form data.
+Measurements exist for one day, though a daily roll-up of Measurement summary statistics called "Trends" persists longer.
+
+When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request wave form data from one or more boxes for a specific time interval (typically just a second or so).
+Note that Makai only has one hour to request high fidelity wave form data before it is lost.
+
+Not every non-nominal power data value is significant, or in other words, meaningful for gaining insight into the grid.
+Over the years, the power community has developed a variety of standards (IEEE 1159, ITIC, SEMA, etc.) for characterizing significant power quality events.
+OPQ Cloud's Mauka service provides classification algorithms to analyze each Event to see if it satisfies any of the standards for significance.
+If so, an Incident is created, indicating that a significant PQ event has occurred at a specific time in a specific location.
+Each Incident can also be annotated with context, which is additional information about the environment or other physical factors present at the time and location of this incident.
+Context can be manually provided by users or automatically associated with Incidents through APIs to online services.
+
+All of the prior levels represent behaviors of a individual Boxes in a single location at a given point in time.
+This final level of OPQ's information architecture is premised on the idea that insightful, actionable information results from the ability to either explain or else predict multiple Incidents.
+First, a Phenomena can be created in order to collect data from multiple Incidents with a common cause.
+An example of this form of phenomena is "the voltage drop experienced by all boxes in Kailua on 5/22/18 at 1:03pm was caused by a transformer trip at the Kailua substation." Second, a Phenomena can collect data from multiple Incidents in order to predict future Incidents.
+For example, "when winds exceed 40 knots in Kahuku, data from prior Incidents leads to the prediction that there will be a voltage surge of approximately 5V for approximately 3 seconds in all boxes located within 1 mile of Kahuku."

### 2020-05-22 literature.tex
update structure.
finished first draft of architecture section. Progress!

@@
 economy is losing between \$104 billion and \$164 billion a year to outages and another \$15 billion to \$24 billion to PQ phenomena'' \cite{elphick_summary_2015} India lost more than \$9.6 billion in 2008 due to power quality, and Europe lost \$150 billion per year \cite{laskar_power_2012}.
+
+A study of the island of Porto Santo in Portugal found that intermittent nature of installed photovoltaic and wind energy could result in a potential drop in frequency of 12 Hz, lasting as long as 7 seconds \cite{delgado_solutions_2011}.
+
+Reduced input voltage can cause excessive power supply heat dissipation, resulting in reduced mean time between failures (MTBF).
+In addition, rectifiers and DC-to-DC converter switching transistors draw high-peak currents, which raise their junction temperatures.
+These temperature excursions take a toll on semiconductor longevity.
+High input voltage can also puncture a power supply's rectifier and switching transistor junctions, causing MTBF reduction and eventual breakdown.
+High-voltage transients lasting microseconds can permanently wreck the power supply and its electronic equipment load.
+\cite{dedad_when_2008}.
+
+Solar panels connected to low voltage networks will result in over-voltages, the switching frequency of the converters in wind turbines causes high-frequency signals flowing into the grid, and harmonics are generated by EV chargers.
+\cite{zavoda_power_2018}

### 2020-05-25 04-related.tex
started on related work

@@
 
-Related work.
+This section explains how OPQ fits into current industry solutions as well as academic research on power quality monitoring and analysis.
+For the purposes of this review, we exclude utility-side power quality monitoring and analysis systems.
+
+Section \ref{sec:commercial-pq-devices} discusses commercial power quality monitoring devices and how they compare to OPQ Box.
+Section \ref{sec:commercial-pq-software} discusses commercial power quality database software and how they compare to OPQ Cloud.
+The final subsection discusses recent research systems and how they compare to OPQ as a whole.
+
+\subsection{Commercial PQ devices}
+\label{sec:commercial-pq-devices}
+
+Fluke \cite{fluke_fluke_2020} provides a variety of Power Quality tools.
+These tools typically measure voltage, current, frequency, dips, swells, power factor, harmonic, current unbalance, inrush current values, and light flicker.
+They generally adhere to power quality standards such as IEEE 1159, and use current transformers rather than plugging into a wall outlet.
+Depending on the model, these devices cost from approximately \$2,000 to \$10,000, and are designed for use by power engineers to diagnose power quality problems resulting from industrial machinery.
+
+Dranetz \cite{dranetz_dranetz_2020} provides the HDPQ family of power quality analyzers.
+These are similar to Fluke devices in that they require connection to mains using current transformers and are intended for use by power quality engineers.
+They cost from \$5,000 to \$17,000 depending upon the model.
+They also sell a set of software answer modules to provide additional analyses regarding the raw power quality data collected by the device.
+For example, the Sag Directivity module detects a voltage sag, identifies its characteristics, and determines whether the source is upstream or downstream of the monitoring point.
+Answer modules cost approximately \$1,000 each.
+
+Elspec \cite{elspec_elspec_2020} provides the BlackBox family of multi functional waveform recorders.
+These devices perform transient recording, disturbance recording, phasor measurements (synchrophasor), power quality analysis, and sequence of events recording.
+They cost from \$6,000 to \$12,000 depending upon the model.
+
+PowerSide (formerly Power Standards Lab) \cite{powerside_powerside_2020} provides the PQube3 family of power monitors.
+Unlike the above devices, PQube is both a revenue-grade energy meter as well as a power quality monitor.
+It can also record environmental data such as temperature, humidity, barometric pressure, and certain process parameters such as fuel level and water flow.
+QubeView is a Windows-based software package that provides real-time monitoring of multiple PQube devices and trends and statistics over the data collected.
+
+ACR Systems \cite{acr_acr_2020} manufactures the PowerWatch Voltage Disturbance Recorder.
+Unlike the above devices, this device connects to a wall outlet.
+It measures surges and sags and is meant for monitoring medical instrumentation, point-of-sale terminals, computer servers, and other sensitive electronic equipment.
+It costs approximately \$625.
+A Windows-based program is available for a separate cost of \$135, and can be used to extract data from the device using a USB cable, and then display graphs of power quality.
+
+\subsection*{Comparison with OPQ Box}
+
+All of these commercially available devices contrast with the OPQ Box in similar ways.
+First, all of them collect a wider variety of power quality measures than OPQ Box, and most have been certified according to one or more industry standards.
+They are generally designed to support industrial applications, where the goal is to ensure that the power being supplied to a building or plant is of adequate quality, and/or that the machinery in the plant is not degrading power quality.
+Apart from the PowerWatch monitor, all of them are attached to electrical mains using current transformers.
+Finally, all of them are designed for "stand alone" operation: each device can independently gather and assess power data.
+
+While the OPQ Box has much more limited functionality, it is designed to be manufactured for approximately \$75, which is 10 to 100 times less expensive than these devices.
+OPQ Boxes are "cloud native", which means that they have no "stand alone" capabilities.
+Instead, they must be connected to an OPQ Cloud instance in order for power quality data to be collected and analyzed.
+
+\subsection{Commercial PQ software}
+\label{sec:commercial-pq-software}
+
+PQView is a database system developed by Electrotek and EPRI to integrate data from a wide variety of digital relays, fault recorders, power quality monitors, smart meters, and SCADA historians into a relational database.
+PQView is not designed for real-time display, instead, the normal mode of operation is to poll the various devices containing power data once per day and upload the results, although polling as fast as once per minute is supported.
+The software can generate reports regarding voltage sags, swells, and interruptions, THD, and overlay voltage sags/swells on CBEMA curves to indicate which ones exceed tolerances.
+PQView is built on Windows software and the data is stored in a SQL Server database.
+PQView costs \$5000 and is available as an EPRI Product.
+
+PQSCADA Sapphire is another vendor-independent system for collecting and displaying data from a variety of power quality monitoring devices.
+It comes in three versions, a free "Express" version with limited import and analysis tools.
+To obtain additional capabilities, user can purchase the professional version for \$2500 or the Enterprise version for \$6400.
+The data visualizations include trend charts, scatter event charts, histograms with summary statistics, and phasor charts.
+
+PQDif, the "Power Quality Data Interchange Format", is a binary file format that specifies voltage, current, power, and energy measurements in standard fashion that allows this data to produced and consumed by devices in a vendor-independent fashion.
+
+The Grid Protection Alliance produces a suite of open source software products for addressing various issues in power data management and analysis, including data collection, archiving, notifications, event analysis, and visualization.
+
+Comparison with OPQ Cloud
+The differences between the way these systems store and manipulate power quality data and the way used by OPQ arise from fundamentally different architectural assumptions and the historical background of the technology.
+PQView and PQSCADA Sapphire are designed to operate in a technology environment consisting of a large number of installed, "stand alone" power quality monitors built by different vendors.
+Their goal is to aggregate the data collected by these devices, and in order to do so, they depend upon the PQDif standard as a way to obtain power quality data independent of the vendor and device generating it.
+This results in a kind of "store and forward" process: power quality data is captured and stored on the device, and then periodically bundled into a PQDif file and sent to the database software.
+
+OPQ Cloud, on the other hand, is designed only to support the capabilities of OPQ Boxes.
+OPQ Boxes, furthermore, have no "stand-alone" capability; they maintain continuous connection to the Internet and upload power quality data to cloud-based services as needed.
+This means that OPQ implements a very different approach to representing and transmitting data than PQDif.
+For details on the representation, see the OPQ Data Model, and for details on communication, see the OPQ Protobuf protocol.
+
+Note that it would be straightforward to develop an export function in OPQ Cloud to provide power quality data in PQDif format, but that functionality is not currently available.
+
+Research systems
+Di Manno et al describes a PQ monitoring system called PiKu.
+Unlike OPQ, PiKu is designed as a hardware device for sensing power quality that is directly integrated into a PC.
+Systems with similar architectures include TRANSIENTMETER, described in Da Ponte et al, BK-ELCOM, described in Bilik et al, and a system described in Xu et al.
+
+There are also research projects based upon leveraging existing monitoring infrastructure.
+Suslov et al describes a distributed power quality monitoring system based upon existing phasor measurement units installed by utilities.
+Sayied et al describes a system designed using existing smart meters.
+Kucuk et al describes a similar system for the Turkish National Grid using utility grid monitoring infrastructure.
+
+The research system most similar to OPQ is FNET, which is described in a recent publication by Liu et al.
+Like OPQ, the FNET system consists of custom hardware that monitors the electrical signal from a wall outlet, and uploads data to the cloud for further processing.
+Unlike OPQ, FNET is designed for monitoring of frequency disturbances, how they propagate across wide area (i.e.
+nation-wide) utility grids, and, if possible, where the frequency disturbance originated.
+This means that FNET devices must be synchronized using GPS, and that the data collected consists of frequency and voltage angle.
+OPQ is designed for more "local" grid analysis, and we are not interested in propagation.
+As a result, OPQ Boxes are synchronized using NTP rather than GPS, which reduces cost and simplifies installation (OPQ Boxes do not need line of site to a GPS satellite).
+Finally, FNET hardware appears to support only "one way" communication from device to the cloud, while OPQ Boxes support "two way" communication (from box to cloud, and from cloud to box).

### 2020-05-27 01-abstract.tex
related work updates

@@
 \abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture with significant distributed, intermittent generation.
-This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) will blur: end-users will both produce and consume energy, making energy management and public policy more complex.
-The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to make useful information about electrical grids available to producers, consumers, and policy makers.
+This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) are blurring: end-users both produce and consume energy, making energy management and public policy more complex.
+The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to make useful information about electrical grids (and their stability) available to producers, consumers, researchers, and policy makers.
 In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.
 Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions.
-We conclude that OPQ provides a new and useful approach to power quality monitoring.}
+We conclude that OPQ provides a new and useful approach to power quality monitoring that can be used in conjunction with other technologies .}
 

### 2020-05-27 02-intro.tex
related work updates

@@
 Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers.
-For example, in Hawaii, utilities are required to make a "best effort" to provide non-harmful voltage and frequency values, but are only required to publicly report on power outages of more than 3 minutes.
+For example, in Hawaii, utilities are only required to make a "best effort" to provide non-harmful voltage and frequency values, and are only required to publicly report about power outages of more than 3 minutes.
 There is no requirement for utilities to report potentially harmful levels of voltage, frequency, THD, or transients as long as they do not lead to outages.
-Indeed, in places like Hawaii, there is not yet infrastructure in place to enable utilities to collect that information, even if they were asked to report on it.
+Indeed, in places like Hawaii, there is not yet even infrastructure in place to enable utilities to collect that information, even if they were asked to report on it.
+Providing a high quality, third party, publicly available source of power quality data not only benefits the public, it makes it much easier to perform research on ways to improve grid stability in the presence of renewable energy generation.
 
@@
 OPQ accomplishes this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services.
-An OPQ sensor network is designed to be useful to consumers who want to better understand the performance of their public utilities, to researchers who wish to design improvements to grid management, and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
-There are several key features of our project emerge from these goals.
+An OPQ sensor network is designed to be useful to: consumers who want to better understand the performance of their public utilities; utilities who desire a low-cost, easily deployable and re-deployable sensor network to gather information about grid stability that complements their existing infrastructure; to researchers who wish to design improvements to grid management and need a low cost mechanism for measuring the impact of their improvements; and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
+Several key features of our project emerge from these goals.
 
@@
 
-While we believe OPQ offers a compelling combination of capabilities, we want to be clear that all network designs involve trade-offs.
+While we believe OPQ offers a compelling combination of capabilities, we want to be clear that all sensor network designs involve trade-offs.
 As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs.
 In some cases, OPQ may not be the appropriate choice.
-In other cases, a hybrid solution consisting of OPQ in combination with other technologies for power monitoring may be most appropriate.
+In other cases, a hybrid solution consisting of OPQ in combination with other technologies for grid monitoring and analysis may be most appropriate.
 

### 2020-05-27 03-architecture.tex
started on related work
related work updates

@@
 
-At its most basic, the OPQ system architecture is extremely simple: OPQ Boxes are plugged into wall outlets, they monitor the quality of power at their wall outlets, and the results are communicated via the Internet to a software system called OPQ Cloud.
+At its most basic, the OPQ system architecture is this: OPQ Boxes are plugged into wall outlets, they monitor the quality of power at their wall outlets, and the results are communicated via the Internet to a software system called OPQ Cloud.
 To see the results, users login to the system using a browser.
 
-A slightly less basic description is this: the OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events.
+A slightly less basic description is this: the OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events:
 
@@
 
-\item OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components.
+\item OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components either at predefined intervals or upon request.
 
@@
 
-\item OPQ Mauka analyzes data, creates "events" when it detects anomolies, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis.
+\item OPQ Mauka analyzes low level data and analyzes it to produce higher-level representations according to the information architecture described below, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis.
 
-\item OPQ View is a visualization platform for displaying the results for data capture and analysis.
+\item OPQ View is a browser-based visualization platform for displaying the results for data capture and analysis.
 
@@
 \center \includegraphics[width=5in]{images/system-diagram.png}
-\caption{High level architecture of the OPQ System}
+\caption{High level system architecture of an OPQ Sensor Network}
 \label{fig:architecture}
@@
 Figure \ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side).
-First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQMakai.
+First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQ Makai.
 OPQ Makai analyzes low fidelity measurements, and requests high fidelity waveforms when desirable.

### 2020-05-27 04-related.tex
related work updates

@@
 
-PQView is a database system developed by Electrotek and EPRI to integrate data from a wide variety of digital relays, fault recorders, power quality monitors, smart meters, and SCADA historians into a relational database.
+PQView \cite{electrotek_concepts_pqview_2020} is a database system developed by Electrotek and EPRI to integrate data from a wide variety of digital relays, fault recorders, power quality monitors, smart meters, and SCADA historians into a relational database.
 PQView is not designed for real-time display, instead, the normal mode of operation is to poll the various devices containing power data once per day and upload the results, although polling as fast as once per minute is supported.
@@
 
-PQSCADA Sapphire is another vendor-independent system for collecting and displaying data from a variety of power quality monitoring devices.
+PQSCADA Sapphire \cite{elspec_ltd_pqscada_2016} is another vendor-independent system for collecting and displaying data from a variety of power quality monitoring devices.
 It comes in three versions, a free "Express" version with limited import and analysis tools.
@@
 
-PQDif, the "Power Quality Data Interchange Format", is a binary file format that specifies voltage, current, power, and energy measurements in standard fashion that allows this data to produced and consumed by devices in a vendor-independent fashion.
+PQDif \cite{sabin_ieee_2020}, the "Power Quality Data Interchange Format", is a binary file format that specifies voltage, current, power, and energy measurements in standard fashion that allows this data to produced and consumed by devices in a vendor-independent fashion.
 
-The Grid Protection Alliance produces a suite of open source software products for addressing various issues in power data management and analysis, including data collection, archiving, notifications, event analysis, and visualization.
+The Grid Protection Alliance \cite{grid_protection_alliance_grid_2020} produces a suite of open source software products for addressing various issues in power data management and analysis, including data collection, archiving, notifications, event analysis, and visualization.
 
-Comparison with OPQ Cloud
-The differences between the way these systems store and manipulate power quality data and the way used by OPQ arise from fundamentally different architectural assumptions and the historical background of the technology.
+\subsection*{Comparison with OPQ Cloud}
+
+The differences between the way OPQ and the above systems store and manipulate power quality data arise from fundamentally different architectural assumptions and the historical background of the technology.
 PQView and PQSCADA Sapphire are designed to operate in a technology environment consisting of a large number of installed, "stand alone" power quality monitors built by different vendors.
@@
 
-Research systems
-Di Manno et al describes a PQ monitoring system called PiKu.
+\subsection{Research systems}
+
+Di Manno et al \cite{di_manno_user_2015} describes a PQ monitoring system called PiKu.
 Unlike OPQ, PiKu is designed as a hardware device for sensing power quality that is directly integrated into a PC.
-Systems with similar architectures include TRANSIENTMETER, described in Da Ponte et al, BK-ELCOM, described in Bilik et al, and a system described in Xu et al.
+Systems with similar architectures include TRANSIENTMETER, described in Da Ponte et al \cite{daponte_transientmeter:_2000}, BK-ELCOM, described in Bilik et al \cite{bilik_modular_2007}, and a system described in Xu et al \cite{xu_distributed_2012}.
 
 There are also research projects based upon leveraging existing monitoring infrastructure.
-Suslov et al describes a distributed power quality monitoring system based upon existing phasor measurement units installed by utilities.
-Sayied et al describes a system designed using existing smart meters.
-Kucuk et al describes a similar system for the Turkish National Grid using utility grid monitoring infrastructure.
+Suslov et al \cite{suslov_distributed_2014} describes a distributed power quality monitoring system based upon existing phasor measurement units installed by utilities.
+Sayied et al \cite{sayied_power_2013} describes a system designed using existing smart meters.
+Kucuk et al \cite{kucuk_extensible_2010} describes a similar system for the Turkish National Grid using utility grid monitoring infrastructure.
 
-The research system most similar to OPQ is FNET, which is described in a recent publication by Liu et al.
+The research system most similar to OPQ is FNET, which is described in a recent publication by Liu et al \cite{liu_distribution_2017}.
 Like OPQ, the FNET system consists of custom hardware that monitors the electrical signal from a wall outlet, and uploads data to the cloud for further processing.

### 2020-05-28 02-intro.tex
Add opq box section

@@
 Section \ref{sec:related-work} presents related research and technology.
-Section \ref{sec:opq-makai} discusses OPQ Box and its triggering system, and Section \ref{sec:opq-makai} discusses our cloud-based analysis services.
+Section \ref{sec:opq-box} discusses OPQ Box and Section \ref{sec:opq-makai} its triggering system.
+Section \ref{sec:opq-mauka} describes our analysis services, and Section \ref{sec:opq-view} our user interface.
 Section \ref{sec:pilot} discusses the pilot deployment of a sensor network and its results.

### 2020-05-28 05-opq-box.tex
Add opq box section

\section{OPQ Box}
\label{sec:opq-box}

OPQ Box is a hardware device designed to provide inexpensive, extensible and accurate residential power quality measurements.
A block diagram of OPQ Box is shown in Figure~\ref{fig:opq:1:1}.
A complete device is shown in Figure~\ref{fig:opq:1:2}.

\begin{figure}[h]
	\centering
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
	  \caption{OPQ Box2 Block Diagram.
	  The power path is in red, signal path is in green and the digital IO is in black.}
	  \label{fig:opq:1:1}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
	  \caption{OPQ Box2 in an ABS plastic enclosure.}
	  \label{fig:opq:1:2}
	\end{subfigure}
	\caption{(a) OPQ Box2 block diagram and (b) production OPQ Box ready for deployment}
	\label{fig:opq:2}
\end{figure}

\subsection{Hardware}\label{subsec:hardware}

The power system of the OPQ Box2 electrically isolates most of the device from the AC mains power.
An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
5V is used to power the Raspberry Pi, equipment connected to the expansion port, 3.3V regulators and voltage reference and an isolated DC/DC converter.
3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
The hot side of the isolation amplifier is powered from the isolated DC/DC converter.
This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power outage.

The analog signal path of the OPQ Box2 is complicated by the fact that the STM32F3 ADC/DSP is electrically isolated from the mains power.
A previous iteration of the OPQ Box, OPQ Box1, overcame this by employing a small circuit board mount isolation transformer.
Unfortunately it was found that the frequency response of these transformers varied wildly between individuals, thus incurring a lengthy calibration process for each device.
Design on the OPQ Box2 solved this issue by using an AMC1100 isolation amplifier as the isolation component.
Internally AMC1100 consists of a single die comprised of a $\Sigma\Delta$ analog to digital and digital to analog converters.
These converters are separated by a silicon glass region on the integrated circuit which acts as a coupling capacitor.
Since the signal passes the isolation boundary as a $\Sigma\Delta$ encoded digital signal, it incurs no distortion and no additional calibration is required.
In order to match the dynamic range of the AMC1100 the $120V_{ac}$ is passed through the resistor divider to attenuate it to $120mV_{ac}$.
The input and output of the isolation amplifier is filtered with a passive first order anti-aliasing filter.
The isolated signal is then digitized via a 16bit ADC of the STM32F3 DSP at $12 KSps$, which gives 200 data samples per grid cycle.
Internally the digitization process runs asynchronously with respect to the DSP CPU, in order to minimize timing jitter.
It was verified that the sampling jitter of the ADC is less than 1us, however due to limited precision of equipment an exact figure was not established.
Data stream in its digital form is transferred to the Raspberry Pi single board computer (SBC) for analysis.

\begin{figure}[h]
  \begin{center}
  \includegraphics[width=0.9\textwidth]{images/opq-box/opqbox_software.png}
  \end{center}
  \caption{Block diagram of the OPQ Box 2 software stack.}
  \label{fig:opq:3}
\end{figure}

Raspberry Pi SBC is responsible for signal analysis and anomaly detection.
The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256MB of main memory and a single core 1GHz ARM11 CPU. Furthermore, Pi Zero W is equipped with an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle used in previous OPQ Box devices.
\subsection{Software}\label{subsec:software}
The software stack of the Raspberry Pi aims to deliver a full featured power quality analysis framework despite its rather limited hardware capabilities.
A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with the Pi acting as the master and the DSP as a slave device.
A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer.
During the initial design of the OPQ Box 2 software, SPI data transfer was attempted in userland.
However due to the lack of support for DMA in the SPI kernel-to-userland bridge, a large portion of the CPU time was spent facilitating data transfer, resulting in degraded analysis performance as well as missed data samples.
Current revision of the OPQ Box 2 software stack utilizes a kernel driver for management of SPI bus.
Internally the OPQ driver maintains a ring buffer of 16 windows each 200 data samples in size.
Upon receiving the interrupt for the DSP, the CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
This scheme requires the CPU to only service 60 interrupts a second, with each interrupt requiring on the order of 100 instructions, for a CPU utilization of less than $1\%$ in normal operation.
Userland applications communicate with the kernel driver using a file descriptor, where every $read$ system call yields 200 samples of raw waveform.
Thus the smallest window that a userland application may process is a single AC cycle of the grid mains.

The userland component of the OPQ Box 2 software is a multi-threaded extensible analysis framework called Triggering.
The reader thread is responsible for transferring and accumulating data from the kernel driver.
The smallest data buffer that the Triggering application processes at any given time is 10 grid cycles or 2k samples.
Once the cycles are transferred to the userland and timestamped, they are passed to the analysis thread for feature extraction, as well as to the Raw Data Ring Buffer (RDRB).
Since internally all data is addressed using shared pointers, during data duplication no copying is required.
RDRS is capable of buffering up to an hour of historic data before it's overwritten resulting in the RDBS maximum size of 100MB.

Analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
Four metrics are extracted from the data stream:
\begin{itemize}
	\item Fundamental frequency.
	\item RMS Voltage.
	\item Total Harmonic Distortion.
	\item Transient.
\end{itemize}

\subsection{Fundamental Frequency}\label{subsec:fundamental-frequency}

Fundamental frequency is calculated by computing the zero crossings of the AC waveform.
Since a sinusoid has two zero crossings for a full cycle the frequency can be calculated as:
\begin{equation} \label{eq:1}
 f = \frac{1}{\frac{2}{n}\sum\limits_{k=0}^{k=n}{\Delta t_{k}}}
\end{equation}

\begin{figure}[ht!]
	\centering
	\begin{subfigure}{.8\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{images/opq-box/filter1_gain.png}
	  \caption{}
	  \label{fig:opq:4:1}
	\end{subfigure}%

	\begin{subfigure}{.8\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{images/opq-box/filter1_response.png}
	  \caption{}
	  \label{fig:opq:4:2}
	\end{subfigure}
\end{figure}
\begin{figure}[ht!]\ContinuedFloat
	\centering
	\begin{subfigure}{.75\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{images/opq-box/filter2_gain.png}
	  \caption{}
	  \label{fig:opq:4:3}
	\end{subfigure}%

	\begin{subfigure}{.75\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{images/opq-box/filter2_response.png}
	  \caption{}
	  \label{fig:opq:4:4}
	\end{subfigure}
	\caption{Filters used for mains frequency calculation. (a) Downsampling filter gain. (b) Downsampling filter impulse response. (c) Lowpass filter gain. (d) Lowpass filter impulse response.}
	\label{fig:opq:4}
\end{figure}

Where the $\Delta t_{k}$ is the k'th time between two adjacent zero crossings.
In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz) it is very computationally expensive to perform this filtering in a single step.
Instead filtering is accomplished via a set of two finite impulse response (FIR) filters shown in Figure~\ref{fig:opq:4:2} and~\ref{fig:opq:4:4}.
First the Down sampling filter is applied to the raw waveform, which results in the frequency response shown in Figure~\ref{fig:opq:4:1}.
As is evident by the plot the frequency content of the result is 0-600Hz, Thus it can be downsampled to the $\frac{N}{10}$, or 200 samples without aliasing.
Next the low pass filter is applied to the downsampled waveform with the frequency response shown in Figure~\ref{fig:opq:4:3}.This resulting frequency content is 0-100Hz, thus all of the higher frequency harmonics and noise are removed.
Finally the twice filtered downsampled waveform is used to estimate the fundamental frequency according to the Equation~\ref{eq:1}.
The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.

All electrical generation systems connected to the grid run synchronously with each other, meaning that while small variations in voltage are common across locations, the fundamental frequency and phase must remain strictly in sync.
This effect is demonstrated in Figure~\ref{fig:opq:5}, which is a frequency fluctuation event recorded on November 8 2017.
While the two devices were separated by ten miles, their frequency measurements track closely together.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.6\linewidth]{images/opq-box/frequency_two_devices.pdf}
	\caption{Frequency measurement across two devices recorded during a lighting strike.}
	\label{fig:opq:5}
\end{figure}

\subsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}

Root mean square voltage ($V_{rms}$) in electrical power is the equivalent value of DC voltage which would dissipate the same power in the resistive load. $V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed value.
For the sinusoidal signal $V_{rms}$ can be calculated from the peak to peak value ($V_{pp}$) as:
\begin{equation} \label{eq:2}
	V_{rms} = \frac{V_{pp}}{2\sqrt{2}}
\end{equation}
It is common for multimeters to employ the equation above for computing $V_{rms}$.
However this equation is only valid for a perfect sinusoid, and thus does not result in a suitable metric for identifying power quality disturbances.
Instead OPQ Box 2 computes $V_{rms}$ as follows:
\begin{equation} \label{eq:3}
	V_{rms} = \sqrt{\frac{1}{n}\sum\limits_{k=0}^{k=n}V_{k}^{2}}
\end{equation}
Similarly to the frequency calculation OPQ Box 2 uses a 10 cycle window for a single $V_{rms}$ calculation, however unlike the frequency calculation the input is not filtered a priori.
An example of a power quality disturbance which exhibits low $V_{rms}$ is shown in Figure~\ref{fig:opq:6:1} and~\ref{fig:opq:6:2}.
These disturbances are the result of a lighting strike recorded by two OPQ Box 2 devices on Nov 1, 2017.

\begin{figure}[h]
		\centering
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=0.9\linewidth]{images/opq-box/voltage_sag.pdf}
	  \caption{}
	  \label{fig:opq:6:1}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=0.9\linewidth]{images/opq-box/voltage_sag_zoomed_in.pdf}
	  \caption{}
	  \label{fig:opq:6:2}
	\end{subfigure}
	\caption{A lightning strike recorded by two OPQ Box 2 devices separated by 10 miles. (a) A lightning strike manifested as a $V_{rms}$ dip which lasted 11 cycles. (b) As a consequence of using NTP these devices have a $\frac{1}{2}$ cycle mismatch in reported timestamps.}
	\label{fig:opq:6}
\end{figure}

\subsection{Total Harmonic Distortion}\label{subsec:thd}
OPQ Box calculates THD using industry standard methodology.
In the power delivery industry THD is defined as:
\begin{equation} \label{eq:4}
THD = \frac{\sqrt{\sum{V_{n}^2}}}{V_{f}}*100\%
\end{equation}
Where $V_{f}$ is the fundamental 60Hz power and $V_{n}$ is the power at $n^{th}$ harmonic.
It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
Operationally, OPQ Box computes THD for 10 cycles of the fundamental frequency.
First an FFT transforms the real voltage samples into its frequency components.
Next, the square of the harmonic bins is accumulated and scaled by the magnitude of the fundamental power.

\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{images/opq-box/thd_two_devices_24_hours.png}
	\caption{A common THD trend across two OPQ Box devices each deployed in the two Flexible Response to Ongoing Growth buildings on UH campus.}
	\label{fig:opq:7}
\end{figure}

Figure~\ref{fig:opq:7} shows a common trend observed by all OPQ Box devices installed on the UH campus.
For clarity only two devices are shown.
It is assumed that the large drop observed daily from approximately 6pm to 6am corresponds to the automatic response of the power delivery system to the reactive power in the grid, by deploying a large capacitor bank to compensate for the current phase lag.

\subsection{Transient Detection}\label{subsec:transient-detection}

OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter and searching for a maximum value in the remainder.
The high pass filter has a cutoff of $400Hz$, and the filter coefficients and response are shown in Figure~\ref{fig:opq:8:2} and Figure~\ref{fig:opq:8:1} respectively.
The result of the high pass filter operation is shown in Figure~\ref{fig:opq:8}.
Figure~\ref{fig:opq:8:3} shows a synthetic signal generated via a SDG1025 signal generator and fed into the OPQ Box.
This signal contains a $5V_{pp}$ transient injected at 11ms.
Filtered signal is shown in Figure~\ref{fig:opq:8:4}, with the fundamental removed and the transient preserved.
OPQ Box scans for the highest sample in the filtered waveform and uses its magnitude as a transient detection metric.

\begin{figure}[h]
	\centering
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{images/opq-box/thd_hpf_resp.pdf}
		\caption{}
		\label{fig:opq:8:1}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\linewidth]{images/opq-box/thd_hpf.pdf}
		\caption{}
		\label{fig:opq:8:2}
	\end{subfigure}

	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=0.86\linewidth]{images/opq-box/box_eval/5v_transient.pdf}
		\caption{}
		\label{fig:opq:8:3}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=0.86\linewidth]{images/opq-box/box_eval/5v_transient_filtered.pdf}
		\caption{}
		\label{fig:opq:8:4}
	\end{subfigure}
	\caption{THD detection filtering. (a) Filter gain. (b) Filter response. (c) A 5V transient superimposed onto a fundamental. (d) Filter result from (c).}
	\label{fig:opq:8}
\end{figure}

It should be noted, that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient detection.
This effect can be seen in Figure~\ref{fig:opq:9}.
This figure shows both the THD and transient detection metric during a transient event.
A small transient of approximately 1.6V was observed occurring at 2600s, while the sensitivity of the transient metric is clearly visible, particularly between 3000s and 4500s.

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=1\textwidth]{images/opq-box/trans_thd_det.pdf}
	\end{center}
	\caption{THD and Transient detection metric.}
	\label{fig:opq:9}
\end{figure}

\subsection{Network Communication}\label{subsec:network-communication}

Computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box 2 to verify its peer.
Internally metrics transmission uses ZeroMq's PUB/SUB protocol.
This protocol is a publish subscribe topology, with each message containing the topic, and a payload.
Additionally ZeroMq pub-sub topology allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
This allows for the aggregation service to be spread across multiple nodes, with minimal network overhead.

If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box 2 RDRB via a separate ZeroMq pub sub channel.
If the RDRB buffer contains data for the requested temporal range, OPQ Box 2 transmits the available data to the aggregation service via a push pull ZeroMq channel.
Protobuf message serialization is used to encode messages across the OPQ ecosystem.

In order to make a distributed measurement, all of the OPQ Boxes on the OPQ network need to maintain an accurate time reference.
Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
The expansion port of the OPQ Box 2 supports a GPS receiver.
However, since GPS receivers require line of sight to the sky it was not used for deployment.
NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a close by NTP server.
This error is visualized in a Figure~\ref{fig:opq:6:2}.
With a large coincidental $V_{pp}$ drop across two devices, a 7ms phase error is clearly visible.

### 2020-05-28 08-opq-view.tex
Add opq box section

\section{OPQ View}
\label{sec:opq-view}

OPQ View.

### 2020-05-29 05-opq-box.tex
OPQ Box section draft done.

@@
 	  \includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
-	  \caption{OPQ Box2 Block Diagram.
+	  \caption{OPQ Box Block Diagram.
 	  The power path is in red, signal path is in green and the digital IO is in black.}
@@
 	  \includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
-	  \caption{OPQ Box2 in an ABS plastic enclosure.}
+	  \caption{OPQ Box in an ABS plastic enclosure.}
 	  \label{fig:opq:1:2}
 	\end{subfigure}
-	\caption{(a) OPQ Box2 block diagram and (b) production OPQ Box ready for deployment}
+	\caption{(a) OPQ Box block diagram and (b) production OPQ Box ready for deployment}
 	\label{fig:opq:2}
@@
 
-The power system of the OPQ Box2 electrically isolates most of the device from the AC mains power.
+The power system of the OPQ Box electrically isolates most of the device from the AC mains power.
 An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
-5V is used to power the Raspberry Pi, equipment connected to the expansion port, 3.3V regulators and voltage reference and an isolated DC/DC converter.
+5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference and an isolated DC/DC converter.
 3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
@@
 
-The analog signal path of the OPQ Box2 is complicated by the fact that the STM32F3 ADC/DSP is electrically isolated from the mains power.
-A previous iteration of the OPQ Box, OPQ Box1, overcame this by employing a small circuit board mount isolation transformer.
+The analog signal path of the OPQ Box is complicated by the fact that the STM32F3 ADC/DSP is electrically isolated from the mains power.
+A previous iteration of the OPQ Box overcame this by employing a small circuit board mount isolation transformer.
 Unfortunately it was found that the frequency response of these transformers varied wildly between individuals, thus incurring a lengthy calibration process for each device.
-Design on the OPQ Box2 solved this issue by using an AMC1100 isolation amplifier as the isolation component.
-Internally AMC1100 consists of a single die comprised of a $\Sigma\Delta$ analog to digital and digital to analog converters.
+The current OPQ Box design solved this issue by using an AMC1100 isolation amplifier as the isolation component.
+Internally, the AMC1100 consists of a single die comprised of a $\Sigma\Delta$ analog to digital and digital to analog converters.
 These converters are separated by a silicon glass region on the integrated circuit which acts as a coupling capacitor.
 Since the signal passes the isolation boundary as a $\Sigma\Delta$ encoded digital signal, it incurs no distortion and no additional calibration is required.
-In order to match the dynamic range of the AMC1100 the $120V_{ac}$ is passed through the resistor divider to attenuate it to $120mV_{ac}$.
+In order to match the dynamic range of the AMC1100, the $120V_{ac}$ is passed through the resistor divider to attenuate it to $120mV_{ac}$.
 The input and output of the isolation amplifier is filtered with a passive first order anti-aliasing filter.
 The isolated signal is then digitized via a 16bit ADC of the STM32F3 DSP at $12 KSps$, which gives 200 data samples per grid cycle.
-Internally the digitization process runs asynchronously with respect to the DSP CPU, in order to minimize timing jitter.
-It was verified that the sampling jitter of the ADC is less than 1us, however due to limited precision of equipment an exact figure was not established.
-Data stream in its digital form is transferred to the Raspberry Pi single board computer (SBC) for analysis.
+Internally, the digitization process runs asynchronously with respect to the DSP CPU, in order to minimize timing jitter.
+We verified that the sampling jitter of the ADC is less than 1us, however, due to the limited precision of our equipment, an exact figure was not established.
+The data stream in its digital form is transferred to the Raspberry Pi single board computer (SBC) for analysis.
 
@@
 
-Raspberry Pi SBC is responsible for signal analysis and anomaly detection.
+A Raspberry Pi single board computer (SBC) is responsible for signal analysis and anomaly detection.
 The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256MB of main memory and a single core 1GHz ARM11 CPU.
-Furthermore, Pi Zero W is equipped with an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle used in previous OPQ Box devices.
+It also contains an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle.
+
 \subsection{Software}\label{subsec:software}
-The software stack of the Raspberry Pi aims to deliver a full featured power quality analysis framework despite its rather limited hardware capabilities.
+
+The software stack of the Raspberry Pi aims to deliver an accurate and precise power quality analysis framework despite the rather limited capabilities of the hardware.
 A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
 Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with the Pi acting as the master and the DSP as a slave device.
-A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer.
-During the initial design of the OPQ Box 2 software, SPI data transfer was attempted in userland.
-However due to the lack of support for DMA in the SPI kernel-to-userland bridge, a large portion of the CPU time was spent facilitating data transfer, resulting in degraded analysis performance as well as missed data samples.
-Current revision of the OPQ Box 2 software stack utilizes a kernel driver for management of SPI bus.
-Internally the OPQ driver maintains a ring buffer of 16 windows each 200 data samples in size.
+A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer, and a kernel driver provides management of the SPI bus.
+Internally, the OPQ driver maintains a ring buffer of 16 windows, each of which is 200 data samples in size.
 Upon receiving the interrupt for the DSP, the CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
@@
 Userland applications communicate with the kernel driver using a file descriptor, where every $read$ system call yields 200 samples of raw waveform.
-Thus the smallest window that a userland application may process is a single AC cycle of the grid mains.
+As a result, the smallest window that a userland application may process is a single AC cycle of the grid mains.
 
-The userland component of the OPQ Box 2 software is a multi-threaded extensible analysis framework called Triggering.
+The userland component of the OPQ Box software is a multi-threaded extensible analysis framework called Triggering.
 The reader thread is responsible for transferring and accumulating data from the kernel driver.
@@
 Since internally all data is addressed using shared pointers, during data duplication no copying is required.
-RDRS is capable of buffering up to an hour of historic data before it's overwritten resulting in the RDBS maximum size of 100MB.
+RDRS is capable of buffering up to an hour of data before it is overwritten, resulting in the RDBS maximum size of 100MB.
 
-Analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
-Four metrics are extracted from the data stream:
-\begin{itemize}
-	\item Fundamental frequency.
-	\item RMS Voltage.
-	\item Total Harmonic Distortion.
-	\item Transient.
-\end{itemize}
+The analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
+Four metrics are extracted from the data stream: (1) Fundamental frequency, (2) RMS Voltage, (3) Total Harmonic Distortion, and (4) Transients. Let's briefly discuss how each of these are computed.
 
-\subsection{Fundamental Frequency}\label{subsec:fundamental-frequency}
+\subsubsection{Fundamental Frequency}\label{subsec:fundamental-frequency}
 
-Fundamental frequency is calculated by computing the zero crossings of the AC waveform.
-Since a sinusoid has two zero crossings for a full cycle the frequency can be calculated as:
+The fundamental frequency is calculated by computing the zero crossings of the AC waveform.
+Since a sinusoid has two zero crossings for a full cycle, the frequency can be calculated as:
 \begin{equation} \label{eq:1}
@@
 Where the $\Delta t_{k}$ is the k'th time between two adjacent zero crossings.
+
 In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
-Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz) it is very computationally expensive to perform this filtering in a single step.
-Instead filtering is accomplished via a set of two finite impulse response (FIR) filters shown in Figure~\ref{fig:opq:4:2} and~\ref{fig:opq:4:4}.
-First the Down sampling filter is applied to the raw waveform, which results in the frequency response shown in Figure~\ref{fig:opq:4:1}.
-As is evident by the plot the frequency content of the result is 0-600Hz, Thus it can be downsampled to the $\frac{N}{10}$, or 200 samples without aliasing.
-Next the low pass filter is applied to the downsampled waveform with the frequency response shown in Figure~\ref{fig:opq:4:3}.This resulting frequency content is 0-100Hz, thus all of the higher frequency harmonics and noise are removed.
-Finally the twice filtered downsampled waveform is used to estimate the fundamental frequency according to the Equation~\ref{eq:1}.
+Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz), it is very computationally expensive to perform this filtering in a single step.
+Instead, filtering is accomplished via a set of two finite impulse response (FIR) filters shown in Figure~\ref{fig:opq:4:2} and~\ref{fig:opq:4:4}.
+First, the Down sampling filter is applied to the raw waveform, which results in the frequency response shown in Figure~\ref{fig:opq:4:1}.
+As is evident by the plot, the frequency content of the result is 0-600Hz, Thus, it can be downsampled to the $\frac{N}{10}$, or 200 samples without aliasing.
+
+Next, the low pass filter is applied to the downsampled waveform with the frequency response shown in Figure~\ref{fig:opq:4:3}.This resulting frequency content is 0-100Hz, thus all of the higher frequency harmonics and noise are removed.
+
+Finally, the twice filtered downsampled waveform is used to estimate the fundamental frequency according to the Equation~\ref{eq:1}.
 The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.
 
-All electrical generation systems connected to the grid run synchronously with each other, meaning that while small variations in voltage are common across locations, the fundamental frequency and phase must remain strictly in sync.
-This effect is demonstrated in Figure~\ref{fig:opq:5}, which is a frequency fluctuation event recorded on November 8 2017.
-While the two devices were separated by ten miles, their frequency measurements track closely together.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.6\linewidth]{images/opq-box/frequency_two_devices.pdf}
-	\caption{Frequency measurement across two devices recorded during a lighting strike.}
-	\label{fig:opq:5}
-\end{figure}
-
-\subsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}
+\subsubsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}
 
@@
 $V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed value.
-For the sinusoidal signal $V_{rms}$ can be calculated from the peak to peak value ($V_{pp}$) as:
+For a sinusoidal signal, $V_{rms}$ can be calculated from the peak to peak value ($V_{pp}$) as:
 \begin{equation} \label{eq:2}
@@
 It is common for multimeters to employ the equation above for computing $V_{rms}$.
-However this equation is only valid for a perfect sinusoid, and thus does not result in a suitable metric for identifying power quality disturbances.
-Instead OPQ Box 2 computes $V_{rms}$ as follows:
+However, this equation is only valid for a perfect sinusoid, and thus does not result in a suitable metric for identifying power quality disturbances.
+Instead, OPQ Box computes $V_{rms}$ as follows:
 \begin{equation} \label{eq:3}
@@
 \end{equation}
-Similarly to the frequency calculation OPQ Box 2 uses a 10 cycle window for a single $V_{rms}$ calculation, however unlike the frequency calculation the input is not filtered a priori.
+Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation, however unlike the frequency calculation the input is not filtered a priori.
+
 An example of a power quality disturbance which exhibits low $V_{rms}$ is shown in Figure~\ref{fig:opq:6:1} and~\ref{fig:opq:6:2}.
-These disturbances are the result of a lighting strike recorded by two OPQ Box 2 devices on Nov 1, 2017.
+These disturbances are the result of a lighting strike recorded by two OPQ Box devices on Nov 1, 2017.
 
@@
 
-\subsection{Total Harmonic Distortion}\label{subsec:thd}
-OPQ Box calculates THD using industry standard methodology.
-In the power delivery industry THD is defined as:
+\subsubsection{Total Harmonic Distortion}\label{subsec:thd}
+
+The OPQ Box calculates total harmonic distortion (THD) using the industry standard methodology.
+In the power delivery industry, THD is defined as:
 \begin{equation} \label{eq:4}
@@
 Where $V_{f}$ is the fundamental 60Hz power and $V_{n}$ is the power at $n^{th}$ harmonic.
+
 It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
@@
 	\includegraphics[width=1\linewidth]{images/opq-box/thd_two_devices_24_hours.png}
-	\caption{A common THD trend across two OPQ Box devices each deployed in the two Flexible Response to Ongoing Growth buildings on UH campus.}
+	\caption{A common THD trend across two OPQ Box devices deployed in two buildings on the University of Hawaii campus.}
 	\label{fig:opq:7}
@@
 
-Figure~\ref{fig:opq:7} shows a common trend observed by all OPQ Box devices installed on the UH campus.
-For clarity only two devices are shown.
+Figure~\ref{fig:opq:7} shows a common trend observed by OPQ Box devices installed on the UH campus.
+For clarity, only two devices are shown.
 It is assumed that the large drop observed daily from approximately 6pm to 6am corresponds to the automatic response of the power delivery system to the reactive power in the grid, by deploying a large capacitor bank to compensate for the current phase lag.
 
-\subsection{Transient Detection}\label{subsec:transient-detection}
+\subsubsection{Transient Detection}\label{subsec:transient-detection}
 
@@
 This signal contains a $5V_{pp}$ transient injected at 11ms.
-Filtered signal is shown in Figure~\ref{fig:opq:8:4}, with the fundamental removed and the transient preserved.
+The filtered signal is shown in Figure~\ref{fig:opq:8:4}, with the fundamental removed and the transient preserved.
 OPQ Box scans for the highest sample in the filtered waveform and uses its magnitude as a transient detection metric.
@@
 
-It should be noted, that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
+It should be noted that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
 However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient detection.
@@
 
-\subsection{Network Communication}\label{subsec:network-communication}
+\subsubsection{Network Communication}\label{subsec:network-communication}
 
-Computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
+The computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
 Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
-Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box 2 to verify its peer.
-Internally metrics transmission uses ZeroMq's PUB/SUB protocol.
+Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box to verify its peer.
+Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
 This protocol is a publish subscribe topology, with each message containing the topic, and a payload.
-Additionally ZeroMq pub-sub topology allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
+Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
 This allows for the aggregation service to be spread across multiple nodes, with minimal network overhead.
 
-If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box 2 RDRB via a separate ZeroMq pub sub channel.
-If the RDRB buffer contains data for the requested temporal range, OPQ Box 2 transmits the available data to the aggregation service via a push pull ZeroMq channel.
+If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box RDRB via a separate ZeroMq pub sub channel.
+If the RDRB buffer contains data for the requested temporal range, OPQ Box transmits the available data to the aggregation service via a push pull ZeroMq channel.
 Protobuf message serialization is used to encode messages across the OPQ ecosystem.
@@
 Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
-The expansion port of the OPQ Box 2 supports a GPS receiver.
-However, since GPS receivers require line of sight to the sky it was not used for deployment.
+The expansion port of the OPQ Box supports a GPS receiver.
+However, since GPS receivers require line of sight to the sky, it was not used for deployment.
 NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a close by NTP server.

### 2020-05-29 05-opq-box.tex
checkpoint

@@
 With a large coincidental $V_{pp}$ drop across two devices, a 7ms phase error is clearly visible.
+
+\subsubsection{Manufacturing}
+
+Currently, there is no mechanism for mass production of OPQ Boxes, but all of the plans are available under an Open Source license, so interested organizations with some basic hardware engineering skills can build their own boxes.
+
+Complete specifications for the OPQ Box hardware, firmware, and software are available \cite{negrashov_opq_2020}.
+As of the time of writing, a single OPQ Box can be manufactured for approximately \$100 in parts.
+The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately $75 in parts.

### 2020-05-30 02-intro.tex
OPQ Makai first draft done.

@@
 The remainder of this paper is organized as follows.
-Sections \ref{sec:system-architecture} and \ref{sec:information-architecture} present an overview of the OPQ system architecture and information architecture.
+Section \ref{sec:architecture} presents an overview of the OPQ system architecture and information architecture.
 Section \ref{sec:related-work} presents related research and technology.
@@
 Section \ref{sec:opq-mauka} describes our analysis services, and Section \ref{sec:opq-view} our user interface.
-Section \ref{sec:pilot} discusses the pilot deployment of a sensor network and its results.
+Section \ref{sec:pilot-study} discusses the pilot deployment of a sensor network and its results.
 Section \ref{sec:conclusions} concludes with some proposals for future research.

### 2020-05-30 06-opq-makai.tex
OPQ Makai first draft done.

\section{OPQ Makai}
\label{sec:opq-makai}

\subsection{Overview}

OPQ Box provides an inexpensive hardware device for collecting four important power quality measures with high fidelity, but realizing its potential requires an innovative approach involving two-way communication between OPQ Boxes and the OPQ cloud-based services. To see why, consider the IEEE 1159 standard for single location power quality monitoring \cite{unruh_ieee_2018}.  For transient monitoring, IEEE 1159 suggests a sampling rate of at least 7680 samples/second, up to 1 Megasample/second. This implies that if the cloud service requires the high fidelity data from all OPQ Boxes, it would incur a very large bandwidth cost. At 20 Ksamples/second with 16bit samples, a single OPQ Box will generate 300Kb/second of network bandwidth. Several thousand devices would easily saturate a 1 GB network link. In addition, collecting and recording all of the raw waveform data from residential power quality meters could lead to security and privacy issues.

Network bandwidth saturation is a common problem for distributed sensor networks, and a common solution is called "self-triggering".  In this approach, each monitoring device is configured with a threshold value for one or more measures of interest. When the threshold for a measure of interest is exceeded, then and only then is data sent over the network to cloud-based services for analysis.

The problem with the self-triggering approach is that grid-wide power quality events do not affect the entire grid in the same way. For example, due to the grid’s hierarchical structure, a voltage sag on one sub-circuit can manifest as as a sag of a different magnitude or even a swell on another \cite{kahle_power_2015}. This may result in a situation where some of the monitoring devices will not consider a power quality anomaly as an event, because it did not surpass the metric threshold, and simply ignore it. From an analysis perspective, however, it can be useful to get raw data from all of the affected devices, not just the ones that were the affected to the point where the box was triggered. This additional information can be used to localize the disturbance, as well as better evaluate its impact.

Since sending all the data is infeasible, and since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized event acquisition scheme which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai. In this scheme, OPQ Boxes use local processing resources to feature extract the incoming waveforms while storing them locally for an hour. Each OPQ Box sends its feature data to OPQ Makai once a second, which we called the "triggering stream". Feature data is very small, on the order of a few kilobytes, and so this approach allows the sensor network to scale to thousands of devices without incurring network bandwidth saturation.  OPQ Makai processes the incoming triggering stream and looks for anomalies. If an anomaly is present in only a single device, it is highly probable that the cause is local and not grid-level. On the other hand, if the triggering stream shows an anomaly temporally collocated across multiple devices, the entire network or a subset of the network may be queried for raw waveform data for a temporal region which corresponds to the disturbance in the triggering stream.

Our pilot study, discussed in Section \ref{sec:pilot-study} will provide examples of the novel analysis capabilities made possible by OPQ Box and OPQ Makai communication. In general, here are the main advantages of our hybrid centralized/decentralized approach over traditional self-triggering and the "naive" approach of sending all of the data:

{\em Bandwidth usage is minimized.} Instead of sending the entirety of raw data, only extracted features are sent. This results in a tiny fraction of the bandwidth requirement when compared to raw waveforms. Furthermore, the temporal window which encompasses a single feature can be adjusted in real time. Thus as soon as an anomalous behavior is observed in a subset of sensors, this window can be adjusted for a finer grained feature extraction.

{\em Effects of latency are minimized.} In this case, "latency" refers to the time required for OPQ Makai to process the incoming feature stream and decide whether or not to request high fidelity data from one or more OPQ Boxes. Even at 1M samples/second at 16 bits of resolution, the memory requirement to store 5 minutes of raw waveform without compression are on the order of 512MB, which is well within the realm of inexpensive single board computers such as Raspberry PI. With compression specifically suited to the signal of interest, the memory requirement can be reduced even further. In the case of the OPQ sensor network, OPQ Makai has an hour to process feature data and request high fidelity data from OPQ Boxes. During our pilot study, OPQ Makai always responded within a second or two, although the sensor network was very small with only 15 OPQ Boxes.

{\em Cloud processing requirements are reduced.} Since feature extraction is already performed at the device level, cloud computation requirements are reduced. With the advent of the Internet of Things, the computational capacity of edge devices is increasing.

{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes. This allows analysis of data from devices which were only mildly affected or even not affected at all by the disturbance. This creates new possibilities for investigation of disturbance propagation across the sensed area, as will be discussed in Section \ref{sec:pilot-study}.

{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features. This allows for a simple robust algorithm which may be deployed at the sink node for anomaly detection.

\subsection{Design}

OPQ Makai is a distributed extensible microservice framework responsible for receiving the triggering stream from the OPQ Boxes, locating anomalous temporal regions and requesting raw waveform for the anomalous time ranges. As shown in Figure \ref{fig:makai-design}, Makai consists of four major components: Acquisition Broker, Triggering Broker, Event Service and the Acquisition Service.

\begin{figure}
\center \includegraphics[width=3in]{images/makai/makai_main.pdf}
\caption{Makai component design}
\label{fig:makai-design}
\end{figure}

The Triggering Broker is perhaps the simplest component of the OPQ Makai system. The triggering stream generated by the OPQ Boxes is encrypted to preserve user privacy. In order to minimize the CPU time spent decrypting the data across multiple OPQ services, the Triggering Broker decrypts the data and sends clear text measurements to other OPQ cloud services. The Triggering Broker uses the ZeroMq subscribe socket to receive data from OPQ Boxes, and sends it via a publish socket to any connected client. Each publish message is published to a topic which corresponds to the ASCII representation of the originating OPQ Box ID. This allows services which utilize the Triggering Broker to select a subset of IDs to operate on. This is useful for load balancing the backend services, or dividing the OPQ network into separate regions with no electrical connections between them.

The Acquisition Broker manages the two-way communication between the OPQ Boxes and the rest of the cloud infrastructure. Unlike the triggering stream which originates from the OPQ Box, two way communication is always initiated by OPQ cloud services. Two way communication is realized via a command response interface, where the OPQ service initiates the communication by sending a clear text command to the Acquisition Broker, which then forwards it in encrypted form to the appropriate OPQ Boxes.

The Acquisition Service resides between the Triggering and Acquisition Brokers. The Acquisition Service is responsible for three tasks:
(1) Computation of statistics of the incoming triggering stream; (2) Hosting plugins for triggering stream analysis; and (3) Generating data event requests for OPQ Boxes. The Acquisition Service accesses the triggering stream by connecting to the publish socket of the Triggering broker. Since the connection is managed through the ZeroMq publish-subscribe socket, several Acquisition Services can be connected to a single Triggering broker endpoint, each servicing a subset of OPQ Boxes by subscribing to only specific devices. The Acquisition Service does not include any analysis capabilities by default. Instead, analysis is performed by shared library loadable plugins. These plugins can be loaded and unloaded at runtime, thus allowing live upgrading and testing of new analysis methods.

The Event service is a microservice which stores raw data generated by OPQ Boxes in the MongoDB Database. On initialization, the Event service queries MongoDB database for the highest event number recorded so far, connects to the Acquisition Broker’s publish port, and subscribes to all messages that start with the prefix “data”. This allows the Event service to capture every response from OPQ Boxes generated from commands issued by the Acquisition service plugins. Once the Event service receives a data response with an identity containing an event token it hasn’t seen before, it will increment the event number, and store it in an internal key value store.

### 2020-05-30 09-pilot.tex
OPQ Makai first draft done.

\section{Results from the Pilot Deployment}
\label{sec:pilot-study}

Pilot study.

### 2020-05-31 06-opq-makai.tex
Started on mauka

@@
 
-Since sending all the data is infeasible, and since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized event acquisition scheme which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai.
+Since sending all the data is infeasible, and since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai.
 In this scheme, OPQ Boxes use local processing resources to feature extract the incoming waveforms while storing them locally for an hour.
 Each OPQ Box sends its feature data to OPQ Makai once a second, which we called the "triggering stream".
-Feature data is very small, on the order of a few kilobytes, and so this approach allows the sensor network to scale to thousands of devices without incurring network bandwidth saturation.
+Feature data is very small, on the order of a few kilobytes, and so this approach allows the sensor network to scale to thousands of devices with acceptable network bandwidth requirements.
 OPQ Makai processes the incoming triggering stream and looks for anomalies.

### 2020-05-31 07-opq-mauka.tex
Started on mauka

\section{OPQ Mauka}
\label{sec:opq-mauka}

The previous sections discussed the design of OPQ Box, a custom hardware device for collecting four important measures of power quality, and OPQ Makai, a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes.  As a result of these two systems, the OPQ sensor network services will have the ability to collect and analyze high fidelity data about power quality anomalies in a cost-effective, scalable fashion.

There are remaining challenges to creating a useful power quality sensor network. First, the data provided by OPQ Boxes is low-level, "primitive" data consisting of either features (i.e. frequency, voltage, THD, and transients) or waveform data. But what we actually want is actionable insights into grid stability. For example, we might want to know if a given anomalous data value is actually detrimental, or we might wnat to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.

A second challenge is data volume. Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data will accumulate. One strategy is to simply store all of the data sent to the cloud forever. This means that data storage requirements will increase monotonically over time, making the sensor network more costly to maintain the longer it is in place. An alternative strategy is to implement an algorithm to identify uninteresting (or no longer interesting) data and discard it.  Ideally, such an algorithm would enable OPQ sensor network designers to calculate an upper bound on the total amount of cloud storage required as a function of the number of nodes (OPQ Boxes) in the network.

OPQ Mauka addresses both of these issues. First, OPQ Mauka provides a multi-leveled representation for structuring and processing DSN data. The structure and processing at each level is designed with the explicit goal of turning low-level data into actionable insights. Second, each level in the framework implements a "time-to-live" (TTL) strategy for data within the level. This strategy states that data must either progress upwards through the levels towards more abstract, useful representations within a fixed time window, or be discarded and lost forever. The TTL strategy is useful because when implemented, it allows DSN designers to calculate upper bounds on data storage at each level of the framework and supports graceful degradation of DSN performance.

### 2020-06-01 07-opq-mauka.tex
checkpoint

@@
 The TTL strategy is useful because when implemented, it allows DSN designers to calculate upper bounds on data storage at each level of the framework and supports graceful degradation of DSN performance.
+
+Figure \ref{fig:mauka-data-model} illustrates the hierarchical data model for OPQ Mauka.
+This data model can be conceptualized as a multi-level hierarchy that adaptively optimizes data storage using a tiered TTL approach and provides a mechanism in which typed aggregated data is continually refined to the point of being of becoming actionable.
+The data model also includes software components called "actors" that both move data upward through the levels and apply optimizations downward through the levels.
+Actors are implemented through a plugin architecture, making it easy to experiment with the data model and improve it over time.
+
+\begin{figure}
+\center \includegraphics[width=4in]{images/mauka/mauka-data-model.png}
+\caption{Mauka data model hierarchy}
+\label{fig:mauka-data-model}
+\end{figure}
+
+OPQ Mauka sits between OPQ Makai and OPQ’s database.
+It receives messages from OPQ Makai when it has triggered OPQ Boxes due to something interesting in the low-fidelity data stream.
+OPQ Mauka then retrieves the raw waveforms stored in the database by OPQ Makai, performs feature extraction, and then forwards the features and/or raw waveform to OPQ Mauka Actors to perform classification and other high level analytics.
+These results are then stored in the database.
+Finally, TTL Actors run at predetermined intervals to remove data whose time-to-live has expired.
+
+\subsection{OPQ Mauka Actors}
+
+The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins.
+
+{\em BasePlugin.} All Actor plugins derive from the BasePlugin, which provides primitives for running plugins as separate processes, serialization and deserialization of type safe messages, communication over ZeroMQ, metrics collection, MongoDB communication, debugging, and plugin health.
+
+{\em MakaiEventPlugin.} The MakaiEventPlugin Actor is responsible for reading Events newly created by OPQ Makai into OPQ Mauka.
+It performs feature extraction on the raw Event data streams and forwards those features (or the raw data) to subscribing plugins.
+This allows OPQ Mauka to perform feature extraction once, and reuse those features in multiple plugins.
+
+{\em FrequencyVariationPlugin.} The FrequencyVariationPlugin is used to classify generic fre- quency sags, swells, and interruptions as defined by the IEEE1159 standard[29].
+Both duration and deviation from nominal are used to perform these classifications.
+Duration classifications in- clude frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
+Classifications for deviations from nominal are performed for values that are up to 40% deviation
+     49
+
+     nominal. The plugin subscribes to messages from the “WindowedFrequency” topic which includes a payload of frequency features over one electrical cycle.
+     When a message is received, the plugin uses configurable thresholds to find frequency fluctu- ations within the windowed frequency streams.
+The plugin is able to classify frequency swells, frequency interruptions, and frequency sags.
+When a new frequency variation is detected, a fre- quency Incident is created

### 2020-06-03 07-opq-mauka.tex
Finished rough draft of Mauka.

@@
 The previous sections discussed the design of OPQ Box, a custom hardware device for collecting four important measures of power quality, and OPQ Makai, a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes.
-As a result of these two systems, the OPQ sensor network services will have the ability to collect and analyze high fidelity data about power quality anomalies in a cost-effective, scalable fashion.
+As a result of these two innovations, an OPQ sensor network has the ability to collect and analyze high fidelity, low level data about power quality anomalies in a cost-effective, scalable fashion.
 
@@
 
-A second challenge is data volume.
-Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data will accumulate.
+A second challenge involves the potentially high volume of data might accumulate in the cloud.
+Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data could accumulate.
 One strategy is to simply store all of the data sent to the cloud forever.
@@
 This strategy states that data must either progress upwards through the levels towards more abstract, useful representations within a fixed time window, or be discarded and lost forever.
-The TTL strategy is useful because when implemented, it allows DSN designers to calculate upper bounds on data storage at each level of the framework and supports graceful degradation of DSN performance.
+The TTL strategy is useful because when implemented, it allows DSN designers to make reasonable predictions of the upper bounds on data storage at each level of the framework, and making possible a "graceful degradation" of network performance if those bounds turn out to be exceeded.
 
@@
 
-OPQ Mauka sits between OPQ Makai and OPQ’s database.
-It receives messages from OPQ Makai when it has triggered OPQ Boxes due to something interesting in the low-fidelity data stream.
-OPQ Mauka then retrieves the raw waveforms stored in the database by OPQ Makai, performs feature extraction, and then forwards the features and/or raw waveform to OPQ Mauka Actors to perform classification and other high level analytics.
-These results are then stored in the database.
-Finally, TTL Actors run at predetermined intervals to remove data whose time-to-live has expired.
+{\bf Anthony: Please review the following paragraphs for accuracy and quality.}
+
+In essence, the raw data sent to the cloud from OPQ Boxes is initially stored in the Instantaneous Measurements Layer (IML).
+IML Actors process this data to determine if additional data from OPQ Boxes are warranted, and request this data if necessary.
+This data, along with the instantaneous data, is then stored at the Aggregate Measurement Layer (AML).
+All data from the IML layer is discarded after one hour if it is not found to be interesting and thus promoted to the AML level.
+Once at the AML Layer, AML actors process the augmented data sets to determine if they constitute an actual power quality anomaly, and if multiple OPQ Boxes appear to be implicated in the anomaly.
+If so, all of the associated AML data is promoted to the Incident Layer (IL).
+AML data must be promoted within one day or else it is discarded.
+Finally, IL Actors are responsible for looking for ways to make Incidents actionable, and if successful, will promote the IL data to the Phenomena Layer (PL).
+For example, if the same boxes are generating the same Incidents on a periodic basis, then these Incidents could be promoted to a Phenomena that can predict when the next occurrence of this Incident might occur.
+
+While IML, AML, DL, and IL Actors all promote data upward, the Phenomena Actor actually works in the opposite direction, using its knowledge to configure lower levels.
+For example, if a periodic Phenomena has been created, then a PL Actor might alter the behavior at the AML, gathering more raw data from Boxes around the time of the next predicted Incident.
 
@@
 
-{\em FrequencyVariationPlugin.} The FrequencyVariationPlugin is used to classify generic fre- quency sags, swells, and interruptions as defined by the IEEE1159 standard[29].
+{\em FrequencyVariationPlugin.} The FrequencyVariationPlugin is used to classify generic frequency sags, swells, and interruptions as defined by the IEEE 1159 standard.
 Both duration and deviation from nominal are used to perform these classifications.
-Duration classifications in- clude frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
-Classifications for deviations from nominal are performed for values that are up to 40% deviation
-     49
+Duration classifications include frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
+Classifications for deviations from nominal are performed for values that are up to 40\% deviation from nominal.
+The plugin is able to classify frequency swells, frequency interruptions, and frequency sags.
+When a new frequency variation is detected, a frequency Incident is created.
 
-     nominal. The plugin subscribes to messages from the “WindowedFrequency” topic which includes a payload of frequency features over one electrical cycle.
-     When a message is received, the plugin uses configurable thresholds to find frequency fluctu- ations within the windowed frequency streams.
-The plugin is able to classify frequency swells, frequency interruptions, and frequency sags.
-When a new frequency variation is detected, a fre- quency Incident is created
+{\em IEEE1159VoltagePlugin.} The IEEE1159VoltagePlugin is used to classify voltage Incidents in accordance with the IEEE 1159 standard[29].
+In general, this standard classifies voltage disturbances by duration and by magnitude.
+Voltage durations are classified from 0.5 to 30 cycles, 30 cycles to 3 seconds, 3 seconds to a minute, and greater than 1 minute.
+Voltage deviations are classified in both the sag and swell directions as a percentage from nominal.
+Sags are generally classified between 10\% and 90\% of nominal while swells are generally classified from 110\% to 180\% of nominal.
+This plugin is capable of classifying voltage sags, swells, and interruptions as defined by the standard.
+The plugin works by identifying sags, swells, and interruptions and determining the duration of those Events.
+If the duration or delta is large enough, a voltage Incident is created.
+
+{\em BoxOptimizationPlugin.} The BoxOptimizationPlugin is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
+This plugin is capable of requesting the state of each OPQ Box (e.g.
+uptime, Measurement rate, security keys, etc).
+This plugin is also capable of adjusting the state of individual OPQ Boxes by changing things such as the Measurement and Trend rate or the sampling rate that the Box is sampling at.
+
+{\em FuturePhenomenaPlugin.} The FuturePhenomenaPlugin is responsible for creating Future or Predictive Phenomena.
+These Phenomena are used to predict Events and Incidents that may occur in the future.
+This plugin does not subscribe to any messages, but instead utilizes timers to perform its work.
+By default, this plugin runs every 10 minutes.
+
+When the plugin runs, it loads any active Periodic Phenomena found in the database.
+If Periodic Phenomena are found, this plugin extrapolates possible Events and Incidents by first examining the timestamps of previous periods and then extrapolating into the future using the mean period and the standard deviation.
+For each timestamp in the periodic phenomena, the mean period is added.
+If the resulting timestamp is in the future, a Future Phenomena is created using the time range of the future timestamp plus or minus the standard deviation of the Periodic Phenomena.
+
+When a Future Phenomena is created, timers are started in a separate thread signifying the start and end timestamps of the Future Phenomena.
+When the first timer runs, messages are sent to the BoxOptimizationPlugin and the ThresholdOptimizationPlugin instructing Event thresholds to be set lower and Box Measurement rates to be set higher.
+This increases the chance of seeing an Event over the predicted time window.
+When the second timer runs, these values are reset to their default values.
+Thus, the plugin increases fidelity and decreases thresholds over the period of a Future Phenomena.
+
+{\em ITICPlugin.} The ITICPlugin analyzes Vrms to determine where it falls within the ITIC curve \cite{thallam_power_2000}.
+The ITIC curve is a power acceptability curve that plots time on the x-axis and Vrms on the y-axis.
+The purpose of the curve is to provide a tolerance envelope for single-phase 120V equipment.
+The curve defines three regions.
+The first region is "No Interruption" and generally includes all voltages with very short sustained durations.
+All events within this region have no noticeable effect on power equipment.
+The second region, the "No Damage Region", occurs during voltage sags for extended periods of time.
+Power Events in this region may cause equipment interruptions, but it will not damage the equipment.
+The final region, the "Prohibited" region, is caused by sustained voltage swells and may cause damage to power equipment.
+This plugin determines if an event falls within the "No Damage" or "Prohibited" regions and if so, creates an Incident to record this.
+
+{\em SemiF47Plugin.} The SemiF47 plugin is another plugin like the IticPlugin that plots voltage and duration against a power acceptability curve.
+In this case, the standard used is the SemiF47 standard \cite{djokic_sensitivity_2005}.
+Rather than using a point-in-polygon approach, this plugin reads the VRMS features sequentially and uses a state machine to keep track of the current classification.
+This plugin only classifies values as a "violation" or as "nominal".
+
+{\em TransientPlugin.] The "TransientPlugin" is responsible for classifying frequency transients in power waveforms.
+The plugin subscribes to messages from the "RawVoltage" topic which contains a calibrated power waveform payload.
+The TransientPlugin is capable of classifying impulsive, arcing, oscillatory, and periodic notching transients.
+A decision tree is utilized to select the most likely transient type and then further analysis is used to perform the actual classification of transients.
+Dickens et al \cite{dickens_transient_2019} provides more details on the transient classification system.
+
+{\em PeriodicityPlugin.} The "Periodicity Plugin" is responsible for detecting periodic signals in power data.
+This plugin does not subscribe to any messages, but instead runs off of a configurable timer.
+The plugin is set to run by default once an hour and every hour it scrapes the last 24 hours worth of data and attempts to find periods in the Measurements over that duration.
+
+For each feature in the Measurement and Trend data (e.g.
+frequency, voltage, and THD), the Periodicity plugin first removes the DC offset from the data by subtracting the mean.
+Next, the plugin filters the signal using a 4th order high-pass filter to filter out noise.
+The plugin then performs autocorrelation on the signal followed by finding the peaks of the autocorrelation.
+The mean distance between the peaks of the autocorrelation provides the period of the signal.
+
+The plugin only classifies data as periodic if at least 3 peaks were found and the standard deviation of the period is less than 600 seconds (10 minutes).
+Once a positive identification has been made, peak detection is performed on the original signal.
+Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the plugin can group Measurements, Trends, Events, and Incidents that were created during the periodic signals together as part of the Periodic Phenomena.

### 2020-06-04 08-opq-view.tex
OPQ View first rough draft

@@
 
-OPQ View.
+The final component of the OPQ Sensor Network system architecture is called OPQ View.
+It is a web application, implemented using the Meteor application framework, which provides a variety of visualization and query services.
+An example of the OPQ View home page is provided in Figure \ref{fig:opq-view-home}.
+
+\begin{figure}
+\center \includegraphics[width=5in]{images/view/homepage.png}
+\caption{OPQ View: home page}
+\label{fig:opq-view-home}
+\end{figure}
+
+The home page provides three components:  a "System Stats" window which indicates the number of data elements currently at each level of the OPQ Mauka data hierarchy, a "System Health" window that indicates whether or not the Cloud services appear to be running correctly and the status of communication with all known OPQ Boxes on the sensor network, and a "Box Trends" visualization that provides the ability to quickly see trends in the four basic measures (Voltage, Frequency, THD, and Transients) over time.
+
+{\em Anthony: a couple of questions about this home page image.
+First, no phenomena?
+What's up with that?
+Second, although the TTL for Measurements is set to 1 day, the System Status window says there are 16M measurements.
+It seems to me that with 1 measurement per second per box, we should only have around 1.2M per day, so at most double that given the TTL.
+Can you explain to me what's happening?
+}
+
+Figure \ref{fig:opq-view-box-map} shows a map-based view of the sensor network.
+This figure shows the location of 11 OPQ Boxes on the University of Hawaii campus at one point during Fall of 2019.
+Depending on the zoom level of the interface, some boxes are collapsed into disks with a number indicating the number of boxes at that location.
+Zooming in reveals more information about the box including near-real time values for voltage and frequency, as is shown for the box in the upper right corner of the figure.
+
+\begin{figure}
+\center \includegraphics[width=5in]{images/view/boxmap-2.png}
+\caption{OPQ View: Box Map}
+\label{fig:opq-view-box-map}
+\end{figure}
+
+Figure \ref{fig:opq-view-incident-summary} shows a visual display of a single Incident.
+This view includes a map-based location of the box whose data was involved in the Incident, the start and end time and duration of the Incident, the classification(s) of the anomalous data, and the waveform associated with the Incident (when applicable).
+
+\begin{figure}
+\center \includegraphics[width=5in]{images/view/incident-summary.png}
+\caption{OPQ View: Incident Summary}
+\label{fig:opq-view-incident-summary}
+\end{figure}
+
+The above images provide a sense for how OPQ View helps users to understand, monitor, and assess an OPQ Sensor Network and the underlying power quality of the grid it is attached to.
+There are several features in OPQ View for user and box management; for details, please see the OPQ View documentation.

### 2020-06-04 08-opq-view.tex
add update

@@
 This view includes a map-based location of the box whose data was involved in the Incident, the start and end time and duration of the Incident, the classification(s) of the anomalous data, and the waveform associated with the Incident (when applicable).
+If additional analysis is desired, the raw data can be downloaded in CSV format.
 
@@
 There are several features in OPQ View for user and box management; for details, please see the OPQ View documentation.
+
+While OPQ View provides a variety of useful visualization and analysis features, users wishing to understand the power quality of their grid are not restricted to its current feature set.
+The OPQ database schema is public, and if a new analysis is desired, it is straightforward to query the database directly for the data of interest and build a new analysis off of it.

### 2020-06-05 07-opq-mauka.tex
OPQ View first rough draft
initial start on pilot

@@
 
-{\bf Anthony: Please review the following paragraphs for accuracy and quality.}
+\begin{tcolorbox}
+Anthony: Please review the following paragraphs for accuracy and quality.
+\end{tcolorbox}
 
@@
 
-{\em TransientPlugin.] The "TransientPlugin" is responsible for classifying frequency transients in power waveforms.
+{\em TransientPlugin.} The "TransientPlugin" is responsible for classifying frequency transients in power waveforms.
 The plugin subscribes to messages from the "RawVoltage" topic which contains a calibrated power waveform payload.

### 2020-06-05 09-pilot.tex
initial start on pilot

@@
 
-Pilot study.
+To evaluate the capabilities of the OPQ Sensor Network, 15 OPQ Boxes were deployed at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019.
+This location was advantageous because it is an isolated microgrid connected to the Oahu powergrid only via a single 46kV feeder as shown in Figure 4.5.
+Another advantage of the UH campus is the high number of smart meters deployed across various levels of the power delivery infrastructure.
+While the purpose of these meters is to monitor the power consumption, they do include some rudimentary power quality monitoring capabilities.
+Data from the campus deployed meters was used as ground truth for comparison against the measurements, and for analysis performed by the OPQ project.
+The location of smart meters in the grid topology is shown in Figure 4.5 as the M nodes.
+As evident by the meter location none of them were monitoring the consumer level power and mainly focused on the higher voltage power delivery.
+This placement was a consequence of the smart meters’ role as a consumption monitor, and thus the deployment of the OPQ Boxes at the residential level complemented UH power quality monitoring capabilities without introducing redundancies.
+                                                                                                                                    The University of Hawaii power grid supplies a highly diverse infrastructure.
+Beyond traditional residential equipment such as computers and consumer grade electronics, the UH power grid powers scientific and laboratory equipment, machine shops, and server farms.
+All of these elements have varying requirements/tolerances for power quality anomalies as well as different levels of power quality “pollution”.
+Furthermore, some of the electricity consumers in the UH campus are entirely unique.
+For example, the free electron laser located in the Watanabe Hall is one of the only free electron lasers in the world, and the impact/sensitivity of power quality on the instrument are completely unstudied.
+
+\subsection{OPQ Boxes provide valid and reliable collection of power quality data}
+
+\subsection{The OPQ Information Architecture provides a means to produce actionable insights}
+
+\subsection{The OPQ Information Architecture provides predictable upper bounds on storage resources}
+
+\subsection{OPQ Mauka provides useful adaptive optimization capabilities}

### 2020-06-06 09-pilot.tex
updates

--- 2020-05-30 09-pilot.tex
@@
 
-Pilot study.
+To evaluate the capabilities of the OPQ Sensor Network, 15 OPQ Boxes were deployed at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019.
+The University of Hawaii campus is an isolated microgrid connected to the Oahu powergrid only via a single 46kV feeder.
+The UH Campus also has traditional, revenue-grade electrical meters deployed across various levels of the power delivery infrastructure.
+While the primary purpose of these meters is to monitor power consumption, they do include power quality monitoring capabilities.
+Data from these meters were used as ground truth for validation studies of the OPQ sensor network.
+
+The University of Hawaii power grid is interesting in that it supplies a highly diverse infrastructure.
+Beyond traditional residential equipment such as computers and consumer grade electronics, the UH power grid powers scientific and laboratory equipment, machine shops, and server farms.
+All of these elements have varying requirements/tolerances for power quality anomalies as well as different levels of power quality “pollution”.
+Furthermore, some of the electricity consumers in the UH campus are entirely unique.
+For example, the free electron laser located in the Watanabe Hall is one of only ten in the world, and the impact/sensitivity of power quality on the instrument are completely unstudied.
+
+The pilot study had the following major goals:
+
+\begin{enumerate}
+
+\item {\em Evaluation of sensor data accuracy and quality.} While we conducted extensive laboratory tests of the sensor network, the pilot study enabled us to evaluate the accuracy and utility of our hardware devices and power quality data collection procedures in a real-world setting.
+Does an OPQ sensor network measure power quality (and detect anomalies) as well or better than conventional building-level electrical meters?
+
+\item {\em Evaluation of the triggering system.} The OPQ Makai triggering system provides a novel two-way communication between sensors and the cloud, resulting in a triggering system with unique capabilities.
+Does the data collected by this triggering system provide analysis opportunities not available through more conventional approaches?
+
+\item {\em Evaluation of the information architecture.} OPQ Mauka implements a tiered, hierarchical information architecture.
+The pilot study enabled us to evaluate this information architecture.
+Would this information architecture prove useful with real world data?
+Could actionable insights emerge from low-level sensor data?
+
+\item {\em Evaluation of cloud data storage management.} OPQ Mauka manages data storage through a "time to live" (TTL) mechanism, which results in data being discarded if it is not found to be useful and thus "promoted" to a higher level of the information hierarchy within a defined period of time.
+The pilot study enables us to evaluate this approach.
+Does this mechanism allow sensor network designers to determine upper bounds on storage requirements?
+Is information thrown away prematurely because it was not found to be useful within the time limits?
+
+\end{enumerate}
+
+\subsection{Descriptive statistics}
+
+\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+
+Could you provide descriptive statistics for this section?
+This includes the start and end dates of the pilot study, the total number of data elements collected at each level of the information hierarchy, and for incidents and phenomena, the total number of each type of incident and event.
+\end{tcolorbox}
+
+\subsection{OPQ Boxes provide valid and reliable collection of power quality data}
+
+\subsection{The OPQ Information Architecture provides a means to produce actionable insights}
+
+\subsection{The OPQ Information Architecture provides predictable upper bounds on storage resources}
+
+\subsection{OPQ Mauka provides useful adaptive optimization capabilities}

### 2020-06-09 01-abstract.tex
Updates to pilot study section

@@
 This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) are blurring: end-users both produce and consume energy, making energy management and public policy more complex.
-The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to make useful information about electrical grids (and their stability) available to producers, consumers, researchers, and policy makers.
-In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.
-Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions.
+The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to provide useful information about electrical grids (and their stability) in this new era to producers, consumers, researchers, and policy makers.
+In 2019, we performed a pilot study involving the deployment of an OPQ sensor network one the University of Hawaii microgrid for three months.
+Results of the pilot study provide evidence that OPQ provides a variety of useful services and that the system could be scaled to service larger geographic regions.
 We conclude that OPQ provides a new and useful approach to power quality monitoring that can be used in conjunction with other technologies .}

### 2020-06-09 02-intro.tex
Updates to pilot study section

@@
 
-Successfully maintaining adequate power quality and providing sufficient amounts of it to meet the rising needs of consumers has been a triumph of electrical utilities for over 100 years.
+Successfully maintaining adequate power quality and providing sufficient amounts of electrical power to meet the rising needs of consumers has been a triumph of electrical utilities for over 100 years.
 In recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing concern and interest.
@@
 Third, modern consumer electronics place more stringent demands on power quality.
-Finally, effective public policy making in this modern context can be aided by public access to power quality data.
-Let's look at each of these in a bit more detail.
+Finally, effective public policy making in this modern context can be aided by increased public access to power quality data.
+Let's examine each of these assertions in more detail.
 

### 2020-06-09 03-architecture.tex
Add opq box section
OPQ Makai first draft done.
Updates to pilot study section

@@
-\section{System Architecture}
-\label{sec:system-architecture}
+\section{OPQ Architecture}
+\label{sec:architecture}
+
+\subsection{System Architecture}
 
@@
 \begin{figure}
-\center \includegraphics[width=5in]{images/system-diagram.png}
+\center \includegraphics[width=5in]{images/architecture/system-diagram.png}
 \caption{High level system architecture of an OPQ Sensor Network}
@@
 
+\begin{tcolorbox}[title=ANTHONY AND SERGE]
+You might notice that there is no mention of "Napali" or "Laha" in this paper.
+I could not think of a good way to add them in that would not additionally complicate any already terminologically complicated topic.
+To me, they feel like "overlays" onto the four component architecture described here and thus not strictly necessary to understand the system, but let me know if you disagree.
+\end{tcolorbox}
+
 This architecture has a number of benefits.
@@
 
-\section{Information Architecture}
-\label{sec:information-architecture}
+\subsection{Information Architecture}
 
@@
 
-\begin{table}[H]
+\begin{table}[ht]
 \caption{OPQ Information Architecture Levels}

### 2020-06-09 09-pilot.tex
Updates to pilot study section

@@
 \item {\em Evaluation of the triggering system.} The OPQ Makai triggering system provides a novel two-way communication between sensors and the cloud, resulting in a triggering system with unique capabilities.
-Does the data collected by this triggering system provide analysis opportunities not available through more conventional approaches?
+Does the triggering system provide efficiencies not available from more conventional approaches?
+Does it provide data enabling analysis opportunities not available through more conventional approaches?
 
@@
 This includes the start and end dates of the pilot study, the total number of data elements collected at each level of the information hierarchy, and for incidents and phenomena, the total number of each type of incident and event.
+Perhaps the size of the database?
 \end{tcolorbox}
@@
 \subsection{OPQ Boxes provide valid and reliable collection of power quality data}
+
+The first goal of the pilot study was to assess how well an OPQ sensor network is able to collect basic power quality data.
+To do this, we compared data on voltage and frequency collected by OPQ Boxes with data on voltage and frequency collected by the revenue-grade building level power monitors.
+The results are shown in Figure \ref{fig:opqbox-f-v-validation}.
+
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}{.5\textwidth}
+	  \centering
+	  \includegraphics[width=0.9\linewidth]{images/pilot/opqbox-frequency-validation.png}
+	  \caption{Frequency differences}
+	  \label{fig:opqbox-validation-1}
+	\end{subfigure}%
+	\begin{subfigure}{.5\textwidth}
+	  \centering
+	  \includegraphics[width=0.9\linewidth]{images/pilot/opqbox-voltage-validation.png}
+	  \caption{Voltage differences}
+	  \label{fig:opqbox-validation-2}
+	\end{subfigure}
+	\caption{Validation of (a) Frequency and (b) Voltage}
+	\label{fig:opqbox-f-v-validation}
+\end{figure}
+
+As the charts illustrate, there is very little difference between the frequency and voltage values collected by OPQ Boxes and the building meters.
+
+Validating THD and transient data, the other two basic power quality measures collected by OPQ Boxes, was more challenging.
+
+Figure \ref{fig:opqbox-thd-validation} shows the results of comparing values of THD collected by OPQ Boxes and building meters.
+
+\begin{figure}[ht]
+  \centering
+	\includegraphics[width=0.8\linewidth]{images/pilot/opqbox-thd-validation.png}
+	\caption{THD differences}
+	\label{fig:opqbox-thd-validation}
+\end{figure}
+
+Unlike frequency and voltage, where there was very close alignment between OPQ Boxes and building meters, THD data shows two peaks, one of which is indicates a kind of "offset" of approximately 0.13% by the OPQ Boxes from the building meters.
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
+Could you augment the above paragraph with a few sentences about why you think the THD validation looked like it did? I know you covered this in your defense but I forgot.
+\end{tcolorbox}
+
+Unfortunately, we unable to perform validation of transient data collection, because building meters did not provide us with access to that information.
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
+Could you add an additional sentence or two to the above paragraph about transient data validation? Maybe more about why we couldn't do it, or what the implications of non-validation might be?
+\end{tcolorbox}
+
+\subsection{OPQ Makai's triggering system provides advantages with respect to bandwidth and computation}
+
+The pilot study provided an opportunity to collect data on the resources required by an OPQ sensor network with respect to cloud-level network resources and server-level computation overhead.
+
+To assess network bandwidth utilization, we analyzed the data stream of frequency and voltage collected by OPQ Boxes, and the resulting network bandwidth utilized by the OPQ Makai triggering algorithm.
+Over the course of a typical day, OPQ Makai requested approximately 136 MB of data from the deployed OPQ Boxes.
+We then calculated how much data would be sent by a power quality meter using more conventional triggering approach in which exceeding a threshold would automatically result in sending waveform data, and found that, under the same conditions, approximately 1025 MB would be sent to the cloud, or eight times the network bandwidth.
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
+Please review and improve the above paragraph.
+\end{tcolorbox}
+
+To assess computational cost, we again analyzed the data stream for a typical day...
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
+Please finish the above paragraph by describing in words the histogram in slide 38 of your thesis defense presentation, where you show that Napali uses much less computational resources.
+\end{tcolorbox}
 

### 2020-06-09 literature.tex
Updates to pilot study section

@@
 \cite{zavoda_power_2018}
+
+OPQ Boxes do not collect PQ measures such as phase angle or reactive power.
+
+The main challenge is to go beyond manual methods based on the intuition and heuristics of human experts...it is crucial to develop the machine intelligence needed to automate and scale up the analytics on billions of µPMU measurements and terabytes of data on a daily basis and in real time.
+\cite{mohsenian-rad_distribution_2018}

### 2020-06-11 09-pilot.tex
working on pilot

@@
 
-To evaluate the capabilities of the OPQ Sensor Network, 15 OPQ Boxes were deployed at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019.
+To evaluate the capabilities of the OPQ Sensor Network, 16 OPQ Boxes were deployed at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019.
 The University of Hawaii campus is an isolated microgrid connected to the Oahu powergrid only via a single 46kV feeder.
-The UH Campus also has traditional, revenue-grade electrical meters deployed across various levels of the power delivery infrastructure.
+The UH Campus also has commercial electrical meters (a mixture of GE PQMII and GE EPM 7000) deployed across various levels of the power delivery infrastructure.
 While the primary purpose of these meters is to monitor power consumption, they do include power quality monitoring capabilities.
@@
 
-\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+The pilot study started on October 7, 2019 and ended on February 4, 2020.
+We deployed 16 OPQ Boxes across campus.
+As noted above, each OPQ Box collects 200 measurements per grid cycle, for a total of approximately 1B raw measurements per day per box.
+Values for the maximum and minimum voltage, frequency, and THD, along with the presence or absence of transients, is sent once a second to the cloud by each box.
+Thus, each box sends approximately 86,400 measures per day to the cloud at a minimum.
+Over the course of the pilot study, a total of approximately 116M aggregate measures were sent by all of the OPQ Boxes.
 
-Could you provide descriptive statistics for this section?
-This includes the start and end dates of the pilot study, the total number of data elements collected at each level of the information hierarchy, and for incidents and phenomena, the total number of each type of incident and event.
-Perhaps the size of the database?
-\end{tcolorbox}
+Figure \ref{fig:statistics} provides two additional sets of summary statistics regarding the pilot study.
+Figure (a) shows the number of anomalous events (i.e.
+where threshold values for frequency, voltage, THD, or transients were exceeded) along with the number of OPQ Boxes that experienced the power quality anomaly during that same time period.
+So, for example, 170,925 power quality anomalies were experienced by only one of the 16 OPQ boxes, and 463 anomalous power quality measurements were experienced by all 16 OPQ Boxes.
 
-\subsection{OPQ Boxes provide valid and reliable collection of power quality data}
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}{.5\textwidth}
+        \begin{tabularx}{\textwidth}{XXXX|}
+            \toprule
+            \textbf{Events} & \textbf{Boxes} & \textbf{Events} & \textbf{Boxes} \\
+            \midrule
+            170,925 & 1 & 203 & 9 \\
+            1,654 & 2 & 160 & 10 \\
+            1,109 & 3 & 162 & 11 \\
+            853 & 4 & 130 & 12 \\
+            593 & 5 & 169 & 13 \\
+            416 & 6 & 210 & 14 \\
+            354 & 7 & 477 & 15 \\
+            246 & 8 & 463 & 16 \\
+             &  &  &  \\
+            \bottomrule
+        \end{tabularx}
+	  \caption{Summary statistics: Events}
+	\end{subfigure}%
+	\begin{subfigure}{.5\textwidth}
+	 \begin{tabularx}{\textwidth}{lX}
+               \toprule
+               \textbf{Incident} & \textbf{Total}  \\
+               \midrule
+               Frequency Swell & 291,235 \\
+               Frequency Sag & 244,286  \\
+               Excessive THD & 21,395 \\
+               Voltage Sag & 620 \\
+               ITIC (No Damage) & 93 \\
+               SEMI F47 (Violation) & 24 \\
+               Voltage Interruption & 16 \\
+               Frequency Interruption & 14 \\
+               Voltage Swell & 8 \\
+               \bottomrule
+           \end{tabularx}
+   	  \caption{Summary statistics: Incidents}
+	\end{subfigure}
+	\caption{Summary statistics of (a) boxes involved in events and (b) incident type occurrences}
+	\label{fig:statistics}
+\end{figure}
+
+\subsection{OPQ provides valid and reliable collection of power quality data}
 
 The first goal of the pilot study was to assess how well an OPQ sensor network is able to collect basic power quality data.
-To do this, we compared data on voltage and frequency collected by OPQ Boxes with data on voltage and frequency collected by the revenue-grade building level power monitors.
+To do this, we compared data on voltage and frequency collected by OPQ Boxes with data on voltage and frequency collected by the existing UH power monitors.
 The results are shown in Figure \ref{fig:opqbox-f-v-validation}.
@@
 
-As the charts illustrate, there is very little difference between the frequency and voltage values collected by OPQ Boxes and the building meters.
+As the charts illustrate, there is very close correspondence between the building meters and the OPQ Boxes.
+For frequency differences, the value of $\sigma$ is 0.0079 Hz, and for voltage, $\sigma$ is 0.1703 V.
+(Note that typical thresholds for frequency and voltage PQ anomalies is 1 Hz and 6 V, so OPQ Box values appear accurate enough for their intended purpose.)
 
@@
 
-\subsection{OPQ Makai's triggering system provides advantages with respect to bandwidth and computation}
+\subsection{OPQs triggering system provides advantages with respect to bandwidth and computation}
 
@@
 
+\subsection{OPQ enables subthreshold event detection based on temporal locality}
+
+One interesting capability enabled by two way communication between OPQ Boxes and their cloud-based services is what we term {\em subthreshold event detection based on temporal locality}.
+In a nutshell, when one OPQ Box determines that a power quality measure has exceeded a threshold, one of the actions of the sensor network is to request high fidelity wave form data from neighboring boxes for the same temporal period, regardless of whether those boxes are reporting over-threshold data.
+In the event that any of those boxes actually do report over-threshold data, then the request ripples outward to the boxes neighboring that box, and so forth.
+
+During the pilot study, we discovered several situations in which the OPQ sensor network was able to characterize power quality anomalies in the UH microgrid in a manner that the building level meters could not.
+For example...
+
+\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
+Please finish the above paragraph by describing one situation of subthreshold event detection and why it was cool.
+\end{tcolorbox}
+
 \subsection{The OPQ Information Architecture provides a means to produce actionable insights}
+
+Figure \ref{fig:level-statistics} provides some descriptive statistics on the number of entities created at each level of the OPQ Information Architecture.
+
+\begin{figure}[ht]
+  \centering
+		 \begin{tabularx}{.4\textwidth}{lX}
+       \toprule
+       \textbf{Level} & \textbf{Total}  \\
+       \midrule
+       Phenomena & 10.4K \\
+       Incidents & 415K \\
+       Detections & 91.4K \\
+       AML (Trends) & 1.94M \\
+       AML (Measurements) & 116 M \\
+       IML  & 100B \\
+       \bottomrule
+     \end{tabularx}
+	\caption{Entities created at each level of the OPQ Information Architecture during the pilot study}
+	\label{fig:level-statistics}
+\end{figure}
 
@@
 
+Figure \ref{fig:data-management-graph} shows a graph of storage requirements both with and without TTL-based data removal.
+
+\begin{figure}[ht]
+  \centering
+	\includegraphics[width=0.4\linewidth]{images/pilot/data-management-graph.png}
+	\caption{Storage requirements with and without TTL}
+	\label{fig:data-management-graph}
+\end{figure}
+
 \subsection{OPQ Mauka provides useful adaptive optimization capabilities}

### 2020-06-11 literature.tex
working on pilot

--- 2020-05-22 literature.tex
@@
 \cite{zavoda_power_2018}
+
+OPQ Boxes do not collect PQ measures such as phase angle or reactive power.
+
+The main challenge is to go beyond manual methods based on the intuition and heuristics of human experts...it is crucial to develop the machine intelligence needed to automate and scale up the analytics on billions of PMU measurements and terabytes of data on a daily basis and in real time.
+\cite{mohsenian-rad_distribution_2018}

### 2020-06-12 07-opq-mauka.tex
updates
Finished pilot study, conclusions drafts.

--- 2020-06-01 07-opq-mauka.tex
@@
 OPQ Mauka addresses both of these issues.
-First, OPQ Mauka provides a multi-leveled representation for structuring and processing DSN data.
-The structure and processing at each level is designed with the explicit goal of turning low-level data into actionable insights.
-Second, each level in the framework implements a "time-to-live" (TTL) strategy for data within the level.
-This strategy states that data must either progress upwards through the levels towards more abstract, useful representations within a fixed time window, or be discarded and lost forever.
-The TTL strategy is useful because when implemented, it allows DSN designers to calculate upper bounds on data storage at each level of the framework and supports graceful degradation of DSN performance.
+First, OPQ Mauka provides a multi-layered representation for structuring and processing DSN data.
+The structure and processing at each layer is designed with the explicit goal of turning low-level data into actionable insights.
+Second, each layer in the framework implements a "time-to-live" (TTL) strategy for data within the level.
+This strategy states that data must either progress upwards through the layers towards more abstract, useful representations within a fixed time window, or else it can be discarded.
+The TTL strategy is useful because when implemented, it allows DSN designers to make reasonable predictions of the upper bounds on data storage at each level of the framework.
+
+ TTL also makes possible a "graceful degradation" of system performance if those bounds turn out to be exceeded.
+For example, consider a situation in which a power network enters a prolonged period of widespread power quality instability, where every OPQ Box is reporting continuous anomalous conditions with respect to voltage, frequency, THD, and transients.
+This "worst case scenario" would lead to the potential situation of every OPQ Box trying to upload raw waveform data all the time.
+The TTL system provides safeguards, in that whatever low-level data has not been processed relatively quickly can be discarded.
+Thus, instead of the system potentially going down entirely, it could instead continue to operate at a reduced capacity.
 
@@
 
-OPQ Mauka sits between OPQ Makai and OPQ’s database.
-It receives messages from OPQ Makai when it has triggered OPQ Boxes due to something interesting in the low-fidelity data stream.
-OPQ Mauka then retrieves the raw waveforms stored in the database by OPQ Makai, performs feature extraction, and then forwards the features and/or raw waveform to OPQ Mauka Actors to perform classification and other high level analytics.
-These results are then stored in the database.
-Finally, TTL Actors run at predetermined intervals to remove data whose time-to-live has expired.
+\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+Please review the following paragraphs for accuracy and quality.
+\end{tcolorbox}
+
+The lowest layer of the hierarchy is the Instantaneous Measurements Layer (IML).
+The IML contains "raw" data, in other words, the digitize waveform.
+IML data exists both on each OPQ Box (where it is available for up to the previous 60 minutes).
+It also exists in the cloud, in the event that OPQ's triggering mechanism has caused a temporal interval of waveform data to be uploaded.
+IML data in the cloud has a TTL of 15 minutes: unless the wave form data is found to be useful by a cloud service within 15 minutes, it can be discarded.
+
+The second layer is the Aggegate Measurements Level (AML).
+The AML stores power quality summary statistics sent once per second by each OPQ Box.
+These summary statistics include the maximum, minimum, and average values of voltage, frequency, and THD over the past minute, as well as the presence or absence of transients.
+It is AML data that is used to initiate the triggering process of uploading IML data from the cloud.
+
+The third layer is the Detections Layer (DL).
+This layer is responsible for processing the IML and AML data to produce a representation of an "event" with a precisely defined start and end time based upon examination of the waveform.
+As will be discussed in Section \ref{sec:subthreshold-events}, knowledge of the start and end time of a power quality anomaly allows investigation of how that anomaly might manifest itself elsewhere in the grid, even if this manifestation is not severe enough to produce over-threshold data values.
+
+The fourth layer is the Incident Layer (IL).
+This layer starts to answer the question of whether or not the data is "actionable" by classifying the detected event according to various industry standards for power quality anomalies: IEEE 1159, ITIC, SEMI F47, and so forth.
+For example, an event that falls into the ITIC "prohibited" region clearly indicates a power quality anomaly that requires further study and intervention.
+
+The fifth and final layer is the Phenomena Layer (PL).
+This layer contains the results of analyses that attempt to identify cyclic, and thus predictable, power quality disturbances.
+It also contains analysis results regarding the similarity of various incidents, which can help uncover causal factors.
+Finally, it provides analyses for adaptive optimization of the OPQ Sensor Network.
+These optimizations can change the thresholds for individual boxes to either increase their sensitivity or decrease their sensitivity over specific intervals of time.
+The ultimate goal of adaptive optimization is to help the network learn to acquire all of the data useful for analyses, and only the data useful for analyses.
+We are still in the early stages of exploring the potential of adaptive optimization in OPQ networks.
 
@@
 
-The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins.
+The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins. Nine of the most important Actors are described below.
 
-{\em BasePlugin.} All Actor plugins derive from the BasePlugin, which provides primitives for running plugins as separate processes, serialization and deserialization of type safe messages, communication over ZeroMQ, metrics collection, MongoDB communication, debugging, and plugin health.
+{\em Makai Event Actor.} The Makai Event Actor is responsible for reading data newly created by OPQ Makai into OPQ Mauka.
+It performs feature extraction on the raw data stream and forwards those features (or the raw data) to subscribing Actor plugins.
+This allows OPQ Mauka to perform feature extraction once, and allow use of those features by multiple Actors.
 
-{\em MakaiEventPlugin.} The MakaiEventPlugin Actor is responsible for reading Events newly created by OPQ Makai into OPQ Mauka.
-It performs feature extraction on the raw Event data streams and forwards those features (or the raw data) to subscribing plugins.
-This allows OPQ Mauka to perform feature extraction once, and reuse those features in multiple plugins.
+{\em Frequency Variation Actor.} The Frequency Variation Actor classifies generic frequency sags, swells, and interruptions as defined by the IEEE 1159 standard.
+Both duration and deviation from nominal are used to perform these classifications.
+Duration classifications include frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
+Classifications for deviations from nominal are performed for values that are up to 40\% deviation from nominal.
+This Actor is able to classify frequency swells, frequency interruptions, and frequency sags, leading to the creation of data at the Incident Layer.
 
-{\em FrequencyVariationPlugin.} The FrequencyVariationPlugin is used to classify generic fre- quency sags, swells, and interruptions as defined by the IEEE1159 standard[29].
-Both duration and deviation from nominal are used to perform these classifications.
-Duration classifications in- clude frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
-Classifications for deviations from nominal are performed for values that are up to 40% deviation
-     49
+{\em IEEE 1159 Voltage Actor.} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard[29].
+In general, this standard classifies voltage disturbances by duration and by magnitude.
+Voltage durations are classified from 0.5 to 30 cycles, 30 cycles to 3 seconds, 3 seconds to a minute, and greater than 1 minute.
+Voltage deviations are classified in both the sag and swell directions as a percentage from nominal.
+Sags are generally classified between 10\% and 90\% of nominal while swells are generally classified from 110\% to 180\% of nominal.
+This Actor is capable of classifying voltage sags, swells, and interruptions as defined by the standard, and creating data at the Incident Layer if appropriate.
 
-     nominal. The plugin subscribes to messages from the “WindowedFrequency” topic which includes a payload of frequency features over one electrical cycle.
-     When a message is received, the plugin uses configurable thresholds to find frequency fluctu- ations within the windowed frequency streams.
-The plugin is able to classify frequency swells, frequency interruptions, and frequency sags.
-When a new frequency variation is detected, a fre- quency Incident is created
+{\em Box Optimization Actor.} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
+This Actor is capable of requesting the state of each OPQ Box (e.g.
+uptime, Measurement rate, security keys, etc).
+It is also capable of adjusting the state of individual OPQ Boxes by changing things such as the Measurement and Trend rate or the sampling rate used by the Box.
+
+{\em Future Phenomena Actor.} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena.
+These Phenomena are used to predict Events and Incidents that may occur in the future.
+This plugin does not subscribe to any messages, but instead utilizes timers to perform its work.
+By default, this plugin runs every 10 minutes.
+
+When a Future Phenomena Actor runs, it loads any active Periodic Phenomena found in the database.
+If Periodic Phenomena are found, this Actor extrapolates possible Detection and Incident Layer data by first examining their timestamps and then extrapolating into the future using the mean period and the standard deviation.
+For each timestamp in a Periodic Phenomena, the mean period is added.
+If the resulting timestamp is in the future, a Future Phenomena is created using the time range of the future timestamp plus or minus the standard deviation of the Periodic Phenomena.
+
+When a Future Phenomena is created, timers are started in a separate thread signifying the start and end timestamps of the Future Phenomena.
+When the first timer runs, messages are sent to the Box Optimization Actor and the Threshold Optimization Actor instructing OPQ Box thresholds to be set lower and measurement rates to be set higher.
+This increases the chance of seeing an anomaly over the predicted time window.
+When the second timer runs, these values are reset to their default values.
+Thus, the plugin increases fidelity and decreases thresholds over the period of a Future Phenomena.
+
+{\em ITIC Actor.} The ITIC Actor analyzes voltage to determine where it falls within the ITIC curve \cite{thallam_power_2000}.
+The ITIC curve is a power acceptability curve that plots time on the x-axis and voltage on the y-axis.
+The purpose of the curve is to provide a tolerance envelope for single-phase 120V equipment.
+The curve defines three regions.
+The first region is "No Interruption" and generally includes all voltages with very short sustained durations.
+All events within this region have no noticeable effect on power equipment.
+The second region, the "No Damage Region", occurs during voltage sags for extended periods of time.
+Power Events in this region may cause equipment interruptions, but it will not damage the equipment.
+The final region, the "Prohibited" region, is caused by sustained voltage swells and may cause damage to power equipment.
+This Actor determines if an event falls within the "No Damage" or "Prohibited" regions and if so, creates an Incident to record this.
+
+{\em SEMI F47 Actor.} The SEMI F47 Actor is similar to the ITIC Actor in that it plots voltage and duration against a power acceptability curve.
+In this case, the standard used is the SEMI F47 standard \cite{djokic_sensitivity_2005}.
+Rather than using a point-in-polygon approach, this plugin reads the voltage features sequentially and uses a state machine to keep track of the current classification.
+This plugin only classifies values as a "violation" or as "nominal".
+
+{\em Transient Actor.} The Transient Actor is responsible for classifying frequency transients in power waveforms.
+The plugin subscribes to messages from a topic which contains a calibrated power waveform payload.
+The Transient Actor is capable of classifying impulsive, arcing, oscillatory, and periodic notching transients.
+A decision tree is utilized to select the most likely transient type and then further analysis is used to perform the actual classification of transients.
+Dickens et al \cite{dickens_transient_2019} provides more details on the transient classification system used by this Actor.
+
+{\em Periodicity Actor.} The Periodicity Actor is responsible for detecting periodic signals in power data.
+This Actor does not subscribe to any messages, but instead runs off of a configurable timer.
+The Actor is set to run by default once an hour and every hour it scrapes the last 24 hours worth of data and attempts to find periods in the Measurements over that duration.
+
+For each feature in the Measurement and Trend data (e.g.
+frequency, voltage, and THD), the Periodicity Actor first removes the DC offset from the data by subtracting the mean.
+Next, the Actor filters the signal using a 4th order high-pass filter to filter out noise.
+The Actor then performs autocorrelation on the signal followed by finding the peaks of the autocorrelation.
+The mean distance between the peaks of the autocorrelation provides the period of the signal.
+
+The Periodicity Actor only classifies data as periodic if at least 3 peaks were found and the standard deviation of the period is less than 600 seconds (10 minutes).
+Once a positive identification has been made, peak detection is performed on the original signal.
+Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the plugin can group Measurements, Trends, Detection Layer Events, and Incidents that were created during the periodic signals together as part of the Periodic Phenomena.

### 2020-06-12 09-pilot.tex
Finished pilot study, conclusions drafts.

@@
 \subsection{OPQ enables subthreshold event detection based on temporal locality}
+\label{sec:subthreshold-events}
 
@@
 
-Figure \ref{fig:level-statistics} provides some descriptive statistics on the number of entities created at each level of the OPQ Information Architecture.
+Figure \ref{fig:level-statistics} provides some descriptive statistics on the total number of entities created at each layer of the OPQ Information Architecture over the course of the pilot study.
+Note that the number of entities at a given layer at a given point in time is typically far less, due to the use of TTL to discard entities not promoted to higher layers.
 
@@
 
+Figure \ref{fig:level-statistics} shows that all layers of the hierarchy were actively used.
+In addition, the system was able to produce actionable insights by automatically detecting periodic phenomena for two boxes, and then predicting future occurrences of power quality anomalies with over a 50\% success rate (in other words, the current implementation of predictive phenomena leads to a significant false positive rate, predicting twice as many future power quality anomalies as actually occurred).
+Improving the predictive capabilities is a topic of future research.
+
 \subsection{The OPQ Information Architecture provides predictable upper bounds on storage resources}
 
-Figure \ref{fig:data-management-graph} shows a graph of storage requirements both with and without TTL-based data removal.
+Figure \ref{fig:data-management-graph} shows a graph that illustrates storage requirements both with and without TTL-based data removal.
+As the graph shows, over the course of the three month period, we estimate that cloud-level storage requirements would have reached approximately 2.5TB if all data was kept.
+However, due to TTL, the amount of data held in the cloud at the conclusion of the case study was only 100 GB.
 
@@
 
-\subsection{OPQ Mauka provides useful adaptive optimization capabilities}
+It is difficult to compute a strict upper bound on storage requirements for an OPQ network, because the amount of storage does depend on the number of power quality anomalies experienced by the grid.
+That said, we did notice that storage requirements were rising much more slowly at the end of the pilot.
+
+\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+Could you add a sentence or two to explain the assumptions under which the 2.5 TB storage requirement was calculated?
+
+Could you provide a paragraph to add more concrete details to this section? For example, how much additional storage was required, say, each week, by the end of the study period?
+
+\end{tcolorbox}
+
+\subsection{OPQ provides useful adaptive optimization capabilities}
+
+The top layer of OPQ's information architecture supports adaptive optimization: the ability to analyze previously collected data and use it to change the settings associated with data collection from OPQ Boxes.
+One way we used this capability was to identify periodic phenomena, then predict when it was going to occur in future, and then change the settings associated with Boxes to support better data collection during periods in which a power quality anomaly was predicted.
+
+\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+Can you provide a "for example" of this with more concrete details, hopefully indicating how this adaptive optimization actually yielded better results?
+\end{tcolorbox}
+
+\begin{tcolorbox}[title=ANTHONY AND SERGE]
+Any other pilot study results that I've forgotten to include?
+\end{tcolorbox}

### 2020-06-12 10-conclusions.tex
Finished pilot study, conclusions drafts.

\section{Conclusions and Future Directions}
\label{sec:conclusions}

This research project has produced both hardware and software for power quality monitoring with a variety of innovations.  Our OPQ Boxes collect frequency, voltage, THD, and transients with high fidelity, and at a cost that is generally 10x to 100x cheaper than current commercial offerings (though these commercial offerings offer a variety of features not available from OPQ Boxes, so the appropriate choice depends upon the needs of the user).

Our sensor network provides real-time, two way communication between the sensor nodes (OPQ Boxes) and their cloud services, providing several innovative features for power quality monitoring. The OPQ triggering system exploits temporal locality to request high fidelity data from neighboring boxes when one exceeds a threshold, enabling our network to detect and analyze data that would have been unreported by a naive, threshold-based approach. The OPQ Information Architecture enables Actors at the Phenomena Layer to control sensitivity settings of individual OPQ Boxes in order to improve data collection when power quality anomalies have been predicted.

Our sensor network is designed to efficiently use network and storage resources. Two way communication enables OPQ Boxes to send low fidelity summary statistics on power quality at one second intervals, which can be used by cloud-based services to decide whether or not to request high fidelity data.  Our TTL mechanism implements a "use it or lose it" approach to cloud-based data, which in our pilot study reduced cloud storage requirements by over ten-fold.

Our pilot study has provided evidence that OPQ sensor networks can provide useful new support for understanding grid stability, particularly in grids with distributed, intermittent renewables. This in part because OPQ Boxes are suitable for installation at the residential level: installation requires only an available wall socket and WiFi connection, no electrician or GPS line-of-sight required. Finally, the low cost of OPQ Boxes means mass deployment across a neighborhood is not financially infeasible.

A primary future goal is to partner with an organization that can enable OPQ Boxes to be produced at volume. We made the OPQ Boxes by hand for the pilot study, and although we have had numerous requests for OPQ Boxes, we do not possess manufacturing capacity.

With manufacturing capacity will come the ability to deploy OPQ sensor networks at higher scale. We believe that a single sensor network can easily scale to hundreds of boxes before bandwidth and processing constraints become an issue. To scale to many thousands of boxes, we believe a federated approach would work in which individual OPQ sensor networks communicate their findings to each other. Those networks could be designed around natural grid boundaries such as substations.

Finally, we believe an important direction for future research is to build Actors that operate not only on power quality data but also environmental data such as wind, temperature, humidity, and insolation.  These additional data streams could yield valuable mechanisms for creating Predictive Phenomena for power quality anomalies associated with renewable energy sources. Ultimately, such understanding could lead to ways to significantly increase the amount of distributed renewable energy that can be incorporated into electrical grids.

### 2020-06-13 01-abstract.tex
Updates to merge literature.tex

--- 2020-05-27 01-abstract.tex
@@
 This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc) are blurring: end-users both produce and consume energy, making energy management and public policy more complex.
-The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network in order to make useful information about electrical grids (and their stability) available to producers, consumers, researchers, and policy makers.
-In 2019, we performed a pilot study where 15 OPQ hardware devices were deployed across the University of Hawaii microgrid for three months.
-Results of the pilot study provide evidence that OPQ provides a variety of useful monitoring services and that the system could be scaled to service larger geographic regions.
-We conclude that OPQ provides a new and useful approach to power quality monitoring that can be used in conjunction with other technologies .}
+The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network that provides useful new forms of information about modern electrical grids to producers, consumers, researchers, and policy makers.
+In 2019, we performed a pilot study involving the deployment of an OPQ sensor network at the University of Hawaii microgrid for three months.
+Results of the pilot study validate the ability of OPQ to collect accurate power quality data in a way that provides useful new insights into electrical grids.}
 

### 2020-06-13 02-intro.tex
Updates to merge literature.tex

@@
 Third, modern consumer electronics place more stringent demands on power quality.
-Finally, effective public policy making in this modern context can be aided by increased public access to power quality data.
+Finally, effective public policy making in this modern context can be aided by increased public access to power quality data reflecting the state of the grid as it is experienced by end-users.
 Let's examine each of these assertions in more detail.
@@
 In the case of Hawaii, consumer demand for grid-tied rooftop solar exceeded the ability of Hawaiian Electric to manage, resulting first in complex and expensive "interconnection studies" \cite{trabish_solar_2014,anastasi_energy_2009}, and later in the requirement for new installations to include batteries and grid-disconnected operation.
+Photovoltaics affect power quality by introducing harmonics \cite{anurangi_effects_2017}.
+The power feed-in of PV generation in rural low-voltage grids can influence power quality (PQ) as well as facility operation and reliability \cite{rita_pinto_impact_2016}.
+In fact, a study of the island of Porto Santo in Portugal found that intermittent nature of installed photovoltaic and wind energy could result in a potential drop in frequency of 12 Hz, lasting as long as 7 seconds \cite{delgado_solutions_2011}.
 
@@
 In these cases, the unpredictable nature of generation can lead to local power quality problems that are not experienced by the substation, and that are not due to any single individual, but rather the bottom-up power generation architecture of the grid.
+Solar panels connected to low voltage networks will result in over-voltages, the switching frequency of the converters in wind turbines causes high-frequency signals flowing into the grid, and harmonics are generated by EV chargers.
+\cite{zavoda_power_2018}
  Nakafuji notes that Hawaiian Electric is contending with PV penetrations in excess of 60\% on certain distribution circuits, for which traditional rules of thumb for design of protection and distribution systems might not hold \cite{nakafuji_back--basics_2011}.
@@
 {\em 3.
-Consumer electronics require higher quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality.
+Consumer electronics require high quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate" power quality.
 Only a few decades ago, computers were a rare presence outside of labs and large institutions.
@@
 Poor power quality can result in electronic devices failing unpredictably, and/or decreasing their lifespan.
+Reduced input voltage can cause excessive power supply heat dissipation, resulting in reduced mean time between failures (MTBF).
+In addition, rectifiers and DC-to-DC converter switching transistors draw high-peak currents, which raise their junction temperatures.
+These temperature excursions take a toll on semiconductor longevity.
+High input voltage can also puncture a power supply's rectifier and switching transistor junctions, causing MTBF reduction and eventual breakdown.
+High-voltage transients lasting microseconds can permanently wreck the power supply and its electronic equipment load.
+\cite{dedad_when_2008}.
 
@@
 Providing a high quality, third party, publicly available source of power quality data not only benefits the public, it makes it much easier to perform research on ways to improve grid stability in the presence of renewable energy generation.
+
+Better public understanding of power quality is important because it is such a significant economic problem.
+Across all business sectors, the U.S.
+economy is estimated to be losing between \$104 billion and \$164 billion a year to outages and another \$15 billion to \$24 billion to power quality phenomena \cite{elphick_summary_2015}.
+India lost more than \$9.6 billion in 2008 due to power quality problems, and Europe is estimated to be losing \$150 billion per year \cite{laskar_power_2012}.
 

### 2020-06-13 04-related.tex
Updates to merge literature.tex

@@
 
-Section \ref{sec:commercial-pq-devices} discusses commercial power quality monitoring devices and how they compare to OPQ Box.
-Section \ref{sec:commercial-pq-software} discusses commercial power quality database software and how they compare to OPQ Cloud.
-The final subsection discusses recent research systems and how they compare to OPQ as a whole.
+Section \ref{sec:commercial-pq-devices} discusses power quality monitoring devices and how they compare to OPQ Box.
+Section \ref{sec:commercial-pq-software} discusses power quality database software and how it compares to OPQ Cloud.
+The final subsection discusses recent research systems comprising both hardware and software and how they compare to OPQ.
 
-\subsection{Commercial PQ devices}
+\subsection{Power quality hardware}
 \label{sec:commercial-pq-devices}
@@
 
+OpenZMeter \cite{viciana_openzmeter_2018} is a low-cost, open source device that connects to household mains and collects both power quality and energy data.
+
 \subsection*{Comparison with OPQ Box}
 
-All of these commercially available devices contrast with the OPQ Box in similar ways.
+All of the above devices contrast with the OPQ Box in similar ways.
 First, all of them collect a wider variety of power quality measures than OPQ Box, and most have been certified according to one or more industry standards.
-They are generally designed to support industrial applications, where the goal is to ensure that the power being supplied to a building or plant is of adequate quality, and/or that the machinery in the plant is not degrading power quality.
+Except for OpenZMeter, they are generally designed to support industrial applications, where the goal is to ensure that the power being supplied to a building or plant is of adequate quality, and/or that the machinery in the plant is not degrading power quality.
 Apart from the PowerWatch monitor, all of them are attached to electrical mains using current transformers.
@@
 
-While the OPQ Box has much more limited functionality, it is designed to be manufactured for approximately \$75, which is 10 to 100 times less expensive than these devices.
-OPQ Boxes are "cloud native", which means that they have no "stand alone" capabilities.
-Instead, they must be connected to an OPQ Cloud instance in order for power quality data to be collected and analyzed.
+While the OPQ Box has much more limited functionality, it is designed to be manufactured for approximately \$75, which is 10 to 100 times less expensive than most commercial devices, and similar is cost to OpenZMeter.
+The most important distinguishing feature of OPQ Box is that it is designed to grid-level, not single point monitoring, and thus incorporates features (such as two way communication with the cloud, and subthreshold triggering) that are not present in devices intended for "stand alone" capabilities.
 
-\subsection{Commercial PQ software}
+\subsection{Power quality software}
 \label{sec:commercial-pq-software}
@@
 
-PQDif \cite{sabin_ieee_2020}, the "Power Quality Data Interchange Format", is a binary file format that specifies voltage, current, power, and energy measurements in standard fashion that allows this data to produced and consumed by devices in a vendor-independent fashion.
+PQDIF \cite{sabin_ieee_2020}, the "Power Quality Data Interchange Format", is a binary file format that specifies voltage, current, power, and energy measurements in standard fashion that allows this data to produced and consumed by devices in a vendor-independent fashion.
 
@@
 
-The research system most similar to OPQ is FNET, which is described in a recent publication by Liu et al \cite{liu_distribution_2017}.
+The research system most similar to OPQ is FNET \cite{liu_distribution_2017}.
 Like OPQ, the FNET system consists of custom hardware that monitors the electrical signal from a wall outlet, and uploads data to the cloud for further processing.

### 2020-06-13 09-pilot.tex
Updates to merge literature.tex

@@
 \end{enumerate}
+
+Our pilot study is significant because much of the literature on power quality assessment relies on models, not actual installations \cite{anurangi_effects_2017,bayindir_effects_2016,farhoodnea_power_2012,shafiullah_experimental_2014}.
+In other cases, data was collected from only one location or for a very short time span \cite{kucuk_assessment_2013,viciana_openzmeter_2018}.
 

### 2020-06-13 10-conclusions.tex
Updates to merge literature.tex

@@
 
-A primary future goal is to partner with an organization that can enable OPQ Boxes to be produced at volume.
+One future goal is to partner with an organization that can enable OPQ Boxes to be produced at volume.
 We made the OPQ Boxes by hand for the pilot study, and although we have had numerous requests for OPQ Boxes, we do not possess manufacturing capacity.
@@
 
-Finally, we believe an important direction for future research is to build Actors that operate not only on power quality data but also environmental data such as wind, temperature, humidity, and insolation.
+A second goal is to explore ways to combine OPQ sensor networks with more traditional power quality monitoring tools and standards, such as PQDIF.
+For example, it is possible that OPQ analyses could be improved with access to additional power data such as current, phase angle, and so forth.
+
+A third goal is to build Actors that operate not only on power quality data but also environmental data such as wind, temperature, humidity, and insolation.
 These additional data streams could yield valuable mechanisms for creating Predictive Phenomena for power quality anomalies associated with renewable energy sources.
 Ultimately, such understanding could lead to ways to significantly increase the amount of distributed renewable energy that can be incorporated into electrical grids.
+
+To conclude, a recent paper by Mohsenian et al \cite{mohsenian-rad_distribution_2018} states, "the main challenge is to go beyond manual methods based on the intuition and heuristics of human experts...it is crucial to develop the machine intelligence needed to automate and scale up the analytics on billions of PMU measurements and terabytes of data on a daily basis and in real time." We believe that the OPQ sensor network represents a small step along the path toward that future.

### 2020-06-13 04-related.tex
draft mostly complete

@@
 
-The research system most similar to OPQ is FNET \cite{liu_distribution_2017}.
+Mohsenian-Rad et al.
+\cite{mohsenian-rad_distribution_2018} designed the $\mu$PMU (phase measurement unit) system which provides distributed power quality measurements over power grid distribution systems.
+The $\mu$PMUs in conjunction with their backend software provide two types of analytics.
+Descriptive analytics provide information about the types and classifications of power quality issues that are observed within the power distribution grid.
+Predictive analytics are used to predict future power quality issues.
+The authors describe their system as providing the ground for for enabling future prescriptive analytics, which is the idea of self-tuning the DSN to prepare for future power quality problems by using a combination of descriptive and predictive analytics.
+In contrast, as we will discuss in Section \ref{sec:adaptive-optimization}, the OPQ sensor network already supports self-tuning through adaptive optimization.
+
+One research system very similar in spirit to OPQ is FNET \cite{liu_distribution_2017}.
 Like OPQ, the FNET system consists of custom hardware that monitors the electrical signal from a wall outlet, and uploads data to the cloud for further processing.

### 2020-06-21 09-pilot.tex
draft mostly complete
Discuss assumption under which 2.5TB data was calculated.

@@
 It is difficult to compute a strict upper bound on storage requirements for an OPQ network, because the amount of storage does depend on the number of power quality anomalies experienced by the grid.
-That said, we did notice that storage requirements were rising much more slowly at the end of the pilot.
+That said, we did notice that storage requirements were rising much more slowly at the end of the pilot when utilizing TTL as compared to not utilizing TTL.
 
-\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
-Could you add a sentence or two to explain the assumptions under which the 2.5 TB storage requirement was calculated?
+The largest contribution to data storage is the raw power data.
+In a system without TTL, raw data storage rapidly grows.
+The bounds on raw storage can be calculated by taking into account the sampling rate and sample size for active OPQ Boxes as described in Equation ~\ref{eq:RawDataBytes}.
 
-Could you provide a paragraph to add more concrete details to this section? For example, how much additional storage was required, say, each week, by the end of the study period?
+\begin{equation}
+    RawDataBytes = NumBoxes * TimeSeconds * SampleRate * SampleSize
+    \label{eq:RawDataBytes}
+\end{equation}
 
-\end{tcolorbox}
+Given that OPQ Boxes were configured to sample at 12 kHz at 2 bytes per sample during the deployment, we can calculate the bounds on raw storage for 16 OPQ Boxes over 3 months without TTL as 3.1 TiB.
+Figure~\ref{fig:data-management-graph} shows slightly less data as OPQ Boxes were added to the deployment over the course of the month and we did not have all 16 OPQ Boxes recording continuously since the begining of the pilot study.
+
+%\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+%Could you add a sentence or two to explain the assumptions under which the 2.5 TB storage requirement was calculated?
+%
+%Could you provide a paragraph to add more concrete details to this section? For example, how much additional storage was required, say, each week, by the end of the study period?
+%
+%\end{tcolorbox}
 
 \subsection{OPQ provides useful adaptive optimization capabilities}
+\label{sec:adaptive-optimization}
 

### 2020-06-22 09-pilot.tex
Expand on optimizations.

@@
 Given that OPQ Boxes were configured to sample at 12 kHz at 2 bytes per sample during the deployment, we can calculate the bounds on raw storage for 16 OPQ Boxes over 3 months without TTL as 3.1 TiB.
-Figure~\ref{fig:data-management-graph} shows slightly less data as OPQ Boxes were added to the deployment over the course of the month and we did not have all 16 OPQ Boxes recording continuously since the begining of the pilot study.
+Figure~\ref{fig:data-management-graph} shows slightly less data as the number of OPQ Boxes was not constant during the pilot study (Figure~\ref{fig:active-opq-boxes})
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.75\linewidth]{images/pilot/active-opq-boxes.png}
+    \caption{Active OPQ Boxes vs. Time}
+    \label{fig:active-opq-boxes}
+\end{figure}
 
@@
 The top layer of OPQ's information architecture supports adaptive optimization: the ability to analyze previously collected data and use it to change the settings associated with data collection from OPQ Boxes.
-One way we used this capability was to identify periodic phenomena, then predict when it was going to occur in future, and then change the settings associated with Boxes to support better data collection during periods in which a power quality anomaly was predicted.
 
-\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
-Can you provide a "for example" of this with more concrete details, hopefully indicating how this adaptive optimization actually yielded better results?
-\end{tcolorbox}
+%One way we used this capability was to identify periodic phenomena, then predict when it was going to occur in future, and then change the settings associated with Boxes to support better data collection during periods in which a power quality anomaly was predicted.
+
+Two areas where adaptive optimizations were utilized are identifying periodic phenomena and predictions of future phenomena.
+Periodic phenomena are PQ related signals that occur at regular intervals.
+One example of periodic phenomena is a cyclic volage sag that was observed at one of our sensors with a period of about 34 minutes during our pilot study (Figure~\ref{fig:periodic-voltage-sags}).
+This periodic phenomena not only provided a classification of an interesting pattern, but also helped optimize the OPQ system for capture of future phenomena.
+Several system wide optimizations are utilized.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.75\linewidth]{images/pilot/periodic-voltage-sags.png}
+    \caption{Detection of Periodic Voltage Sags}
+    \label{fig:periodic-voltage-sags}
+\end{figure}
+
+First, classified periodic phenomena are used to generate future phenomena.
+Future phenomena are a prediction of future signals of interest.
+When a prediction is made, the system will automatically increase sampling fidelity for the predicted sensor during the time window of the prediction.
+In particular, the measurement rate is increased from 1 measurement per second to 6 measurements per second providing higher fidelity data.
+Detection thresholds are also decreased, making it more likely that the predicted signal will be observed even if it is of low magnitude.
+
+As an example, Figure~\ref{fig:future-phenomena} shows the percentage of correctly predicted future phenomena as a function of time.
+The self-optimizing nature of the system allows the system to make more accurate predictions due to acting on more accurate periodic phenomena.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.60\linewidth]{images/pilot/realized-future-phenomena.png}
+    \caption{Future Phenomena Prediction Accuracy Over Time}
+    \label{fig:future-phenomena}
+\end{figure}
+
+Second, the increase in detection sensitivity makes it more likely that sub-threshold signals of interest will be classified.
+As example, the increased detection sensitivity allowed us to observe and classify a periodic voltage sag that would not have have been detected by our normal classifier due to the small magnitude of the sag.
+The detection is provided in Figure~\ref{fig:sub-threshold-detection}.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.85\linewidth]{images/pilot/subthreshold-signal.png}
+    \caption{Sub-threshold Detection}
+    \label{fig:sub-threshold-detection}
+\end{figure}
+
+This creates a positive feedback loop where as periodic phenomena become more accurate, future phenomena become more accurate.
+As future phenomena become more accurate, periodic phenomena become more accurate.
+
+%\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+%Can you provide a "for example" of this with more concrete details, hopefully indicating how this adaptive optimization actually yielded better results?
+%\end{tcolorbox}
 

### 2020-06-22 03-architecture.tex
Some updates to table 1.

@@
 Similarly, data points in the Measurement layer will be deleted after one day if they are not associated with an Event.
+Details of this hierarchy are further explained in Section~\ref{sec:opq-mauka}.
 
@@
 \midrule
-Box		& 1 hour			& Collects and holds a rolling one hour window of "high fidelity" wave form data for a single location.\\
-Measurement		& 1 day			& Measurements provide "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients). \\
-Event & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': wave form data from one or more boxes for a specific time interval (typically just a second or so).
+Box		& 1 hour			& Collects and holds a rolling one hour window of "high fidelity" instantaneous wave form data measurements for a single location.\\
+Measurement		& 1 day/2 weeks		& Aggregate measurements provide "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients) at a rate of either 1/second (Measurement) or 1/day (Trend).
+\\
+Event/Detection & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': wave form data from one or more boxes for a specific time interval (typically just a second or so).
 \\

### 2020-06-22 07-opq-mauka.tex
Updates to OPQ Mauka.

@@
 
-\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
-Please review the following paragraphs for accuracy and quality.
-\end{tcolorbox}
+%\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
+%Please review the following paragraphs for accuracy and quality.
+%\end{tcolorbox}
 
@@
 
-The second layer is the Aggegate Measurements Level (AML).
-The AML stores power quality summary statistics sent once per second by each OPQ Box.
-These summary statistics include the maximum, minimum, and average values of voltage, frequency, and THD over the past minute, as well as the presence or absence of transients.
+The second layer is the Aggregate Measurements Level (AML).
+The AML stores power quality summary statistics sent either once per second or once per minute by each OPQ Box.
+These summary statistics include the maximum, minimum, and average values of voltage, frequency, and THD over the time window, as well as the presence or absence of transients.
 It is AML data that is used to initiate the triggering process of uploading IML data from the cloud.
 
-The third layer is the Detections Layer (DL).
+The third layer is the Detections Level (DL).
 This layer is responsible for processing the IML and AML data to produce a representation of an "event" with a precisely defined start and end time based upon examination of the waveform.
@@
 
-The fourth layer is the Incident Layer (IL).
+The fourth layer is the Incident Level (IL).
 This layer starts to answer the question of whether or not the data is "actionable" by classifying the detected event according to various industry standards for power quality anomalies: IEEE 1159, ITIC, SEMI F47, and so forth.
@@
 
-The fifth and final layer is the Phenomena Layer (PL).
+The fifth and final level is the Phenomena Layer (PL).
 This layer contains the results of analyses that attempt to identify cyclic, and thus predictable, power quality disturbances.

### 2020-06-22 08-opq-view.tex
initial start on pilot
updates
More updates to the pilot and OPQ view clarifications.

@@
 
-{\em Anthony: a couple of questions about this home page image.
-First, no phenomena?
-What's up with that?
-Second, although the TTL for Measurements is set to 1 day, the System Status window says there are 16M measurements.
-It seems to me that with 1 measurement per second per box, we should only have around 1.2M per day, so at most double that given the TTL.
-Can you explain to me what's happening?
-}
+The home page ``System Stats'' widget shows the number of items stored at each OPQ storage level.
+The right hand column shows items that have been stored all time (not including items that have been discarded due to TTL).
+Thus, the left column of system data shows items that have been saved by TTL from either events, incidents, or phenomena.
+The left column shows items that have been collected in a single day (not including items that have been discarded by TTL).
 

### 2020-06-22 09-pilot.tex
More updates to the pilot and OPQ view clarifications.

@@
 Given that OPQ Boxes were configured to sample at 12 kHz at 2 bytes per sample during the deployment, we can calculate the bounds on raw storage for 16 OPQ Boxes over 3 months without TTL as 3.1 TiB.
-Figure~\ref{fig:data-management-graph} shows slightly less data as the number of OPQ Boxes was not constant during the pilot study (Figure~\ref{fig:active-opq-boxes})
+Figure~\ref{fig:data-management-graph} shows slightly less data as the number of OPQ Boxes was not constant during the pilot study (Figure~\ref{fig:active-opq-boxes}).
+By implementing TTL, significant data savings are made by discarding of IML data that is mostly noise.
 
@@
 
-%\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
-%Could you add a sentence or two to explain the assumptions under which the 2.5 TB storage requirement was calculated?
-%
-%Could you provide a paragraph to add more concrete details to this section? For example, how much additional storage was required, say, each week, by the end of the study period?
-%
-%\end{tcolorbox}
+Given a network of 16 OPQ Boxes over the pilot deployment period with TTL, the combined data storage from all levels reached an asymptotic limit of near 100 GB.
+The layout of data storage is provided per data storage level in Figure~\ref{fig:opq-data-storage}.
+The high rate of detections made by OPQ utilize most of the space.
+The reason for this is that Events store associated IML data along with each detection for high-fidelity analysis.
+It's also possible to observe how the AML and IML levels level off asymptomatically over time.
+This is a result of TTL discarding not useful data.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.75\linewidth]{images/pilot/laha-levels-storage.png}
+    \caption{Combined OPQ Data Storage}
+    \label{fig:opq-data-storage}
+\end{figure}
 
@@
 
-%One way we used this capability was to identify periodic phenomena, then predict when it was going to occur in future, and then change the settings associated with Boxes to support better data collection during periods in which a power quality anomaly was predicted.
-
 Two areas where adaptive optimizations were utilized are identifying periodic phenomena and predictions of future phenomena.
 Periodic phenomena are PQ related signals that occur at regular intervals.
-One example of periodic phenomena is a cyclic volage sag that was observed at one of our sensors with a period of about 34 minutes during our pilot study (Figure~\ref{fig:periodic-voltage-sags}).
+One example of periodic phenomena is a cyclic voltage sag that was observed at one of our sensors with a period of about 34 minutes during our pilot study (Figure~\ref{fig:periodic-voltage-sags}).
 This periodic phenomena not only provided a classification of an interesting pattern, but also helped optimize the OPQ system for capture of future phenomena.
@@
 Second, the increase in detection sensitivity makes it more likely that sub-threshold signals of interest will be classified.
-As example, the increased detection sensitivity allowed us to observe and classify a periodic voltage sag that would not have have been detected by our normal classifier due to the small magnitude of the sag.
-The detection is provided in Figure~\ref{fig:sub-threshold-detection}.
+As example, the increased detection sensitivity allowed us to observe and classify a periodic voltage sag that would not have been detected by our normal classifier due to the small magnitude of the sag.
+The sub-threshold detection is shown in Figure~\ref{fig:sub-threshold-detection}.
 
@@
 
-This creates a positive feedback loop where as periodic phenomena become more accurate, future phenomena become more accurate.
+This creates a positive feedback loop whereas periodic phenomena become more accurate, future phenomena become more accurate.
 As future phenomena become more accurate, periodic phenomena become more accurate.
-
-%\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
-%Can you provide a "for example" of this with more concrete details, hopefully indicating how this adaptive optimization actually yielded better results?
-%\end{tcolorbox}
 

### 2020-06-22 03-architecture.tex
Some updates to table 1.
Get rid of colorbox.

@@
 
-\begin{tcolorbox}[title=ANTHONY AND SERGE]
-You might notice that there is no mention of "Napali" or "Laha" in this paper.
-I could not think of a good way to add them in that would not additionally complicate any already terminologically complicated topic.
-To me, they feel like "overlays" onto the four component architecture described here and thus not strictly necessary to understand the system, but let me know if you disagree.
-\end{tcolorbox}
-
 This architecture has a number of benefits.
@@
 Similarly, data points in the Measurement layer will be deleted after one day if they are not associated with an Event.
-Details of this hierarchy are further explained in Section~\ref{sec:opq-mauka}.
 
@@
 For example, "when winds exceed 40 knots in Kahuku, data from prior Incidents leads to the prediction that there will be a voltage surge of approximately 5V for approximately 3 seconds in all boxes located within 1 mile of Kahuku."
+
+Further details of this hierarchy are provided in Section~\ref{sec:opq-mauka}.

### 2020-06-22 09-pilot.tex
First pass

--- 2020-06-22 09-pilot.tex
@@
 
-Unlike frequency and voltage, where there was very close alignment between OPQ Boxes and building meters, THD data shows two peaks, one of which is indicates a kind of "offset" of approximately 0.13% by the OPQ Boxes from the building meters.
-
-\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
-Could you augment the above paragraph with a few sentences about why you think the THD validation looked like it did? I know you covered this in your defense but I forgot.
-\end{tcolorbox}
+During the hours of 6pm and 6am, OPQBox and the building level meter displayed a high level of agreement as shown in the blue histogram in Figure \ref{fig:opqbox-thd-validaton}.
+On the other hand, during the hours of 6pm and 6am, there was a static disparity of 0.13\% between the two meters as portrayed  in the red histogram.
+This is likely attributed to the meter location in the power transmission hierarchy.
+While the OPQ Box is plugged the $120V_{ac}$ line, the building meter is monitoring the $480V_{ac}$ three phase line.
+An additional active conditioning system installed along side the transformer responsible for compensating for reactive power in the building is the likely culprit in the disparity.
 
 Unfortunately, we unable to perform validation of transient data collection, because building meters did not provide us with access to that information. 
-
-\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
-Could you add an additional sentence or two to the above paragraph about transient data validation? Maybe more about why we couldn't do it, or what the implications of non-validation might be?
-\end{tcolorbox}
+However, synthetic test performed in the lab against a calibration source, showed that OPQ Box is able to measure transient magnitude with $\sigma=0.125V$, significantly higher then the triggering threshold.
+It should be noted that the internal transient metric provided by the OPQ Box is only used for event detection, while higher level analytics compute their own transient classification parameters.
+As such, the transient detection capabilities of the OPQ Box are more then sufficient for its role in the OPQ ecosystem.
 

### 2020-06-22 07-opq-mauka.tex
More updates to the pilot and OPQ view clarifications.

@@
 But what we actually want is actionable insights into grid stability.
-For example, we might want to know if a given anomalous data value is actually detrimental, or we might wnat to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.
+For example, we might want to know if a given anomalous data value is actually detrimental, or we might want to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.
 
@@
 \end{figure}
-
-%\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=ANTHONY]
-%Please review the following paragraphs for accuracy and quality.
-%\end{tcolorbox}
 

### 2020-06-22 08-opq-view.tex

@@
 
-The home page ``System Stats'' widget shows the number of items stored at each OPQ storage level.
+The home page's ``System Stats'' widget shows the number of items stored at each OPQ storage level.
 The right hand column shows items that have been stored all time (not including items that have been discarded due to TTL).
-Thus, the left column of system data shows items that have been saved by TTL from either events, incidents, or phenomena.
 The left column shows items that have been collected in a single day (not including items that have been discarded by TTL).

### 2020-06-22 09-pilot.tex
More edits from serge

@@
 
-\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
-Please review and improve the above paragraph.
-\end{tcolorbox}
-
-To assess computational cost, we again analyzed the data stream for a typical day...
-
-\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
-Please finish the above paragraph by describing in words the histogram in slide 38 of your thesis defense presentation, where you show that Napali uses much less computational resources.
-\end{tcolorbox}
+Computational cost, is proportional to the amount of date acquired.
+Since the meter level analytics are at best only good enough for event detection, further analysis is required for event classification.
+This implies that the OPQ Makai triggering algorithm is at least 8 times more efficient then the self triggered counterpart. 
+Furthermore, in the self triggered case, data at the lowest granularity consists of waveforms from an individual device.
+These waveforms need to be matched together in order to extract grid-wide anomalies.
+On the other hand, OPQ Makai events contain data from all devices which experienced an anomaly, thus saving additional CPU time.
 
@@
 In the event that any of those boxes actually do report over-threshold data, then the request ripples outward to the boxes neighboring that box, and so forth.
+This is accomplished by maintaining an model of the OPQ Box metrics for each device.
+Once a single device passes a threshold, OPQ Makai will search the network for all other devices which exhibit measurements which deviate significantly from the device model.
+If more devices are located, data from all the affected devices will be captured.
+Otherwise the event is deemed local and ignored.
 
-During the pilot study, we discovered several situations in which the OPQ sensor network was able to characterize power quality anomalies in the UH microgrid in a manner that the building level meters could not.
-For example...
-
-\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=SERGE]
-Please finish the above paragraph by describing one situation of subthreshold event detection and why it was cool.
-\end{tcolorbox}
+Sub-threshold triggering allowed us to cluster devices based on severity of the disturbances they experienced.
+These clusters closely corresponded to the topology of the power grid.
+This in turn made it possible to perform limited localization of PQ incidents without prior knowledge of the grid layout.
 

### 2020-06-24 05-opq-box.tex
OPQ Makai first draft done.
Updated clustering and localization

@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \begin{center}
@@
 
-\begin{figure}[ht!]
-	\centering
-	\begin{subfigure}{.8\textwidth}
-	  \centering
-	  \includegraphics[width=1\linewidth]{images/opq-box/filter1_gain.png}
-	  \caption{}
-	  \label{fig:opq:4:1}
-	\end{subfigure}%
-
-	\begin{subfigure}{.8\textwidth}
-	  \centering
-	  \includegraphics[width=1\linewidth]{images/opq-box/filter1_response.png}
-	  \caption{}
-	  \label{fig:opq:4:2}
-	\end{subfigure}
-\end{figure}
-\begin{figure}[ht!]\ContinuedFloat
-	\centering
-	\begin{subfigure}{.75\textwidth}
-	  \centering
-	  \includegraphics[width=1\linewidth]{images/opq-box/filter2_gain.png}
-	  \caption{}
-	  \label{fig:opq:4:3}
-	\end{subfigure}%
-
-	\begin{subfigure}{.75\textwidth}
-	  \centering
-	  \includegraphics[width=1\linewidth]{images/opq-box/filter2_response.png}
-	  \caption{}
-	  \label{fig:opq:4:4}
-	\end{subfigure}
-	\caption{Filters used for mains frequency calculation. (a) Downsampling filter gain. (b) Downsampling filter impulse response. (c) Lowpass filter gain. (d) Lowpass filter impulse response.}
-	\label{fig:opq:4}
-\end{figure}
-
 Where the $\Delta t_{k}$ is the k'th time between two adjacent zero crossings.
@@
 Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz), it is very computationally expensive to perform this filtering in a single step.
-Instead, filtering is accomplished via a set of two finite impulse response (FIR) filters shown in Figure~\ref{fig:opq:4:2} and~\ref{fig:opq:4:4}.
-First, the Down sampling filter is applied to the raw waveform, which results in the frequency response shown in Figure~\ref{fig:opq:4:1}.
-As is evident by the plot, the frequency content of the result is 0-600Hz, Thus, it can be downsampled to the $\frac{N}{10}$, or 200 samples without aliasing.
-
-Next, the low pass filter is applied to the downsampled waveform with the frequency response shown in Figure~\ref{fig:opq:4:3}.This resulting frequency content is 0-100Hz, thus all of the higher frequency harmonics and noise are removed.
-
-Finally, the twice filtered downsampled waveform is used to estimate the fundamental frequency according to the Equation~\ref{eq:1}.
+Instead, filtering is accomplished via a set of two low pass finite impulse response (FIR) filters.
+First filter has a passband of 0-600Hz allowing us to downsample the waveform to 1200 samples per second.
+Next filter has a passband of 0-100Hz allowing for further removal of high frequency noise.
+Finally zero crossing are extracted and used for the frequency calculation.
 The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 		\centering
@@
 
-OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter and searching for a maximum value in the remainder.
-The high pass filter has a cutoff of $400Hz$, and the filter coefficients and response are shown in Figure~\ref{fig:opq:8:2} and Figure~\ref{fig:opq:8:1} respectively.
+OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter with a cutoff frequency of $400Hz$ and searching for a maximum value in the remainder.
 The result of the high pass filter operation is shown in Figure~\ref{fig:opq:8}.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=1\linewidth]{images/opq-box/thd_hpf_resp.pdf}
-		\caption{}
-		\label{fig:opq:8:1}
-	\end{subfigure}%
-	\begin{subfigure}{.5\textwidth}
-		\centering
-		\includegraphics[width=1\linewidth]{images/opq-box/thd_hpf.pdf}
-		\caption{}
-		\label{fig:opq:8:2}
-	\end{subfigure}
-
 	\begin{subfigure}{.5\textwidth}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\begin{center}
@@
 As of the time of writing, a single OPQ Box can be manufactured for approximately \$100 in parts.
-The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately $75 in parts.
+The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately \$75 in parts per device.

### 2020-06-24 09-pilot.tex
Updated clustering and localization

@@
 We deployed 16 OPQ Boxes across campus.
-As noted above, each OPQ Box collects 200 measurements per grid cycle, for a total of approximately 1B raw measurements per day per box.
+As noted above, each OPQ Box collects 200 measurements per grid cycle, for a total of approximately 17M raw measurements per day per box.
 Values for the maximum and minimum voltage, frequency, and THD, along with the presence or absence of transients, is sent once a second to the cloud by each box.
@@
 
-During the hours of 6pm and 6am, OPQBox and the building level meter displayed a high level of agreement as shown in the blue histogram in Figure \ref{fig:opqbox-thd-validaton}.
+During the hours of 6pm and 6am, OPQBox and the building level meter displayed a high level of agreement as shown in the blue histogram in Figure \ref{fig:opqbox-thd-validation}.
 On the other hand, during the hours of 6pm and 6am, there was a static disparity of 0.13\% between the two meters as portrayed  in the red histogram.
@@
 
-Sub-threshold triggering allowed us to cluster devices based on severity of the disturbances they experienced.
+Sub-threshold triggering allows for clustering of OPQ Boxes based on their electrical distance without a priori knowledge of the power grid topology.
+A distance metric comprised of the magnitude of the $V_{rms}$  disturbance between every pair of OPQ Boxes was used as the clustering parameter.
+This distance metric describes how much local disturbances observed by a device affect every other device on the network.
+Hierarchical clustering employed on the aforementioned distance metric results in a the hierarchy shown in Figure \ref{fig:clustering}.
 These clusters closely corresponded to the topology of the power grid.
-This in turn made it possible to perform limited localization of PQ incidents without prior knowledge of the grid layout.
+For example the two top level clusters correspond to the two 12kV feeders which power the university campus.
+As such, disturbances which originate on one feeder, have a minimal impact on the other.
+On the other hand, devices 1009 and 1008 are located in adjacent buildings, and a common 480V feeder, resulting in large level of commonality in observed anomalies.
+Sub-threshold triggering along with clustering in turn made it possible to perform limited localization of PQ incidents without prior knowledge of the grid layout.
+In example, consider a $V_{rms}$ sag observed by the OPQ network shown in Figure \ref{fig:localisation}.
+This event had the largest impact on devices 1008 and 1009 which as shown via hierarchical clustering are closely coupled.
+From this information it is possible to conclude that the event depicted in Figure \ref{fig:localisation} originated from inside the UH power grid relatively close to devices 1008 and 1009.
+
+\begin{figure}[ht]
+
+    \centering
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=0.9\linewidth]{images/pilot/clustring.png}
+        \caption{Hierarchical clustering of OPQ Boxes.}
+        \label{fig:clustering}
+    \end{subfigure}%
+    \begin{subfigure}{.5\textwidth}
+        \centering
+        \includegraphics[width=0.9\linewidth]{images/pilot/localisation.png}
+        \caption{Localisation of a disturbance.}
+        \label{fig:localisation}
+    \end{subfigure}
+    \caption{Hierarchical clustering of OPQ Boxes, and localization of a disturbance.}
+    \label{fig:localization-clustering}
+\end{figure}
 

### 2020-06-24 02-intro.tex
Fix some typos in the introduction.

@@
 We need more renewable energy.} Concerns including pollution, environmental degradation, and climate change have produced a global movement away from centralized, fossil-fuel based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar.
-But the economic, environmental, and political advantages of renewable energy comes with significant new technical challenges.
+But the economic, environmental, and political advantages of renewable energy come with significant new technical challenges.
 Wind and solar are intermittent (for example, solar energy cannot be harvested at night) and unpredictable (for example, wind and solar energy fluctuate based upon cloud cover and wind speed).
@@
 The goal of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
-OPQ accomplishes this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services.
+OPQ accomplishes this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based services.
 An OPQ sensor network is designed to be useful to: consumers who want to better understand the performance of their public utilities; utilities who desire a low-cost, easily deployable and re-deployable sensor network to gather information about grid stability that complements their existing infrastructure; to researchers who wish to design improvements to grid management and need a low cost mechanism for measuring the impact of their improvements; and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
@@
 Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and with what fidelity in order to most efficiently produce actionable information.
-Two-way communication between the OPQ Boxes and could services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
+Two-way communication between the OPQ Boxes and cloud services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
 As a result, OPQ cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.

### 2020-06-24 03-architecture.tex
Updates to architecture.

@@
 The combination of low and high fidelity data reduces both network overhead and storage requirements, which increases the scalability of the system in terms of the number of OPQ Boxes that can be tied to a single OPQ Cloud instance.
-OPQ Makai and OPQ Mauka have a plugin architecture, making it easier to extend their functionality to incorporate new triggers for high quality data (in the case of OPQ Makai) and new events and analyses (in the case of OPQ Mauka).
+OPQ Makai and OPQ Mauka have a plugin architecture, making it easier to extend their functionality to incorporate new triggers for high quality data (in the case of OPQ Makai) and new events and power analyses (in the case of OPQ Mauka).
 Finally, the open source licensing of both hardware and software makes it possible to incorporate new ideas, bug fixes, and enhancements from technologists across the power quality community.
@@
 The information architecture must address several requirements.
-First, it must support analyses that require access to high fidelity data, which in the case of power quality is wave form data produced by sampling the wave form 256 times per cycle, which means there are approximately 1.3 billion data points generated per day per OPQ Box.
+First, it must support analyses that require access to high fidelity data, which in the case of power quality is wave form data produced by sampling the wave form 256 times per cycle (or 12,000 times per second), which means there are approximately 1 billion data points generated per day per OPQ Box.
 Second, for typical grids that are generally stable, only an extremely small fraction of these data points are actually interesting.
@@
 Measurement		& 1 day/2 weeks		& Aggregate measurements provide "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients) at a rate of either 1/second (Measurement) or 1/day (Trend).
+Measurements and Trends both collect summary statistics, but at different window lengths and with different TTLs.\\
+Event/Detection & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': wave form data from one or more boxes for a specific time interval (typically, just a second or so).
 \\
-Event/Detection & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': wave form data from one or more boxes for a specific time interval (typically just a second or so).
-\\
-Incident & 6 months & An Incident is created when an Event satisfies a standard (such as IEEE 1159, ITIC, or SEMA) for a significant power quality problem.\\
+Incident & 6 months & An Incident is generated by OPQ Cloud's Mauka when an Event satisfies a standard (such as IEEE 1159, ITIC, or SEMA) for a significant power quality problem.\\
 Phenomena & 1 year & Phenomena classify Incidents when they are predictable and/or have a known causal explanation. \\
@@
 Note that Makai only has one hour to request high fidelity wave form data before it is lost.
+Mauka may also request Event data for specific Phenomena, which will be discussed in more detail below.
 
@@
 
-All of the prior levels represent behaviors of a individual Boxes in a single location at a given point in time.
+All of the prior levels represent behaviors of individual Boxes in a single location at a given point in time.
 This final level of OPQ's information architecture is premised on the idea that insightful, actionable information results from the ability to either explain or else predict multiple Incidents.
@@
 
-Further details of this hierarchy are provided in Section~\ref{sec:opq-mauka}.
+Further details of this information hierarchy are provided in Section~\ref{sec:opq-mauka}.

### 2020-06-24 04-related.tex
Updates to related.

@@
 While the OPQ Box has much more limited functionality, it is designed to be manufactured for approximately \$75, which is 10 to 100 times less expensive than most commercial devices, and similar is cost to OpenZMeter.
-The most important distinguishing feature of OPQ Box is that it is designed to grid-level, not single point monitoring, and thus incorporates features (such as two way communication with the cloud, and subthreshold triggering) that are not present in devices intended for "stand alone" capabilities.
+The most important distinguishing feature of OPQ Box is that it is designed to grid-level, not single point monitoring, and thus incorporates features (such as two-way communication with the cloud, and subthreshold triggering) that are not present in devices intended for "stand alone" capabilities.
 

### 2020-06-24 05-opq-box.tex
Updates to opq box.

@@
 An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
-5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference and an isolated DC/DC converter.
+5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference, and an isolated DC/DC converter.
 3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
@@
 A previous iteration of the OPQ Box overcame this by employing a small circuit board mount isolation transformer.
-Unfortunately it was found that the frequency response of these transformers varied wildly between individuals, thus incurring a lengthy calibration process for each device.
+Unfortunately, it was found that the frequency response of these transformers varied wildly between individuals, thus incurring a lengthy calibration process for each device.
 The current OPQ Box design solved this issue by using an AMC1100 isolation amplifier as the isolation component.
@@
 Next filter has a passband of 0-100Hz allowing for further removal of high frequency noise.
-Finally zero crossing are extracted and used for the frequency calculation.
+Finally, zero crossings are extracted and used for the frequency calculation.
 The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.
@@
 \end{equation}
-Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation, however unlike the frequency calculation the input is not filtered a priori.
+Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation. Unlike the frequency calculation, the input is not filtered a priori.
 
@@
 First an FFT transforms the real voltage samples into its frequency components.
-Next, the square of the harmonic bins is accumulated and scaled by the magnitude of the fundamental power.
+Next, the square of the harmonic bins are accumulated and scaled by the magnitude of the fundamental power.
 
@@
 Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
-This protocol is a publish subscribe topology, with each message containing the topic, and a payload.
+This protocol is a publish subscribe topology, with each message containing the topic and a payload.
 Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
@@
 However, since GPS receivers require line of sight to the sky, it was not used for deployment.
-NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a close by NTP server.
+NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a nearby NTP server.
 This error is visualized in a Figure~\ref{fig:opq:6:2}.

### 2020-06-24 02-intro.tex
wave form -> waveform.

@@
 Two-way communication between the OPQ Boxes and cloud services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
-As a result, OPQ cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.
+As a result, OPQ cloud services can obtain and analyze high fidelity waveform data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.
 This enables OPQ installations to be simultaneously responsive and scalable.

### 2020-06-24 03-architecture.tex
wave form -> waveform.

@@
 The information architecture must address several requirements.
-First, it must support analyses that require access to high fidelity data, which in the case of power quality is wave form data produced by sampling the wave form 256 times per cycle (or 12,000 times per second), which means there are approximately 1 billion data points generated per day per OPQ Box.
+First, it must support analyses that require access to high fidelity data, which in the case of power quality is waveform data produced by sampling the waveform 256 times per cycle (or 12,000 times per second), which means there are approximately 1 billion data points generated per day per OPQ Box.
 Second, for typical grids that are generally stable, only an extremely small fraction of these data points are actually interesting.
@@
 \midrule
-Box		& 1 hour			& Collects and holds a rolling one hour window of "high fidelity" instantaneous wave form data measurements for a single location.\\
+Box		& 1 hour			& Collects and holds a rolling one hour window of "high fidelity" instantaneous waveform data measurements for a single location.\\
 Measurement		& 1 day/2 weeks		& Aggregate measurements provide "low fidelity" summary statistics regarding four power measurements (Frequency, Voltage, THD, Transients) at a rate of either 1/second (Measurement) or 1/day (Trend).
 Measurements and Trends both collect summary statistics, but at different window lengths and with different TTLs.\\
-Event/Detection & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': wave form data from one or more boxes for a specific time interval (typically, just a second or so).
+Event/Detection & 1 month & When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request an ``event'': waveform data from one or more boxes for a specific time interval (typically, just a second or so).
 \\
@@
 Measurements are the second layer of the information architecture, and provide "low fidelity" summary statistics regarding four power quality measurements (Frequency, Voltage, THD, and Transients).
-Measurements give OPQ Cloud basic situational awareness of the grid without the overhead of transmitting, storing, and analyzing high fidelity wave form data.
+Measurements give OPQ Cloud basic situational awareness of the grid without the overhead of transmitting, storing, and analyzing high fidelity waveform data.
 Measurements exist for one day, though a daily roll-up of Measurement summary statistics called "Trends" persists longer.
 
-When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request wave form data from one or more boxes for a specific time interval (typically just a second or so).
-Note that Makai only has one hour to request high fidelity wave form data before it is lost.
+When OPQ Cloud's Makai service detects non-nominal values in the stream of Measurements, it can decide to request waveform data from one or more boxes for a specific time interval (typically just a second or so).
+Note that Makai only has one hour to request high fidelity waveform data before it is lost.
 Mauka may also request Event data for specific Phenomena, which will be discussed in more detail below.

### 2020-06-24 07-opq-mauka.tex
wave form -> waveform.

@@
 It also exists in the cloud, in the event that OPQ's triggering mechanism has caused a temporal interval of waveform data to be uploaded.
-IML data in the cloud has a TTL of 15 minutes: unless the wave form data is found to be useful by a cloud service within 15 minutes, it can be discarded.
+IML data in the cloud has a TTL of 15 minutes: unless the waveform data is found to be useful by a cloud service within 15 minutes, it can be discarded.
 

### 2020-06-24 09-pilot.tex
wave form -> waveform.

@@
 One interesting capability enabled by two way communication between OPQ Boxes and their cloud-based services is what we term {\em subthreshold event detection based on temporal locality}.
-In a nutshell, when one OPQ Box determines that a power quality measure has exceeded a threshold, one of the actions of the sensor network is to request high fidelity wave form data from neighboring boxes for the same temporal period, regardless of whether those boxes are reporting over-threshold data.
+In a nutshell, when one OPQ Box determines that a power quality measure has exceeded a threshold, one of the actions of the sensor network is to request high fidelity waveform data from neighboring boxes for the same temporal period, regardless of whether those boxes are reporting over-threshold data.
 In the event that any of those boxes actually do report over-threshold data, then the request ripples outward to the boxes neighboring that box, and so forth.

### 2020-06-24 09-pilot.tex
Updates to pilot.

@@
 \item {\em Evaluation of sensor data accuracy and quality.} While we conducted extensive laboratory tests of the sensor network, the pilot study enabled us to evaluate the accuracy and utility of our hardware devices and power quality data collection procedures in a real-world setting.
-Does an OPQ sensor network measure power quality (and detect anomalies) as well or better than conventional building-level electrical meters?
+Does an OPQ sensor network measure power quality (and detect anomalies) as well (or better) than conventional building-level electrical meters?
 
@@
 
-\item {\em Evaluation of cloud data storage management.} OPQ Mauka manages data storage through a "time to live" (TTL) mechanism, which results in data being discarded if it is not found to be useful and thus "promoted" to a higher level of the information hierarchy within a defined period of time.
+\item {\em Evaluation of cloud data storage management.} OPQ Mauka manages data storage through a "time-to-live" (TTL) mechanism, which results in data being discarded if it is not found to be useful and thus ``promoted" to a higher level of the information hierarchy within a defined period of time.
 The pilot study enables us to evaluate this approach.
@@
 We deployed 16 OPQ Boxes across campus.
-As noted above, each OPQ Box collects 200 measurements per grid cycle, for a total of approximately 17M raw measurements per day per box.
+As noted above, each OPQ Box collects 200 measurements per grid cycle, for a total of approximately 1B raw measurements per day per box.
 Values for the maximum and minimum voltage, frequency, and THD, along with the presence or absence of transients, is sent once a second to the cloud by each box.
@@
 Unfortunately, we unable to perform validation of transient data collection, because building meters did not provide us with access to that information.
-However, synthetic test performed in the lab against a calibration source, showed that OPQ Box is able to measure transient magnitude with $\sigma=0.125V$, significantly higher then the triggering threshold.
+However, synthetic tests performed in the lab against a calibration source, showed that OPQ Box is able to measure transient magnitude with $\sigma=0.125V$, significantly higher then the triggering threshold.
 It should be noted that the internal transient metric provided by the OPQ Box is only used for event detection, while higher level analytics compute their own transient classification parameters.
-As such, the transient detection capabilities of the OPQ Box are more then sufficient for its role in the OPQ ecosystem.
+As such, the transient detection capabilities of the OPQ Box are more than sufficient for its role in the OPQ ecosystem.
 
@@
 
-One interesting capability enabled by two way communication between OPQ Boxes and their cloud-based services is what we term {\em subthreshold event detection based on temporal locality}.
+One interesting capability enabled by two-way communication between OPQ Boxes and their cloud-based services is what we term {\em subthreshold event detection based on temporal locality}.
 In a nutshell, when one OPQ Box determines that a power quality measure has exceeded a threshold, one of the actions of the sensor network is to request high fidelity waveform data from neighboring boxes for the same temporal period, regardless of whether those boxes are reporting over-threshold data.
 In the event that any of those boxes actually do report over-threshold data, then the request ripples outward to the boxes neighboring that box, and so forth.
-This is accomplished by maintaining an model of the OPQ Box metrics for each device.
+This is accomplished by maintaining a model of the OPQ Box metrics for each device.
 Once a single device passes a threshold, OPQ Makai will search the network for all other devices which exhibit measurements which deviate significantly from the device model.
 If more devices are located, data from all the affected devices will be captured.
-Otherwise the event is deemed local and ignored.
+Otherwise, the event is deemed local and ignored.
 
@@
 This distance metric describes how much local disturbances observed by a device affect every other device on the network.
-Hierarchical clustering employed on the aforementioned distance metric results in a the hierarchy shown in Figure \ref{fig:clustering}.
+Hierarchical clustering employed on the aforementioned distance metric results in the hierarchy shown in Figure \ref{fig:clustering}.
 These clusters closely corresponded to the topology of the power grid.
-For example the two top level clusters correspond to the two 12kV feeders which power the university campus.
+For example the two top-level clusters correspond to the two 12kV feeders which power the university campus.
 As such, disturbances which originate on one feeder, have a minimal impact on the other.
@@
 As future phenomena become more accurate, periodic phenomena become more accurate.
-
-\begin{tcolorbox}[title=ANTHONY AND SERGE]
-Any other pilot study results that I've forgotten to include?
-\end{tcolorbox}

### 2020-06-24 10-conclusions.tex
wave form -> waveform.
Updates to conclusions.

@@
 
-Our sensor network provides real-time, two way communication between the sensor nodes (OPQ Boxes) and their cloud services, providing several innovative features for power quality monitoring.
+Our sensor network provides real-time, two-way communication between the sensor nodes (OPQ Boxes) and their cloud services, providing several innovative features for power quality monitoring.
 The OPQ triggering system exploits temporal locality to request high fidelity data from neighboring boxes when one exceeds a threshold, enabling our network to detect and analyze data that would have been unreported by a naive, threshold-based approach.
 The OPQ Information Architecture enables Actors at the Phenomena Layer to control sensitivity settings of individual OPQ Boxes in order to improve data collection when power quality anomalies have been predicted.
-
+`
 Our sensor network is designed to efficiently use network and storage resources.
-Two way communication enables OPQ Boxes to send low fidelity summary statistics on power quality at one second intervals, which can be used by cloud-based services to decide whether or not to request high fidelity data.
+Two way communication enables OPQ Boxes to send low fidelity summary statistics on power quality at one second intervals, which can be used by cloud-based services to decide whether to request high fidelity data.
 Our TTL mechanism implements a "use it or lose it" approach to cloud-based data, which in our pilot study reduced cloud storage requirements by over ten-fold.

### 2020-06-27 09-pilot.tex
final tweaks

@@
 
-Computational cost, is proportional to the amount of date acquired.
+The computational cost of an OPQ sensor network is proportional to the amount of data acquired.
 Since the meter level analytics are at best only good enough for event detection, further analysis is required for event classification.
-This implies that the OPQ Makai triggering algorithm is at least 8 times more efficient then the self triggered counterpart.
-Furthermore, in the self triggered case, data at the lowest granularity consists of waveforms from an individual device.
-These waveforms need to be matched together in order to extract grid-wide anomalies.
-On the other hand, OPQ Makai events contain data from all devices which experienced an anomaly, thus saving additional CPU time.
+This implies that the OPQ Makai triggering algorithm is at least 8 times more efficient then the self triggered counterpart, since 8 times less data is sent by OPQ Boxes to the cloud for analysis.
 

### 2020-07-16 cover-letter/reviewer-1.tex
save reviewer letters.

\documentclass[12pt]{letter}
\usepackage{times}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\signature{Philip Johnson}
\address{Information \& Computer Sciences \\ University of Hawaii \\ Honolulu, HI USA 96822}
\longindentation=0pt
\begin{document}

\begin{letter}{}

{\bf Comments from Reviewer 1:}

{\em (1) I think this article should be shortened significantly. The first sections are very descriptive. It's more a project report than a scientific article. Only the section on the results (section 8) is more research oriented. Therefore I suggest to rework the article such that there is much more focus on the scientific aspects and results.}

Fixed.

\end{letter}
\end{document}

### 2020-07-16 cover-letter/reviewer-2.tex
save reviewer letters.

\documentclass[12pt]{letter}
\usepackage{times}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\signature{Philip Johnson}
\address{Information \& Computer Sciences \\ University of Hawaii \\ Honolulu, HI USA 96822}
\longindentation=0pt
\begin{document}

\begin{letter}{}

{\bf Comments from Reviewer 2:}

{\em (1) In my opinion the paper is too long. Getting through the whole paper is very tiring. It seems that the maximum length of a scientific article (not including appendix) should be 20–22 pages. Maybe some of the descriptions can be moved to the appendix.}

{\em (2) I suggest a typical layout used in the articles: Introduction, Methods, Results and Discussion, Conclusions. This template is also recommended by Journal (see Instructions for Authors).}

{\em (3) Because this is a scientific article, the aims of the work and the research gap should be more clearly highlighted. I also suggest to improve the conclusions, delete the description of the work carried out (it is already well known) and focus on the most important conclusions and the future research. }

{\em (4) The quality of Figures should be corrected, their descriptions, especially axis descriptions are unreadable.}

Fixed.  The issue is incorporating modern web application development as the target domain.  The tech stack associated with modern web application development is complex enough that the semester can easily devolve into ``learning the tech stack", not ``experience software engineering".  Historically, the target domains for project, studio, and flipped pedagogies required only a simple tech stack and so this was not an issue.

\end{letter}
\end{document}

### 2020-07-16 cover-letter/reviewer-3.tex
save reviewer letters.

--- 2020-07-16 cover-letter/reviewer-2.tex
@@
 
-{\bf Comments from Reviewer 2:}
+{\bf Comments from Reviewer 3:}
 
-{\em (1) In my opinion the paper is too long.
-Getting through the whole paper is very tiring.
-It seems that the maximum length of a scientific article (not including appendix) should be 20–22 pages.
-Maybe some of the descriptions can be moved to the appendix.}
+{\em (1) Ideas presented in the article are well described, including their advantages and limitations.
+The introduction covers the basic overview of the topic.
+It is hard to point out a single specific problem of the article since authors are very precisely defining the structure and following this in the article body (some abbreviations are used before explanation (e.g.
+THD).
+Figures 6c and 6d are missing).}
 
-{\em (2) I suggest a typical layout used in the articles: Introduction, Methods, Results and Discussion, Conclusions. This template is also recommended by Journal (see Instructions for Authors).}
+Thanks for the kind words.
 
-{\em (3) Because this is a scientific article, the aims of the work and the research gap should be more clearly highlighted.
-I also suggest to improve the conclusions, delete the description of the work carried out (it is already well known) and focus on the most important conclusions and the future research.
-}
-
-{\em (4) The quality of Figures should be corrected, their descriptions, especially axis descriptions are unreadable.}
+{\em (2) Generally, I am looking for a scientific background of the article – and it is tough to find.
+Authors have already realized a lot of engineering work (and their presentation is precise); nevertheless, a deep (or deeper) scientific level is missing from my point of view.
+Anyway, the paper might be useful for future readers and can inspire them in their work.
+Therefore I recommend it for publishing.}
 

### 2020-07-16 cover-letter/make-all.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+# echo script execution
+set -x
+
+pdflatex reviewer-1.tex
+pdflatex reviewer-2.tex
+pdflatex reviewer-3.tex

### 2020-07-16 cover-letter/reviewer-3.tex
checkpoint

@@
 
-Thanks for the kind words.
+Thanks for the kind words. As part of the paper reorganization and shortening, Figure 6 has been removed.
 
@@
 
-Fixed.
-The issue is incorporating modern web application development as the target domain.
-The tech stack associated with modern web application development is complex enough that the semester can easily devolve into ``learning the tech stack", not ``experience software engineering".
-Historically, the target domains for project, studio, and flipped pedagogies required only a simple tech stack and so this was not an issue.
+Thanks again!  The rewrite attempts to inject more scientific background.
 

### 2020-07-17 02-intro.tex
checkpoint

@@
 Finally, collecting the data is not useful if it cannot be converted into actionable information in a reasonable amount of time.
-
-\subsection{Goals of the Open Power Quality Project}
-
-The goal of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
-OPQ accomplishes this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based services.
-An OPQ sensor network is designed to be useful to: consumers who want to better understand the performance of their public utilities; utilities who desire a low-cost, easily deployable and re-deployable sensor network to gather information about grid stability that complements their existing infrastructure; to researchers who wish to design improvements to grid management and need a low cost mechanism for measuring the impact of their improvements; and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.
-Several key features of our project emerge from these goals.
-
-{\em OPQ Box.} First, our custom power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75.
-This is from 10 to 100 times less expensive than commercial power quality monitors.
-This cost differential means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building.
-Similarly, instead of monitoring power quality at the substation level of the grid, OPQ Boxes make it feasible to deploy dozens or hundreds into a community to obtain fine-grained data about the impact of solar or other renewables as directly perceived by the customer.
-
-{\em Cloud-native information architecture.} Second, our information architecture is "cloud native", which means that OPQ Boxes are not designed for stand-alone, autonomous use, unlike current commercially available power quality monitors.
-Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and with what fidelity in order to most efficiently produce actionable information.
-Two-way communication between the OPQ Boxes and cloud services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network.
-As a result, OPQ cloud services can obtain and analyze high fidelity waveform data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud.
-This enables OPQ installations to be simultaneously responsive and scalable.
-
-{\em Plug-in analysis architecture.} Third, all OPQ software services are designed for extensibility and interoperation.
-Our middleware components (Mauka and Makai) provide a plugin architecture.
-Our visualization component (View) is built from modular UI elements using React.
-This means that OPQ is not a monolithic, closed system with a frozen feature set, but rather a design environment for experimentation with advanced classification and analysis of power quality data.
-
-{\em Open Source.} Fourth, our software and hardware designs and implementations are made available using open source licenses.
-This means that organizations choosing OPQ do not face the business risk of committing resources to a single vendor with proprietary hardware and software.
-Our goal is to facilitate the creation of a community of researchers and industry practitioners to replicate, extend, and apply the insights gained from the OPQ system.
-
-While we believe OPQ offers a compelling combination of capabilities, we want to be clear that all sensor network designs involve trade-offs.
-As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs.
-In some cases, OPQ may not be the appropriate choice.
-In other cases, a hybrid solution consisting of OPQ in combination with other technologies for grid monitoring and analysis may be most appropriate.
-
-\subsection{Structure of this paper}
-
-The remainder of this paper is organized as follows.
-Section \ref{sec:architecture} presents an overview of the OPQ system architecture and information architecture.
-Section \ref{sec:related-work} presents related research and technology.
-Section \ref{sec:opq-box} discusses OPQ Box and Section \ref{sec:opq-makai} its triggering system.
-Section \ref{sec:opq-mauka} describes our analysis services, and Section \ref{sec:opq-view} our user interface.
-Section \ref{sec:pilot-study} discusses the pilot deployment of a sensor network and its results.
-Section \ref{sec:conclusions} concludes with some proposals for future research.
+This paper describes results to date of the Open Power Quality project, which attempts to address these problems.

### 2020-07-17 02-methods.tex
checkpoint

\section{Methods}

The central aim of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.

Our research methodology consists of three (potentially interleaved) phases. First, we specify, design, and implement novel hardware and software for collecting, transmitting, and analyzing four essential measures of power quality data: voltage, frequency, THD, and transients. Second, we verify that our implementation matches our specifications through laboratory experiments.  Third, we validate our implementation by testing a variety of hypotheses through a real-world deployment. In this paper, we report on results from a pilot deployment at the University of Hawaii over the course of three months.

Our central hypothesis is that we can accomplish this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based services.

We will test this central hypothesis through six subhypotheses: (1) OPQ provides valid and reliable collection of power quality data (Section \ref{hyp:01}), (2) OPQ's triggering system provides advantages with respect to bandwidth and computation (Section \ref{hyp:02}, (3) OPQ enables subthreshold event detection based upon temporal locality (Section \ref{sec:subthreshold-events}), (4) The OPQ information architecture provides a means to produce actionable insights (Section \ref{hyp:04}), (5) the OPQ information architecture provides predictable upper bounds on storage resources (Section \ref{hyp:05}), and (6) OPQ provides useful adaptive optimization capabilites (Section \ref{sec:adaptive-optimization}).

The remainder of this section presents the overall architecture of OPQ, along with details on the design and implementation of each of the principal architectural components. We will also discuss our verification activities for each component. Our validation and hypothesis testing activities will be described in Section \ref{sec:pilot-study}.

\subsection{OPQ Architecture}

The OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events.  OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components either at predefined intervals or upon request. OPQ Makai monitors incoming low fidelity data from OPQ Boxes, requests high fidelity data when necessary, and stores the results in a MongoDB database. OPQ Mauka analyzes low level data and analyzes it to produce higher-level representations according to the information architecture described below, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis. OPQ View is a browser-based visualization platform for displaying the results for data capture and analysis.

\begin{figure}
\center \includegraphics[width=5in]{images/architecture/system-diagram.png}
\caption{High level system architecture of an OPQ Sensor Network}
\label{fig:architecture}
\end{figure}

Figure \ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side).  First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQ Makai. OPQ Makai analyzes low fidelity measurements, and requests high fidelity waveforms when desirable. Both measurements and waveforms are saved in a MongoDB database. OPQ Mauka analyzes low and high fidelity data, and creates "events" to represent anomalies. OPQ View notifies users of events and allows them to drill down into low and high fidelity data.

OPQ Makai, OPQ Mauka, and OPQ View are all cloud-based software services that collectively form a single "instance" with respect to data transmission, storage, analysis, and visualization. We refer to this collection of software-side components as OPQ Cloud. Every OPQ Box connects to a single instance of an OPQ Cloud. It is possible to have multiple OPQ Cloud instances. For example, a company might install an OPQ Cloud instance behind their firewall along with OPQ Boxes to provide a private mechanism for collecting and analyzing power quality data.

\subsection{OPQ Box}
\label{sec:opq-box}

OPQ Box is a hardware device designed to provide inexpensive, extensible and accurate residential power quality measurements.
A block diagram of OPQ Box is shown in Figure~\ref{fig:opq:1:1}.
A complete device is shown in Figure~\ref{fig:opq:1:2}.

\begin{figure}[ht]
	\centering
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
	  \caption{OPQ Box Block Diagram.
	  The power path is in red, signal path is in green and the digital IO is in black.}
	  \label{fig:opq:1:1}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
	  \caption{OPQ Box in an ABS plastic enclosure.}
	  \label{fig:opq:1:2}
	\end{subfigure}
	\caption{(a) OPQ Box block diagram and (b) production OPQ Box ready for deployment}
	\label{fig:opq:2}
\end{figure}

\subsubsection{Hardware}\label{subsec:hardware}

The power system of the OPQ Box electrically isolates most of the device from the AC mains power.
An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference, and an isolated DC/DC converter.
3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
The hot side of the isolation amplifier is powered from the isolated DC/DC converter.
This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power outage.

\begin{figure}[ht]
  \begin{center}
  \includegraphics[width=0.9\textwidth]{images/opq-box/opqbox_software.png}
  \end{center}
  \caption{Block diagram of the OPQ Box 2 software stack.}
  \label{fig:opq:3}
\end{figure}

A Raspberry Pi single board computer (SBC) is responsible for signal analysis and anomaly detection.
The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256MB of main memory and a single core 1GHz ARM11 CPU. It also contains an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle.

\subsubsection{Software}\label{subsec:software}

The software stack of the Raspberry Pi aims to deliver an accurate and precise power quality analysis framework despite the rather limited capabilities of the hardware.
A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with the Pi acting as the master and the DSP as a slave device.
A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer, and a kernel driver provides management of the SPI bus.
Internally, the OPQ driver maintains a ring buffer of 16 windows, each of which is 200 data samples in size.
Upon receiving the interrupt for the DSP, the CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
This scheme requires the CPU to only service 60 interrupts a second, with each interrupt requiring on the order of 100 instructions, for a CPU utilization of less than $1\%$ in normal operation.
Userland applications communicate with the kernel driver using a file descriptor, where every $read$ system call yields 200 samples of raw waveform.
As a result, the smallest window that a userland application may process is a single AC cycle of the grid mains.

The userland component of the OPQ Box software is a multi-threaded extensible analysis framework called Triggering.
The reader thread is responsible for transferring and accumulating data from the kernel driver.
The smallest data buffer that the Triggering application processes at any given time is 10 grid cycles or 2k samples.
Once the cycles are transferred to the userland and timestamped, they are passed to the analysis thread for feature extraction, as well as to the Raw Data Ring Buffer (RDRB).
Since internally all data is addressed using shared pointers, during data duplication no copying is required.
RDRS is capable of buffering up to an hour of data before it is overwritten, resulting in the RDBS maximum size of 100MB.

The analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
Four metrics are extracted from the data stream: (1) Fundamental frequency, (2) RMS Voltage, (3) Total Harmonic Distortion, and (4) Transients. Let's briefly discuss how each of these are computed.

\subsubsection{Fundamental Frequency}\label{subsec:fundamental-frequency}

The fundamental frequency is calculated by computing the zero crossings of the AC waveform.
In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz), it is very computationally expensive to perform this filtering in a single step.
Instead, filtering is accomplished via a set of two low pass finite impulse response (FIR) filters.
First filter has a passband of 0-600Hz allowing us to downsample the waveform to 1200 samples per second.
Next filter has a passband of 0-100Hz allowing for further removal of high frequency noise.
Finally, zero crossings are extracted and used for the frequency calculation.
The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.

\subsubsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}

Root mean square voltage ($V_{rms}$) in electrical power is the equivalent value of DC voltage which would dissipate the same power in the resistive load. $V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed value.

Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation. Unlike the frequency calculation, the input is not filtered a priori.

\subsubsection{Total Harmonic Distortion}\label{subsec:thd}

The OPQ Box calculates total harmonic distortion (THD) using the industry standard methodology.
It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
Operationally, OPQ Box computes THD for 10 cycles of the fundamental frequency.
First an FFT transforms the real voltage samples into its frequency components.
Next, the square of the harmonic bins are accumulated and scaled by the magnitude of the fundamental power.

\subsubsection{Transient Detection}\label{subsec:transient-detection}

OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter with a cutoff frequency of $400Hz$ and searching for a maximum value in the remainder.

It should be noted that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient detection.

\subsubsection{Network Communication}\label{subsec:network-communication}

The computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box to verify its peer.
Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
This protocol is a publish subscribe topology, with each message containing the topic and a payload.
Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
This allows for the aggregation service to be spread across multiple nodes, with minimal network overhead.

If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box RDRB via a separate ZeroMq pub sub channel.
If the RDRB buffer contains data for the requested temporal range, OPQ Box transmits the available data to the aggregation service via a push pull ZeroMq channel.
Protobuf message serialization is used to encode messages across the OPQ ecosystem.

In order to make a distributed measurement, all of the OPQ Boxes on the OPQ network need to maintain an accurate time reference.
Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
The expansion port of the OPQ Box supports a GPS receiver.
However, since GPS receivers require line of sight to the sky, it was not used for deployment.
NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a nearby NTP server.

\subsubsection{Manufacturing}

Currently, there is no mechanism for mass production of OPQ Boxes, but all of the plans are available under an Open Source license, so interested organizations with some basic hardware engineering skills can build their own boxes.

Complete specifications for the OPQ Box hardware, firmware, and software are available \cite{negrashov_opq_2020}. As of the time of writing, a single OPQ Box can be manufactured for approximately \$100 in parts. The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately \$75 in parts per device.

\subsection{OPQ Makai}
\label{sec:opq-makai}

OPQ Box provides an inexpensive hardware device for collecting four important power quality measures with high fidelity, but realizing its potential requires an innovative approach involving two-way communication between OPQ Boxes and the OPQ cloud-based services. To see why, consider the IEEE 1159 standard for single location power quality monitoring \cite{unruh_ieee_2018}.  For transient monitoring, IEEE 1159 suggests a sampling rate of at least 7680 samples/second, up to 1 Megasample/second. This implies that if the cloud service requires the high fidelity data from all OPQ Boxes, it would incur a very large bandwidth cost. At 20 Ksamples/second with 16bit samples, a single OPQ Box will generate 300Kb/second of network bandwidth. Several thousand devices would easily saturate a 1 GB network link. In addition, collecting and recording all of the raw waveform data from residential power quality meters could lead to security and privacy issues.

Network bandwidth saturation is a common problem for distributed sensor networks, and a common solution is called "self-triggering".  In this approach, each monitoring device is configured with a threshold value for one or more measures of interest. When the threshold for a measure of interest is exceeded, then and only then is data sent over the network to cloud-based services for analysis.

The problem with the self-triggering approach is that grid-wide power quality events do not affect the entire grid in the same way. For example, due to the grid’s hierarchical structure, a voltage sag on one sub-circuit can manifest as a sag of a different magnitude or even a swell on another \cite{kahle_power_2015}. This may result in a situation where some of the monitoring devices will not consider a power quality anomaly as an event, because it did not surpass the metric threshold, and simply ignore it. From an analysis perspective, however, it can be useful to get raw data from all of the affected devices, not just the ones that were the affected to the point where the box was triggered. This additional information can be used to localize the disturbance, as well as better evaluate its impact.

Since sending all the data is infeasible, and since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai. In this scheme, OPQ Boxes use local processing resources to feature extract the incoming waveforms while storing them locally for an hour. Each OPQ Box sends its feature data to OPQ Makai once a second, which we called the "triggering stream". Feature data is very small, on the order of a few kilobytes, and so this approach allows the sensor network to scale to thousands of devices with acceptable network bandwidth requirements.  OPQ Makai processes the incoming triggering stream and looks for anomalies. If an anomaly is present in only a single device, it is highly probable that the cause is local and not grid-level. On the other hand, if the triggering stream shows an anomaly temporally collocated across multiple devices, the entire network or a subset of the network may be queried for raw waveform data for a temporal region which corresponds to the disturbance in the triggering stream.

Our pilot study, discussed in Section \ref{sec:pilot-study} will provide examples of the novel analysis capabilities made possible by OPQ Box and OPQ Makai communication. In general, here are the main advantages of our hybrid centralized/decentralized approach over traditional self-triggering and the ``naive" approach of sending all of the data:

{\em Bandwidth usage is minimized.} Instead of sending the entirety of raw data, only extracted features are sent. This results in a tiny fraction of the bandwidth requirement when compared to raw waveforms. Furthermore, the temporal window which encompasses a single feature can be adjusted in real time. Thus, as soon as an anomalous behavior is observed in a subset of sensors, this window can be adjusted for a finer grained feature extraction.

{\em Effects of latency are minimized.} In this case, "latency" refers to the time required for OPQ Makai to process the incoming feature stream and decide whether to request high fidelity data from one or more OPQ Boxes. Even at 1M samples/second at 16 bits of resolution, the memory requirement to store 5 minutes of raw waveform without compression are on the order of 512MB, which is well within the realm of inexpensive single board computers such as Raspberry PI. With compression specifically suited to the signal of interest, the memory requirement can be reduced even further. In the case of the OPQ sensor network, OPQ Makai has an hour to process feature data and request high fidelity data from OPQ Boxes. During our pilot study, OPQ Makai always responded within a second or two.

{\em Cloud processing requirements are reduced.} Since feature extraction is already performed at the device level, cloud computation requirements are reduced. With the advent of the Internet of Things, the computational capacity of edge devices is increasing.

{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes. This allows analysis of data from devices which were only mildly affected or even not affected at all by the disturbance. This creates new possibilities for investigation of disturbance propagation across the sensed area, as will be discussed in Section \ref{sec:pilot-study}.

{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features. This allows for a simple robust algorithm which may be deployed at the sink node for anomaly detection.

\subsubsection{Design}

OPQ Makai is a distributed extensible microservice framework responsible for receiving the triggering stream from the OPQ Boxes, locating anomalous temporal regions and requesting raw waveform for the anomalous time ranges. As shown in Figure \ref{fig:makai-design}, Makai consists of four major components: Acquisition Broker, Triggering Broker, Event Service and the Acquisition Service.

\begin{figure}
\center \includegraphics[width=3in]{images/makai/makai_main.pdf}
\caption{Makai component design}
\label{fig:makai-design}
\end{figure}

The Triggering Broker is perhaps the simplest component of the OPQ Makai system. The triggering stream generated by the OPQ Boxes is encrypted to preserve user privacy. In order to minimize the CPU time spent decrypting the data across multiple OPQ services, the Triggering Broker decrypts the data and sends clear text measurements to other OPQ cloud services. The Triggering Broker uses the ZeroMq subscribe socket to receive data from OPQ Boxes, and sends it via a publish socket to any connected client. Each publish message is published to a topic which corresponds to the ASCII representation of the originating OPQ Box ID. This allows services which utilize the Triggering Broker to select a subset of IDs to operate on. This is useful for load balancing the backend services, or dividing the OPQ network into separate regions with no electrical connections between them.

The Acquisition Broker manages the two-way communication between the OPQ Boxes and the rest of the cloud infrastructure. Unlike the triggering stream which originates from the OPQ Box, two-way communication is always initiated by OPQ cloud services. Two way communication is realized via a command response interface, where the OPQ service initiates the communication by sending a clear text command to the Acquisition Broker, which then forwards it in encrypted form to the appropriate OPQ Boxes.

The Acquisition Service resides between the Triggering and Acquisition Brokers. The Acquisition Service is responsible for three tasks:
(1) Computation of statistics of the incoming triggering stream; (2) Hosting plugins for triggering stream analysis; and (3) Generating data event requests for OPQ Boxes. The Acquisition Service accesses the triggering stream by connecting to the publish socket of the Triggering broker. Since the connection is managed through the ZeroMq publish-subscribe socket, several Acquisition Services can be connected to a single Triggering broker endpoint, each servicing a subset of OPQ Boxes by subscribing to only specific devices. The Acquisition Service does not include any analysis capabilities by default. Instead, analysis is performed by shared library loadable plugins. These plugins can be loaded and unloaded at runtime, thus allowing live upgrading and testing of new analysis methods.

The Event service is a microservice which stores raw data generated by OPQ Boxes in the MongoDB Database. On initialization, the Event service queries MongoDB database for the highest event number recorded so far, connects to the Acquisition Broker’s publish port, and subscribes to all messages that start with the prefix ``data”. This allows the Event service to capture every response from OPQ Boxes generated from commands issued by the Acquisition service plugins. Once the Event service receives a data response with an identity containing an event token it hasn’t seen before, it will increment the event number, and store it in an internal key value store.

\subsection{OPQ Mauka}
\label{sec:opq-mauka}

The previous sections discussed the design of OPQ Box, a custom hardware device for collecting four important measures of power quality, and OPQ Makai, a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes.  As a result of these two innovations, an OPQ sensor network has the ability to collect and analyze high fidelity, low level data about power quality anomalies in a cost-effective, scalable fashion.

There are remaining challenges to creating a useful power quality sensor network. First, the data provided by OPQ Boxes is low-level, "primitive" data consisting of either features (i.e. frequency, voltage, THD, and transients) or waveform data. But what we actually want is actionable insights into grid stability. For example, we might want to know if a given anomalous data value is actually detrimental, or we might want to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.

A second challenge involves the potentially high volume of data might accumulate in the cloud. Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data could accumulate. One strategy is to simply store all of the data sent to the cloud forever. This means that data storage requirements will increase monotonically over time, making the sensor network more costly to maintain the longer it is in place. An alternative strategy is to implement an algorithm to identify uninteresting (or no longer interesting) data and discard it.  Ideally, such an algorithm would enable OPQ sensor network designers to calculate an upper bound on the total amount of cloud storage required as a function of the number of nodes (OPQ Boxes) in the network.

OPQ Mauka addresses both of these issues. First, OPQ Mauka provides a multi-layered representation for structuring and processing DSN data. The structure and processing at each layer is designed with the explicit goal of turning low-level data into actionable insights. Second, each layer in the framework implements a "time-to-live" (TTL) strategy for data within the level. This strategy states that data must either progress upwards through the layers towards more abstract, useful representations within a fixed time window, or else it can be discarded. The TTL strategy is useful because when implemented, it allows DSN designers to make reasonable predictions of the upper bounds on data storage at each level of the framework adjusting for the number of sensors and power anomaly probability.

 TTL also makes possible a ``graceful degradation" of system performance if those bounds turn out to be exceeded. For example, consider a situation in which a power network enters a prolonged period of widespread power quality instability, where every OPQ Box is reporting continuous anomalous conditions with respect to voltage, frequency, THD, and transients.  This ``worst case scenario" would lead to the potential situation of every OPQ Box trying to upload raw waveform data all the time. The TTL system provides safeguards, in that whatever low-level data has not been processed relatively quickly can be discarded.  Thus, instead of the system potentially going down entirely, it could instead continue to operate at a reduced capacity.

Figure \ref{fig:mauka-data-model} illustrates the hierarchical data model for OPQ Mauka. This data model can be conceptualized as a multi-level hierarchy that adaptively optimizes data storage using a tiered TTL approach and provides a mechanism in which typed aggregated data is continually refined to the point of being of becoming actionable. The data model also includes software components called "actors" that both move data upward through the levels and also apply optimizations downward through the levels. Actors are implemented through a plugin architecture, making it easy to experiment with the data model and improve it over time.

\begin{figure}
\center \includegraphics[width=4in]{images/mauka/mauka-data-model.png}
\caption{Mauka data model hierarchy}
\label{fig:mauka-data-model}
\end{figure}

The lowest layer of the hierarchy is the Instantaneous Measurements Layer (IML). The IML contains "raw" data, in other words, the digitized waveform.  IML data exists both on each OPQ Box (where it is available for up to the previous 60 minutes). It also exists in the cloud, in the event that OPQ's triggering mechanism has caused a temporal interval of waveform data to be uploaded. IML data in the cloud has a TTL of 15 minutes: unless the waveform data is found to be useful by a cloud service within 15 minutes, it can be discarded.

The second layer is the Aggregate Measurements Level (AML). The AML stores power quality summary statistics sent either once per second or once per minute by each OPQ Box. These summary statistics include the maximum, minimum, and average values of voltage, frequency, THD, and transient metrics over the time window. It is AML data that is used to initiate the triggering process of uploading IML data from the cloud.

The third layer is the Detections Level (DL). This layer is responsible for processing the IML and AML data to produce a representation of an "event" with a precisely defined start and end time based upon examination of the waveform.  As will be discussed in Section \ref{sec:subthreshold-events}, knowledge of the start and end time of a power quality anomaly allows investigation of how that anomaly might manifest itself elsewhere in the grid, even if this manifestation is not severe enough to produce over-threshold data values.

The fourth layer is the Incident Level (IL).  This layer starts to answer the question of whether the data is "actionable" by classifying the detected event according to various industry standards for power quality anomalies: IEEE 1159, ITIC, SEMI F47, and so forth.  For example, an event that falls into the ITIC "prohibited" region clearly indicates a power quality anomaly that requires further study and intervention.

The fifth and final level is the Phenomena Level (PL). This layer contains the results of analyses that attempt to identify cyclic, and thus predictable, power quality disturbances. It also contains analysis results regarding the similarity of various incidents, which can help uncover causal factors.  Finally, it provides analyses for adaptive optimization of the OPQ Sensor Network. These optimizations can change the thresholds for individual boxes to either increase their sensitivity or decrease their sensitivity over specific intervals of time. The ultimate goal of adaptive optimization is to help the network learn to acquire all of the data useful for analyses, and only the data useful for analyses.  We are still in the early stages of exploring the potential of adaptive optimization in OPQ networks.

\subsubsection{OPQ Mauka Actors}

The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins. Nine of the most important Actors are described below.

{\em Makai Event Actor.} The Makai Event Actor is responsible for reading data newly created by OPQ Makai into OPQ Mauka. It performs feature extraction on the raw data stream and forwards those features (or the raw data) to subscribing Actor plugins. This allows OPQ Mauka to perform feature extraction once, and allow use of those features by multiple Actors.

{\em Frequency Variation Actor.} The Frequency Variation Actor classifies generic frequency sags, swells, and interruptions as defined by the IEEE 1159 standard. Both duration and deviation from nominal are used to perform these classifications. Duration classifications include frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms. Classifications for deviations from nominal are performed for values that are up to 40\% deviation from nominal. This Actor is able to classify frequency swells, frequency interruptions, and frequency sags, leading to the creation of data at the Incident Layer.

{\em IEEE 1159 Voltage Actor.} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard[29]. In general, this standard classifies voltage disturbances by duration and by magnitude. Voltage durations are classified from 0.5 to 30 cycles, 30 cycles to 3 seconds, 3 seconds to a minute, and greater than 1 minute. Voltage deviations are classified in both the sag and swell directions as a percentage from nominal. Sags are generally classified between 10\% and 90\% of nominal while swells are generally classified from 110\% to 180\% of nominal. This Actor is capable of classifying voltage sags, swells, and interruptions as defined by the standard, and creating data at the Incident Layer if appropriate.

{\em Box Optimization Actor.} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka. This Actor is capable of requesting the state of each OPQ Box (e.g. uptime, Measurement rate, security keys, etc). It is also capable of adjusting the state of individual OPQ Boxes by changing things such as the Measurement and Trend rate or the sampling rate used by the Box.

{\em Future Phenomena Actor.} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena. These Phenomena are used to predict Events and Incidents that may occur in the future. This plugin does not subscribe to any messages, but instead utilizes timers to perform its work. By default, this plugin runs every 10 minutes.

When a Future Phenomena Actor runs, it loads any active Periodic Phenomena found in the database. If Periodic Phenomena are found, this Actor extrapolates possible Detection and Incident Layer data by first examining their timestamps and then extrapolating into the future using the mean period and the standard deviation. For each timestamp in a Periodic Phenomena, the mean period is added. If the resulting timestamp is in the future, a Future Phenomena is created using the time range of the future timestamp plus or minus the standard deviation of the Periodic Phenomena.

When a Future Phenomena is created, timers are started in a separate thread signifying the start and end timestamps of the Future Phenomena. When the first timer runs, messages are sent to the Box Optimization Actor and the Threshold Optimization Actor instructing OPQ Box thresholds to be set lower and measurement rates to be set higher. This increases the chance of seeing an anomaly over the predicted time window. When the second timer runs, these values are reset to their default values. Thus, the plugin increases fidelity and decreases thresholds over the period of a Future Phenomena.

{\em ITIC Actor.} The ITIC Actor analyzes voltage to determine where it falls within the ITIC curve \cite{thallam_power_2000}. The ITIC curve is a power acceptability curve that plots time on the x-axis and voltage on the y-axis.  The purpose of the curve is to provide a tolerance envelope for single-phase 120V equipment. The curve defines three regions. The first region is "No Interruption" and generally includes all voltages with very short sustained durations. All events within this region have no noticeable effect on power equipment. The second region, the "No Damage Region", occurs during voltage sags for extended periods of time. Power Events in this region may cause equipment interruptions, but it will not damage the equipment. The final region, the "Prohibited" region, is caused by sustained voltage swells and may cause damage to power equipment. This Actor determines if an event falls within the "No Damage" or "Prohibited" regions and if so, creates an Incident to record this.

{\em SEMI F47 Actor.} The SEMI F47 Actor is similar to the ITIC Actor in that it plots voltage and duration against a power acceptability curve. In this case, the standard used is the SEMI F47 standard \cite{djokic_sensitivity_2005}. Rather than using a point-in-polygon approach, this plugin reads the voltage features sequentially and uses a state machine to keep track of the current classification. This plugin only classifies values as a "violation" or as "nominal".

{\em Transient Actor.} The Transient Actor is responsible for classifying frequency transients in power waveforms. The plugin subscribes to messages from a topic which contains a calibrated power waveform payload. The Transient Actor is capable of classifying impulsive, arcing, oscillatory, and periodic notching transients. A decision tree is utilized to select the most likely transient type and then further analysis is used to perform the actual classification of transients. Dickens et al \cite{dickens_transient_2019} provides more details on the transient classification system used by this Actor.

{\em Periodicity Actor.} The Periodicity Actor is responsible for detecting periodic signals in power data. This Actor does not subscribe to any messages, but instead runs off of a configurable timer. The Actor is set to run by default once an hour and every hour it scrapes the last 24 hours worth of data and attempts to find periods in the Measurements over that duration.

For each feature in the Measurement and Trend data (e.g. frequency, voltage, and THD), the Periodicity Actor first removes the DC offset from the data by subtracting the mean. Next, the Actor filters the signal using a 4th order high-pass filter to filter out noise. The Actor then performs autocorrelation on the signal followed by finding the peaks of the autocorrelation. The mean distance between the peaks of the autocorrelation provides the period of the signal.

The Periodicity Actor only classifies data as periodic if at least 3 peaks were found and the standard deviation of the period is less than 600 seconds (10 minutes). Once a positive identification has been made, peak detection is performed on the original signal. Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the plugin can group Measurements, Trends, Detection Layer Events, and Incidents that were created during the periodic signals together as part of the Periodic Phenomena.

\subsection{OPQ View}
\label{sec:opq-view}

The final component of the OPQ Sensor Network system architecture is called OPQ View. It is a web application, implemented using the Meteor application framework, which provides a variety of visualization and query services.  An example of the OPQ View home page is provided in Figure \ref{fig:opq-view-home}.

\begin{figure}
\center \includegraphics[width=5in]{images/view/homepage.png}
\caption{OPQ View: home page}
\label{fig:opq-view-home}
\end{figure}

The home page provides three components:  a "System Stats" window which indicates the number of data elements currently at each level of the OPQ Mauka data hierarchy, a "System Health" window that indicates whether the Cloud services appear to be running correctly and the status of communication with all known OPQ Boxes on the sensor network, and a "Box Trends" visualization that provides the ability to quickly see trends in the four basic measures (Voltage, Frequency, THD, and Transients) over time.

Figure \ref{fig:opq-view-box-map} shows a map-based view of the sensor network. This figure shows the location of 11 OPQ Boxes on the University of Hawaii campus at one point during Fall of 2019. Depending on the zoom level of the interface, some boxes are collapsed into disks with a number indicating the number of boxes at that location. Zooming in reveals more information about the box including near-real time values for voltage and frequency, as is shown for the box in the upper right corner of the figure.

\begin{figure}
\center \includegraphics[width=5in]{images/view/boxmap-2.png}
\caption{OPQ View: Box Map}
\label{fig:opq-view-box-map}
\end{figure}

Figure \ref{fig:opq-view-incident-summary} shows a visual display of a single Incident. This view includes a map-based location of the box whose data was involved in the Incident, the start and end time and duration of the Incident, the classification(s) of the anomalous data, and the waveform associated with the Incident (when applicable). If additional analysis is desired, the raw data can be downloaded in CSV format.

\begin{figure}
\center \includegraphics[width=5in]{images/view/incident-summary.png}
\caption{OPQ View: Incident Summary}
\label{fig:opq-view-incident-summary}
\end{figure}

The above images provide a sense for how OPQ View helps users to understand, monitor, and assess an OPQ Sensor Network and the underlying power quality of the grid it is attached to. There are several features in OPQ View for user and box management; for details, please see the OPQ View documentation.

While OPQ View provides a variety of useful visualization and analysis features, users wishing to understand the power quality of their grid are not restricted to its current feature set. The OPQ database schema is public, and if a new analysis is desired, it is straightforward to query the database directly for the data of interest and build a new analysis off of it.

### 2020-07-17 04-related.tex
checkpoint

@@
 
-Section \ref{sec:commercial-pq-devices} discusses power quality monitoring devices and how they compare to OPQ Box.
-Section \ref{sec:commercial-pq-software} discusses power quality database software and how it compares to OPQ Cloud.
-The final subsection discusses recent research systems comprising both hardware and software and how they compare to OPQ.
-
 \subsection{Power quality hardware}
@@
 
-Fluke \cite{fluke_fluke_2020} provides a variety of Power Quality tools.
-These tools typically measure voltage, current, frequency, dips, swells, power factor, harmonic, current unbalance, inrush current values, and light flicker.
-They generally adhere to power quality standards such as IEEE 1159, and use current transformers rather than plugging into a wall outlet.
-Depending on the model, these devices cost from approximately \$2,000 to \$10,000, and are designed for use by power engineers to diagnose power quality problems resulting from industrial machinery.
-
-Dranetz \cite{dranetz_dranetz_2020} provides the HDPQ family of power quality analyzers.
-These are similar to Fluke devices in that they require connection to mains using current transformers and are intended for use by power quality engineers.
-They cost from \$5,000 to \$17,000 depending upon the model.
-They also sell a set of software answer modules to provide additional analyses regarding the raw power quality data collected by the device.
-For example, the Sag Directivity module detects a voltage sag, identifies its characteristics, and determines whether the source is upstream or downstream of the monitoring point.
-Answer modules cost approximately \$1,000 each.
-
-Elspec \cite{elspec_elspec_2020} provides the BlackBox family of multi functional waveform recorders.
-These devices perform transient recording, disturbance recording, phasor measurements (synchrophasor), power quality analysis, and sequence of events recording.
-They cost from \$6,000 to \$12,000 depending upon the model.
-
-PowerSide (formerly Power Standards Lab) \cite{powerside_powerside_2020} provides the PQube3 family of power monitors.
-Unlike the above devices, PQube is both a revenue-grade energy meter as well as a power quality monitor.
-It can also record environmental data such as temperature, humidity, barometric pressure, and certain process parameters such as fuel level and water flow.
-QubeView is a Windows-based software package that provides real-time monitoring of multiple PQube devices and trends and statistics over the data collected.
-
-ACR Systems \cite{acr_acr_2020} manufactures the PowerWatch Voltage Disturbance Recorder.
-Unlike the above devices, this device connects to a wall outlet.
-It measures surges and sags and is meant for monitoring medical instrumentation, point-of-sale terminals, computer servers, and other sensitive electronic equipment.
-It costs approximately \$625.
-A Windows-based program is available for a separate cost of \$135, and can be used to extract data from the device using a USB cable, and then display graphs of power quality.
-
-OpenZMeter \cite{viciana_openzmeter_2018} is a low-cost, open source device that connects to household mains and collects both power quality and energy data.
-
-\subsection*{Comparison with OPQ Box}
+There exists a very wide variety of power quality hardware devices, including those made by Fluke \cite{fluke_fluke_2020}, Dranetz \cite{dranetz_dranetz_2020}, Elspec \cite{elspec_elspec_2020}, PowerSide (formerly Power Standards Lab) \cite{powerside_powerside_2020}, ACR Systems \cite{acr_acr_2020}, and OpenZMeter \cite{viciana_openzmeter_2018}.
 
@@
 
-PQView \cite{electrotek_concepts_pqview_2020} is a database system developed by Electrotek and EPRI to integrate data from a wide variety of digital relays, fault recorders, power quality monitors, smart meters, and SCADA historians into a relational database.
-PQView is not designed for real-time display, instead, the normal mode of operation is to poll the various devices containing power data once per day and upload the results, although polling as fast as once per minute is supported.
-The software can generate reports regarding voltage sags, swells, and interruptions, THD, and overlay voltage sags/swells on CBEMA curves to indicate which ones exceed tolerances.
-PQView is built on Windows software and the data is stored in a SQL Server database.
-PQView costs \$5000 and is available as an EPRI Product.
-
-PQSCADA Sapphire \cite{elspec_ltd_pqscada_2016} is another vendor-independent system for collecting and displaying data from a variety of power quality monitoring devices.
-It comes in three versions, a free "Express" version with limited import and analysis tools.
-To obtain additional capabilities, user can purchase the professional version for \$2500 or the Enterprise version for \$6400.
-The data visualizations include trend charts, scatter event charts, histograms with summary statistics, and phasor charts.
-
-PQDIF \cite{sabin_ieee_2020}, the "Power Quality Data Interchange Format", is a binary file format that specifies voltage, current, power, and energy measurements in standard fashion that allows this data to produced and consumed by devices in a vendor-independent fashion.
-
-The Grid Protection Alliance \cite{grid_protection_alliance_grid_2020} produces a suite of open source software products for addressing various issues in power data management and analysis, including data collection, archiving, notifications, event analysis, and visualization.
-
-\subsection*{Comparison with OPQ Cloud}
+PQView \cite{electrotek_concepts_pqview_2020},  PQSCADA Sapphire \cite{elspec_ltd_pqscada_2016}, PQDIF \cite{sabin_ieee_2020}, and Grid Protection Alliance \cite{grid_protection_alliance_grid_2020}  are examples of software and/or software standards for manipulating power quality data.
 
@@
 For details on the representation, see the OPQ Data Model, and for details on communication, see the OPQ Protobuf protocol.
-
-Note that it would be straightforward to develop an export function in OPQ Cloud to provide power quality data in PQDif format, but that functionality is not currently available.
 

### 2020-07-17 09-pilot.tex
checkpoint

@@
-\section{Results from the Pilot Deployment}
+\section{Results and Discussion}
 \label{sec:pilot-study}
 
-To evaluate the capabilities of the OPQ Sensor Network, 16 OPQ Boxes were deployed at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019.
+To evaluate the capabilities of the OPQ Sensor Network, we deployed 16 OPQ Boxes at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019.
 The University of Hawaii campus is an isolated microgrid connected to the Oahu powergrid only via a single 46kV feeder.
@@
 Data from these meters were used as ground truth for validation studies of the OPQ sensor network.
-
-The University of Hawaii power grid is interesting in that it supplies a highly diverse infrastructure.
-Beyond traditional residential equipment such as computers and consumer grade electronics, the UH power grid powers scientific and laboratory equipment, machine shops, and server farms.
-All of these elements have varying requirements/tolerances for power quality anomalies as well as different levels of power quality “pollution”.
-Furthermore, some of the electricity consumers in the UH campus are entirely unique.
-For example, the free electron laser located in the Watanabe Hall is one of only ten in the world, and the impact/sensitivity of power quality on the instrument are completely unstudied.
-
-The pilot study had the following major goals:
-
-\begin{enumerate}
-
-\item {\em Evaluation of sensor data accuracy and quality.} While we conducted extensive laboratory tests of the sensor network, the pilot study enabled us to evaluate the accuracy and utility of our hardware devices and power quality data collection procedures in a real-world setting.
-Does an OPQ sensor network measure power quality (and detect anomalies) as well (or better) than conventional building-level electrical meters?
-
-\item {\em Evaluation of the triggering system.} The OPQ Makai triggering system provides a novel two-way communication between sensors and the cloud, resulting in a triggering system with unique capabilities.
-Does the triggering system provide efficiencies not available from more conventional approaches?
-Does it provide data enabling analysis opportunities not available through more conventional approaches?
-
-\item {\em Evaluation of the information architecture.} OPQ Mauka implements a tiered, hierarchical information architecture.
-The pilot study enabled us to evaluate this information architecture.
-Would this information architecture prove useful with real world data?
-Could actionable insights emerge from low-level sensor data?
-
-\item {\em Evaluation of cloud data storage management.} OPQ Mauka manages data storage through a "time-to-live" (TTL) mechanism, which results in data being discarded if it is not found to be useful and thus ``promoted" to a higher level of the information hierarchy within a defined period of time.
-The pilot study enables us to evaluate this approach.
-Does this mechanism allow sensor network designers to determine upper bounds on storage requirements?
-Is information thrown away prematurely because it was not found to be useful within the time limits?
-
-\end{enumerate}
 
@@
 \subsection{OPQ provides valid and reliable collection of power quality data}
+\label{hyp:01}
 
@@
 \subsection{OPQs triggering system provides advantages with respect to bandwidth and computation}
+\label{hyp:02}
 
@@
 \subsection{The OPQ Information Architecture provides a means to produce actionable insights}
+\label{hyp:04}
 
@@
 \subsection{The OPQ Information Architecture provides predictable upper bounds on storage resources}
+\label{hyp:05}
 

### 2020-07-17 cover-letter/reviewer-2.tex
checkpoint

@@
 
+Thank you for your comments.
+
+The article has been shortened from 32 pages to XX pages.
+
 {\em (2) I suggest a typical layout used in the articles: Introduction, Methods, Results and Discussion, Conclusions. This template is also recommended by Journal (see Instructions for Authors).}
+
+The structure has been modified to provide a more traditional scientific format: Introduction, Related Work, Methods, Results and Discussion, Conclusions.
 
@@
 
+The introduction now highlights the aims of the work and the research gap.
+
+Conclusion section has been improved following your suggestions.
+
 {\em (4) The quality of Figures should be corrected, their descriptions, especially axis descriptions are unreadable.}
 
-Fixed.
-The issue is incorporating modern web application development as the target domain.
-The tech stack associated with modern web application development is complex enough that the semester can easily devolve into ``learning the tech stack", not ``experience software engineering".
-Historically, the target domains for project, studio, and flipped pedagogies required only a simple tech stack and so this was not an issue.
+Some figures have been deleted as part of the shortening process. Those that remain have been checked for readability.
 

### 2020-07-17 graveyard.tex
checkpoint

An OPQ sensor network is designed to be useful to: consumers who want to better understand the performance of their public utilities; utilities who desire a low-cost, easily deployable and re-deployable sensor network to gather information about grid stability that complements their existing infrastructure; to researchers who wish to design improvements to grid management and need a low cost mechanism for measuring the impact of their improvements; and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.  Several key features of our project emerge from these goals.

{\em OPQ Box.} First, our custom power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75. This is from 10 to 100 times less expensive than commercial power quality monitors. This cost differential means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building. Similarly, instead of monitoring power quality at the substation level of the grid, OPQ Boxes make it feasible to deploy dozens or hundreds into a community to obtain fine-grained data about the impact of solar or other renewables as directly perceived by the customer.

{\em Cloud-native information architecture.} Second, our information architecture is "cloud native", which means that OPQ Boxes are not designed for stand-alone, autonomous use, unlike current commercially available power quality monitors. Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and with what fidelity in order to most efficiently produce actionable information. Two-way communication between the OPQ Boxes and cloud services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network. As a result, OPQ cloud services can obtain and analyze high fidelity waveform data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud. This enables OPQ installations to be simultaneously responsive and scalable.

{\em Plug-in analysis architecture.} Third, all OPQ software services are designed for extensibility and interoperation. Our middleware components (Mauka and Makai) provide a plugin architecture. Our visualization component (View) is built from modular UI elements using React.  This means that OPQ is not a monolithic, closed system with a frozen feature set, but rather a design environment for experimentation with advanced classification and analysis of power quality data.

{\em Open Source.} Fourth, our software and hardware designs and implementations are made available using open source licenses. This means that organizations choosing OPQ do not face the business risk of committing resources to a single vendor with proprietary hardware and software. Our goal is to facilitate the creation of a community of researchers and industry practitioners to replicate, extend, and apply the insights gained from the OPQ system.

While we believe OPQ offers a compelling combination of capabilities, we want to be clear that all sensor network designs involve trade-offs. As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs. In some cases, OPQ may not be the appropriate choice. In other cases, a hybrid solution consisting of OPQ in combination with other technologies for grid monitoring and analysis may be most appropriate.

\subsection{Structure of this paper}

The remainder of this paper is organized as follows. Section \ref{sec:architecture} presents an overview of the OPQ system architecture and information architecture. Section \ref{sec:related-work} presents related research and technology.  Section \ref{sec:opq-box} discusses OPQ Box and Section \ref{sec:opq-makai} its triggering system. Section \ref{sec:opq-mauka} describes our analysis services, and Section \ref{sec:opq-view} our user interface.  Section \ref{sec:pilot-study} discusses the pilot deployment of a sensor network and its results. Section \ref{sec:conclusions} concludes with some proposals for future research.

### 2020-07-18 03-methods.tex
revision complete.

--- 2020-07-17 02-methods.tex
@@
 
-\subsubsection{Hardware}\label{subsec:hardware}
+\subsubsection{OPQ Box Hardware}\label{subsec:hardware}
 
@@
 
-\subsubsection{Software}\label{subsec:software}
+\subsubsection{OPQ Box Software}\label{subsec:software}
 
@@
 
-\subsubsection{Fundamental Frequency}\label{subsec:fundamental-frequency}
+\subsubsection{OPQ Box: Calculating Fundamental Frequency}\label{subsec:fundamental-frequency}
 
@@
 
-\subsubsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}
+\subsubsection{OPQ Box: Calculating RMS Voltage}\label{subsec:root-mean-square-voltage}
 
@@
 
-\subsubsection{Total Harmonic Distortion}\label{subsec:thd}
+\subsubsection{OPQ Box: Calculating THD}\label{subsec:thd}
 
@@
 
-\subsubsection{Transient Detection}\label{subsec:transient-detection}
+\subsubsection{OPQ Box: Transient Detection}\label{subsec:transient-detection}
 
@@
 
-\subsubsection{Network Communication}\label{subsec:network-communication}
+\subsubsection{OPQ Box: Network Communication}\label{subsec:network-communication}
 
@@
 
-\subsubsection{Manufacturing}
+\subsubsection{OPQ Box: Manufacturing}
 
@@
 
-\subsubsection{Design}
+\subsubsection{OPQ Makai: Design}
 
@@
 
-\subsubsection{OPQ Mauka Actors}
+\subsubsection{OPQ Mauka: Actors}
 

### 2020-07-18 05-pilot.tex
revision complete.

--- 2020-07-17 09-pilot.tex
@@
 In a system without TTL, raw data storage rapidly grows.
-The bounds on raw storage can be calculated by taking into account the sampling rate and sample size for active OPQ Boxes as described in Equation ~\ref{eq:RawDataBytes}.
-
-\begin{equation}
-    RawDataBytes = NumBoxes * TimeSeconds * SampleRate * SampleSize
-    \label{eq:RawDataBytes}
-\end{equation}
+The bounds on raw storage can be calculated by taking into account the sampling rate and sample size for active OPQ Boxes.
 
 Given that OPQ Boxes were configured to sample at 12 kHz at 2 bytes per sample during the deployment, we can calculate the bounds on raw storage for 16 OPQ Boxes over 3 months without TTL as 3.1 TiB.
-Figure~\ref{fig:data-management-graph} shows slightly less data as the number of OPQ Boxes was not constant during the pilot study (Figure~\ref{fig:active-opq-boxes}).
-By implementing TTL, significant data savings are made by discarding of IML data that is mostly noise.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.75\linewidth]{images/pilot/active-opq-boxes.png}
-    \caption{Active OPQ Boxes vs. Time}
-    \label{fig:active-opq-boxes}
-\end{figure}
 
 Given a network of 16 OPQ Boxes over the pilot deployment period with TTL, the combined data storage from all levels reached an asymptotic limit of near 100 GB.
-The layout of data storage is provided per data storage level in Figure~\ref{fig:opq-data-storage}.
 The high rate of detections made by OPQ utilize most of the space.
@@
 This is a result of TTL discarding not useful data.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=0.75\linewidth]{images/pilot/laha-levels-storage.png}
-    \caption{Combined OPQ Data Storage}
-    \label{fig:opq-data-storage}
-\end{figure}
 
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{images/pilot/periodic-voltage-sags.png}
+    \includegraphics[width=0.90\linewidth]{images/pilot/periodic-voltage-sags.png}
     \caption{Detection of Periodic Voltage Sags}
@@
 
-As an example, Figure~\ref{fig:future-phenomena} shows the percentage of correctly predicted future phenomena as a function of time.
-The self-optimizing nature of the system allows the system to make more accurate predictions due to acting on more accurate periodic phenomena.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.60\linewidth]{images/pilot/realized-future-phenomena.png}
-    \caption{Future Phenomena Prediction Accuracy Over Time}
-    \label{fig:future-phenomena}
-\end{figure}
-
 Second, the increase in detection sensitivity makes it more likely that sub-threshold signals of interest will be classified.
-As example, the increased detection sensitivity allowed us to observe and classify a periodic voltage sag that would not have been detected by our normal classifier due to the small magnitude of the sag.
-The sub-threshold detection is shown in Figure~\ref{fig:sub-threshold-detection}.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.85\linewidth]{images/pilot/subthreshold-signal.png}
-    \caption{Sub-threshold Detection}
-    \label{fig:sub-threshold-detection}
-\end{figure}
+For example, the increased detection sensitivity allowed us to observe and classify a periodic voltage sag that would not have been detected by our normal classifier due to the small magnitude of the sag.
 

### 2020-07-18 06-conclusions.tex
revision complete.

--- 2020-06-24 10-conclusions.tex
@@
 
-This research project has produced both hardware and software for power quality monitoring with a variety of innovations.
+This project has produced both hardware and software for power quality monitoring with a variety of innovations.
 Our OPQ Boxes collect frequency, voltage, THD, and transients with high fidelity, and at a cost that is generally 10x to 100x cheaper than current commercial offerings (though these commercial offerings offer a variety of features not available from OPQ Boxes, so the appropriate choice depends upon the needs of the user).
@@
 The OPQ Information Architecture enables Actors at the Phenomena Layer to control sensitivity settings of individual OPQ Boxes in order to improve data collection when power quality anomalies have been predicted.
-`
+
 Our sensor network is designed to efficiently use network and storage resources.

### 2020-07-18 cover-letter/reviewer-2.tex
revision complete.

@@
 
-The article has been shortened from 32 pages to XX pages.
+We shortened the article from 32 pages to 24 pages.
 
@@
 
-The structure has been modified to provide a more traditional scientific format: Introduction, Related Work, Methods, Results and Discussion, Conclusions.
+We modified the structure to provide a more traditional scientific format: Introduction, Methods, Related Work, Results and Discussion, Conclusions and Future Directions.
 
@@
 
-The introduction now highlights the aims of the work and the research gap.
+The introduction now highlights the aims of the work and the research gap (hypotheses to be tested).
 
-Conclusion section has been improved following your suggestions.
+Conclusion section has been improved.
 

### 2020-07-18 cover-letter/reviewer-3.tex
revision complete.

@@
 
-Thanks for the kind words. As part of the paper reorganization and shortening, Figure 6 has been removed.
+Thank you for your comments.
+
+As part of the paper reorganization and shortening, Figure 6 has been removed.
 
@@
 
-Thanks again!  The rewrite attempts to inject more scientific background.
+We modified the structure to provide a more traditional scientific format: Introduction, Methods, Related Work, Results and Discussion, Conclusions and Future Directions.
+We shortened the article from 32 pages to 24 pages.
+We believe this provides more focus on scientific aspects and results.
 

### 2020-07-18 graveyard/06-opq-makai.tex
revision complete.

--- 2020-07-17 02-methods.tex
@@
-\section{Methods}
+\section{OPQ Makai}
+\label{sec:opq-makai}
 
-The central aim of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
-
-Our research methodology consists of three (potentially interleaved) phases.
-First, we specify, design, and implement novel hardware and software for collecting, transmitting, and analyzing four essential measures of power quality data: voltage, frequency, THD, and transients.
-Second, we verify that our implementation matches our specifications through laboratory experiments.
-Third, we validate our implementation by testing a variety of hypotheses through a real-world deployment.
-In this paper, we report on results from a pilot deployment at the University of Hawaii over the course of three months.
-
-Our central hypothesis is that we can accomplish this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based services.
-
-We will test this central hypothesis through six subhypotheses: (1) OPQ provides valid and reliable collection of power quality data (Section \ref{hyp:01}), (2) OPQ's triggering system provides advantages with respect to bandwidth and computation (Section \ref{hyp:02}, (3) OPQ enables subthreshold event detection based upon temporal locality (Section \ref{sec:subthreshold-events}), (4) The OPQ information architecture provides a means to produce actionable insights (Section \ref{hyp:04}), (5) the OPQ information architecture provides predictable upper bounds on storage resources (Section \ref{hyp:05}), and (6) OPQ provides useful adaptive optimization capabilites (Section \ref{sec:adaptive-optimization}).
-
-The remainder of this section presents the overall architecture of OPQ, along with details on the design and implementation of each of the principal architectural components.
-We will also discuss our verification activities for each component.
-Our validation and hypothesis testing activities will be described in Section \ref{sec:pilot-study}.
-
-\subsection{OPQ Architecture}
-
-The OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events.
-OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components either at predefined intervals or upon request.
-OPQ Makai monitors incoming low fidelity data from OPQ Boxes, requests high fidelity data when necessary, and stores the results in a MongoDB database.
-OPQ Mauka analyzes low level data and analyzes it to produce higher-level representations according to the information architecture described below, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis.
-OPQ View is a browser-based visualization platform for displaying the results for data capture and analysis.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/architecture/system-diagram.png}
-\caption{High level system architecture of an OPQ Sensor Network}
-\label{fig:architecture}
-\end{figure}
-
-Figure \ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side).
-First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQ Makai.
-OPQ Makai analyzes low fidelity measurements, and requests high fidelity waveforms when desirable.
-Both measurements and waveforms are saved in a MongoDB database.
-OPQ Mauka analyzes low and high fidelity data, and creates "events" to represent anomalies.
-OPQ View notifies users of events and allows them to drill down into low and high fidelity data.
-
-OPQ Makai, OPQ Mauka, and OPQ View are all cloud-based software services that collectively form a single "instance" with respect to data transmission, storage, analysis, and visualization.
-We refer to this collection of software-side components as OPQ Cloud.
-Every OPQ Box connects to a single instance of an OPQ Cloud.
-It is possible to have multiple OPQ Cloud instances.
-For example, a company might install an OPQ Cloud instance behind their firewall along with OPQ Boxes to provide a private mechanism for collecting and analyzing power quality data.
-
-\subsection{OPQ Box}
-\label{sec:opq-box}
-
-OPQ Box is a hardware device designed to provide inexpensive, extensible and accurate residential power quality measurements.
-A block diagram of OPQ Box is shown in Figure~\ref{fig:opq:1:1}.
-A complete device is shown in Figure~\ref{fig:opq:1:2}.
-
-\begin{figure}[ht]
-	\centering
-	\begin{subfigure}{.5\textwidth}
-	  \centering
-	  \includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
-	  \caption{OPQ Box Block Diagram.
-	  The power path is in red, signal path is in green and the digital IO is in black.}
-	  \label{fig:opq:1:1}
-	\end{subfigure}%
-	\begin{subfigure}{.5\textwidth}
-	  \centering
-	  \includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
-	  \caption{OPQ Box in an ABS plastic enclosure.}
-	  \label{fig:opq:1:2}
-	\end{subfigure}
-	\caption{(a) OPQ Box block diagram and (b) production OPQ Box ready for deployment}
-	\label{fig:opq:2}
-\end{figure}
-
-\subsubsection{Hardware}\label{subsec:hardware}
-
-The power system of the OPQ Box electrically isolates most of the device from the AC mains power.
-An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
-5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference, and an isolated DC/DC converter.
-3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
-The hot side of the isolation amplifier is powered from the isolated DC/DC converter.
-This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power outage.
-
-\begin{figure}[ht]
-  \begin{center}
-  \includegraphics[width=0.9\textwidth]{images/opq-box/opqbox_software.png}
-  \end{center}
-  \caption{Block diagram of the OPQ Box 2 software stack.}
-  \label{fig:opq:3}
-\end{figure}
-
-A Raspberry Pi single board computer (SBC) is responsible for signal analysis and anomaly detection.
-The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256MB of main memory and a single core 1GHz ARM11 CPU.
-It also contains an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle.
-
-\subsubsection{Software}\label{subsec:software}
-
-The software stack of the Raspberry Pi aims to deliver an accurate and precise power quality analysis framework despite the rather limited capabilities of the hardware.
-A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
-Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with the Pi acting as the master and the DSP as a slave device.
-A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer, and a kernel driver provides management of the SPI bus.
-Internally, the OPQ driver maintains a ring buffer of 16 windows, each of which is 200 data samples in size.
-Upon receiving the interrupt for the DSP, the CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
-This scheme requires the CPU to only service 60 interrupts a second, with each interrupt requiring on the order of 100 instructions, for a CPU utilization of less than $1\%$ in normal operation.
-Userland applications communicate with the kernel driver using a file descriptor, where every $read$ system call yields 200 samples of raw waveform.
-As a result, the smallest window that a userland application may process is a single AC cycle of the grid mains.
-
-The userland component of the OPQ Box software is a multi-threaded extensible analysis framework called Triggering.
-The reader thread is responsible for transferring and accumulating data from the kernel driver.
-The smallest data buffer that the Triggering application processes at any given time is 10 grid cycles or 2k samples.
-Once the cycles are transferred to the userland and timestamped, they are passed to the analysis thread for feature extraction, as well as to the Raw Data Ring Buffer (RDRB).
-Since internally all data is addressed using shared pointers, during data duplication no copying is required.
-RDRS is capable of buffering up to an hour of data before it is overwritten, resulting in the RDBS maximum size of 100MB.
-
-The analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
-Four metrics are extracted from the data stream: (1) Fundamental frequency, (2) RMS Voltage, (3) Total Harmonic Distortion, and (4) Transients. Let's briefly discuss how each of these are computed.
-
-\subsubsection{Fundamental Frequency}\label{subsec:fundamental-frequency}
-
-The fundamental frequency is calculated by computing the zero crossings of the AC waveform.
-In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
-Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz), it is very computationally expensive to perform this filtering in a single step.
-Instead, filtering is accomplished via a set of two low pass finite impulse response (FIR) filters.
-First filter has a passband of 0-600Hz allowing us to downsample the waveform to 1200 samples per second.
-Next filter has a passband of 0-100Hz allowing for further removal of high frequency noise.
-Finally, zero crossings are extracted and used for the frequency calculation.
-The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.
-
-\subsubsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}
-
-Root mean square voltage ($V_{rms}$) in electrical power is the equivalent value of DC voltage which would dissipate the same power in the resistive load.
-$V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed value.
-
-Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation. Unlike the frequency calculation, the input is not filtered a priori.
-
-\subsubsection{Total Harmonic Distortion}\label{subsec:thd}
-
-The OPQ Box calculates total harmonic distortion (THD) using the industry standard methodology.
-It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
-Operationally, OPQ Box computes THD for 10 cycles of the fundamental frequency.
-First an FFT transforms the real voltage samples into its frequency components.
-Next, the square of the harmonic bins are accumulated and scaled by the magnitude of the fundamental power.
-
-\subsubsection{Transient Detection}\label{subsec:transient-detection}
-
-OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter with a cutoff frequency of $400Hz$ and searching for a maximum value in the remainder.
-
-It should be noted that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
-However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient detection.
-
-\subsubsection{Network Communication}\label{subsec:network-communication}
-
-The computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
-Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
-Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box to verify its peer.
-Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
-This protocol is a publish subscribe topology, with each message containing the topic and a payload.
-Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
-This allows for the aggregation service to be spread across multiple nodes, with minimal network overhead.
-
-If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box RDRB via a separate ZeroMq pub sub channel.
-If the RDRB buffer contains data for the requested temporal range, OPQ Box transmits the available data to the aggregation service via a push pull ZeroMq channel.
-Protobuf message serialization is used to encode messages across the OPQ ecosystem.
-
-In order to make a distributed measurement, all of the OPQ Boxes on the OPQ network need to maintain an accurate time reference.
-Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
-The expansion port of the OPQ Box supports a GPS receiver.
-However, since GPS receivers require line of sight to the sky, it was not used for deployment.
-NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a nearby NTP server.
-
-\subsubsection{Manufacturing}
-
-Currently, there is no mechanism for mass production of OPQ Boxes, but all of the plans are available under an Open Source license, so interested organizations with some basic hardware engineering skills can build their own boxes.
-
-Complete specifications for the OPQ Box hardware, firmware, and software are available \cite{negrashov_opq_2020}.
-As of the time of writing, a single OPQ Box can be manufactured for approximately \$100 in parts.
-The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately \$75 in parts per device.
-
-\subsection{OPQ Makai}
-\label{sec:opq-makai}
+\subsection{Overview}
 
@@
 
-\subsubsection{Design}
+\subsection{Design}
 
@@
 Once the Event service receives a data response with an identity containing an event token it hasn’t seen before, it will increment the event number, and store it in an internal key value store.
-
-\subsection{OPQ Mauka}
-\label{sec:opq-mauka}
-
-The previous sections discussed the design of OPQ Box, a custom hardware device for collecting four important measures of power quality, and OPQ Makai, a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes.
-As a result of these two innovations, an OPQ sensor network has the ability to collect and analyze high fidelity, low level data about power quality anomalies in a cost-effective, scalable fashion.
-
-There are remaining challenges to creating a useful power quality sensor network.
-First, the data provided by OPQ Boxes is low-level, "primitive" data consisting of either features (i.e.
-frequency, voltage, THD, and transients) or waveform data.
-But what we actually want is actionable insights into grid stability.
-For example, we might want to know if a given anomalous data value is actually detrimental, or we might want to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.
-
-A second challenge involves the potentially high volume of data might accumulate in the cloud.
-Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data could accumulate.
-One strategy is to simply store all of the data sent to the cloud forever.
-This means that data storage requirements will increase monotonically over time, making the sensor network more costly to maintain the longer it is in place.
-An alternative strategy is to implement an algorithm to identify uninteresting (or no longer interesting) data and discard it.
-Ideally, such an algorithm would enable OPQ sensor network designers to calculate an upper bound on the total amount of cloud storage required as a function of the number of nodes (OPQ Boxes) in the network.
-
-OPQ Mauka addresses both of these issues.
-First, OPQ Mauka provides a multi-layered representation for structuring and processing DSN data.
-The structure and processing at each layer is designed with the explicit goal of turning low-level data into actionable insights.
-Second, each layer in the framework implements a "time-to-live" (TTL) strategy for data within the level.
-This strategy states that data must either progress upwards through the layers towards more abstract, useful representations within a fixed time window, or else it can be discarded.
-The TTL strategy is useful because when implemented, it allows DSN designers to make reasonable predictions of the upper bounds on data storage at each level of the framework adjusting for the number of sensors and power anomaly probability.
-
- TTL also makes possible a ``graceful degradation" of system performance if those bounds turn out to be exceeded.
-For example, consider a situation in which a power network enters a prolonged period of widespread power quality instability, where every OPQ Box is reporting continuous anomalous conditions with respect to voltage, frequency, THD, and transients.
-This ``worst case scenario" would lead to the potential situation of every OPQ Box trying to upload raw waveform data all the time.
-The TTL system provides safeguards, in that whatever low-level data has not been processed relatively quickly can be discarded.
-Thus, instead of the system potentially going down entirely, it could instead continue to operate at a reduced capacity.
-
-Figure \ref{fig:mauka-data-model} illustrates the hierarchical data model for OPQ Mauka.
-This data model can be conceptualized as a multi-level hierarchy that adaptively optimizes data storage using a tiered TTL approach and provides a mechanism in which typed aggregated data is continually refined to the point of being of becoming actionable.
-The data model also includes software components called "actors" that both move data upward through the levels and also apply optimizations downward through the levels.
-Actors are implemented through a plugin architecture, making it easy to experiment with the data model and improve it over time.
-
-\begin{figure}
-\center \includegraphics[width=4in]{images/mauka/mauka-data-model.png}
-\caption{Mauka data model hierarchy}
-\label{fig:mauka-data-model}
-\end{figure}
-
-The lowest layer of the hierarchy is the Instantaneous Measurements Layer (IML).
-The IML contains "raw" data, in other words, the digitized waveform.
-IML data exists both on each OPQ Box (where it is available for up to the previous 60 minutes).
-It also exists in the cloud, in the event that OPQ's triggering mechanism has caused a temporal interval of waveform data to be uploaded.
-IML data in the cloud has a TTL of 15 minutes: unless the waveform data is found to be useful by a cloud service within 15 minutes, it can be discarded.
-
-The second layer is the Aggregate Measurements Level (AML).
-The AML stores power quality summary statistics sent either once per second or once per minute by each OPQ Box.
-These summary statistics include the maximum, minimum, and average values of voltage, frequency, THD, and transient metrics over the time window.
-It is AML data that is used to initiate the triggering process of uploading IML data from the cloud.
-
-The third layer is the Detections Level (DL).
-This layer is responsible for processing the IML and AML data to produce a representation of an "event" with a precisely defined start and end time based upon examination of the waveform.
-As will be discussed in Section \ref{sec:subthreshold-events}, knowledge of the start and end time of a power quality anomaly allows investigation of how that anomaly might manifest itself elsewhere in the grid, even if this manifestation is not severe enough to produce over-threshold data values.
-
-The fourth layer is the Incident Level (IL).
-This layer starts to answer the question of whether the data is "actionable" by classifying the detected event according to various industry standards for power quality anomalies: IEEE 1159, ITIC, SEMI F47, and so forth.
-For example, an event that falls into the ITIC "prohibited" region clearly indicates a power quality anomaly that requires further study and intervention.
-
-The fifth and final level is the Phenomena Level (PL).
-This layer contains the results of analyses that attempt to identify cyclic, and thus predictable, power quality disturbances.
-It also contains analysis results regarding the similarity of various incidents, which can help uncover causal factors.
-Finally, it provides analyses for adaptive optimization of the OPQ Sensor Network.
-These optimizations can change the thresholds for individual boxes to either increase their sensitivity or decrease their sensitivity over specific intervals of time.
-The ultimate goal of adaptive optimization is to help the network learn to acquire all of the data useful for analyses, and only the data useful for analyses.
-We are still in the early stages of exploring the potential of adaptive optimization in OPQ networks.
-
-\subsubsection{OPQ Mauka Actors}
-
-The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins. Nine of the most important Actors are described below.
-
-{\em Makai Event Actor.} The Makai Event Actor is responsible for reading data newly created by OPQ Makai into OPQ Mauka.
-It performs feature extraction on the raw data stream and forwards those features (or the raw data) to subscribing Actor plugins.
-This allows OPQ Mauka to perform feature extraction once, and allow use of those features by multiple Actors.
-
-{\em Frequency Variation Actor.} The Frequency Variation Actor classifies generic frequency sags, swells, and interruptions as defined by the IEEE 1159 standard.
-Both duration and deviation from nominal are used to perform these classifications.
-Duration classifications include frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
-Classifications for deviations from nominal are performed for values that are up to 40\% deviation from nominal.
-This Actor is able to classify frequency swells, frequency interruptions, and frequency sags, leading to the creation of data at the Incident Layer.
-
-{\em IEEE 1159 Voltage Actor.} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard[29].
-In general, this standard classifies voltage disturbances by duration and by magnitude.
-Voltage durations are classified from 0.5 to 30 cycles, 30 cycles to 3 seconds, 3 seconds to a minute, and greater than 1 minute.
-Voltage deviations are classified in both the sag and swell directions as a percentage from nominal.
-Sags are generally classified between 10\% and 90\% of nominal while swells are generally classified from 110\% to 180\% of nominal.
-This Actor is capable of classifying voltage sags, swells, and interruptions as defined by the standard, and creating data at the Incident Layer if appropriate.
-
-{\em Box Optimization Actor.} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
-This Actor is capable of requesting the state of each OPQ Box (e.g.
-uptime, Measurement rate, security keys, etc).
-It is also capable of adjusting the state of individual OPQ Boxes by changing things such as the Measurement and Trend rate or the sampling rate used by the Box.
-
-{\em Future Phenomena Actor.} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena.
-These Phenomena are used to predict Events and Incidents that may occur in the future.
-This plugin does not subscribe to any messages, but instead utilizes timers to perform its work.
-By default, this plugin runs every 10 minutes.
-
-When a Future Phenomena Actor runs, it loads any active Periodic Phenomena found in the database.
-If Periodic Phenomena are found, this Actor extrapolates possible Detection and Incident Layer data by first examining their timestamps and then extrapolating into the future using the mean period and the standard deviation.
-For each timestamp in a Periodic Phenomena, the mean period is added.
-If the resulting timestamp is in the future, a Future Phenomena is created using the time range of the future timestamp plus or minus the standard deviation of the Periodic Phenomena.
-
-When a Future Phenomena is created, timers are started in a separate thread signifying the start and end timestamps of the Future Phenomena.
-When the first timer runs, messages are sent to the Box Optimization Actor and the Threshold Optimization Actor instructing OPQ Box thresholds to be set lower and measurement rates to be set higher.
-This increases the chance of seeing an anomaly over the predicted time window.
-When the second timer runs, these values are reset to their default values.
-Thus, the plugin increases fidelity and decreases thresholds over the period of a Future Phenomena.
-
-{\em ITIC Actor.} The ITIC Actor analyzes voltage to determine where it falls within the ITIC curve \cite{thallam_power_2000}.
-The ITIC curve is a power acceptability curve that plots time on the x-axis and voltage on the y-axis.
-The purpose of the curve is to provide a tolerance envelope for single-phase 120V equipment.
-The curve defines three regions.
-The first region is "No Interruption" and generally includes all voltages with very short sustained durations.
-All events within this region have no noticeable effect on power equipment.
-The second region, the "No Damage Region", occurs during voltage sags for extended periods of time.
-Power Events in this region may cause equipment interruptions, but it will not damage the equipment.
-The final region, the "Prohibited" region, is caused by sustained voltage swells and may cause damage to power equipment.
-This Actor determines if an event falls within the "No Damage" or "Prohibited" regions and if so, creates an Incident to record this.
-
-{\em SEMI F47 Actor.} The SEMI F47 Actor is similar to the ITIC Actor in that it plots voltage and duration against a power acceptability curve.
-In this case, the standard used is the SEMI F47 standard \cite{djokic_sensitivity_2005}.
-Rather than using a point-in-polygon approach, this plugin reads the voltage features sequentially and uses a state machine to keep track of the current classification.
-This plugin only classifies values as a "violation" or as "nominal".
-
-{\em Transient Actor.} The Transient Actor is responsible for classifying frequency transients in power waveforms.
-The plugin subscribes to messages from a topic which contains a calibrated power waveform payload.
-The Transient Actor is capable of classifying impulsive, arcing, oscillatory, and periodic notching transients.
-A decision tree is utilized to select the most likely transient type and then further analysis is used to perform the actual classification of transients.
-Dickens et al \cite{dickens_transient_2019} provides more details on the transient classification system used by this Actor.
-
-{\em Periodicity Actor.} The Periodicity Actor is responsible for detecting periodic signals in power data.
-This Actor does not subscribe to any messages, but instead runs off of a configurable timer.
-The Actor is set to run by default once an hour and every hour it scrapes the last 24 hours worth of data and attempts to find periods in the Measurements over that duration.
-
-For each feature in the Measurement and Trend data (e.g.
-frequency, voltage, and THD), the Periodicity Actor first removes the DC offset from the data by subtracting the mean.
-Next, the Actor filters the signal using a 4th order high-pass filter to filter out noise.
-The Actor then performs autocorrelation on the signal followed by finding the peaks of the autocorrelation.
-The mean distance between the peaks of the autocorrelation provides the period of the signal.
-
-The Periodicity Actor only classifies data as periodic if at least 3 peaks were found and the standard deviation of the period is less than 600 seconds (10 minutes).
-Once a positive identification has been made, peak detection is performed on the original signal.
-Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the plugin can group Measurements, Trends, Detection Layer Events, and Incidents that were created during the periodic signals together as part of the Periodic Phenomena.
-
-\subsection{OPQ View}
-\label{sec:opq-view}
-
-The final component of the OPQ Sensor Network system architecture is called OPQ View.
-It is a web application, implemented using the Meteor application framework, which provides a variety of visualization and query services.
-An example of the OPQ View home page is provided in Figure \ref{fig:opq-view-home}.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/view/homepage.png}
-\caption{OPQ View: home page}
-\label{fig:opq-view-home}
-\end{figure}
-
-The home page provides three components:  a "System Stats" window which indicates the number of data elements currently at each level of the OPQ Mauka data hierarchy, a "System Health" window that indicates whether the Cloud services appear to be running correctly and the status of communication with all known OPQ Boxes on the sensor network, and a "Box Trends" visualization that provides the ability to quickly see trends in the four basic measures (Voltage, Frequency, THD, and Transients) over time.
-
-Figure \ref{fig:opq-view-box-map} shows a map-based view of the sensor network.
-This figure shows the location of 11 OPQ Boxes on the University of Hawaii campus at one point during Fall of 2019.
-Depending on the zoom level of the interface, some boxes are collapsed into disks with a number indicating the number of boxes at that location.
-Zooming in reveals more information about the box including near-real time values for voltage and frequency, as is shown for the box in the upper right corner of the figure.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/view/boxmap-2.png}
-\caption{OPQ View: Box Map}
-\label{fig:opq-view-box-map}
-\end{figure}
-
-Figure \ref{fig:opq-view-incident-summary} shows a visual display of a single Incident.
-This view includes a map-based location of the box whose data was involved in the Incident, the start and end time and duration of the Incident, the classification(s) of the anomalous data, and the waveform associated with the Incident (when applicable).
-If additional analysis is desired, the raw data can be downloaded in CSV format.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/view/incident-summary.png}
-\caption{OPQ View: Incident Summary}
-\label{fig:opq-view-incident-summary}
-\end{figure}
-
-The above images provide a sense for how OPQ View helps users to understand, monitor, and assess an OPQ Sensor Network and the underlying power quality of the grid it is attached to.
-There are several features in OPQ View for user and box management; for details, please see the OPQ View documentation.
-
-While OPQ View provides a variety of useful visualization and analysis features, users wishing to understand the power quality of their grid are not restricted to its current feature set.
-The OPQ database schema is public, and if a new analysis is desired, it is straightforward to query the database directly for the data of interest and build a new analysis off of it.

### 2020-07-18 graveyard/07-opq-mauka.tex
revision complete.

--- 2020-07-17 02-methods.tex
@@
-\section{Methods}
-
-The central aim of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
-
-Our research methodology consists of three (potentially interleaved) phases.
-First, we specify, design, and implement novel hardware and software for collecting, transmitting, and analyzing four essential measures of power quality data: voltage, frequency, THD, and transients.
-Second, we verify that our implementation matches our specifications through laboratory experiments.
-Third, we validate our implementation by testing a variety of hypotheses through a real-world deployment.
-In this paper, we report on results from a pilot deployment at the University of Hawaii over the course of three months.
-
-Our central hypothesis is that we can accomplish this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based services.
-
-We will test this central hypothesis through six subhypotheses: (1) OPQ provides valid and reliable collection of power quality data (Section \ref{hyp:01}), (2) OPQ's triggering system provides advantages with respect to bandwidth and computation (Section \ref{hyp:02}, (3) OPQ enables subthreshold event detection based upon temporal locality (Section \ref{sec:subthreshold-events}), (4) The OPQ information architecture provides a means to produce actionable insights (Section \ref{hyp:04}), (5) the OPQ information architecture provides predictable upper bounds on storage resources (Section \ref{hyp:05}), and (6) OPQ provides useful adaptive optimization capabilites (Section \ref{sec:adaptive-optimization}).
-
-The remainder of this section presents the overall architecture of OPQ, along with details on the design and implementation of each of the principal architectural components.
-We will also discuss our verification activities for each component.
-Our validation and hypothesis testing activities will be described in Section \ref{sec:pilot-study}.
-
-\subsection{OPQ Architecture}
-
-The OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events.
-OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components either at predefined intervals or upon request.
-OPQ Makai monitors incoming low fidelity data from OPQ Boxes, requests high fidelity data when necessary, and stores the results in a MongoDB database.
-OPQ Mauka analyzes low level data and analyzes it to produce higher-level representations according to the information architecture described below, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis.
-OPQ View is a browser-based visualization platform for displaying the results for data capture and analysis.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/architecture/system-diagram.png}
-\caption{High level system architecture of an OPQ Sensor Network}
-\label{fig:architecture}
-\end{figure}
-
-Figure \ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side).
-First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQ Makai.
-OPQ Makai analyzes low fidelity measurements, and requests high fidelity waveforms when desirable.
-Both measurements and waveforms are saved in a MongoDB database.
-OPQ Mauka analyzes low and high fidelity data, and creates "events" to represent anomalies.
-OPQ View notifies users of events and allows them to drill down into low and high fidelity data.
-
-OPQ Makai, OPQ Mauka, and OPQ View are all cloud-based software services that collectively form a single "instance" with respect to data transmission, storage, analysis, and visualization.
-We refer to this collection of software-side components as OPQ Cloud.
-Every OPQ Box connects to a single instance of an OPQ Cloud.
-It is possible to have multiple OPQ Cloud instances.
-For example, a company might install an OPQ Cloud instance behind their firewall along with OPQ Boxes to provide a private mechanism for collecting and analyzing power quality data.
-
-\subsection{OPQ Box}
-\label{sec:opq-box}
-
-OPQ Box is a hardware device designed to provide inexpensive, extensible and accurate residential power quality measurements.
-A block diagram of OPQ Box is shown in Figure~\ref{fig:opq:1:1}.
-A complete device is shown in Figure~\ref{fig:opq:1:2}.
-
-\begin{figure}[ht]
-	\centering
-	\begin{subfigure}{.5\textwidth}
-	  \centering
-	  \includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
-	  \caption{OPQ Box Block Diagram.
-	  The power path is in red, signal path is in green and the digital IO is in black.}
-	  \label{fig:opq:1:1}
-	\end{subfigure}%
-	\begin{subfigure}{.5\textwidth}
-	  \centering
-	  \includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
-	  \caption{OPQ Box in an ABS plastic enclosure.}
-	  \label{fig:opq:1:2}
-	\end{subfigure}
-	\caption{(a) OPQ Box block diagram and (b) production OPQ Box ready for deployment}
-	\label{fig:opq:2}
-\end{figure}
-
-\subsubsection{Hardware}\label{subsec:hardware}
-
-The power system of the OPQ Box electrically isolates most of the device from the AC mains power.
-An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
-5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference, and an isolated DC/DC converter.
-3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
-The hot side of the isolation amplifier is powered from the isolated DC/DC converter.
-This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power outage.
-
-\begin{figure}[ht]
-  \begin{center}
-  \includegraphics[width=0.9\textwidth]{images/opq-box/opqbox_software.png}
-  \end{center}
-  \caption{Block diagram of the OPQ Box 2 software stack.}
-  \label{fig:opq:3}
-\end{figure}
-
-A Raspberry Pi single board computer (SBC) is responsible for signal analysis and anomaly detection.
-The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256MB of main memory and a single core 1GHz ARM11 CPU.
-It also contains an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle.
-
-\subsubsection{Software}\label{subsec:software}
-
-The software stack of the Raspberry Pi aims to deliver an accurate and precise power quality analysis framework despite the rather limited capabilities of the hardware.
-A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
-Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with the Pi acting as the master and the DSP as a slave device.
-A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer, and a kernel driver provides management of the SPI bus.
-Internally, the OPQ driver maintains a ring buffer of 16 windows, each of which is 200 data samples in size.
-Upon receiving the interrupt for the DSP, the CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
-This scheme requires the CPU to only service 60 interrupts a second, with each interrupt requiring on the order of 100 instructions, for a CPU utilization of less than $1\%$ in normal operation.
-Userland applications communicate with the kernel driver using a file descriptor, where every $read$ system call yields 200 samples of raw waveform.
-As a result, the smallest window that a userland application may process is a single AC cycle of the grid mains.
-
-The userland component of the OPQ Box software is a multi-threaded extensible analysis framework called Triggering.
-The reader thread is responsible for transferring and accumulating data from the kernel driver.
-The smallest data buffer that the Triggering application processes at any given time is 10 grid cycles or 2k samples.
-Once the cycles are transferred to the userland and timestamped, they are passed to the analysis thread for feature extraction, as well as to the Raw Data Ring Buffer (RDRB).
-Since internally all data is addressed using shared pointers, during data duplication no copying is required.
-RDRS is capable of buffering up to an hour of data before it is overwritten, resulting in the RDBS maximum size of 100MB.
-
-The analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
-Four metrics are extracted from the data stream: (1) Fundamental frequency, (2) RMS Voltage, (3) Total Harmonic Distortion, and (4) Transients. Let's briefly discuss how each of these are computed.
-
-\subsubsection{Fundamental Frequency}\label{subsec:fundamental-frequency}
-
-The fundamental frequency is calculated by computing the zero crossings of the AC waveform.
-In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
-Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz), it is very computationally expensive to perform this filtering in a single step.
-Instead, filtering is accomplished via a set of two low pass finite impulse response (FIR) filters.
-First filter has a passband of 0-600Hz allowing us to downsample the waveform to 1200 samples per second.
-Next filter has a passband of 0-100Hz allowing for further removal of high frequency noise.
-Finally, zero crossings are extracted and used for the frequency calculation.
-The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.
-
-\subsubsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}
-
-Root mean square voltage ($V_{rms}$) in electrical power is the equivalent value of DC voltage which would dissipate the same power in the resistive load.
-$V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed value.
-
-Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation. Unlike the frequency calculation, the input is not filtered a priori.
-
-\subsubsection{Total Harmonic Distortion}\label{subsec:thd}
-
-The OPQ Box calculates total harmonic distortion (THD) using the industry standard methodology.
-It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
-Operationally, OPQ Box computes THD for 10 cycles of the fundamental frequency.
-First an FFT transforms the real voltage samples into its frequency components.
-Next, the square of the harmonic bins are accumulated and scaled by the magnitude of the fundamental power.
-
-\subsubsection{Transient Detection}\label{subsec:transient-detection}
-
-OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter with a cutoff frequency of $400Hz$ and searching for a maximum value in the remainder.
-
-It should be noted that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
-However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient detection.
-
-\subsubsection{Network Communication}\label{subsec:network-communication}
-
-The computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
-Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
-Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box to verify its peer.
-Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
-This protocol is a publish subscribe topology, with each message containing the topic and a payload.
-Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
-This allows for the aggregation service to be spread across multiple nodes, with minimal network overhead.
-
-If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box RDRB via a separate ZeroMq pub sub channel.
-If the RDRB buffer contains data for the requested temporal range, OPQ Box transmits the available data to the aggregation service via a push pull ZeroMq channel.
-Protobuf message serialization is used to encode messages across the OPQ ecosystem.
-
-In order to make a distributed measurement, all of the OPQ Boxes on the OPQ network need to maintain an accurate time reference.
-Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
-The expansion port of the OPQ Box supports a GPS receiver.
-However, since GPS receivers require line of sight to the sky, it was not used for deployment.
-NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a nearby NTP server.
-
-\subsubsection{Manufacturing}
-
-Currently, there is no mechanism for mass production of OPQ Boxes, but all of the plans are available under an Open Source license, so interested organizations with some basic hardware engineering skills can build their own boxes.
-
-Complete specifications for the OPQ Box hardware, firmware, and software are available \cite{negrashov_opq_2020}.
-As of the time of writing, a single OPQ Box can be manufactured for approximately \$100 in parts.
-The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately \$75 in parts per device.
-
-\subsection{OPQ Makai}
-\label{sec:opq-makai}
-
-OPQ Box provides an inexpensive hardware device for collecting four important power quality measures with high fidelity, but realizing its potential requires an innovative approach involving two-way communication between OPQ Boxes and the OPQ cloud-based services.
-To see why, consider the IEEE 1159 standard for single location power quality monitoring \cite{unruh_ieee_2018}.
-For transient monitoring, IEEE 1159 suggests a sampling rate of at least 7680 samples/second, up to 1 Megasample/second.
-This implies that if the cloud service requires the high fidelity data from all OPQ Boxes, it would incur a very large bandwidth cost.
-At 20 Ksamples/second with 16bit samples, a single OPQ Box will generate 300Kb/second of network bandwidth.
-Several thousand devices would easily saturate a 1 GB network link.
-In addition, collecting and recording all of the raw waveform data from residential power quality meters could lead to security and privacy issues.
-
-Network bandwidth saturation is a common problem for distributed sensor networks, and a common solution is called "self-triggering".
-In this approach, each monitoring device is configured with a threshold value for one or more measures of interest.
-When the threshold for a measure of interest is exceeded, then and only then is data sent over the network to cloud-based services for analysis.
-
-The problem with the self-triggering approach is that grid-wide power quality events do not affect the entire grid in the same way.
-For example, due to the grid’s hierarchical structure, a voltage sag on one sub-circuit can manifest as a sag of a different magnitude or even a swell on another \cite{kahle_power_2015}.
-This may result in a situation where some of the monitoring devices will not consider a power quality anomaly as an event, because it did not surpass the metric threshold, and simply ignore it.
-From an analysis perspective, however, it can be useful to get raw data from all of the affected devices, not just the ones that were the affected to the point where the box was triggered.
-This additional information can be used to localize the disturbance, as well as better evaluate its impact.
-
-Since sending all the data is infeasible, and since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai.
-In this scheme, OPQ Boxes use local processing resources to feature extract the incoming waveforms while storing them locally for an hour.
-Each OPQ Box sends its feature data to OPQ Makai once a second, which we called the "triggering stream".
-Feature data is very small, on the order of a few kilobytes, and so this approach allows the sensor network to scale to thousands of devices with acceptable network bandwidth requirements.
-OPQ Makai processes the incoming triggering stream and looks for anomalies.
-If an anomaly is present in only a single device, it is highly probable that the cause is local and not grid-level.
-On the other hand, if the triggering stream shows an anomaly temporally collocated across multiple devices, the entire network or a subset of the network may be queried for raw waveform data for a temporal region which corresponds to the disturbance in the triggering stream.
-
-Our pilot study, discussed in Section \ref{sec:pilot-study} will provide examples of the novel analysis capabilities made possible by OPQ Box and OPQ Makai communication.
-In general, here are the main advantages of our hybrid centralized/decentralized approach over traditional self-triggering and the ``naive" approach of sending all of the data:
-
-{\em Bandwidth usage is minimized.} Instead of sending the entirety of raw data, only extracted features are sent.
-This results in a tiny fraction of the bandwidth requirement when compared to raw waveforms.
-Furthermore, the temporal window which encompasses a single feature can be adjusted in real time.
-Thus, as soon as an anomalous behavior is observed in a subset of sensors, this window can be adjusted for a finer grained feature extraction.
-
-{\em Effects of latency are minimized.} In this case, "latency" refers to the time required for OPQ Makai to process the incoming feature stream and decide whether to request high fidelity data from one or more OPQ Boxes.
-Even at 1M samples/second at 16 bits of resolution, the memory requirement to store 5 minutes of raw waveform without compression are on the order of 512MB, which is well within the realm of inexpensive single board computers such as Raspberry PI.
-With compression specifically suited to the signal of interest, the memory requirement can be reduced even further.
-In the case of the OPQ sensor network, OPQ Makai has an hour to process feature data and request high fidelity data from OPQ Boxes.
-During our pilot study, OPQ Makai always responded within a second or two.
-
-{\em Cloud processing requirements are reduced.} Since feature extraction is already performed at the device level, cloud computation requirements are reduced.
-With the advent of the Internet of Things, the computational capacity of edge devices is increasing.
-
-{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes.
-This allows analysis of data from devices which were only mildly affected or even not affected at all by the disturbance.
-This creates new possibilities for investigation of disturbance propagation across the sensed area, as will be discussed in Section \ref{sec:pilot-study}.
-
-{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features.
-This allows for a simple robust algorithm which may be deployed at the sink node for anomaly detection.
-
-\subsubsection{Design}
-
-OPQ Makai is a distributed extensible microservice framework responsible for receiving the triggering stream from the OPQ Boxes, locating anomalous temporal regions and requesting raw waveform for the anomalous time ranges.
-As shown in Figure \ref{fig:makai-design}, Makai consists of four major components: Acquisition Broker, Triggering Broker, Event Service and the Acquisition Service.
-
-\begin{figure}
-\center \includegraphics[width=3in]{images/makai/makai_main.pdf}
-\caption{Makai component design}
-\label{fig:makai-design}
-\end{figure}
-
-The Triggering Broker is perhaps the simplest component of the OPQ Makai system.
-The triggering stream generated by the OPQ Boxes is encrypted to preserve user privacy.
-In order to minimize the CPU time spent decrypting the data across multiple OPQ services, the Triggering Broker decrypts the data and sends clear text measurements to other OPQ cloud services.
-The Triggering Broker uses the ZeroMq subscribe socket to receive data from OPQ Boxes, and sends it via a publish socket to any connected client.
-Each publish message is published to a topic which corresponds to the ASCII representation of the originating OPQ Box ID.
-This allows services which utilize the Triggering Broker to select a subset of IDs to operate on.
-This is useful for load balancing the backend services, or dividing the OPQ network into separate regions with no electrical connections between them.
-
-The Acquisition Broker manages the two-way communication between the OPQ Boxes and the rest of the cloud infrastructure.
-Unlike the triggering stream which originates from the OPQ Box, two-way communication is always initiated by OPQ cloud services.
-Two way communication is realized via a command response interface, where the OPQ service initiates the communication by sending a clear text command to the Acquisition Broker, which then forwards it in encrypted form to the appropriate OPQ Boxes.
-
-The Acquisition Service resides between the Triggering and Acquisition Brokers. The Acquisition Service is responsible for three tasks:
-(1) Computation of statistics of the incoming triggering stream; (2) Hosting plugins for triggering stream analysis; and (3) Generating data event requests for OPQ Boxes.
-The Acquisition Service accesses the triggering stream by connecting to the publish socket of the Triggering broker.
-Since the connection is managed through the ZeroMq publish-subscribe socket, several Acquisition Services can be connected to a single Triggering broker endpoint, each servicing a subset of OPQ Boxes by subscribing to only specific devices.
-The Acquisition Service does not include any analysis capabilities by default.
-Instead, analysis is performed by shared library loadable plugins.
-These plugins can be loaded and unloaded at runtime, thus allowing live upgrading and testing of new analysis methods.
-
-The Event service is a microservice which stores raw data generated by OPQ Boxes in the MongoDB Database.
-On initialization, the Event service queries MongoDB database for the highest event number recorded so far, connects to the Acquisition Broker’s publish port, and subscribes to all messages that start with the prefix ``data”.
-This allows the Event service to capture every response from OPQ Boxes generated from commands issued by the Acquisition service plugins.
-Once the Event service receives a data response with an identity containing an event token it hasn’t seen before, it will increment the event number, and store it in an internal key value store.
-
-\subsection{OPQ Mauka}
+\section{OPQ Mauka}
 \label{sec:opq-mauka}
@@
 
-\subsubsection{OPQ Mauka Actors}
+\subsection{OPQ Mauka Actors}
 
@@
 Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the plugin can group Measurements, Trends, Detection Layer Events, and Incidents that were created during the periodic signals together as part of the Periodic Phenomena.
-
-\subsection{OPQ View}
-\label{sec:opq-view}
-
-The final component of the OPQ Sensor Network system architecture is called OPQ View.
-It is a web application, implemented using the Meteor application framework, which provides a variety of visualization and query services.
-An example of the OPQ View home page is provided in Figure \ref{fig:opq-view-home}.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/view/homepage.png}
-\caption{OPQ View: home page}
-\label{fig:opq-view-home}
-\end{figure}
-
-The home page provides three components:  a "System Stats" window which indicates the number of data elements currently at each level of the OPQ Mauka data hierarchy, a "System Health" window that indicates whether the Cloud services appear to be running correctly and the status of communication with all known OPQ Boxes on the sensor network, and a "Box Trends" visualization that provides the ability to quickly see trends in the four basic measures (Voltage, Frequency, THD, and Transients) over time.
-
-Figure \ref{fig:opq-view-box-map} shows a map-based view of the sensor network.
-This figure shows the location of 11 OPQ Boxes on the University of Hawaii campus at one point during Fall of 2019.
-Depending on the zoom level of the interface, some boxes are collapsed into disks with a number indicating the number of boxes at that location.
-Zooming in reveals more information about the box including near-real time values for voltage and frequency, as is shown for the box in the upper right corner of the figure.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/view/boxmap-2.png}
-\caption{OPQ View: Box Map}
-\label{fig:opq-view-box-map}
-\end{figure}
-
-Figure \ref{fig:opq-view-incident-summary} shows a visual display of a single Incident.
-This view includes a map-based location of the box whose data was involved in the Incident, the start and end time and duration of the Incident, the classification(s) of the anomalous data, and the waveform associated with the Incident (when applicable).
-If additional analysis is desired, the raw data can be downloaded in CSV format.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/view/incident-summary.png}
-\caption{OPQ View: Incident Summary}
-\label{fig:opq-view-incident-summary}
-\end{figure}
-
-The above images provide a sense for how OPQ View helps users to understand, monitor, and assess an OPQ Sensor Network and the underlying power quality of the grid it is attached to.
-There are several features in OPQ View for user and box management; for details, please see the OPQ View documentation.
-
-While OPQ View provides a variety of useful visualization and analysis features, users wishing to understand the power quality of their grid are not restricted to its current feature set.
-The OPQ database schema is public, and if a new analysis is desired, it is straightforward to query the database directly for the data of interest and build a new analysis off of it.

### 2020-07-18 graveyard/08-opq-view.tex
revision complete.

--- 2020-07-17 02-methods.tex
@@
-\section{Methods}
-
-The central aim of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation.
-
-Our research methodology consists of three (potentially interleaved) phases.
-First, we specify, design, and implement novel hardware and software for collecting, transmitting, and analyzing four essential measures of power quality data: voltage, frequency, THD, and transients.
-Second, we verify that our implementation matches our specifications through laboratory experiments.
-Third, we validate our implementation by testing a variety of hypotheses through a real-world deployment.
-In this paper, we report on results from a pilot deployment at the University of Hawaii over the course of three months.
-
-Our central hypothesis is that we can accomplish this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based services.
-
-We will test this central hypothesis through six subhypotheses: (1) OPQ provides valid and reliable collection of power quality data (Section \ref{hyp:01}), (2) OPQ's triggering system provides advantages with respect to bandwidth and computation (Section \ref{hyp:02}, (3) OPQ enables subthreshold event detection based upon temporal locality (Section \ref{sec:subthreshold-events}), (4) The OPQ information architecture provides a means to produce actionable insights (Section \ref{hyp:04}), (5) the OPQ information architecture provides predictable upper bounds on storage resources (Section \ref{hyp:05}), and (6) OPQ provides useful adaptive optimization capabilites (Section \ref{sec:adaptive-optimization}).
-
-The remainder of this section presents the overall architecture of OPQ, along with details on the design and implementation of each of the principal architectural components.
-We will also discuss our verification activities for each component.
-Our validation and hypothesis testing activities will be described in Section \ref{sec:pilot-study}.
-
-\subsection{OPQ Architecture}
-
-The OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and reporting of consumer level local and global PQ events.
-OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components either at predefined intervals or upon request.
-OPQ Makai monitors incoming low fidelity data from OPQ Boxes, requests high fidelity data when necessary, and stores the results in a MongoDB database.
-OPQ Mauka analyzes low level data and analyzes it to produce higher-level representations according to the information architecture described below, and can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis.
-OPQ View is a browser-based visualization platform for displaying the results for data capture and analysis.
-
-\begin{figure}
-\center \includegraphics[width=5in]{images/architecture/system-diagram.png}
-\caption{High level system architecture of an OPQ Sensor Network}
-\label{fig:architecture}
-\end{figure}
-
-Figure \ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side).
-First, OPQ Boxes analyze power from wall outlets, and send low fidelity measurements to OPQ Makai.
-OPQ Makai analyzes low fidelity measurements, and requests high fidelity waveforms when desirable.
-Both measurements and waveforms are saved in a MongoDB database.
-OPQ Mauka analyzes low and high fidelity data, and creates "events" to represent anomalies.
-OPQ View notifies users of events and allows them to drill down into low and high fidelity data.
-
-OPQ Makai, OPQ Mauka, and OPQ View are all cloud-based software services that collectively form a single "instance" with respect to data transmission, storage, analysis, and visualization.
-We refer to this collection of software-side components as OPQ Cloud.
-Every OPQ Box connects to a single instance of an OPQ Cloud.
-It is possible to have multiple OPQ Cloud instances.
-For example, a company might install an OPQ Cloud instance behind their firewall along with OPQ Boxes to provide a private mechanism for collecting and analyzing power quality data.
-
-\subsection{OPQ Box}
-\label{sec:opq-box}
-
-OPQ Box is a hardware device designed to provide inexpensive, extensible and accurate residential power quality measurements.
-A block diagram of OPQ Box is shown in Figure~\ref{fig:opq:1:1}.
-A complete device is shown in Figure~\ref{fig:opq:1:2}.
-
-\begin{figure}[ht]
-	\centering
-	\begin{subfigure}{.5\textwidth}
-	  \centering
-	  \includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
-	  \caption{OPQ Box Block Diagram.
-	  The power path is in red, signal path is in green and the digital IO is in black.}
-	  \label{fig:opq:1:1}
-	\end{subfigure}%
-	\begin{subfigure}{.5\textwidth}
-	  \centering
-	  \includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
-	  \caption{OPQ Box in an ABS plastic enclosure.}
-	  \label{fig:opq:1:2}
-	\end{subfigure}
-	\caption{(a) OPQ Box block diagram and (b) production OPQ Box ready for deployment}
-	\label{fig:opq:2}
-\end{figure}
-
-\subsubsection{Hardware}\label{subsec:hardware}
-
-The power system of the OPQ Box electrically isolates most of the device from the AC mains power.
-An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
-5V is used to power the Raspberry Pi, the equipment connected to the expansion port, the 3.3V regulators and voltage reference, and an isolated DC/DC converter.
-3.3V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
-The hot side of the isolation amplifier is powered from the isolated DC/DC converter.
-This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power outage.
-
-\begin{figure}[ht]
-  \begin{center}
-  \includegraphics[width=0.9\textwidth]{images/opq-box/opqbox_software.png}
-  \end{center}
-  \caption{Block diagram of the OPQ Box 2 software stack.}
-  \label{fig:opq:3}
-\end{figure}
-
-A Raspberry Pi single board computer (SBC) is responsible for signal analysis and anomaly detection.
-The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256MB of main memory and a single core 1GHz ARM11 CPU.
-It also contains an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI dongle.
-
-\subsubsection{Software}\label{subsec:software}
-
-The software stack of the Raspberry Pi aims to deliver an accurate and precise power quality analysis framework despite the rather limited capabilities of the hardware.
-A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
-Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with the Pi acting as the master and the DSP as a slave device.
-A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer, and a kernel driver provides management of the SPI bus.
-Internally, the OPQ driver maintains a ring buffer of 16 windows, each of which is 200 data samples in size.
-Upon receiving the interrupt for the DSP, the CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
-This scheme requires the CPU to only service 60 interrupts a second, with each interrupt requiring on the order of 100 instructions, for a CPU utilization of less than $1\%$ in normal operation.
-Userland applications communicate with the kernel driver using a file descriptor, where every $read$ system call yields 200 samples of raw waveform.
-As a result, the smallest window that a userland application may process is a single AC cycle of the grid mains.
-
-The userland component of the OPQ Box software is a multi-threaded extensible analysis framework called Triggering.
-The reader thread is responsible for transferring and accumulating data from the kernel driver.
-The smallest data buffer that the Triggering application processes at any given time is 10 grid cycles or 2k samples.
-Once the cycles are transferred to the userland and timestamped, they are passed to the analysis thread for feature extraction, as well as to the Raw Data Ring Buffer (RDRB).
-Since internally all data is addressed using shared pointers, during data duplication no copying is required.
-RDRS is capable of buffering up to an hour of data before it is overwritten, resulting in the RDBS maximum size of 100MB.
-
-The analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
-Four metrics are extracted from the data stream: (1) Fundamental frequency, (2) RMS Voltage, (3) Total Harmonic Distortion, and (4) Transients. Let's briefly discuss how each of these are computed.
-
-\subsubsection{Fundamental Frequency}\label{subsec:fundamental-frequency}
-
-The fundamental frequency is calculated by computing the zero crossings of the AC waveform.
-In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
-Since our sampling rate is quite high (12kSps) and the fundamental frequency is quite low (60Hz), it is very computationally expensive to perform this filtering in a single step.
-Instead, filtering is accomplished via a set of two low pass finite impulse response (FIR) filters.
-First filter has a passband of 0-600Hz allowing us to downsample the waveform to 1200 samples per second.
-Next filter has a passband of 0-100Hz allowing for further removal of high frequency noise.
-Finally, zero crossings are extracted and used for the frequency calculation.
-The zero crossings themselves were calculated by using linear interpolation between two points which bracket the time axis.
-
-\subsubsection{Root Mean Square Voltage}\label{subsec:root-mean-square-voltage}
-
-Root mean square voltage ($V_{rms}$) in electrical power is the equivalent value of DC voltage which would dissipate the same power in the resistive load.
-$V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed value.
-
-Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation. Unlike the frequency calculation, the input is not filtered a priori.
-
-\subsubsection{Total Harmonic Distortion}\label{subsec:thd}
-
-The OPQ Box calculates total harmonic distortion (THD) using the industry standard methodology.
-It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
-Operationally, OPQ Box computes THD for 10 cycles of the fundamental frequency.
-First an FFT transforms the real voltage samples into its frequency components.
-Next, the square of the harmonic bins are accumulated and scaled by the magnitude of the fundamental power.
-
-\subsubsection{Transient Detection}\label{subsec:transient-detection}
-
-OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass pass filter with a cutoff frequency of $400Hz$ and searching for a maximum value in the remainder.
-
-It should be noted that this transient detection method is susceptible to THD fluctuations, since any harmonic above $400Hz$ will remain in the filtered waveform.
-However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient detection.
-
-\subsubsection{Network Communication}\label{subsec:network-communication}
-
-The computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
-Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
-Each device holds a unique private and public key, as well as the servers' public key, allowing both the Makai service and the OPQ Box to verify its peer.
-Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
-This protocol is a publish subscribe topology, with each message containing the topic and a payload.
-Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
-This allows for the aggregation service to be spread across multiple nodes, with minimal network overhead.
-
-If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box RDRB via a separate ZeroMq pub sub channel.
-If the RDRB buffer contains data for the requested temporal range, OPQ Box transmits the available data to the aggregation service via a push pull ZeroMq channel.
-Protobuf message serialization is used to encode messages across the OPQ ecosystem.
-
-In order to make a distributed measurement, all of the OPQ Boxes on the OPQ network need to maintain an accurate time reference.
-Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
-The expansion port of the OPQ Box supports a GPS receiver.
-However, since GPS receivers require line of sight to the sky, it was not used for deployment.
-NTP performance has been verified against GPS resulting in time error of $8ms\pm 5ms$ which is typical for NTP running over the Internet with a nearby NTP server.
-
-\subsubsection{Manufacturing}
-
-Currently, there is no mechanism for mass production of OPQ Boxes, but all of the plans are available under an Open Source license, so interested organizations with some basic hardware engineering skills can build their own boxes.
-
-Complete specifications for the OPQ Box hardware, firmware, and software are available \cite{negrashov_opq_2020}.
-As of the time of writing, a single OPQ Box can be manufactured for approximately \$100 in parts.
-The cost drops significantly with scale, for example, 100 OPQ Boxes can be manufactured for a cost of approximately \$75 in parts per device.
-
-\subsection{OPQ Makai}
-\label{sec:opq-makai}
-
-OPQ Box provides an inexpensive hardware device for collecting four important power quality measures with high fidelity, but realizing its potential requires an innovative approach involving two-way communication between OPQ Boxes and the OPQ cloud-based services.
-To see why, consider the IEEE 1159 standard for single location power quality monitoring \cite{unruh_ieee_2018}.
-For transient monitoring, IEEE 1159 suggests a sampling rate of at least 7680 samples/second, up to 1 Megasample/second.
-This implies that if the cloud service requires the high fidelity data from all OPQ Boxes, it would incur a very large bandwidth cost.
-At 20 Ksamples/second with 16bit samples, a single OPQ Box will generate 300Kb/second of network bandwidth.
-Several thousand devices would easily saturate a 1 GB network link.
-In addition, collecting and recording all of the raw waveform data from residential power quality meters could lead to security and privacy issues.
-
-Network bandwidth saturation is a common problem for distributed sensor networks, and a common solution is called "self-triggering".
-In this approach, each monitoring device is configured with a threshold value for one or more measures of interest.
-When the threshold for a measure of interest is exceeded, then and only then is data sent over the network to cloud-based services for analysis.
-
-The problem with the self-triggering approach is that grid-wide power quality events do not affect the entire grid in the same way.
-For example, due to the grid’s hierarchical structure, a voltage sag on one sub-circuit can manifest as a sag of a different magnitude or even a swell on another \cite{kahle_power_2015}.
-This may result in a situation where some of the monitoring devices will not consider a power quality anomaly as an event, because it did not surpass the metric threshold, and simply ignore it.
-From an analysis perspective, however, it can be useful to get raw data from all of the affected devices, not just the ones that were the affected to the point where the box was triggered.
-This additional information can be used to localize the disturbance, as well as better evaluate its impact.
-
-Since sending all the data is infeasible, and since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai.
-In this scheme, OPQ Boxes use local processing resources to feature extract the incoming waveforms while storing them locally for an hour.
-Each OPQ Box sends its feature data to OPQ Makai once a second, which we called the "triggering stream".
-Feature data is very small, on the order of a few kilobytes, and so this approach allows the sensor network to scale to thousands of devices with acceptable network bandwidth requirements.
-OPQ Makai processes the incoming triggering stream and looks for anomalies.
-If an anomaly is present in only a single device, it is highly probable that the cause is local and not grid-level.
-On the other hand, if the triggering stream shows an anomaly temporally collocated across multiple devices, the entire network or a subset of the network may be queried for raw waveform data for a temporal region which corresponds to the disturbance in the triggering stream.
-
-Our pilot study, discussed in Section \ref{sec:pilot-study} will provide examples of the novel analysis capabilities made possible by OPQ Box and OPQ Makai communication.
-In general, here are the main advantages of our hybrid centralized/decentralized approach over traditional self-triggering and the ``naive" approach of sending all of the data:
-
-{\em Bandwidth usage is minimized.} Instead of sending the entirety of raw data, only extracted features are sent.
-This results in a tiny fraction of the bandwidth requirement when compared to raw waveforms.
-Furthermore, the temporal window which encompasses a single feature can be adjusted in real time.
-Thus, as soon as an anomalous behavior is observed in a subset of sensors, this window can be adjusted for a finer grained feature extraction.
-
-{\em Effects of latency are minimized.} In this case, "latency" refers to the time required for OPQ Makai to process the incoming feature stream and decide whether to request high fidelity data from one or more OPQ Boxes.
-Even at 1M samples/second at 16 bits of resolution, the memory requirement to store 5 minutes of raw waveform without compression are on the order of 512MB, which is well within the realm of inexpensive single board computers such as Raspberry PI.
-With compression specifically suited to the signal of interest, the memory requirement can be reduced even further.
-In the case of the OPQ sensor network, OPQ Makai has an hour to process feature data and request high fidelity data from OPQ Boxes.
-During our pilot study, OPQ Makai always responded within a second or two.
-
-{\em Cloud processing requirements are reduced.} Since feature extraction is already performed at the device level, cloud computation requirements are reduced.
-With the advent of the Internet of Things, the computational capacity of edge devices is increasing.
-
-{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes.
-This allows analysis of data from devices which were only mildly affected or even not affected at all by the disturbance.
-This creates new possibilities for investigation of disturbance propagation across the sensed area, as will be discussed in Section \ref{sec:pilot-study}.
-
-{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features.
-This allows for a simple robust algorithm which may be deployed at the sink node for anomaly detection.
-
-\subsubsection{Design}
-
-OPQ Makai is a distributed extensible microservice framework responsible for receiving the triggering stream from the OPQ Boxes, locating anomalous temporal regions and requesting raw waveform for the anomalous time ranges.
-As shown in Figure \ref{fig:makai-design}, Makai consists of four major components: Acquisition Broker, Triggering Broker, Event Service and the Acquisition Service.
-
-\begin{figure}
-\center \includegraphics[width=3in]{images/makai/makai_main.pdf}
-\caption{Makai component design}
-\label{fig:makai-design}
-\end{figure}
-
-The Triggering Broker is perhaps the simplest component of the OPQ Makai system.
-The triggering stream generated by the OPQ Boxes is encrypted to preserve user privacy.
-In order to minimize the CPU time spent decrypting the data across multiple OPQ services, the Triggering Broker decrypts the data and sends clear text measurements to other OPQ cloud services.
-The Triggering Broker uses the ZeroMq subscribe socket to receive data from OPQ Boxes, and sends it via a publish socket to any connected client.
-Each publish message is published to a topic which corresponds to the ASCII representation of the originating OPQ Box ID.
-This allows services which utilize the Triggering Broker to select a subset of IDs to operate on.
-This is useful for load balancing the backend services, or dividing the OPQ network into separate regions with no electrical connections between them.
-
-The Acquisition Broker manages the two-way communication between the OPQ Boxes and the rest of the cloud infrastructure.
-Unlike the triggering stream which originates from the OPQ Box, two-way communication is always initiated by OPQ cloud services.
-Two way communication is realized via a command response interface, where the OPQ service initiates the communication by sending a clear text command to the Acquisition Broker, which then forwards it in encrypted form to the appropriate OPQ Boxes.
-
-The Acquisition Service resides between the Triggering and Acquisition Brokers. The Acquisition Service is responsible for three tasks:
-(1) Computation of statistics of the incoming triggering stream; (2) Hosting plugins for triggering stream analysis; and (3) Generating data event requests for OPQ Boxes.
-The Acquisition Service accesses the triggering stream by connecting to the publish socket of the Triggering broker.
-Since the connection is managed through the ZeroMq publish-subscribe socket, several Acquisition Services can be connected to a single Triggering broker endpoint, each servicing a subset of OPQ Boxes by subscribing to only specific devices.
-The Acquisition Service does not include any analysis capabilities by default.
-Instead, analysis is performed by shared library loadable plugins.
-These plugins can be loaded and unloaded at runtime, thus allowing live upgrading and testing of new analysis methods.
-
-The Event service is a microservice which stores raw data generated by OPQ Boxes in the MongoDB Database.
-On initialization, the Event service queries MongoDB database for the highest event number recorded so far, connects to the Acquisition Broker’s publish port, and subscribes to all messages that start with the prefix ``data”.
-This allows the Event service to capture every response from OPQ Boxes generated from commands issued by the Acquisition service plugins.
-Once the Event service receives a data response with an identity containing an event token it hasn’t seen before, it will increment the event number, and store it in an internal key value store.
-
-\subsection{OPQ Mauka}
-\label{sec:opq-mauka}
-
-The previous sections discussed the design of OPQ Box, a custom hardware device for collecting four important measures of power quality, and OPQ Makai, a novel, hybrid centralized/decentralized data acquisition scheme which involves two-way communication between the OPQ Boxes.
-As a result of these two innovations, an OPQ sensor network has the ability to collect and analyze high fidelity, low level data about power quality anomalies in a cost-effective, scalable fashion.
-
-There are remaining challenges to creating a useful power quality sensor network.
-First, the data provided by OPQ Boxes is low-level, "primitive" data consisting of either features (i.e.
-frequency, voltage, THD, and transients) or waveform data.
-But what we actually want is actionable insights into grid stability.
-For example, we might want to know if a given anomalous data value is actually detrimental, or we might want to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical data.
-
-A second challenge involves the potentially high volume of data might accumulate in the cloud.
-Although OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over time, a substantial amount of data could accumulate.
-One strategy is to simply store all of the data sent to the cloud forever.
-This means that data storage requirements will increase monotonically over time, making the sensor network more costly to maintain the longer it is in place.
-An alternative strategy is to implement an algorithm to identify uninteresting (or no longer interesting) data and discard it.
-Ideally, such an algorithm would enable OPQ sensor network designers to calculate an upper bound on the total amount of cloud storage required as a function of the number of nodes (OPQ Boxes) in the network.
-
-OPQ Mauka addresses both of these issues.
-First, OPQ Mauka provides a multi-layered representation for structuring and processing DSN data.
-The structure and processing at each layer is designed with the explicit goal of turning low-level data into actionable insights.
-Second, each layer in the framework implements a "time-to-live" (TTL) strategy for data within the level.
-This strategy states that data must either progress upwards through the layers towards more abstract, useful representations within a fixed time window, or else it can be discarded.
-The TTL strategy is useful because when implemented, it allows DSN designers to make reasonable predictions of the upper bounds on data storage at each level of the framework adjusting for the number of sensors and power anomaly probability.
-
- TTL also makes possible a ``graceful degradation" of system performance if those bounds turn out to be exceeded.
-For example, consider a situation in which a power network enters a prolonged period of widespread power quality instability, where every OPQ Box is reporting continuous anomalous conditions with respect to voltage, frequency, THD, and transients.
-This ``worst case scenario" would lead to the potential situation of every OPQ Box trying to upload raw waveform data all the time.
-The TTL system provides safeguards, in that whatever low-level data has not been processed relatively quickly can be discarded.
-Thus, instead of the system potentially going down entirely, it could instead continue to operate at a reduced capacity.
-
-Figure \ref{fig:mauka-data-model} illustrates the hierarchical data model for OPQ Mauka.
-This data model can be conceptualized as a multi-level hierarchy that adaptively optimizes data storage using a tiered TTL approach and provides a mechanism in which typed aggregated data is continually refined to the point of being of becoming actionable.
-The data model also includes software components called "actors" that both move data upward through the levels and also apply optimizations downward through the levels.
-Actors are implemented through a plugin architecture, making it easy to experiment with the data model and improve it over time.
-
-\begin{figure}
-\center \includegraphics[width=4in]{images/mauka/mauka-data-model.png}
-\caption{Mauka data model hierarchy}
-\label{fig:mauka-data-model}
-\end{figure}
-
-The lowest layer of the hierarchy is the Instantaneous Measurements Layer (IML).
-The IML contains "raw" data, in other words, the digitized waveform.
-IML data exists both on each OPQ Box (where it is available for up to the previous 60 minutes).
-It also exists in the cloud, in the event that OPQ's triggering mechanism has caused a temporal interval of waveform data to be uploaded.
-IML data in the cloud has a TTL of 15 minutes: unless the waveform data is found to be useful by a cloud service within 15 minutes, it can be discarded.
-
-The second layer is the Aggregate Measurements Level (AML).
-The AML stores power quality summary statistics sent either once per second or once per minute by each OPQ Box.
-These summary statistics include the maximum, minimum, and average values of voltage, frequency, THD, and transient metrics over the time window.
-It is AML data that is used to initiate the triggering process of uploading IML data from the cloud.
-
-The third layer is the Detections Level (DL).
-This layer is responsible for processing the IML and AML data to produce a representation of an "event" with a precisely defined start and end time based upon examination of the waveform.
-As will be discussed in Section \ref{sec:subthreshold-events}, knowledge of the start and end time of a power quality anomaly allows investigation of how that anomaly might manifest itself elsewhere in the grid, even if this manifestation is not severe enough to produce over-threshold data values.
-
-The fourth layer is the Incident Level (IL).
-This layer starts to answer the question of whether the data is "actionable" by classifying the detected event according to various industry standards for power quality anomalies: IEEE 1159, ITIC, SEMI F47, and so forth.
-For example, an event that falls into the ITIC "prohibited" region clearly indicates a power quality anomaly that requires further study and intervention.
-
-The fifth and final level is the Phenomena Level (PL).
-This layer contains the results of analyses that attempt to identify cyclic, and thus predictable, power quality disturbances.
-It also contains analysis results regarding the similarity of various incidents, which can help uncover causal factors.
-Finally, it provides analyses for adaptive optimization of the OPQ Sensor Network.
-These optimizations can change the thresholds for individual boxes to either increase their sensitivity or decrease their sensitivity over specific intervals of time.
-The ultimate goal of adaptive optimization is to help the network learn to acquire all of the data useful for analyses, and only the data useful for analyses.
-We are still in the early stages of exploring the potential of adaptive optimization in OPQ networks.
-
-\subsubsection{OPQ Mauka Actors}
-
-The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins. Nine of the most important Actors are described below.
-
-{\em Makai Event Actor.} The Makai Event Actor is responsible for reading data newly created by OPQ Makai into OPQ Mauka.
-It performs feature extraction on the raw data stream and forwards those features (or the raw data) to subscribing Actor plugins.
-This allows OPQ Mauka to perform feature extraction once, and allow use of those features by multiple Actors.
-
-{\em Frequency Variation Actor.} The Frequency Variation Actor classifies generic frequency sags, swells, and interruptions as defined by the IEEE 1159 standard.
-Both duration and deviation from nominal are used to perform these classifications.
-Duration classifications include frequency deviations that last for less than 50 ns, between 50 ns to 1 ms, and 1 ms to 50 ms.
-Classifications for deviations from nominal are performed for values that are up to 40\% deviation from nominal.
-This Actor is able to classify frequency swells, frequency interruptions, and frequency sags, leading to the creation of data at the Incident Layer.
-
-{\em IEEE 1159 Voltage Actor.} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard[29].
-In general, this standard classifies voltage disturbances by duration and by magnitude.
-Voltage durations are classified from 0.5 to 30 cycles, 30 cycles to 3 seconds, 3 seconds to a minute, and greater than 1 minute.
-Voltage deviations are classified in both the sag and swell directions as a percentage from nominal.
-Sags are generally classified between 10\% and 90\% of nominal while swells are generally classified from 110\% to 180\% of nominal.
-This Actor is capable of classifying voltage sags, swells, and interruptions as defined by the standard, and creating data at the Incident Layer if appropriate.
-
-{\em Box Optimization Actor.} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
-This Actor is capable of requesting the state of each OPQ Box (e.g.
-uptime, Measurement rate, security keys, etc).
-It is also capable of adjusting the state of individual OPQ Boxes by changing things such as the Measurement and Trend rate or the sampling rate used by the Box.
-
-{\em Future Phenomena Actor.} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena.
-These Phenomena are used to predict Events and Incidents that may occur in the future.
-This plugin does not subscribe to any messages, but instead utilizes timers to perform its work.
-By default, this plugin runs every 10 minutes.
-
-When a Future Phenomena Actor runs, it loads any active Periodic Phenomena found in the database.
-If Periodic Phenomena are found, this Actor extrapolates possible Detection and Incident Layer data by first examining their timestamps and then extrapolating into the future using the mean period and the standard deviation.
-For each timestamp in a Periodic Phenomena, the mean period is added.
-If the resulting timestamp is in the future, a Future Phenomena is created using the time range of the future timestamp plus or minus the standard deviation of the Periodic Phenomena.
-
-When a Future Phenomena is created, timers are started in a separate thread signifying the start and end timestamps of the Future Phenomena.
-When the first timer runs, messages are sent to the Box Optimization Actor and the Threshold Optimization Actor instructing OPQ Box thresholds to be set lower and measurement rates to be set higher.
-This increases the chance of seeing an anomaly over the predicted time window.
-When the second timer runs, these values are reset to their default values.
-Thus, the plugin increases fidelity and decreases thresholds over the period of a Future Phenomena.
-
-{\em ITIC Actor.} The ITIC Actor analyzes voltage to determine where it falls within the ITIC curve \cite{thallam_power_2000}.
-The ITIC curve is a power acceptability curve that plots time on the x-axis and voltage on the y-axis.
-The purpose of the curve is to provide a tolerance envelope for single-phase 120V equipment.
-The curve defines three regions.
-The first region is "No Interruption" and generally includes all voltages with very short sustained durations.
-All events within this region have no noticeable effect on power equipment.
-The second region, the "No Damage Region", occurs during voltage sags for extended periods of time.
-Power Events in this region may cause equipment interruptions, but it will not damage the equipment.
-The final region, the "Prohibited" region, is caused by sustained voltage swells and may cause damage to power equipment.
-This Actor determines if an event falls within the "No Damage" or "Prohibited" regions and if so, creates an Incident to record this.
-
-{\em SEMI F47 Actor.} The SEMI F47 Actor is similar to the ITIC Actor in that it plots voltage and duration against a power acceptability curve.
-In this case, the standard used is the SEMI F47 standard \cite{djokic_sensitivity_2005}.
-Rather than using a point-in-polygon approach, this plugin reads the voltage features sequentially and uses a state machine to keep track of the current classification.
-This plugin only classifies values as a "violation" or as "nominal".
-
-{\em Transient Actor.} The Transient Actor is responsible for classifying frequency transients in power waveforms.
-The plugin subscribes to messages from a topic which contains a calibrated power waveform payload.
-The Transient Actor is capable of classifying impulsive, arcing, oscillatory, and periodic notching transients.
-A decision tree is utilized to select the most likely transient type and then further analysis is used to perform the actual classification of transients.
-Dickens et al \cite{dickens_transient_2019} provides more details on the transient classification system used by this Actor.
-
-{\em Periodicity Actor.} The Periodicity Actor is responsible for detecting periodic signals in power data.
-This Actor does not subscribe to any messages, but instead runs off of a configurable timer.
-The Actor is set to run by default once an hour and every hour it scrapes the last 24 hours worth of data and attempts to find periods in the Measurements over that duration.
-
-For each feature in the Measurement and Trend data (e.g.
-frequency, voltage, and THD), the Periodicity Actor first removes the DC offset from the data by subtracting the mean.
-Next, the Actor filters the signal using a 4th order high-pass filter to filter out noise.
-The Actor then performs autocorrelation on the signal followed by finding the peaks of the autocorrelation.
-The mean distance between the peaks of the autocorrelation provides the period of the signal.
-
-The Periodicity Actor only classifies data as periodic if at least 3 peaks were found and the standard deviation of the period is less than 600 seconds (10 minutes).
-Once a positive identification has been made, peak detection is performed on the original signal.
-Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the plugin can group Measurements, Trends, Detection Layer Events, and Incidents that were created during the periodic signals together as part of the Periodic Phenomena.
-
-\subsection{OPQ View}
+\section{OPQ View}
 \label{sec:opq-view}
@@
 The home page provides three components:  a "System Stats" window which indicates the number of data elements currently at each level of the OPQ Mauka data hierarchy, a "System Health" window that indicates whether the Cloud services appear to be running correctly and the status of communication with all known OPQ Boxes on the sensor network, and a "Box Trends" visualization that provides the ability to quickly see trends in the four basic measures (Voltage, Frequency, THD, and Transients) over time.
+
+The home page's ``System Stats'' widget shows the number of items stored at each OPQ storage level.
+The right hand column shows items that have been stored all time (not including items that have been discarded due to TTL).
+The left column shows items that have been collected in a single day (not including items that have been discarded by TTL).
 

### 2020-07-31 manuscript.v5/energies-864131.tex
camera ready

% LaTeX support: latex@mdpi.com
% In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

%=================================================================
\documentclass[energies,article,accept,moreauthors,pdftex]{Definitions/mdpi}
% \documentclass[preprints,article,accept,moreauthors,pdftex]{Definitions/mdpi}

%=================================================================
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{xx}
\issuenum{1}
\articlenumber{5}
\pubyear{2020}
\copyrightyear{2020}
%\externaleditor{Academic Editor: name}
\history{Received: 27 June 2020; Accepted: 28 July 2020; Published: date}
\updates{yes} % If there is an update available, un-comment this line

%% MDPI internal command: uncomment if new journal that already uses continuous page numbers
%\continuouspages{yes}

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx,epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption

\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{\alph{subfigure}}
\DeclareCaptionLabelFormat{subcaptionlabel}{\normalfont(\textbf{#2}\normalfont)}
\captionsetup[subfigure]{labelformat=subcaptionlabel}

\usepackage{tcolorbox}
\usepackage{tabularx}

%=================================================================
%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{%The paper was slightly modified by our English editor, please check the whole text and confirm if your meaning is retained; 2.	Do not delete any comment we left for you and reply to each comment so that we can understand your meaning clearly; 3.	If you need to revise somewhere in your paper, please highlight the revisions to let us know. (Thank you for your cooperation in advance.)
\highlight{Design}, Implementation, and Evaluation of Open Power Quality}

% Authors, for the paper (add full first names)
\Author{\hl{Anthony} %Please carefully check the accuracy of names and affiliations. 
J. Christe, Sergey Negrashov and Philip M. Johnson *}

% Authors, for metadata in PDF
\AuthorNames{Anthony J. Christe, Sergey Negrashov and Philip M. Johnson}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
Department of Information and Computer Sciences, University of Hawaii at Manoa, Honolulu, HI 96822, USA; achriste@hawaii.edu (A.J.C.); sin8@hawaii.edu (S.N.)
}

\corres{Correspondence: johnson@hawaii.edu}

% Abstract (single paragraph, 200 words max, background, methods, results, conclusions. Currently 193 words.)

\abstract{Modern electrical grids are transitioning from a centralized generation architecture to an architecture with significant distributed, intermittent generation. This transition means that the formerly sharp distinction between energy producers (utility companies) and consumers (residences, businesses, etc.) are blurring: end-users both produce and consume energy, making energy management and public policy more complex. The goal of the Open Power Quality (OPQ) project is to design and implement a low cost, distributed power quality sensor network that provides useful new forms of information about modern electrical grids to producers, consumers, researchers, and~policy makers. In 2019, we performed a pilot study involving the deployment of an OPQ sensor network at the University of Hawaii microgrid for three months. Results of the pilot study validate the ability of OPQ to collect accurate power quality data in a way that provides useful new insights into electrical grids.}

% Keywords
\keyword{power quality; open source; renewable energy; grid stability}

\begin{document}

\section{Introduction}

%{\em The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions.}

Power quality is not currently a concern for most people in developed nations. Just like most people in developed nations assume that their tap water is of adequate quality to drink, most also assume that their electricity is of adequate quality to power their homes and appliances without causing harm. In~both cases, most people assume that public utilities will appropriately monitor and correct any quality problems if they~occur.

Successfully maintaining adequate power quality and providing sufficient amounts of electrical power to meet the rising needs of consumers has been a triumph of electrical utilities for over 100~years. In~recent times, however, there have been changes to the nature of electrical generation and consumption that make power quality of increasing concern and interest. First, there is a global need to shift to renewable energy sources. Second, traditional approaches to top-down grid power quality monitoring may not be well suited to bottom-up, distributed generation associated with renewable energy sources such as rooftop photovoltaic panels (PV). Third, modern consumer electronics place more stringent demands on power quality. Finally, effective public policy making in this modern context can be aided by increased public access to power quality data reflecting the state of the grid as it is experienced by end-users. Let us examine each of these assertions in more~detail.

\highlight{{\em 1. We need more renewable energy.}} %Is italics necessary?
 Concerns including pollution, environmental degradation, and~climate change have produced a global movement away from centralized, fossil-fuel-based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar. However,~the economic, environmental, and~political advantages of renewable energy come with significant new technical challenges. Wind and solar are intermittent (for example, solar energy cannot be harvested at night) and unpredictable (for example, wind and solar energy fluctuate based upon cloud cover and wind speed). In~addition, renewable energy generation is often distributed throughout the grid (such as in the case of residential rooftop photovoltaic (PV) systems).

One impact of adding renewable energy generation to an electrical grid is that maintaining adequate power quality is much more challenging. This problem is inversely proportional to the size of the electrical grid. For~example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai. (In contrast, the~size of the European electrical grid is approximately 600 GW and the size of the U.S. continental grid is over 1000 GW.) For small grids, the~unpredictable nature of power generation by distributed, intermittent renewables can quickly create problems for grid stability. In~the case of Hawaii, consumer demand for grid-tied rooftop solar exceeded the ability of Hawaiian Electric to manage, resulting first in complex and expensive ``interconnection studies''~\cite{trabish_solar_2014,anastasi_energy_2009}, and~later in the requirement for new installations to include batteries and grid-disconnected operation. Photovoltaics affect power quality by introducing harmonics~\cite{anurangi_effects_2017}. The~power feed-in of PV generation in rural low-voltage grids can influence power quality (PQ) as well as facility operation and reliability~\cite{rita_pinto_impact_2016}. In~fact, a~study of the island of Porto Santo in Portugal found that the intermittent nature of installed photovoltaic and wind energy could result in a potential drop in frequency of 12 Hz, lasting as long as 7 s~\cite{delgado_solutions_2011}.

\highlight{{\em 2. Traditional top-down monitoring is not well-suited to bottom-up energy generation.}} For traditional grid architectures where generation is centralized and under the complete control of the utility, it is common to monitor power quality only to the substation level, because~it can be assumed that the power quality experienced at the substation is a reasonably accurate proxy for the power quality experienced by the 1000 or so end-users serviced by the substation. Furthermore, if~an end-user experiences a power quality problem not experienced by the substation, then it is most likely due to equipment or electrical issues local to that end-user and not a grid-level~problem.

These assumptions might not hold in grids with distributed, intermittent generation by end-users, such as is the case with roof top solar. In~these cases, the~unpredictable nature of generation can lead to local power quality problems that are not experienced by the substation, and~that are not due to any single individual, but~rather the bottom-up power generation architecture of the grid. Solar~panels connected to low voltage networks will result in over-voltages, the~switching frequency of the converters in wind turbines causes high-frequency signals flowing into the grid, and~harmonics are generated by EV chargers.~\cite{zavoda_power_2018}
Nakafuji notes that Hawaiian Electric is contending with PV penetrations in excess of 60\% on certain distribution circuits, for~which traditional rules of thumb for the design of protection and distribution systems might not hold~\cite{nakafuji_back--basics_2011}.

\highlight{{\em 3. Consumer electronics require high quality power.}} The rise of consumer electronics has raised the bar for what constitutes ``adequate'' power quality. Only a few decades ago, computers were a rare presence outside of labs and large institutions. Today, computers are everywhere: embedded in phones, washers, refrigerators, thermostats, and~so forth. These electronic devices not only have higher power quality requirements, but~some of them actually introduce power quality problems in the form of harmonic distortion. Poor power quality can result in electronic devices failing unpredictably, and/or decreasing their lifespan. Reduced input voltage can cause excessive power supply heat dissipation, resulting in reduced mean time between failures (MTBF). In~addition, rectifiers and DC-to-DC converter switching transistors draw high-peak currents, which raise their junction temperatures. These temperature excursions take a toll on semiconductor longevity. High input voltage can also puncture a power supply's rectifier and switching transistor junctions, causing MTBF reduction and eventual breakdown. High-voltage transients lasting microseconds can permanently wreck the power supply and its electronic equipment load.~\cite{dedad_when_2008}.

\highlight{{\em 4. Power quality data should be publicly available.}} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers. For~example, in~Hawaii, utilities are only required to make a ``best effort'' to provide non-harmful voltage and frequency values, and~are only required to publicly report about power outages of more than 3 min. There is no requirement for utilities to report potentially harmful levels of voltage, frequency, THD, or~transients as long as they do not lead to outages. Indeed, in~places like Hawaii, there is not yet even infrastructure in place to enable utilities to collect that information, even if they were asked to report on it. Providing a high quality, third party, publicly available source of power quality data not only benefits the public, it makes it much easier to perform research on ways to improve grid stability in the presence of renewable energy~generation.

Better public understanding of power quality is important because it is such a significant economic problem. Across all business sectors, the~U.S. economy is estimated to be losing between \$104 and \$164 billion a year to outages and another \$15 to \$24 billion to power quality phenomena~\cite{elphick_summary_2015}. India lost more than \$9.6 billion in 2008 due to power quality problems, and~Europe is estimated to be losing \$150 billion per year~\cite{laskar_power_2012}.

Providing power quality data to address the above issues requires solving several difficult technical problems. First, monitoring below the substation level increases the number of required monitoring devices by two to three orders of magnitude. This has significant implications for the cost of monitoring in terms of the devices, and~the sheer amount of low-level power quality data that will be produced. Second, this low-level data must be processed in a manner that does not require inordinate amounts of processing, network bandwidth, or~storage. Finally, collecting the data is not useful if it cannot be converted into actionable information in a reasonable amount of time. This paper describes results to date of the Open Power Quality project, which attempts to address these~problems.

\section{Methods}

The central aim of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power~generation.

Our research methodology consists of three (potentially interleaved) phases. First, we specify, design, and~implement novel hardware and software for collecting, transmitting, and~analyzing four essential measures of power quality data: voltage, frequency, THD, and~transients. Second,~we~verify that our implementation matches our specifications through laboratory experiments. Third, we~validate our implementation by testing a variety of hypotheses through a real-world deployment. In~this paper, we report on results from a pilot deployment at the University of Hawaii over the course of three~months.

Our central hypothesis is that we can accomplish this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two-way communication to a set of cloud-based~services.

We will test this central hypothesis through six subhypotheses: (1) OPQ provides valid and reliable collection of power quality data (Section \ref{hyp:01}), (2) OPQ's triggering system provides advantages with respect to bandwidth and computation (Section \ref{hyp:02}), (3) OPQ enables subthreshold event detection based upon temporal locality (Section \ref{sec:subthreshold-events}), (4) the OPQ information architecture provides a means to produce actionable insights (Section \ref{hyp:04}), (5) the OPQ information architecture provides predictable upper bounds on storage resources (Section \ref{hyp:05}), and~(6) OPQ provides useful adaptive optimization capabilities (Section \ref{sec:adaptive-optimization}).

The remainder of this section presents the overall architecture of OPQ, along with details on the design and implementation of each of the principal architectural components. We will also discuss our verification activities for each component. Our validation and hypothesis testing activities will be described in Section~\ref{sec:pilot-study}.

\subsection{OPQ~Architecture}

The OPQ system architecture consists of four major open source hardware and software components that provide end-to-end support for the capture, triggering, analysis, and~reporting of consumer level local and global PQ events. OPQ Box is a hardware device that detects the electrical waveform from a standard residential outlet and communicates both low and high fidelity representations of the waveform to other OPQ system components either at predefined intervals or upon request. OPQ Makai monitors incoming low fidelity data from OPQ Boxes, requests high fidelity data when necessary, and~stores the results in a MongoDB database. OPQ Mauka analyzes low level data and analyzes it to produce higher-level representations according to the information architecture described below, and~can tell OPQ Makai to request high fidelity data from one or more OPQ Boxes to facilitate analysis. OPQ View is a browser-based visualization platform for displaying the results for data capture and~analysis.

Figure~\ref{fig:architecture} illustrates how these components work together to take information from wall outlets (on the left side) to the display of analyses in a browser (on the right hand side). First, OPQ Boxes analyze power from wall outlets, and~send low fidelity measurements to OPQ Makai. OPQ Makai analyzes low fidelity measurements, and~requests high fidelity waveforms when desirable. Both measurements and waveforms are saved in a MongoDB database. OPQ Mauka analyzes low and high fidelity data, and~creates ``events'' to represent anomalies. OPQ View notifies users of events and allows them to drill down into low and high fidelity~data.

OPQ Makai, OPQ Mauka, and~OPQ View are all cloud-based software services that collectively form a single ``instance'' with respect to data transmission, storage, analysis, and~visualization. We~refer to this collection of software-side components as OPQ Cloud. Every OPQ Box connects to a single instance of an OPQ Cloud. It is possible to have multiple OPQ Cloud instances. For~example, a~company might install an OPQ Cloud instance behind their firewall along with OPQ Boxes to provide a private mechanism for collecting and analyzing power quality~data.

\begin{figure}[H]
\center \includegraphics[width=5in]{images/architecture/system-diagram.png}
\caption{High level system architecture of an Open Power Quality (OPQ) Sensor~Network.}
\label{fig:architecture}
\end{figure}

\subsection{OPQ~Box}
\label{sec:opq-box}

OPQ Box is a hardware device designed to provide inexpensive, extensible and accurate residential power quality measurements.
A block diagram of OPQ Box is shown in Figure~\ref{fig:opq:2}a.
A complete device is shown in Figure~\ref{fig:opq:2}b.

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=0.9\linewidth]{images/opq-box/opqbox_diagram.png}
\caption{OPQ Box Block Diagram.
The power path is in red, signal path is in green and the digital IO is in black.}
\label{fig:opq:1:1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=0.7\linewidth]{images/opq-box/opqbox_photo.jpg}
\caption{OPQ Box in an ABS plastic~enclosure.}
\label{fig:opq:1:2}
\end{subfigure}
\caption{(\textbf{a}) OPQ Box block diagram and (\textbf{b}) production OPQ Box ready for~deployment.}
\label{fig:opq:2}
\end{figure}
\unskip

\subsubsection{OPQ Box~Hardware}\label{subsec:hardware}

The power system of the OPQ Box electrically isolates most of the device from the AC mains power.
An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
A total of \hl{5 V} is used to power the Raspberry Pi, the~equipment connected to the expansion port, the~3.3 V regulators and voltage reference, and~an isolated DC/DC converter.
A total of 3.3 V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
The~hot side of the isolation amplifier is powered from the isolated DC/DC converter.
This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power~outage.

%%% We added a space between 5 and V. Please confirm.

A Raspberry Pi single board computer (SBC) is responsible for signal analysis and anomaly detection.
The Raspberry Pi model used in OPQ Box is the Pi Zero W equipped with 256 MB of main memory and a single core 1 GHz ARM11 CPU. It also contains an on-board 802.11n WIFI transceiver, which removes the need for an external WIFI~dongle.

\subsubsection{OPQ Box~Software}\label{subsec:software}

The software stack of the Raspberry Pi aims to deliver an accurate and precise power quality analysis framework despite the rather limited capabilities of the hardware.
A block diagram of the software stack is shown in Figure~\ref{fig:opq:3}.
Digital data is transferred from the DSP to the Raspberry Pi via Serial Peripheral Interface, with~the Pi acting as the master and the DSP as a slave device.
A hardware interrupt line is used to inform Pi software that the DSP is ready for the data transfer, and~a kernel driver provides management of the SPI bus.
Internally, the~OPQ driver maintains a ring buffer of 16 windows, each of which is 200 data samples in size.
Upon receiving the interruption for the DSP, the~CPU sets up the DMA transfer and the DMA engine transfers a 200 sample window into the kernel memory without CPU interaction.
This scheme requires the CPU to only service 60 interruptions a second, with~each interruption requiring on the order of 100 instructions, for~a CPU utilization of less than $1\%$ in normal operation.
Userland applications communicate with the kernel driver using a file descriptor, where every \highlight{$read$} %Is italics necessary here?
 system call yields 200 samples of raw waveform.
As a result, the~smallest window that a userland application may process is a single AC cycle of the grid~mains.

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.9\textwidth]{images/opq-box/opqbox_software.png}
\end{center}
\caption{Block diagram of the OPQ Box 2 software~stack.}
\label{fig:opq:3}
\end{figure}

The userland component of the OPQ Box software is a multi-threaded extensible analysis framework called Triggering.
The reader thread is responsible for transferring and accumulating data from the kernel driver.
The smallest data buffer that the Triggering application processes at any given time is 10 grid cycles or 2k samples.
Once the cycles are transferred to the userland and timestamped, they are passed to the analysis thread for feature extraction, as~well as to the Raw Data Ring Buffer (RDRB).
Since internally all data is addressed using shared pointers, during~data duplication no copying is required.
RDRS is capable of buffering up to an hour of data before it is overwritten, resulting in the RDBS maximum size of~100 MB.

The analysis thread of the Triggering application performs feature extraction of the raw data windows of 2000 samples.
Four metrics are extracted from the data stream: (1) Fundamental frequency, (2) RMS Voltage, (3) Total Harmonic Distortion, and~(4) Transients. Let us briefly discuss how each of these are~computed.

\subsubsection{OPQ Box: Calculating Fundamental~Frequency}\label{subsec:fundamental-frequency}

The fundamental frequency is calculated by computing the zero crossings of the AC waveform.
In order to improve the accuracy of the frequency calculation one must first filter out as much noise as possible.
Since our sampling rate is quite high (12 kSps) and the fundamental frequency is quite low (60~Hz), it is very computationally expensive to perform this filtering in a single step.
Instead, filtering is accomplished via a set of two low pass finite impulse response (FIR) filters.
The first filter has a passband of 0--600 Hz allowing us to downsample the waveform to 1200 samples per second.
The next~filter has a passband of 0--100 Hz allowing for further removal of high frequency noise.
Finally, zero crossings are extracted and used for the frequency calculation.
The zero crossings themselves were calculated by using linear interpolation between two points that bracket the time~axis.

\subsubsection{OPQ Box: Calculating RMS~Voltage}\label{subsec:root-mean-square-voltage}

Root mean square voltage ($V_{rms}$) in electrical power is the equivalent value of DC voltage, which would dissipate the same power in the resistive load. $V_{rms}$ is a convenient measure for detecting voltage sags and swells, since they result in nominally higher and lower computed~value.

Similarly to the frequency calculation, OPQ Box uses a 10 cycle window for a single $V_{rms}$ calculation. Unlike the frequency calculation, the~input is not filtered a~priori.

\subsubsection{OPQ Box: Calculating~THD}\label{subsec:thd}

The OPQ Box calculates total harmonic distortion (THD) using the industry standard methodology.
It should be noted that in the power quality domain THD is expressed as a percentage as opposed to $\frac{dB}{\sqrt{Hz}}$ as used in other disciplines.
Operationally, OPQ Box computes THD for 10 cycles of the fundamental frequency.
First an FFT transforms the real voltage samples into its frequency components.
Next, the~square of the harmonic bins are accumulated and scaled by the magnitude of the fundamental~power.

\subsubsection{OPQ Box: Transient~Detection}\label{subsec:transient-detection}

OPQ Box transient detection is performed via filtering out of the fundamental frequency via an FIR high pass filter with a cutoff frequency of 400 Hz and searching for a maximum value in the~remainder.

It should be noted that this transient detection method is susceptible to THD fluctuations, since any harmonic above 400 Hz will remain in the filtered waveform.
However, since the THD information is transmitted along with the transient detection metric, they can be correlated in downstream transient~detection.

\subsubsection{OPQ Box: Network~Communication}\label{subsec:network-communication}

The computed fundamental frequency and $V_{rms}$ are transmitted to the Makai service for aggregation.
Data transmission is handled using ZeroMq software stack with Curve25519 elliptic curve encryption.
Each device holds a unique private and public key, as~well as the servers' public key, allowing both the Makai service and the OPQ Box to verify its peer.
Internally, metrics transmission uses ZeroMq's PUB/SUB protocol.
This protocol is a publish/subscribe topology, with~each message containing the topic and a payload.
Additionally, ZeroMq allows for multiple sub peers with subscriptions forwarded to the publisher automatically via a side channel.
This allows for the aggregation service to be spread across multiple nodes, with~minimal network~overhead.

If the aggregation service determines that an anomaly has occurred, it is able to request raw waveform from the OPQ Box RDRB via a separate ZeroMq pub sub channel.
If the RDRB buffer contains data for the requested temporal range, OPQ Box transmits the available data to the aggregation service via a push pull ZeroMq channel.
Protobuf message serialization is used to encode messages across the OPQ~ecosystem.

In order to make a distributed measurement, all of the OPQ Boxes on the OPQ network need to maintain an accurate time reference.
Time synchronization across multiple OPQ Boxes is accomplished using the Network Time Protocol.
The expansion port of the OPQ Box supports a GPS receiver.
However, since GPS receivers require line of sight to the sky, it was not used for deployment.
NTP performance has been verified against GPS resulting in time error of 8 $\pm$ 5 ms, which is typical for NTP running over the Internet with a nearby NTP~server.

\subsubsection{OPQ Box: Manufacturing}

Currently, there is no mechanism for mass production of OPQ Boxes, but~all of the plans are available under an Open Source license, so interested organizations with some basic hardware engineering skills can build their own~boxes.

Complete specifications for the OPQ Box hardware, firmware, and~software are available~\cite{negrashov_opq_2020}. As~of the time of writing, a~single OPQ Box can be manufactured for approximately \$100 in parts. The~cost drops significantly with scale, for~example, 100 OPQ Boxes can be manufactured for a cost of approximately \$75 in parts per~device.

\subsection{OPQ~Makai}
\label{sec:opq-makai}

OPQ Box provides an inexpensive hardware device for collecting four important power quality measures with high fidelity, but~realizing its potential requires an innovative approach involving two-way communication between OPQ Boxes and the OPQ cloud-based services. To~see why, consider the IEEE 1159 standard for single location power quality monitoring~\cite{unruh_ieee_2018}. For~transient monitoring, IEEE 1159 suggests a sampling rate of at least 7680 samples/second, up~to 1 Megasample/second. This~implies that if the cloud service requires the high fidelity data from all OPQ Boxes, it would incur a very large bandwidth cost. At~20 Ksamples/second with 16bit samples, a~single OPQ Box will generate 300 Kb/s of network bandwidth. Several thousand devices would easily saturate a 1~GB network link. In~addition, collecting and recording all of the raw waveform data from residential power quality meters could lead to security and privacy~issues.

Network bandwidth saturation is a common problem for distributed sensor networks, and~a common solution is called ``self-triggering''. In~this approach, each monitoring device is configured with a threshold value for one or more measures of interest. When the threshold for a measure of interest is exceeded, then and only then is data sent over the network to cloud-based services for~analysis.

The problem with the self-triggering approach is that grid-wide power quality events do not affect the entire grid in the same way. For~example, due to the grid’s hierarchical structure, a~voltage sag on one sub-circuit can manifest as a sag of a different magnitude or even a swell on another~\cite{kahle_power_2015}. This may result in a situation where some of the monitoring devices will not consider a power quality anomaly as an event, because~it did not surpass the metric threshold, and~simply ignore it. From~an analysis perspective, however, it can be useful to get raw data from all of the affected devices, not just the ones that were affected to the point where the box was triggered. This additional information can be used to localize the disturbance, as~well as better evaluate its~impact.

Since sending all the data is infeasible, and~since the self-triggering approach can potentially miss important forms of information, OPQ implements a novel, hybrid centralized/decentralized data acquisition scheme, which involves two-way communication between the OPQ Boxes and a cloud service called OPQ Makai. In~this scheme, OPQ Boxes use local processing resources to feature extract the incoming waveforms while storing them locally for an hour. Each OPQ Box sends its feature data to OPQ Makai once a second, which we called the ``triggering stream''. Feature data is very small, on~the order of a few kilobytes, and~so this approach allows the sensor network to scale to thousands of devices with acceptable network bandwidth requirements. OPQ Makai processes the incoming triggering stream and looks for anomalies. If~an anomaly is present in only a single device, it is highly probable that the cause is local and not grid-level. On~the other hand, if~the triggering stream shows an anomaly temporally collocated across multiple devices, the~entire network or a subset of the network may be queried for raw waveform data for a temporal region which corresponds to the disturbance in the triggering~stream.

Our pilot study, discussed in Section~\ref{sec:pilot-study} will provide examples of the novel analysis capabilities made possible by OPQ Box and OPQ Makai communication. In~general, here are the main advantages of our hybrid centralized/decentralized approach over traditional self-triggering and the ``naive'' approach of sending all of the data:

\highlight{{\em Bandwidth usage is minimized.}} % Is italics necessary?
 Instead of sending the entirety of raw data, only extracted features are sent. This results in a tiny fraction of the bandwidth requirement when compared to raw waveforms. Furthermore, the~temporal window, which encompasses a single feature, can be adjusted in real time. Thus, as~soon as an anomalous behavior is observed in a subset of sensors, this window can be adjusted for a finer grained feature~extraction.

\highlight{{\em Effects of latency are minimized.}} In this case, ``latency'' refers to the time required for OPQ Makai to process the incoming feature stream and decide whether to request high fidelity data from one or more OPQ Boxes. Even at 1 M samples/second at 16 bits of resolution, the~memory requirement to store 5 min of raw waveform without compression are on the order of 512 MB, which is well within the realm of inexpensive single board computers such as Raspberry PI. With~compression specifically suited to the signal of interest, the~memory requirement can be reduced even further. In~the case of the OPQ sensor network, OPQ Makai has an hour to process feature data and request high fidelity data from OPQ Boxes. During~our pilot study, OPQ Makai always responded within a second or~two.

\highlight{{\em Cloud processing requirements are reduced.}} Since feature extraction is already performed at the device level, cloud computation requirements are reduced. With~the advent of the Internet of Things, the~computational capacity of edge devices is~increasing.

\highlight{{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.}} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes. This allows analysis of data from devices that were only mildly affected or even not affected at all by the disturbance. This creates new possibilities for investigation of disturbance propagation across the sensed area, as~will be discussed in Section~\ref{sec:pilot-study}.

\highlight{{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.}} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features. This allows for a simple robust algorithm, which may be deployed at the sink node for anomaly~detection.

\subsubsection*{OPQ Makai: Design}

OPQ Makai is a distributed extensible microservice framework responsible for receiving the triggering stream from the OPQ Boxes, locating anomalous temporal regions and requesting raw waveform for the anomalous time ranges. As~shown in Figure~\ref{fig:makai-design}, Makai consists of four major components: Acquisition Broker, Triggering Broker, Event Service and the Acquisition~Service.

\begin{figure}[H]
\center \includegraphics[width=3in]{images/makai/makai_main.pdf}
\caption{Makai component~design.}
\label{fig:makai-design}
\end{figure}

The Triggering Broker is perhaps the simplest component of the OPQ Makai system. The~triggering stream generated by the OPQ Boxes is encrypted to preserve user privacy. In~order to minimize the CPU time spent decrypting the data across multiple OPQ services, the~Triggering Broker decrypts the data and sends clear text measurements to other OPQ cloud services. The~Triggering Broker uses the ZeroMq subscribe socket to receive data from OPQ Boxes, and~sends it via a publish socket to any connected client. Each publish message is published to a topic, which corresponds to the ASCII representation of the originating OPQ Box ID. This allows services that utilize the Triggering Broker to select a subset of IDs to operate on. This is useful for load balancing the backend services, or~dividing the OPQ network into separate regions with no electrical connections between~them.

The Acquisition Broker manages the two-way communication between the OPQ Boxes and the rest of the cloud infrastructure. Unlike the triggering stream, which originates from the OPQ Box, two-way communication is always initiated by OPQ cloud services. Two way communication is realized via a command response interface, where the OPQ service initiates the communication by sending a clear text command to the Acquisition Broker, which then forwards it in encrypted form to the appropriate OPQ~Boxes.

The Acquisition Service resides between the Triggering and Acquisition Brokers. The~Acquisition Service is responsible for three tasks:
(1) Computation of statistics of the incoming triggering stream; (2) Hosting plugins for triggering stream analysis; and (3) Generating data event requests for OPQ Boxes. The~Acquisition Service accesses the triggering stream by connecting to the publish socket of the Triggering broker. Since the connection is managed through the ZeroMq publish-subscribe socket, several Acquisition Services can be connected to a single Triggering broker endpoint, each servicing a subset of OPQ Boxes by subscribing to only specific devices. The~Acquisition Service does not include any analysis capabilities by default. Instead, analysis is performed by shared library loadable plugins. These plugins can be loaded and unloaded at runtime, thus allowing live upgrading and testing of new analysis~methods.

The Event service is a microservice that stores raw data generated by OPQ Boxes in the MongoDB Database. On~initialization, the~Event service queries MongoDB database for the highest event number recorded so far, connects to the Acquisition Broker’s publish port, and~subscribes to all messages that start with the prefix ``data”. This allows the Event service to capture every response from OPQ Boxes generated from commands issued by the Acquisition service plugins. Once the Event service receives a data response with an identity containing an event token it has not seen before, it will increment the event number, and~store it in an internal key value~store.

\subsection{OPQ~Mauka}
\label{sec:opq-mauka}

The previous sections discussed the design of OPQ Box, a~custom hardware device for collecting four important measures of power quality, and~OPQ Makai, a~novel, hybrid centralized/decentralized data acquisition scheme, which involves two-way communication between the OPQ Boxes. As~a result of these two innovations, an~OPQ sensor network has the ability to collect and analyze high fidelity, low level data about power quality anomalies in a cost-effective, scalable~fashion.

There are remaining challenges to creating a useful power quality sensor network. First, the~data provided by OPQ Boxes is low-level, ``primitive'' data consisting of either features (i.e., frequency, voltage, THD, and~transients) or waveform data. However,~what we actually want is actionable insights into grid stability. For~example, we might want to know if a given anomalous data value is actually detrimental, or~we might want to be able to predict when a power quality event might occur in the future based upon the recognition of cyclical events in the historical~data.

A second challenge involves the potentially high volume of data that might accumulate in the cloud. Although~OPQ Box and OPQ Makai provide a scalable mechanism for communicating power quality data to the cloud services, it is still the case that, over~time, a~substantial amount of data could accumulate. One strategy is to simply store all of the data sent to the cloud forever. This means that data storage requirements will increase monotonically over time, making the sensor network more costly to maintain the longer it is in place. An~alternative strategy is to implement an algorithm to identify uninteresting (or no longer interesting) data and discard it. Ideally, such an algorithm would enable OPQ sensor network designers to calculate an upper bound on the total amount of cloud storage required as a function of the number of nodes (OPQ Boxes) in the~network.

OPQ Mauka addresses both of these issues. First, OPQ Mauka provides a multi-layered representation for structuring and processing DSN data. The~structure and processing at each layer is designed with the explicit goal of turning low-level data into actionable insights. Second, each layer in the framework implements a ``time-to-live'' (TTL) strategy for data within the level. This~strategy states that data must either progress upwards through the layers towards more abstract, useful representations within a fixed time window, or~else it can be discarded. The~TTL strategy is useful because when implemented, it allows DSN designers to make reasonable predictions of the upper bounds on data storage at each level of the framework adjusting for the number of sensors and power anomaly~probability.

TTL also makes possible a ``graceful degradation'' of system performance if those bounds turn out to be exceeded. For~example, consider a situation in which a power network enters a prolonged period of widespread power quality instability, where every OPQ Box is reporting continuous anomalous conditions with respect to voltage, frequency, THD, and~transients. This ``worst case scenario'' would lead to the potential situation of every OPQ Box trying to upload raw waveform data all the time. The~TTL system provides safeguards, in~that whatever low-level data has not been processed relatively quickly can be discarded. Thus, instead of the system potentially going down entirely, it could instead continue to operate at a reduced~capacity.

Figure~\ref{fig:mauka-data-model} illustrates the hierarchical data model for OPQ Mauka. This data model can be conceptualized as a multi-level hierarchy that adaptively optimizes data storage using a tiered TTL approach and provides a mechanism in which typed aggregated data is continually refined to the point of becoming actionable. The~data model also includes software components called ``actors'' that both move data upward through the levels and also apply optimizations downward through the levels. Actors are implemented through a plugin architecture, making it easy to experiment with the data model and improve it over~time.

The lowest layer of the hierarchy is the Instantaneous Measurements Layer (IML). The~IML contains “raw” data, in~other words, the~digitized waveform. IML data exists both on each OPQ Box (where it is available for up to the previous 60 min). It also exists in the cloud, in~the event that OPQ's triggering mechanism has caused a temporal interval of waveform data to be uploaded. IML data in the cloud has a TTL of 15 min: unless the waveform data is found to be useful by a cloud service within 15 min, it can be~discarded.

\begin{figure}[H]
\center \includegraphics[width=4in]{images/mauka/mauka-data-model.png}
\caption{Mauka data model~hierarchy.}
\label{fig:mauka-data-model}
\end{figure}

The second layer is the Aggregate Measurements Level (AML). The~AML stores power quality summary statistics sent either once per second or once per minute by each OPQ Box. These~summary statistics include the maximum, minimum, and~average values of voltage, frequency, THD, and~transient metrics over the time window. It is AML data that is used to initiate the triggering process of uploading IML data from the~cloud.

The third layer is the Detections Level (DL). This layer is responsible for processing the IML and AML data to produce a representation of an ``event'' with a precisely defined start and end time based upon examination of the waveform. As~will be discussed in Section~\ref{sec:subthreshold-events}, knowledge of the start and end time of a power quality anomaly allows investigation of how that anomaly might manifest itself elsewhere in the grid, even if this manifestation is not severe enough to produce over-threshold data~values.

The fourth layer is the Incident Level (IL). This layer starts to answer the question of whether the data is ``actionable'' by classifying the detected event according to various industry standards for power quality anomalies: IEEE 1159, ITIC, SEMI F47, and~so forth. For~example, an~event that falls into the ITIC ``prohibited'' region clearly indicates a power quality anomaly that requires further study and~intervention.

The fifth and final level is the Phenomena Level (PL). This layer contains the results of analyses that attempt to identify cyclic, and~thus predictable, power quality disturbances. It also contains analysis results regarding the similarity of various incidents, which can help uncover causal factors. Finally, it provides analyses for adaptive optimization of the OPQ Sensor Network. These optimizations can change the thresholds for individual boxes to either increase their sensitivity or decrease their sensitivity over specific intervals of time. The~ultimate goal of adaptive optimization is to help the network learn to acquire all of the data useful for analyses, and~only the data useful for analyses. We~are still in the early stages of exploring the potential of adaptive optimization in OPQ~networks.

\subsubsection*{OPQ Mauka: Actors}

The current capabilities of OPQ Mauka can be summarized in terms of its Actors, which are implemented as plugins. Nine of the most important Actors are described~below.

\highlight{{\em Makai Event Actor.}} % Please confirm if this is necessary.
 The Makai Event Actor is responsible for reading data newly created by OPQ Makai into OPQ Mauka. It performs feature extraction on the raw data stream and forwards those features (or the raw data) to subscribing Actor plugins. This allows OPQ Mauka to perform feature extraction once, and~allow use of those features by multiple~Actors.

\highlight{{\em Frequency Variation Actor.}} The Frequency Variation Actor classifies generic frequency sags, swells, and~interruptions as defined by the IEEE 1159 standard. Both duration and deviation from nominal are used to perform these classifications. Duration classifications include frequency deviations that last for less than 50 ns, between~50 ns to 1 ms, and~1 to 50 ms. Classifications for deviations from nominal are performed for values that are up to 40\% deviation from nominal. This Actor is able to classify frequency swells, frequency interruptions, and~frequency sags, leading to the creation of data at the Incident~Layer.

\highlight{{\em IEEE 1159 Voltage Actor.}} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard \hl{[29]}. In~general, this standard classifies voltage disturbances by duration and by magnitude. Voltage durations are classified from 0.5 to 30 cycles, 30 cycles to 3~s, 3 s to a minute, and~greater than 1 minute. Voltage deviations are classified in both the sag and swell directions as a percentage from nominal. Sags are generally classified between 10\% and 90\% of nominal while swells are generally classified from 110\% to 180\% of nominal. This Actor is capable of classifying voltage sags, swells, and~interruptions as defined by the standard, and~creating data at the Incident Layer if~appropriate.
%%% Should it be reference citation? 
\highlight{{\em Box Optimization Actor.}} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka. This Actor is capable of requesting the state of each OPQ Box (e.g., uptime, Measurement rate, security keys). It is also capable of adjusting the state of individual OPQ Boxes by changing things such as the Measurement and Trend rate or the sampling rate used by the~Box.

\highlight{{\em Future Phenomena Actor.}} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena. These Phenomena are used to predict Events and Incidents that may occur in the future. This plugin does not subscribe to any messages, but~instead utilizes timers to perform its work. By~default, this plugin runs every 10~min.

When a Future Phenomena Actor runs, it loads any active Periodic Phenomena found in the database. If~Periodic Phenomena are found, this Actor extrapolates possible Detection and Incident Layer data by first examining their timestamps and then extrapolating into the future using the mean period and the standard deviation. For~each timestamp in a Periodic Phenomena, the~mean period is added. If~the resulting timestamp is in the future, a~Future Phenomena is created using the time range of the future timestamp plus or minus the standard deviation of the Periodic~Phenomena.

When a Future Phenomena is created, timers are started in a separate thread signifying the start and end timestamps of the Future Phenomena. When the first timer runs, messages are sent to the Box Optimization Actor and the Threshold Optimization Actor instructing OPQ Box thresholds to be set lower and measurement rates to be set higher. This increases the chance of seeing an anomaly over the predicted time window. When the second timer runs, these values are reset to their default values. Thus, the~plugin increases fidelity and decreases thresholds over the period of a Future~Phenomena.

{\em ITIC Actor.} The ITIC Actor analyzes voltage to determine where it falls within the ITIC curve~\cite{thallam_power_2000}. The~ITIC curve is a power acceptability curve that plots time on the x-axis and voltage on the y-axis. The~purpose of the curve is to provide a tolerance envelope for single-phase 120V equipment. The~curve defines three regions. The~first region is ``No Interruption'' and generally includes all voltages with very short sustained durations. All events within this region have no noticeable effect on power equipment. The~second region, the~''No Damage Region'', occurs during voltage sags for extended periods of time. Power Events in this region may cause equipment interruptions, but~it will not damage the equipment. The~final region, the~''Prohibited'' region, is caused by sustained voltage swells and may cause damage to power equipment. This Actor determines if an event falls within the ``No Damage'' or ``Prohibited'' regions and if so, creates an Incident to record~this.

{\em SEMI F47 Actor.} The SEMI F47 Actor is similar to the ITIC Actor in that it plots voltage and duration against a power acceptability curve. In~this case, the~standard used is the SEMI F47 standard~\cite{djokic_sensitivity_2005}. Rather than using a point-in-polygon approach, this plugin reads the voltage features sequentially and uses a state machine to keep track of the current classification. This plugin only classifies values as a ``violation'' or as~''nominal''.

{\em Transient Actor.} The Transient Actor is responsible for classifying frequency transients in power waveforms. The~plugin subscribes to messages from a topic that contains a calibrated power waveform payload. The~Transient Actor is capable of classifying impulsive, arcing, oscillatory, and~periodic notching transients. A~decision tree is utilized to select the most likely transient type and then further analysis is used to perform the actual classification of transients. Dickens et al.~\cite{dickens_transient_2019} provide more details on the transient classification system used by this~Actor.

{\em Periodicity Actor.} The Periodicity Actor is responsible for detecting periodic signals in power data. This Actor does not subscribe to any messages, but~instead runs off of a configurable timer. The~Actor is set to run by default once an hour and every hour it scrapes the last 24 h worth of data and attempts to find periods in the Measurements over that~duration.

For each feature in the Measurement and Trend data (e.g., frequency, voltage, and~THD), the~Periodicity Actor first removes the DC offset from the data by subtracting the mean. Next, the~Actor filters the signal using a fourth order high-pass filter to filter out noise. The~Actor then performs autocorrelation on the signal followed by finding the peaks of the autocorrelation. The~mean distance between the peaks of the autocorrelation provides the period of the~signal.

The Periodicity Actor only classifies data as periodic if at least three peaks were found and the standard deviation of the period is less than 600 s (10 min). Once a positive identification has been made, peak detection is performed on the original signal. Once the plugin has the timestamps and deviations from nominal of the periodic signal of interest, the~plugin can group Measurements, Trends, Detection Layer Events, and~Incidents that were created during the periodic signals together as part of the Periodic~Phenomena.

\subsection{OPQ~View}
\label{sec:opq-view}

The final component of the OPQ Sensor Network system architecture is called OPQ View. It is a web application, implemented using the Meteor application framework, which provides a variety of visualization and query services. An~example of the OPQ View home page is provided in Figure~\ref{fig:opq-view-home}.

\begin{figure}[H]
\center \includegraphics[width=5in]{images/view/homepage.png}
\caption{OPQ View: home~page.}
\label{fig:opq-view-home}
\end{figure}

The home page provides three components: a ``System Stats'' window, which indicates the number of data elements currently at each level of the OPQ Mauka data hierarchy, a~''System Health'' window that indicates whether the Cloud services appear to be running correctly and the status of communication with all known OPQ Boxes on the sensor network, and~a ``Box Trends'' visualization that provides the ability to quickly see trends in the four basic measures (Voltage, Frequency, THD, and~Transients) over~time.

Figure~\ref{fig:opq-view-box-map} shows a map-based view of the sensor network. This figure shows the location of 11 OPQ Boxes on the University of Hawaii campus at one point during Fall of 2019. Depending on the zoom level of the interface, some boxes are collapsed into disks with a number indicating the number of boxes at that location. Zooming in reveals more information about the box including near-real time values for voltage and frequency, as~is shown for the box in the upper right corner of the~figure.

\begin{figure}[H]
\center \includegraphics[width=5in]{images/view/boxmap-2.png}
\caption{OPQ View: Box~map.}
\label{fig:opq-view-box-map}
\end{figure}

Figure~\ref{fig:opq-view-incident-summary} shows a visual display of a single Incident. This view includes a map-based location of the box that was involved in the Incident, the~start and end time and duration of the Incident, the~classification(s) of the anomalous data, and~the waveform associated with the Incident (when applicable). If~additional analysis is desired, the~raw data can be downloaded in CSV~format.

\begin{figure}[H]
\center \includegraphics[width=5in]{images/view/incident-summary.png}
\caption{OPQ View: Incident~summary.}
\label{fig:opq-view-incident-summary}
\end{figure}

The above images provide a sense for how OPQ View helps users to understand, monitor, and~assess an OPQ Sensor Network and the underlying power quality of the grid it is attached to. There are several features in OPQ View for user and box management; for details, please see the OPQ View~documentation.

While OPQ View provides a variety of useful visualization and analysis features, users wishing to understand the power quality of their grid are not restricted to its current feature set. The~OPQ database schema is public, and~if a new analysis is desired, it is straightforward to query the database directly for the data of interest and build a new analysis off of~it.

\section{Related~Work}
\label{sec:related-work}

This section explains how OPQ fits into current industry solutions as well as academic research on power quality monitoring and analysis. For~the purposes of this review, we exclude utility-side power quality monitoring and analysis~systems.

\subsection{Power Quality~Hardware}
\label{sec:commercial-pq-devices}

There exists a very wide variety of power quality hardware devices, including those made by Fluke~\cite{fluke_fluke_2020}, Dranetz~\cite{dranetz_dranetz_2020}, Elspec~\cite{elspec_elspec_2020}, PowerSide (formerly Power Standards Lab) \cite{powerside_powerside_2020}, ACR Systems~\cite{acr_acr_2020}, and~OpenZMeter~\cite{viciana_openzmeter_2018}.

All of the above devices contrast with the OPQ Box in similar ways. First, all of them collect a wider variety of power quality measures than OPQ Box, and~most have been certified according to one or more industry standards. Except~for OpenZMeter, they are generally designed to support industrial applications, where the goal is to ensure that the power being supplied to a building or plant is of adequate quality, and/or that the machinery in the plant is not degrading power quality. Apart from the PowerWatch monitor, all of them are attached to electrical mains using current transformers. Finally, all of them are designed for “stand alone” operation: each device can independently gather and assess power~data.

While the OPQ Box has much more limited functionality, it is designed to be manufactured for approximately \$75, which is 10 to 100 times less expensive than most commercial devices, and~similar is cost to OpenZMeter. The~most important distinguishing feature of OPQ Box is that it is designed to grid-level, not single point monitoring, and~thus incorporates features (such as two-way communication with the cloud, and~subthreshold triggering) that are not present in devices intended for ``stand alone''~capabilities.

\subsection{Power Quality~Software}
\label{sec:commercial-pq-software}

PQView~\cite{electrotek_concepts_pqview_2020}, PQSCADA Sapphire~\cite{elspec_ltd_pqscada_2016}, PQDIF~\cite{sabin_ieee_2020}, and~Grid Protection Alliance~\cite{grid_protection_alliance_grid_2020} are examples of software and/or software standards for manipulating power quality~data.

The differences between the way OPQ and the above systems store and manipulate power quality data arise from fundamentally different architectural assumptions and the historical background of the technology. PQView and PQSCADA Sapphire are designed to operate in a technology environment consisting of a large number of installed, “stand alone” power quality monitors built by different vendors. Their goal is to aggregate the data collected by these devices, and~in order to do so, they depend upon the PQDif standard as a way to obtain power quality data independent of the vendor and device generating it. This results in a kind of “store and forward” process: power quality data is captured and stored on the device, and~then periodically bundled into a PQDif file and sent to the database~software.

OPQ Cloud, on~the other hand, is designed only to support the capabilities of OPQ Boxes. OPQ Boxes, furthermore, have no “stand-alone” capability; they maintain continuous connection to the Internet and upload power quality data to cloud-based services as needed. This means that OPQ implements a very different approach to representing and transmitting data than PQDif. For~details on the representation, see the OPQ Data Model, and~for details on communication, see the OPQ Protobuf~protocol.

\subsection{Research~Systems}

Di Manno et al.~\cite{di_manno_user_2015} describes a PQ monitoring system called PiKu. Unlike OPQ, PiKu is designed as a hardware device for sensing power quality that is directly integrated into a PC. Systems with similar architectures include TRANSIENTMETER, described in Da Ponte et al.~\cite{daponte_transientmeter:_2000}, BK-ELCOM, described in Bilik et al.~\cite{bilik_modular_2007}, and~a system described in Xu et al.~\cite{xu_distributed_2012}.

There are also research projects based upon leveraging existing monitoring infrastructure. \mbox{Suslov et al.~\cite{suslov_distributed_2014}} describe a distributed power quality monitoring system based upon existing phasor measurement units installed by utilities. Sayied et al.~\cite{sayied_power_2013} describe a system designed using existing smart meters. Kucuk et al.~\cite{kucuk_extensible_2010} describe a similar system for the Turkish National Grid using utility grid monitoring~infrastructure.

Mohsenian-Rad~et~al.~\cite{mohsenian-rad_distribution_2018} designed the $\mu$PMU (phase measurement unit) system, which provides distributed power quality measurements over power grid distribution systems. The~$\mu$PMUs in conjunction with their backend software provide two types of analytics. Descriptive analytics provide information about the types and classifications of power quality issues that are observed within the power distribution grid. Predictive analytics are used to predict future power quality issues. The~authors describe their system as providing the ground for enabling future prescriptive analytics, which is the idea of self-tuning the DSN to prepare for future power quality problems by using a combination of descriptive and predictive analytics. In~contrast, as~we will discuss in Section~\ref{sec:adaptive-optimization}, the~OPQ sensor network already supports self-tuning through adaptive~optimization.

One research system very similar in spirit to OPQ is FNET~\cite{liu_distribution_2017}. Like OPQ, the~FNET system consists of custom hardware that monitors the electrical signal from a wall outlet, and~uploads data to the cloud for further processing. Unlike OPQ, FNET is designed for monitoring of frequency disturbances, how they propagate across wide area (i.e., nation-wide) utility grids, and, if~possible, where the frequency disturbance originated. This means that FNET devices must be synchronized using GPS, and~that the data collected consists of frequency and voltage angle. OPQ is designed for more “local” grid analysis, and~we are not interested in propagation. As~a result, OPQ Boxes are synchronized using NTP rather than GPS, which reduces cost and simplifies installation (OPQ Boxes do not need line of site to a GPS satellite). Finally, FNET hardware appears to support only “one way” communication from device to the cloud, while OPQ Boxes support ``two way'' communication (from box to cloud, and~from cloud to box).

\section{Results and~Discussion}
\label{sec:pilot-study}

To evaluate the capabilities of the OPQ Sensor Network, we deployed 16 OPQ Boxes at the University of Hawaii Manoa campus over the course of three months in the Fall of 2019. The~University of Hawaii campus is an isolated microgrid connected to the Oahu powergrid only via a single 46 kV feeder. The~UH Campus also has commercial electrical meters (a mixture of GE PQMII and GE EPM 7000) deployed across various levels of the power delivery infrastructure. While the primary purpose of these meters is to monitor power consumption, they do include power quality monitoring capabilities. Data from these meters were used as ground truth for validation studies of the OPQ sensor~network.

Our pilot study is significant because much of the literature on power quality assessment relies on models, not actual installations~\cite{anurangi_effects_2017,bayindir_effects_2016,farhoodnea_power_2012,shafiullah_experimental_2014}. In~other cases, data was collected from only one location or for a very short time span~\cite{kucuk_assessment_2013,viciana_openzmeter_2018}.

\subsection{Descriptive~Statistics}

The pilot study started on 7 October 2019 and ended on 4 February 2020. We deployed 16 OPQ Boxes across campus. As~noted above, each OPQ Box collects 200 measurements per grid cycle, for~a total of approximately 1B raw measurements per day per box. Values for the maximum and minimum voltage, frequency, and~THD, along with the presence or absence of transients, is sent once a second to the cloud by each box. Thus, each box sends approximately 86,400 measures per day to the cloud at a minimum. Over~the course of the pilot study, a~total of approximately 116M aggregate measures were sent by all of the OPQ~Boxes.

Figure~\ref{fig:statistics} provides two additional sets of summary statistics regarding the pilot study. Figure 9a shows the number of anomalous events (i.e., where threshold values for frequency, voltage, THD, or~transients were exceeded) along with the number of OPQ Boxes that experienced the power quality anomaly during that same time period. So, for~example, 170,925 power quality anomalies were experienced by only one of the 16 OPQ boxes, and~463 anomalous power quality measurements were experienced by all 16 OPQ~Boxes.

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
\begin{tabularx}{\textwidth}{XXXX|}
\toprule
\textbf{Events} & \textbf{Boxes} & \textbf{Events} & \textbf{Boxes} \\
\midrule
170,925 & 1 & 203 & 9 \\
1654 & 2 & 160 & 10 \\
1109 & 3 & 162 & 11 \\
853 & 4 & 130 & 12 \\
593 & 5 & 169 & 13 \\
416 & 6 & 210 & 14 \\
354 & 7 & 477 & 15 \\
246 & 8 & 463 & 16 \\
& & & \\
\bottomrule
\end{tabularx}
\caption{Summary statistics: Events}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Incident} & \textbf{Total} \\
\midrule
Frequency Swell & 291,235 \\
Frequency Sag & 244,286 \\
Excessive THD & 21,395 \\
Voltage Sag & 620 \\
ITIC (No Damage) & 93 \\
SEMI F47 (Violation) & 24 \\
Voltage Interruption & 16 \\
Frequency Interruption & 14 \\
Voltage Swell & 8 \\
\bottomrule
\end{tabularx}
\caption{Summary statistics: Incidents}
\end{subfigure}
\caption{Summary statistics of (\textbf{a}) boxes involved in events and (\textbf{b}) incident type~occurrences.}
\label{fig:statistics}
\end{figure}
\unskip

\subsection{OPQ Provides Valid and Reliable Collection of Power Quality~Data}
\label{hyp:01}

The first goal of the pilot study was to assess how well an OPQ sensor network is able to collect basic power quality data. To~do this, we compared data on voltage and frequency collected by OPQ Boxes with data on voltage and frequency collected by the existing UH power monitors. The~results are shown in Figure~\ref{fig:opqbox-f-v-validation}.

\begin{figure}[H]
\centering
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=0.9\linewidth]{images/pilot/opqbox-frequency-validation.png}
\caption{Frequency~differences.}
\label{fig:opqbox-validation-1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=0.9\linewidth]{images/pilot/opqbox-voltage-validation.png}
\caption{Voltage~differences}
\label{fig:opqbox-validation-2}
\end{subfigure}
\caption{Validation of (\textbf{a}) frequency and (\textbf{b}) voltage.}
\label{fig:opqbox-f-v-validation}
\end{figure}

As the charts illustrate, there is very close correspondence between the building meters and the OPQ Boxes. For~frequency differences, the~value of $\sigma$ is 0.0079 Hz, and~for voltage, $\sigma$ is 0.1703 V. (Note that typical thresholds for frequency and voltage PQ anomalies is 1 Hz and 6 V, so OPQ Box values appear accurate enough for their intended purpose.)

Validating THD and transient data, the~other two basic power quality measures collected by OPQ Boxes, was more~challenging.

Figure~\ref{fig:opqbox-thd-validation} shows the results of comparing values of THD collected by OPQ Boxes and building~meters.

\begin{figure}[H]
\centering
\includegraphics[width=0.8\linewidth]{images/pilot/opqbox-thd-validation.png}
\caption{Total harmonic distortion (THD)~differences.}
\label{fig:opqbox-thd-validation}
\end{figure}

During the hours of 6 p.m. and 6 a.m., OPQBox and the building level meter displayed a high level of agreement, as shown in the blue histogram in Figure~\ref{fig:opqbox-thd-validation}.
On the other hand, during~the hours of 6 p.m. and 6 a.m., there was a static disparity of 0.13\% between the two meters as portrayed in the red histogram.
This is likely attributed to the meter location in the power transmission hierarchy.
While~the OPQ Box is plugged the $120V_{ac}$ line, the~building meter is monitoring the $480V_{ac}$ three phase line.
An additional active conditioning system installed along side the transformer responsible for compensating for reactive power in the building is the likely culprit in the~disparity.

Unfortunately, we were unable to perform validation of transient data collection, because~building meters did not provide us with access to that information.
However, synthetic tests performed in the lab against a calibration source showed that OPQ Box is able to measure transient magnitude with $\sigma=0.125$ V, significantly higher than the triggering threshold.
It should be noted that the internal transient metric provided by the OPQ Box is only used for event detection, while higher level analytics compute their own transient classification parameters.
As such, the~transient detection capabilities of the OPQ Box are more than sufficient for its role in the OPQ~ecosystem.

\subsection{OPQs Triggering System Provides Advantages with Respect to Bandwidth and~Computation}
\label{hyp:02}

The pilot study provided an opportunity to collect data on the resources required by an OPQ sensor network with respect to cloud-level network resources and server-level computation~overhead.

To assess network bandwidth utilization, we analyzed the data stream of frequency and voltage collected by OPQ Boxes, and~the resulting network bandwidth utilized by the OPQ Makai triggering algorithm. Over~the course of a typical day, OPQ Makai requested approximately 136 MB of data from the deployed OPQ Boxes. We then calculated how much data would be sent by a power quality meter using a more conventional triggering approach in which exceeding a threshold would automatically result in sending waveform data, and~found that, under~the same conditions, approximately 1025 MB would be sent to the cloud, or~eight times the network~bandwidth.

The computational cost of an OPQ sensor network is proportional to the amount of data acquired. Since the meter level analytics are at best only good enough for event detection, further analysis is required for event classification.
This implies that the OPQ Makai triggering algorithm is at least eight times more efficient than the self triggered counterpart, since eight times less data is sent by OPQ Boxes to the cloud for~analysis.

\subsection{OPQ Enables Subthreshold Event Detection Based on Temporal~Locality}
\label{sec:subthreshold-events}

One interesting capability enabled by two-way communication between OPQ Boxes and their cloud-based services is what we term {\em subthreshold event detection based on temporal locality}.
In a nutshell, when one OPQ Box determines that a power quality measure has exceeded a threshold, one of the actions of the sensor network is to request high fidelity waveform data from neighboring boxes for the same temporal period, regardless of whether those boxes are reporting over-threshold data.
In the event that any of those boxes actually do report over-threshold data, then the request ripples outward to the boxes neighboring that box, and~so forth.
This is accomplished by maintaining a model of the OPQ Box metrics for each device.
Once a single device passes a threshold, OPQ Makai will search the network for all other devices, which exhibit measurements that deviate significantly from the device model.
If more devices are located, data from all the affected devices will be captured.
Otherwise, the~event is deemed local and~ignored.

Sub-threshold triggering allows for clustering of OPQ Boxes based on their electrical distance without a priori knowledge of the power grid topology.
A distance metric comprised of the magnitude of the $V_{rms}$ disturbance between every pair of OPQ Boxes was used as the clustering parameter.
This~distance metric describes how much local disturbances observed by a device affect every other device on the network.
Hierarchical clustering employed on the aforementioned distance metric results in the hierarchy shown in Figure~\ref{fig:localization-clustering}a.
These clusters closely corresponded to the topology of the power grid.
For example, the two top-level clusters correspond to the two 12 kV feeders, which power the university campus.
As such, disturbances that originate on one feeder, have a minimal impact on the other.
On the other hand, devices 1009 and 1008 are located in adjacent buildings, and~a common 480 V feeder, resulting in a large level of commonality in observed anomalies.
Sub-threshold triggering along with clustering in turn made it possible to perform limited localization of PQ incidents without prior knowledge of the grid layout.
For example, consider a $V_{rms}$ sag observed by the OPQ network shown in Figure~\ref{fig:localization-clustering}b.
This event had the largest impact on devices 1008 and 1009, which, as shown via hierarchical clustering, are closely coupled.
From this information it is possible to conclude that the event depicted in Figure~\ref{fig:localization-clustering}b originated from inside the UH power grid relatively close to devices 1008 and~1009.

\begin{figure}[H]

\centering
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=0.9\linewidth]{images/pilot/clustring.png}
\caption{Hierarchical clustering of OPQ~Boxes.}
\label{fig:clustering}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[width=0.9\linewidth]{images/pilot/localisation.png}
\caption{Localization of a~disturbance.}
\label{fig:localisation}
\end{subfigure}
\caption{Hierarchical clustering of OPQ Boxes, and~localization of a~disturbance.}
\label{fig:localization-clustering}
\end{figure}
\unskip

\subsection{The OPQ Information Architecture Provides a Means to Produce Actionable~Insights}
\label{hyp:04}

Figure~\ref{fig:level-statistics} provides some descriptive statistics on the total number of entities created at each layer of the OPQ Information Architecture over the course of the pilot study. Note that the number of entities at a given layer at a given point in time is typically far less, due to the use of TTL to discard entities not promoted to higher~layers.

Figure~\ref{fig:level-statistics} shows that all layers of the hierarchy were actively used. In~addition, the~system was able to produce actionable insights by automatically detecting periodic phenomena for two boxes, and~then predicting future occurrences of power quality anomalies with over a 50\% success rate (in other words, the~current implementation of predictive phenomena leads to a significant false positive rate, predicting twice as many future power quality anomalies as actually occurred). Improving the predictive capabilities is a topic of future~research.

\begin{figure}[H]
\centering
\begin{tabularx}{.4\textwidth}{lX}
\toprule
\textbf{Level} & \textbf{Total} \\
\midrule
Phenomena & 10.4K \\
Incidents & 415K \\
Detections & 91.4K \\
AML (Trends) & 1.94M \\
AML (Measurements) & 116~M \\
IML & 100B \\
\bottomrule
\end{tabularx}
\caption{Entities created at each level of the OPQ Information Architecture during the pilot~study.}
\label{fig:level-statistics}
\end{figure}

\subsection{The OPQ Information Architecture Provides Predictable Upper Bounds on Storage~Resources}
\label{hyp:05}

Figure~\ref{fig:data-management-graph} shows a graph that illustrates storage requirements both with and without TTL-based data removal. As~the graph shows, over~the course of the three month period, we estimate that cloud-level storage requirements would have reached approximately 2.5 TB if all data was kept. However, due to TTL, the~amount of data held in the cloud at the conclusion of the case study was only 100~GB.

\begin{figure}[H]
\centering
\includegraphics[width=0.4\linewidth]{images/pilot/data-management-graph.png}
\caption{\hl{Storage} requirements with and without~time-to-live (TTL).} %%% It seems that the figure isn't complete, please modify.
\label{fig:data-management-graph}
\end{figure}

It is difficult to compute a strict upper bound on storage requirements for an OPQ network, because~the amount of storage does depend on the number of power quality anomalies experienced by the grid. That said, we did notice that storage requirements were rising much more slowly at the end of the pilot when utilizing TTL as compared to not utilizing~TTL.

The largest contribution to data storage is the raw power data. In~a system without TTL, raw data storage rapidly grows. The~bounds on raw storage can be calculated by taking into account the sampling rate and sample size for active OPQ~Boxes.

Given that OPQ Boxes were configured to sample at 12 kHz at 2 bytes per sample during the deployment, we can calculate the bounds on raw storage for 16 OPQ Boxes over three months without TTL as 3.1~TiB.

Given a network of 16 OPQ Boxes over the pilot deployment period with TTL, the~combined data storage from all levels reached an asymptotic limit of near 100 GB. The~high rate of detections made by OPQ utilizes most of the space. The~reason for this is that Events store associated IML data along with each detection for high-fidelity analysis. It is also possible to observe how the AML and IML levels level off asymptomatically over time. This is a result of TTL discarding not useful~data.

\subsection{OPQ Provides Useful Adaptive Optimization~Capabilities}
\label{sec:adaptive-optimization}

The top layer of OPQ's information architecture supports adaptive optimization: the ability to analyze previously collected data and use it to change the settings associated with data collection from OPQ~Boxes.

Two areas where adaptive optimizations were utilized are identifying periodic phenomena and predictions of future phenomena. Periodic phenomena are PQ related signals that occur at regular intervals. One example of periodic phenomena is a cyclic voltage sag that was observed at one of our sensors with a period of about 34 min during our pilot study (Figure~\ref{fig:periodic-voltage-sags}). This periodic phenomena not only provided a classification of an interesting pattern, but~also helped optimize the OPQ system for the capture of future phenomena. Several system wide optimizations are~utilized.

\begin{figure}[H]
\centering
\includegraphics[width=0.90\linewidth]{images/pilot/periodic-voltage-sags.png}
\caption{Detection of periodic voltage~sags.}
\label{fig:periodic-voltage-sags}
\end{figure}

First, classified periodic phenomena are used to generate future phenomena. Future phenomena are a prediction of future signals of interest. When a prediction is made, the~system will automatically increase sampling fidelity for the predicted sensor during the time window of the prediction. In~particular, the~measurement rate is increased from one measurement per second to six measurements per second, providing higher fidelity data. Detection thresholds are also decreased, making it more likely that the predicted signal will be observed even if it is of low~magnitude.

Second, the~increase in detection sensitivity makes it more likely that sub-threshold signals of interest will be classified. For~example, the~increased detection sensitivity allowed us to observe and classify a periodic voltage sag that would not have been detected by our normal classifier due to the small magnitude of the~sag.

This creates a positive feedback loop whereas periodic phenomena become more accurate, future phenomena become more accurate. As~future phenomena become more accurate, periodic phenomena become more~accurate.

\section{Conclusions and Future~Directions}
\label{sec:conclusions}

This project has produced both hardware and software for power quality monitoring with a variety of innovations. Our OPQ Boxes collect frequency, voltage, THD, and~transients with high fidelity, and~at a cost that is generally 10$\times$ to 100$\times$ cheaper than current commercial offerings (though these commercial offerings offer a variety of features not available from OPQ Boxes, so the appropriate choice depends upon the needs of the user).

Our sensor network provides real-time, two-way communication between the sensor nodes (OPQ Boxes) and their cloud services, providing several innovative features for power quality monitoring. The~OPQ triggering system exploits temporal locality to request high fidelity data from neighboring boxes when one exceeds a threshold, enabling our network to detect and analyze data that would have been unreported by a naive, threshold-based approach. The~OPQ Information Architecture enables Actors at the Phenomena Layer to control sensitivity settings of individual OPQ Boxes in order to improve data collection when power quality anomalies have been~predicted.

Our sensor network is designed to efficiently use network and storage resources. Two way communication enables OPQ Boxes to send low fidelity summary statistics on power quality at one second intervals, which can be used by cloud-based services to decide whether to request high fidelity data. Our TTL mechanism implements a “use it or lose it” approach to cloud-based data, which in our pilot study reduced cloud storage requirements by over~ten-fold.

Our pilot study has provided evidence that OPQ sensor networks can provide useful new support for understanding grid stability, particularly in grids with distributed, intermittent renewables. This~in part because OPQ Boxes are suitable for installation at the residential level: installation requires only an available wall socket and WiFi connection, no electrician or GPS line-of-sight required. Finally, the~low cost of OPQ Boxes means mass deployment across a neighborhood is not financially~infeasible.

One future goal is to partner with an organization that can enable OPQ Boxes to be produced at volume. We made the OPQ Boxes by hand for the pilot study, and~although we have had numerous requests for OPQ Boxes, we do not possess manufacturing~capacity.

With manufacturing capacity will come the ability to deploy OPQ sensor networks at a higher scale. We believe that a single sensor network can easily scale to hundreds of boxes before bandwidth and processing constraints become an issue. To~scale to many thousands of boxes, we believe a federated approach would work in which individual OPQ sensor networks communicate their findings to each other. Those networks could be designed around natural grid boundaries such as~substations.

A second goal is to explore ways to combine OPQ sensor networks with more traditional power quality monitoring tools and standards, such as PQDIF. For~example, it is possible that OPQ analyses could be improved with access to additional power data such as current, phase angle, and~so~forth.

A third goal is to build Actors that operate not only on power quality data but also environmental data such as wind, temperature, humidity, and~insolation. These additional data streams could yield valuable mechanisms for creating Predictive Phenomena for power quality anomalies associated with renewable energy sources. Ultimately, such understanding could lead to ways to significantly increase the amount of distributed renewable energy that can be incorporated into electrical~grids.

To conclude, a~recent paper by Mohsenian et al.~\cite{mohsenian-rad_distribution_2018} states, “the main challenge is to go beyond manual methods based on the intuition and heuristics of human experts...it is crucial to develop the machine intelligence needed to automate and scale up the analytics on billions of PMU measurements and terabytes of data on a daily basis and in real time.” We believe that the OPQ sensor network represents a small step along the path toward that~future.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{\hl{For} }%% research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work reported.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\funding{\hl{Please} }%add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{\hl{Declare} } %conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.

% Bibliography
\reftitle{References}
\begin{thebibliography}{999}
\providecommand{\natexlab}[1]{#1}

\bibitem[Trabish(2014)]{trabish_solar_2014}
Trabish, H.
\newblock Solar Installers flee {Hawaii} as Interconnection Queue Backs Up.
\newblock {\em Utility Dive}, {\hl{2014}}.%% Please add day month.

\bibitem[Anastasi \em{et~al.}(2009)Anastasi, Conti, Di~Francesco, and
Passarella]{anastasi_energy_2009}
Anastasi, G.; Conti, M.; Di~Francesco, M.; Passarella, A.
\newblock Energy conservation in wireless sensor networks: {A} survey.
\newblock {\em Ad~Hoc Networks} {\bf 2009}, {\em 7},~537--568, doi:10.1016/j.adhoc.2008.06.003.

\bibitem[Anurangi \em{et~al.}(2017)Anurangi, Rodrigo, and
Jayatunga]{anurangi_effects_2017}
Anurangi, R.O.; Rodrigo, A.S.; Jayatunga, U.
\newblock Effects of high levels of harmonic penetration in distribution
networks with photovoltaic inverters.
\newblock In {Proceedings of the } 2017 {IEEE} {International} {Conference} on IEEE Industrial and {Information} {Systems} ({ICIIS}),  \hl{2017}; pp. 1--6, doi:10.1109/ICIINFS.2017.8300335.%% Please add conference location and date.

\bibitem[{Rita Pinto} \em{et~al.}(2016){Rita Pinto}, {Sílvio Mariano}, {Maria
Do Rosário Calado}, and {José Felippe de Souza}]{rita_pinto_impact_2016}
Pinto, R.; Mariano, S.; Calado, M.D.R.;  De Souza, J.F. 
\newblock Impact of {Rural} {Grid}-{Connected} {Photovoltaic} {Generation}
{Systems} on {Power} {Quality}.
\newblock {\em Energies} {\bf 2016}, {\em 9},~739, doi:10.3390/en9090739.

\bibitem[Delgado \em{et~al.}(2011)Delgado, Santos, de~Almeida, and
Figueira]{delgado_solutions_2011}
Delgado, J.; Santos, B.; de~Almeida, A.T.; Figueira, A.
\newblock Solutions to mitigate power quality disturbances resulting from
integrating intermittent renewable energy in the grid of {Porto} {Santo}.
\newblock In {Proceedings of the } 11th {International} {Conference} on {Electrical} {Power} {Quality}
and {Utilisation}, \hl{2011}; pp. 1--6, 
\newblock ISSN 2150-6655,
doi:{\changeurlcolor{black}\href{https://doi.org/10.1109/EPQU.2011.6128870}{\detokenize{10.1109/EPQU.2011.6128870}}}.

\bibitem[Zavoda(2018)]{zavoda_power_2018}
Zavoda, F.
\newblock { Power {Quality} and {EMC} {Issues} with {Future} {Electricity}
{Networks}}. In {Proceedings of the } CIGRE, \hl{2018}.

\bibitem[Nakafuji \em{et~al.}(2011)Nakafuji, Aukai, Dangelmaier, Reynolds,
Yoshimura, and {Ying Hu}]{nakafuji_back--basics_2011}
Nakafuji, D.; Aukai, T.; Dangelmaier, L.; Reynolds, C.; Yoshimura, J.; {
Hu, Y}.
\newblock “{Back}-to-basics”: {Operationalizing} data mining and visualization
techniques for utilities.
\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, \hl{2011}; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.

\bibitem[Dedad(2008)]{dedad_when_2008}
Dedad, J.
\newblock When {Does} {Poor} {Power} {Quality} {Cause} {Electronics}
{Failures}?
\newblock {\em Electrical Construction and Maintenance Magazine}, {\hl{2008}}.
\newblock Library Catalog.  Available online:  \url{www.ecmweb.com} (accessed on \hl{day month year}).

\bibitem[Elphick \em{et~al.}(2015)Elphick, Ciufo, Smith, and
Parera]{elphick_summary_2015}
Elphick, S.; Ciufo, P.; Smith, V.; Parera, S.
\newblock Summary of the economic impacts of power quality on consumers.
\newblock  In {Proceedings of the }2015 {Australasian} {Universities} {Power} {Engineering}
{Conference}, \hl{2015}.

\bibitem[Laskar(2012)]{laskar_power_2012}
Laskar, S.H.
\newblock Power quality issues and need of intelligent {PQ} monitoring in the
smart grid environment.
\newblock In~{Proceedings of the }2012 47th
{International} Universities {Power} {Engineering} {Conference} ({UPEC}), \hl{2012}; pp.~1--6.

\bibitem[Negrashov(2020)]{negrashov_opq_2020}
Negrashov, S.
\newblock {OPQ} {Box} {Specifications} and {Design}, 2020.
\newblock Library Catalog.  Available online:  \url{github.com} (accessed on \hl{day month year}).

\bibitem[Unruh(2018)]{unruh_ieee_2018}
Unruh, T.
\newblock \emph{IEEE P1159/D3: {Draft} {Recommended} {Practice} for
{Monitoring} {Electric} {Power} {Quality}}; \hl{2018}. %% Please provide publisher and its location.

\bibitem[Kahle(2015)]{kahle_power_2015}
Kahle, K.
\newblock Power {Converters} and {Power} {Quality}.
\newblock \emph{arXiv} \textbf{2015}, arXiv:1607.01556. doi:10.5170/CERN-2015-003.57.

\bibitem[Thallam and Heydt(2000)]{thallam_power_2000}
Thallam, R.; Heydt, G.
\newblock Power acceptability and voltage sag indices in the three phase sense.
\newblock In {Proceedings of the } 2000 {Power} {Engineering} {Society} {Summer} {Meeting}, 
 \hl{2000}; Volume~2, pp. 905--910, doi:10.1109/PESS.2000.867482.

\bibitem[Djokic \em{et~al.}(2005)Djokic, Desmet, Vanalme, Milanovic, and
Stockman]{djokic_sensitivity_2005}
Djokic, S.; Desmet, J.; Vanalme, G.; Milanovic, J.; Stockman, K.
\newblock Sensitivity of personal computers to voltage sags and short
interruptions.
\newblock {\em IEEE Trans. Power Deliv.} {\bf 2005}, {\em
20},~375--383.
\newblock 
doi:{\changeurlcolor{black}\href{https://doi.org/10.1109/TPWRD.2004.837828}{\detokenize{10.1109/TPWRD.2004.837828}}}.

\bibitem[Dickens \em{et~al.}(2019)Dickens, Christe, and
Johnson]{dickens_transient_2019}
Dickens, C.; Christe, A.; Johnson, P.
\newblock A {Transient} {Classification} {System} {Implementation} on an {Open}
{Source} {Distributed} {Power} {Quality} {Network}.
\newblock  In Proceedings of the {Ninth} {International} {Conference} on {Smart}
{Grids}, {Green} {Communications} and {IT} {Energy}-aware {Technologies},
\hl{2019}.

\bibitem[Fluke(2020)]{fluke_fluke_2020}
Fluke, I.
\newblock \emph{Fluke {Corporation}: {Fluke} {Electronics}, {Biomedical},
{Calibration} and {Networks}}; \hl{2020}.

\bibitem[Dranetz(2020)]{dranetz_dranetz_2020}
Dranetz, I.
\newblock Dranetz {Power} {Monitoring}, 2020.
\newblock Library Catalog. Available online:  \url{ www.dranetz.com} (accessed on \hl{day month year}).

\bibitem[Elspec(2020)]{elspec_elspec_2020}
Elspec, I.
\newblock Elspec---{Power} Quality Analyzers and Solutions, 2020.
\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com} (accessed on \hl{day month year}).

\bibitem[Powerside(2020)]{powerside_powerside_2020}
Powerside, I.
\newblock Powerside, {Inc}., 2020.
\newblock Library Catalog. Available online:  \url{Powerside.com}  (accessed on \hl{day month year}).

\bibitem[ACR(2020)]{acr_acr_2020}
ACR, S.
\newblock {ACR} {Systems}, 2020.
\newblock Library Catalog. Available online:  \url{www.acrdatasolutions.com}  (accessed on \hl{day month year}).

\bibitem[Viciana \em{et~al.}(2018)Viciana, Alcayde, Montoya, Baños,
Arrabal-Campos, Zapata-Sierra, and
Manzano-Agugliaro]{viciana_openzmeter_2018}
Viciana, E.; Alcayde, A.; Montoya, F.; Baños, R.; Arrabal-Campos, F.;
Zapata-Sierra, A.; Manzano-Agugliaro, F.
\newblock {OpenZmeter}: {An} {Efficient} {Low}-{Cost} {Energy} {Smart} {Meter}
and {Power} {Quality} {Analyzer}.
\newblock {\em Sustainability} {\bf 2018}, {\em 10}, 4038.
\newblock 
doi:{\changeurlcolor{black}\href{https://doi.org/10.3390/su10114038}{\detokenize{10.3390/su10114038}}}.

\bibitem[Concepts(2020)]{electrotek_concepts_pqview_2020}
Concepts, E.
\newblock {PQView} 4 {\textbar} {Electrotek} {Concepts}, 2020.
\newblock Library Catalog. Available online:  \url{www.electrotek.com}  (accessed on \hl{day month year}).

\bibitem[{Elspec, Ltd}(2016)]{elspec_ltd_pqscada_2016}
{Elspec, Ltd}.
\newblock {PQSCADA} {Sapphire}, 2016.
\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com}  (accessed on \hl{day month year}).

\bibitem[Sabin(2020)]{sabin_ieee_2020}
Sabin, D.
\newblock \emph{{IEEE} 1159.3 {PQDIF} {Task} {Force}}; \hl{2020}.

\bibitem[Alliance(2020)]{grid_protection_alliance_grid_2020}
Alliance, G.P.
\newblock \emph{Grid {Protection} {Alliance}---{Home}}; \hl{2020}.

\bibitem[Di~Manno \em{et~al.}(2015)Di~Manno, Varalone, Verde, De~Santis,
Di~Perna, and Salemme]{di_manno_user_2015}
Di~Manno, M.; Varalone, P.; Verde, P.; De~Santis, M.; Di~Perna, C.; Salemme, M.
\newblock User friendly smart distributed measurement system for monitoring and
assessing the electrical power quality.
\newblock In Proceedings of the 2015 {AEIT} {International} {Annual}
{Conference}, \hl{2015}.

\bibitem[Daponte \em{et~al.}(2000)Daponte, Di~Penta, and
Mercurio]{daponte_transientmeter:_2000}
Daponte, P.; Di~Penta, M.; Mercurio, G.
\newblock {TRANSIENTMETER}: A distributed measurement system for power quality
monitoring.
\newblock In {Proceedings of the } Ninth {International} {Conference} on {Harmonics} and {Quality} of
{Power}, \hl{2000}.

\bibitem[Bilik \em{et~al.}(2007)Bilik, Koval, and Hula]{bilik_modular_2007}
Bilik, P.; Koval, L.; Hula, J.
\newblock Modular system for distributed power quality monitoring.
\newblock In {Proceedings of the } 9th {International} {Conference} on {Electrical} {Power} {Quality}
and {Utilization}, \hl{2007}.

\bibitem[Xu \em{et~al.}(2012)Xu, Xu, Xi, and Zhang]{xu_distributed_2012}
Xu, W.; Xu, G.; Xi, Z.; Zhang, C.
\newblock Distributed power quality monitoring system based on {EtherCAT}.
\newblock In {Proceedings of the }2012 {China} {International} {Conference} on {Electricity}
{Distribution}, \hl{2012}.

\bibitem[Suslov \em{et~al.}(2014)Suslov, Solonina, and
Smirnov]{suslov_distributed_2014}
Suslov, K.; Solonina, N.; Smirnov, A.
\newblock Distributed power quality monitoring.
\newblock In {Proceedings of the } {IEEE} 16th {International} {Conference} on {Harmonics} and
{Quality} of {Power}, \hl{2014}.

\bibitem[Sayied \em{et~al.}(2013)Sayied, Spaulding, and
Akbary]{sayied_power_2013}
Sayied, O.; Spaulding, J.; Akbary, B.
\newblock \hl{Power quality metrics calculations for the smart grid}. %%% Please add more detailed information.

\bibitem[Kucuk \em{et~al.}(2010)Kucuk, Inan, Salor, Demirci, Akkaya, Buhan,
Boyrazoglu, Unsar, Altintas, Haliloglu, Cadirci, and
Ermis]{kucuk_extensible_2010}
Kucuk, D.; Inan, T.; Salor, O.; Demirci, T.; Akkaya, Y.; Buhan, S.; Boyrazoglu,
B.; Unsar, O.; Altintas, E.; Haliloglu, B.; et al.
\newblock An extensible database architecture for nationwide power quality
monitoring.
\newblock {\em Electr. Power Energy Syst.} {\bf 2010}, {\em 32},  559--570.

\bibitem[Mohsenian-Rad \em{et~al.}(2018)Mohsenian-Rad, Stewart, and
Cortez]{mohsenian-rad_distribution_2018}
Mohsenian-Rad, H.; Stewart, E.; Cortez, E.
\newblock Distribution {Synchrophasors}: {Pairing} {Big} {Data} with
{Analytics} to {Create} {Actionable} {Information}.
\newblock {\em IEEE Power Energy Mag.} {\bf 2018}, {\em 16},~26--34.
\newblock
doi:{\changeurlcolor{black}\href{https://doi.org/10.1109/MPE.2018.2790818}{\detokenize{10.1109/MPE.2018.2790818}}}.

\bibitem[Liu \em{et~al.}(2017)Liu, You, Yao, Cui, Wu, Zhou, Zhao, Liu, and
Liu]{liu_distribution_2017}
Liu, Y.; You, S.; Yao, W.; Cui, Y.; Wu, L.; Zhou, D.; Zhao, J.; Liu, H.; Liu,
Y.
\newblock A {Distribution} {Level} {Wide} {Area} {Monitoring} {System} for the
{Electric} {Power} {Grid}: {FNET}/{GridEye}.
\newblock {\em IEEE Access} {\bf 2017}, {\em 5},~2329--2338, doi:10.1109/ACCESS.2017.2666541.

\bibitem[Bayindir \em{et~al.}(2016)Bayindir, Demirbas, Irmak, Cetinkaya, Ova,
and Yesil]{bayindir_effects_2016}
Bayindir, R.; Demirbas, S.; Irmak, E.; Cetinkaya, U.; Ova, A.; Yesil, M.
\newblock Effects of renewable energy sources on the power system.
\newblock In {Proceedings of the } 2016 {IEEE} {International} {Power} {Electronics} and {Motion}
{Control} {Conference} ({PEMC}), \hl{2016}; pp. 388--393, doi:10.1109/EPEPEMC.2016.7752029.

\bibitem[Farhoodnea \em{et~al.}(2012)Farhoodnea, Mohamed, Shareef, and
Zayandehroodi]{farhoodnea_power_2012}
Farhoodnea, M.; Mohamed, A.; Shareef, H.; Zayandehroodi, H.
\newblock Power quality impact of grid-connected photovoltaic generation system
in distribution networks.
\newblock In {Proceedings of the } 2012 {IEEE} {Student} {Conference} on {Research} and {Development}
({SCOReD}), \hl{2012}; pp. 1--6, doi:10.1109/SCOReD.2012.6518600.

\bibitem[Shafiullah \em{et~al.}(2014)Shafiullah, Oo, Ali, Wolfs, and
Stojcevski]{shafiullah_experimental_2014}
Shafiullah, G.M.; Oo, A.M.T.; Ali, A.B.M.S.; Wolfs, P.; Stojcevski, A.
\newblock Experimental and simulation study of the impact of increased
photovoltaic integration with the grid.
\newblock {\em J. Renew. Sustain. Energy} {\bf 2014}, {\em
6}, doi:10.1063/1.4885105.

\bibitem[Kucuk and Salor(2013)]{kucuk_assessment_2013}
Kucuk, D.; Salor, O.
\newblock Assessment of extensive countrywide electrical power quality
measurements through a database architecture.
\newblock {\em Electr. Eng.} {\bf 2013}, {\em 95}, 1--19.

\end{thebibliography}

\end{document}

### 2020-07-31 manuscript.v5/energies-864131.tex
checkpoint

@@
 \Title{%The paper was slightly modified by our English editor, please check the whole text and confirm if your meaning is retained; 2.	Do not delete any comment we left for you and reply to each comment so that we can understand your meaning clearly; 3.	If you need to revise somewhere in your paper, please highlight the revisions to let us know.
-(Thank you for your cooperation in advance.)
+(Thank you for your cooperation in advance.)  (CONFIRMED)
 \highlight{Design}, Implementation, and Evaluation of Open Power Quality}
@@
 % Authors, for the paper (add full first names)
-\Author{\hl{Anthony} %Please carefully check the accuracy of names and affiliations. 
+\Author{\hl{Anthony} %Please carefully check the accuracy of names and affiliations. (CONFIRMED)
 J. Christe, Sergey Negrashov and Philip M. Johnson *}
@@
 
-%%% We added a space between 5 and V. Please confirm.
+%%% We added a space between 5 and V. Please confirm. CONFIRMED.
 
@@
 This Actor is capable of classifying voltage sags, swells, and~interruptions as defined by the standard, and~creating data at the Incident Layer if~appropriate.
-%%% Should it be reference citation? 
+%%% Should it be reference citation? YES
 \highlight{{\em Box Optimization Actor.}} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
@@
 \includegraphics[width=0.4\linewidth]{images/pilot/data-management-graph.png}
-\caption{\hl{Storage} requirements with and without~time-to-live (TTL).} %%% It seems that the figure isn't complete, please modify.
+\caption{\hl{Storage} requirements with and without~time-to-live (TTL).} %%% It seems that the figure isn't complete, please modify. FIGURE IS COMPLETE.
 \label{fig:data-management-graph}
@@
 Authorship must be limited to those who have contributed substantially to the work reported.
+SEE BELOW
+
+\authorcontributions{
+Conceptualization, Johnson, Christe, Negrashov; methodology, Johnson, Christe, Negrashov; software, Johnson, Christe, Negrashov; validation, Johnson, Christe, Negrashov; formal analysis, Johnson, Christe, Negrashov; investigation, Johnson, Christe, Negrashov; resources, Johnson; data curation, Christe, Negrashov; writing--original draft preparation, Johnson, Christe, Negrashov; writing--review and editing, Johnson, Christe, Negrashov; visualization, Johnson, Christe, Negrashov; supervision, Johnson; project administration, Johnson; funding acquisition, Johnson.
+All authors have read and agreed to the published version of the manuscript.
+}
 
@@
 Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.
+
+\funding{This research was funded in part by a grant from the University of Hawaii President's Green Award program.}
 
@@
 
+\conflictsofinterest{The authors declare no conflict of interest.}
+
 % Bibliography
@@
 \newblock Solar Installers flee {Hawaii} as Interconnection Queue Backs Up.
-\newblock {\em Utility Dive}, {\hl{2014}}.%% Please add day month.
+\newblock {\em Utility Dive}, {\hl{September 29, 2014}}.%% Please add day month. DONE
 
@@
 networks with photovoltaic inverters.
-\newblock In {Proceedings of the } 2017 {IEEE} {International} {Conference} on IEEE Industrial and {Information} {Systems} ({ICIIS}),  \hl{2017}; pp.
+\newblock In {Proceedings of the } 2017 {IEEE} {International} {Conference} on IEEE Industrial and {Information} {Systems} ({ICIIS}), Jayapura, Indonesia \hl{November 2, 2017}; pp.
 1--6, doi:10.1109/ICIINFS.2017.8300335.%% Please add conference location and date.
@@
 \newblock In {Proceedings of the } 11th {International} {Conference} on {Electrical} {Power} {Quality}
-and {Utilisation}, \hl{2011}; pp. 1--6, 
+and {Utilisation}, \hl{October 17, 2011}; pp. 1--6,
 \newblock ISSN 2150-6655,
@@
 \newblock { Power {Quality} and {EMC} {Issues} with {Future} {Electricity}
-{Networks}}. In {Proceedings of the } CIGRE, \hl{2018}.
+{Networks}}. In {Proceedings of the } CIGRE, \hl{May 15, 2018}.
 
@@
 techniques for utilities.
-\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, \hl{2011}; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.
+\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, \hl{July 31, 2011}; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.
 
@@
 {Failures}?
-\newblock {\em Electrical Construction and Maintenance Magazine}, {\hl{2008}}.
-\newblock Library Catalog.  Available online:  \url{www.ecmweb.com} (accessed on \hl{day month year}).
+\newblock {\em Electrical Construction and Maintenance Magazine}, {\hl{November 1, 2008}}.
+\newblock Library Catalog.  Available online:  \url{www.ecmweb.com} (accessed on \hl{31 July 2020}).
 
@@
 \newblock  In {Proceedings of the }2015 {Australasian} {Universities} {Power} {Engineering}
-{Conference}, \hl{2015}.
+{Conference}, \hl{September 27, 2015}.
 
@@
 \newblock In~{Proceedings of the }2012 47th
-{International} Universities {Power} {Engineering} {Conference} ({UPEC}), \hl{2012}; pp.~1--6.
+{International} Universities {Power} {Engineering} {Conference} ({UPEC}), \hl{September 4, 2012}; pp.~1--6.
 
@@
 \newblock {OPQ} {Box} {Specifications} and {Design}, 2020.
-\newblock Library Catalog.  Available online:  \url{github.com} (accessed on \hl{day month year}).
+\newblock Library Catalog.  Available online:  \url{github.com} (accessed on \hl{31 July 2020}).
 
@@
 \newblock \emph{IEEE P1159/D3: {Draft} {Recommended} {Practice} for
-{Monitoring} {Electric} {Power} {Quality}}; \hl{2018}. %% Please provide publisher and its location.
+{Monitoring} {Electric} {Power} {Quality}}; IEEE, New York, NY \hl{February 1, 2018}. %% Please provide publisher and its location. DONE
 
@@
 \newblock In {Proceedings of the } 2000 {Power} {Engineering} {Society} {Summer} {Meeting},
- \hl{2000}; Volume~2, pp. 905--910, doi:10.1109/PESS.2000.867482.
+ \hl{July 16, 2000}; Volume~2, pp. 905--910, doi:10.1109/PESS.2000.867482.
 
@@
 {Grids}, {Green} {Communications} and {IT} {Energy}-aware {Technologies},
-\hl{2019}.
+\hl{June 2, 2019}.
 
@@
 \newblock \emph{Fluke {Corporation}: {Fluke} {Electronics}, {Biomedical},
-{Calibration} and {Networks}}; \hl{2020}.
+{Calibration} and {Networks}}; \hl{July 31, 2020}.
 
@@
 \newblock Dranetz {Power} {Monitoring}, 2020.
-\newblock Library Catalog. Available online:  \url{ www.dranetz.com} (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{ www.dranetz.com} (accessed on \hl{31 July 2020}).
 
@@
 \newblock Elspec---{Power} Quality Analyzers and Solutions, 2020.
-\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com} (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com} (accessed on \hl{31 July 2020}).
 
@@
 \newblock Powerside, {Inc}., 2020.
-\newblock Library Catalog. Available online:  \url{Powerside.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{Powerside.com}  (accessed on \hl{31 July 2020}).
 
@@
 \newblock {ACR} {Systems}, 2020.
-\newblock Library Catalog. Available online:  \url{www.acrdatasolutions.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.acrdatasolutions.com}  (accessed on \hl{31 July 2020}).
 
@@
 \newblock {PQView} 4 {\textbar} {Electrotek} {Concepts}, 2020.
-\newblock Library Catalog. Available online:  \url{www.electrotek.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.electrotek.com}  (accessed on \hl{31 July 2020}).
 
@@
 \newblock {PQSCADA} {Sapphire}, 2016.
-\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com}  (accessed on \hl{31 July 2020}).
 
@@
 Sabin, D.
-\newblock \emph{{IEEE} 1159.3 {PQDIF} {Task} {Force}}; \hl{2020}.
+\newblock \emph{{IEEE} 1159.3 {PQDIF} {Task} {Force}}; \hl{July 31, 2020}.
 
@@
 Alliance, G.P.
-\newblock \emph{Grid {Protection} {Alliance}---{Home}}; \hl{2020}.
+\newblock \emph{Grid {Protection} {Alliance}---{Home}}; \hl{July 31, 2020}.
 
@@
 \newblock In Proceedings of the 2015 {AEIT} {International} {Annual}
-{Conference}, \hl{2015}.
+{Conference}, \hl{October 14, 2015}.
 
@@
 \newblock In {Proceedings of the } Ninth {International} {Conference} on {Harmonics} and {Quality} of
-{Power}, \hl{2000}.
+{Power}, Orlando, Florida \hl{October 1, 2000}.
 

### 2020-08-03 manuscript.v5/energies-864131.tex
finished v5

@@
 
+%%% I HAVE ADDED LOCATIONS AND DATES TO FOLLOWING CITATIONS
+
 \bibitem[Trabish(2014)]{trabish_solar_2014}
@@
 \newblock In {Proceedings of the } 11th {International} {Conference} on {Electrical} {Power} {Quality}
-and {Utilisation}, \hl{October 17, 2011}; pp. 1--6,
+and {Utilisation}, Lisbon, Portugal, \hl{October 17, 2011}; pp. 1--6,
 \newblock ISSN 2150-6655,
@@
 \newblock { Power {Quality} and {EMC} {Issues} with {Future} {Electricity}
-{Networks}}. In {Proceedings of the } CIGRE, \hl{May 15, 2018}.
+{Networks}}. In {Proceedings of the } CIGRE, Paris, France, \hl{May 15, 2018}.
 
@@
 techniques for utilities.
-\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, \hl{July 31, 2011}; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.
+\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, San Jose, California, \hl{July 31, 2011}; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.
 
@@
 \newblock  In {Proceedings of the }2015 {Australasian} {Universities} {Power} {Engineering}
-{Conference}, \hl{September 27, 2015}.
+{Conference}, Wollongong, Australia, \hl{September 27, 2015}.
 
@@
 \newblock In~{Proceedings of the }2012 47th
-{International} Universities {Power} {Engineering} {Conference} ({UPEC}), \hl{September 4, 2012}; pp.~1--6.
+{International} Universities {Power} {Engineering} {Conference} ({UPEC}), London, England, \hl{September 4, 2012}; pp.~1--6.
 
@@
 \newblock Power acceptability and voltage sag indices in the three phase sense.
-\newblock In {Proceedings of the } 2000 {Power} {Engineering} {Society} {Summer} {Meeting},
+\newblock In {Proceedings of the } 2000 {Power} {Engineering} {Society} {Summer} {Meeting}, Seattle, Washington,
  \hl{July 16, 2000}; Volume~2, pp. 905--910, doi:10.1109/PESS.2000.867482.
@@
 \newblock  In Proceedings of the {Ninth} {International} {Conference} on {Smart}
-{Grids}, {Green} {Communications} and {IT} {Energy}-aware {Technologies},
+{Grids}, {Green} {Communications} and {IT} {Energy}-aware {Technologies}, Athens, Greece,
 \hl{June 2, 2019}.
@@
 \newblock In Proceedings of the 2015 {AEIT} {International} {Annual}
-{Conference}, \hl{October 14, 2015}.
+{Conference}, Naples, Italy \hl{October 14, 2015}.
 
@@
 \newblock In {Proceedings of the } 9th {International} {Conference} on {Electrical} {Power} {Quality}
-and {Utilization}, \hl{2007}.
+and {Utilization}, Barcelona, Spain, \hl{October 9, 2007}.
 
@@
 \newblock In {Proceedings of the }2012 {China} {International} {Conference} on {Electricity}
-{Distribution}, \hl{2012}.
+{Distribution}, Shanghai, China, \hl{September 10, 2012}.
 
@@
 \newblock In {Proceedings of the } {IEEE} 16th {International} {Conference} on {Harmonics} and
-{Quality} of {Power}, \hl{2014}.
+{Quality} of {Power}, Orlando, Florida, \hl{October 1, 2014}.
 
@@
 Sayied, O.; Spaulding, J.; Akbary, B.
-\newblock \hl{Power quality metrics calculations for the smart grid}. %%% Please add more detailed information.
+\newblock \hl{Power quality metrics calculations for the smart grid}, Technical Report, ECE 401, Department of Electrical and Computer Engineering, Colorado State University, December 12, 2012.
+%%% Please add more detailed information.
 
@@
 \newblock In {Proceedings of the } 2016 {IEEE} {International} {Power} {Electronics} and {Motion}
-{Control} {Conference} ({PEMC}), \hl{2016}; pp. 388--393, doi:10.1109/EPEPEMC.2016.7752029.
+{Control} {Conference} ({PEMC}), Varna, Bulgaria, \hl{September 25, 2016}; pp. 388--393, doi:10.1109/EPEPEMC.2016.7752029.
 
@@
 \newblock In {Proceedings of the } 2012 {IEEE} {Student} {Conference} on {Research} and {Development}
-({SCOReD}), \hl{2012}; pp. 1--6, doi:10.1109/SCOReD.2012.6518600.
+({SCOReD}), Kuala Lumpur, \hl{December 13, 2012}; pp. 1--6, doi:10.1109/SCOReD.2012.6518600.
 

### 2020-08-03 manuscript.v6/energies-864131.tex
v6 changes.

--- 2020-07-31 manuscript.v5/energies-864131.tex
@@
 % Full title of the paper (Capitalized)
-\Title{%The paper was slightly modified by our English editor, please check the whole text and confirm if your meaning is retained; 2.	Do not delete any comment we left for you and reply to each comment so that we can understand your meaning clearly; 3.	If you need to revise somewhere in your paper, please highlight the revisions to let us know.
-(Thank you for your cooperation in advance.)
-\highlight{Design}, Implementation, and Evaluation of Open Power Quality}
+\Title{
+Design, Implementation, and Evaluation of Open Power Quality}
 
 % Authors, for the paper (add full first names)
-\Author{\hl{Anthony} %Please carefully check the accuracy of names and affiliations. 
+\Author{Anthony
 J. Christe, Sergey Negrashov and Philip M. Johnson *}
@@
 
-\highlight{{\em 1. We need more renewable energy.}} %Is italics necessary?
+\highlight{{\em 1. We need more renewable energy.}}
+%%%Is italics necessary? NO. YOU CAN REMOVE THEM.
  Concerns including pollution, environmental degradation, and~climate change have produced a global movement away from centralized, fossil-fuel-based forms of electrical energy generation and toward distributed, renewable alternatives such as wind and solar.
@@
 
-\highlight{{\em 2.
-Traditional top-down monitoring is not well-suited to bottom-up energy generation.}} For traditional grid architectures where generation is centralized and under the complete control of the utility, it is common to monitor power quality only to the substation level, because~it can be assumed that the power quality experienced at the substation is a reasonably accurate proxy for the power quality experienced by the 1000 or so end-users serviced by the substation.
+{\em 2.
+Traditional top-down monitoring is not well-suited to bottom-up energy generation.} For traditional grid architectures where generation is centralized and under the complete control of the utility, it is common to monitor power quality only to the substation level, because~it can be assumed that the power quality experienced at the substation is a reasonably accurate proxy for the power quality experienced by the 1000 or so end-users serviced by the substation.
 Furthermore, if~an end-user experiences a power quality problem not experienced by the substation, then it is most likely due to equipment or electrical issues local to that end-user and not a grid-level~problem.
@@
 
-\highlight{{\em 3.
-Consumer electronics require high quality power.}} The rise of consumer electronics has raised the bar for what constitutes ``adequate'' power quality.
+{\em 3.
+Consumer electronics require high quality power.} The rise of consumer electronics has raised the bar for what constitutes ``adequate'' power quality.
 Only a few decades ago, computers were a rare presence outside of labs and large institutions.
@@
 
-\highlight{{\em 4.
-Power quality data should be publicly available.}} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers.
+{\em 4.
+Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers.
 For~example, in~Hawaii, utilities are only required to make a ``best effort'' to provide non-harmful voltage and frequency values, and~are only required to publicly report about power outages of more than 3 min.
@@
 An isolated AC-DC converter generates $5V_{dc}$ from the mains $120V_{ac}$.
-A total of \hl{5 V} is used to power the Raspberry Pi, the~equipment connected to the expansion port, the~3.3 V regulators and voltage reference, and~an isolated DC/DC converter.
+A total of 5 V is used to power the Raspberry Pi, the~equipment connected to the expansion port, the~3.3 V regulators and voltage reference, and~an isolated DC/DC converter.
 A total of 3.3 V is used to power the isolated end of the isolation amplifier as well as the STM32F3 analog to digital converter/digital signal processor (ADC/DSP).
@@
 This allows OPQ Box to function with a battery attached to the expansion port, so that it may retain data and continue to operate during a power~outage.
-
-%%% We added a space between 5 and V. Please confirm.
 
@@
 This scheme requires the CPU to only service 60 interruptions a second, with~each interruption requiring on the order of 100 instructions, for~a CPU utilization of less than $1\%$ in normal operation.
-Userland applications communicate with the kernel driver using a file descriptor, where every \highlight{$read$} %Is italics necessary here?
+Userland applications communicate with the kernel driver using a file descriptor, where every \highlight{$read$}
+%%%Is italics necessary here? NO, YOU CAN REMOVE THEM.
  system call yields 200 samples of raw waveform.
@@
 
-\highlight{{\em Bandwidth usage is minimized.}} % Is italics necessary?
+\highlight{{\em Bandwidth usage is minimized.}}
+%%% Is italics necessary? NO, YOU CAN REMOVE THEM.
  Instead of sending the entirety of raw data, only extracted features are sent.
@@
 
-\highlight{{\em Effects of latency are minimized.}} In this case, ``latency'' refers to the time required for OPQ Makai to process the incoming feature stream and decide whether to request high fidelity data from one or more OPQ Boxes.
+{\em Effects of latency are minimized.} In this case, ``latency'' refers to the time required for OPQ Makai to process the incoming feature stream and decide whether to request high fidelity data from one or more OPQ Boxes.
 Even at 1 M samples/second at 16 bits of resolution, the~memory requirement to store 5 min of raw waveform without compression are on the order of 512 MB, which is well within the realm of inexpensive single board computers such as Raspberry PI.
@@
 
-\highlight{{\em Cloud processing requirements are reduced.}} Since feature extraction is already performed at the device level, cloud computation requirements are reduced.
+{\em Cloud processing requirements are reduced.} Since feature extraction is already performed at the device level, cloud computation requirements are reduced.
 With~the advent of the Internet of Things, the~computational capacity of edge devices is~increasing.
 
-\highlight{{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.}} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes.
+{\em Subthreshold data acquisition can improve understanding of grid-local anomalies.} OPQ Makai makes the decision to acquire raw waveform from OPQ Boxes.
 This allows analysis of data from devices that were only mildly affected or even not affected at all by the disturbance.
@@
 
-\highlight{{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.}} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features.
+{\em Temporal locality allows OPQ to provide improved insights into power quality anomalies over traditional triggering algorithms.} By exploiting the idea of temporal locality, it is possible to ascertain the geographical extent of an anomaly with only coarse features.
 This allows for a simple robust algorithm, which may be deployed at the sink node for anomaly~detection.
@@
 
-\highlight{{\em Makai Event Actor.}} % Please confirm if this is necessary.
+\highlight{{\em Makai Event Actor.}}
+%%% Please confirm if this is necessary. YOU CAN REMOVE ITALICS.
  The Makai Event Actor is responsible for reading data newly created by OPQ Makai into OPQ Mauka.
@@
 
-\highlight{{\em Frequency Variation Actor.}} The Frequency Variation Actor classifies generic frequency sags, swells, and~interruptions as defined by the IEEE 1159 standard.
+{\em Frequency Variation Actor.} The Frequency Variation Actor classifies generic frequency sags, swells, and~interruptions as defined by the IEEE 1159 standard.
 Both duration and deviation from nominal are used to perform these classifications.
@@
 
-\highlight{{\em IEEE 1159 Voltage Actor.}} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard \hl{[29]}.
+{\em IEEE 1159 Voltage Actor.} The IEEE 1159 Voltage Actor is used to classify voltage Incidents in accordance with the IEEE 1159 standard \highlight{\cite{bilik_modular_2007}}.
 In~general, this standard classifies voltage disturbances by duration and by magnitude.
@@
 This Actor is capable of classifying voltage sags, swells, and~interruptions as defined by the standard, and~creating data at the Incident Layer if~appropriate.
-%%% Should it be reference citation? 
-\highlight{{\em Box Optimization Actor.}} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
+%%% Please revise the reference order.
+Reference 29 appears before reference 14.
+THIS IS A FUNCTION OF LATEX BIBLIOGRAPHY STYLE.
+PLEASE GO AHEAD AND CHANGE THE REFERENCE ORDER TO YOUR OWN SATISFACTION.
+I AM NOT CLEAR WHAT IS NEEDED.
+{\em Box Optimization Actor.} The Box Optimization Actor is responsible for sending and receiving typed messages to and from OPQ Boxes from OPQ Mauka.
 This Actor is capable of requesting the state of each OPQ Box (e.g., uptime, Measurement rate, security keys).
@@
 
-\highlight{{\em Future Phenomena Actor.}} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena.
+{\em Future Phenomena Actor.} The Future Phenomena Actor is responsible for creating Future or Predictive Phenomena.
 These Phenomena are used to predict Events and Incidents that may occur in the future.
@@
 \midrule
-Phenomena & 10.4K \\
-Incidents & 415K \\
-Detections & 91.4K \\
-AML (Trends) & 1.94M \\
+Phenomena & 10.4~K \\
+Incidents & 415~K \\
+Detections & 91.4~K \\
+AML (Trends) & 1.94~M \\
 AML (Measurements) & 116~M \\
-IML & 100B \\
+IML & 100~B \\
 \bottomrule
@@
 \includegraphics[width=0.4\linewidth]{images/pilot/data-management-graph.png}
-\caption{\hl{Storage} requirements with and without~time-to-live (TTL).} %%% It seems that the figure isn't complete, please modify.
+\caption{Storage requirements with and without~time-to-live (TTL).}
 \label{fig:data-management-graph}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{\hl{For} }%% research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``Conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
-All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work reported.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\funding{\hl{Please} }%add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
-Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\conflictsofinterest{\hl{Declare} } %conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
-Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section.
-If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, or in the decision to publish the results''.
+\authorcontributions{
+Conceptualization, P.M.J., A.J.C., S.N.; methodology, P.M.J., A.J.C., S.N.; software, P.M.J., A.J.C., S.N.; validation, P.M.J., A.J.C., S.N.; formal analysis, P.M.J., A.J.C., S.N.; investigation, P.M.J., A.J.C., S.N.; resources, P.M.J.; data curation, A.J.C., S.N.; writing--original draft preparation, P.M.J., A.J.C., S.N.; writing--review and editing, P.M.J., A.J.C., S.N.; visualization, P.M.J., A.J.C., S.N.; supervision, P.M.J.; project administration, P.M.J.; funding acquisition, P.M.J.
+All authors have read and agreed to the published version of the manuscript.
+}
+
+\funding{This research was funded in part by a grant from the University of Hawaii President's Green Award program.}
+
+\conflictsofinterest{The authors declare no conflict of interest.}
 
@@
 \newblock Solar Installers flee {Hawaii} as Interconnection Queue Backs Up.
-\newblock {\em Utility Dive}, {\hl{2014}}.%% Please add day month.
+\newblock {\em Utility Dive}, {September 29, 2014}.
 
@@
 networks with photovoltaic inverters.
-\newblock In {Proceedings of the } 2017 {IEEE} {International} {Conference} on IEEE Industrial and {Information} {Systems} ({ICIIS}),  \hl{2017}; pp.
-1--6, doi:10.1109/ICIINFS.2017.8300335.%% Please add conference location and date.
+\newblock In {Proceedings of the } 2017 {IEEE} {International} {Conference} on IEEE Industrial and {Information} {Systems} ({ICIIS}), Jayapura, Indonesia November 2, 2017; pp.
+1--6, doi:10.1109/ICIINFS.2017.8300335.
 
@@
 \newblock In {Proceedings of the } 11th {International} {Conference} on {Electrical} {Power} {Quality}
-and {Utilisation}, \hl{2011}; pp. 1--6, 
+and {Utilisation}, Lisbon, Portugal, October 17, 2011; pp. 1--6,
 \newblock ISSN 2150-6655,
@@
 \newblock { Power {Quality} and {EMC} {Issues} with {Future} {Electricity}
-{Networks}}. In {Proceedings of the } CIGRE, \hl{2018}.
+{Networks}}. In {Proceedings of the } CIGRE, Paris, France, May 15, 2018.
 
@@
 techniques for utilities.
-\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, \hl{2011}; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.
+\newblock In {Proceedings of the }The 2011 International Joint Conference on Neural Networks, San Jose, California, July 31, 2011; pp. 3093--3098, doi:10.1109/IJCNN.2011.6033630.
 
@@
 {Failures}?
-\newblock {\em Electrical Construction and Maintenance Magazine}, {\hl{2008}}.
-\newblock Library Catalog.  Available online:  \url{www.ecmweb.com} (accessed on \hl{day month year}).
+\newblock {\em Electrical Construction and Maintenance Magazine}, {November 1, 2008}.
+\newblock Library Catalog.  Available online:  \url{www.ecmweb.com} (accessed on 31 July 2020).
 
@@
 \newblock  In {Proceedings of the }2015 {Australasian} {Universities} {Power} {Engineering}
-{Conference}, \hl{2015}.
+{Conference}, Wollongong, Australia, September 27, 2015.
 
@@
 \newblock In~{Proceedings of the }2012 47th
-{International} Universities {Power} {Engineering} {Conference} ({UPEC}), \hl{2012}; pp.~1--6.
+{International} Universities {Power} {Engineering} {Conference} ({UPEC}), London, England, September 4, 2012; pp.~1--6.
 
@@
 \newblock {OPQ} {Box} {Specifications} and {Design}, 2020.
-\newblock Library Catalog.  Available online:  \url{github.com} (accessed on \hl{day month year}).
+\newblock Library Catalog.  Available online:  \url{github.com} (accessed on 31 July 2020).
 
@@
 \newblock \emph{IEEE P1159/D3: {Draft} {Recommended} {Practice} for
-{Monitoring} {Electric} {Power} {Quality}}; \hl{2018}. %% Please provide publisher and its location.
-
+{Monitoring} {Electric} {Power} {Quality}}; IEEE, New York, NY February 1, 2018.
 \bibitem[Kahle(2015)]{kahle_power_2015}
@@
 \newblock Power acceptability and voltage sag indices in the three phase sense.
-\newblock In {Proceedings of the } 2000 {Power} {Engineering} {Society} {Summer} {Meeting}, 
- \hl{2000}; Volume~2, pp. 905--910, doi:10.1109/PESS.2000.867482.
+\newblock In {Proceedings of the } 2000 {Power} {Engineering} {Society} {Summer} {Meeting}, Seattle, Washington,
+ July 16, 2000; Volume~2, pp. 905--910, doi:10.1109/PESS.2000.867482.
 
@@
 \newblock  In Proceedings of the {Ninth} {International} {Conference} on {Smart}
-{Grids}, {Green} {Communications} and {IT} {Energy}-aware {Technologies},
-\hl{2019}.
+{Grids}, {Green} {Communications} and {IT} {Energy}-aware {Technologies}, Athens, Greece,
+June 2, 2019.
 
@@
 \newblock \emph{Fluke {Corporation}: {Fluke} {Electronics}, {Biomedical},
-{Calibration} and {Networks}}; \hl{2020}.
+{Calibration} and {Networks}}; \hl{July 31, 2020}.
+%%% Please confirm whether this is a book. Please add published if it is the case.
 
@@
 \newblock Dranetz {Power} {Monitoring}, 2020.
-\newblock Library Catalog. Available online:  \url{ www.dranetz.com} (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{ www.dranetz.com} (accessed on 31 July 2020).
 
@@
 \newblock Elspec---{Power} Quality Analyzers and Solutions, 2020.
-\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com} (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com} (accessed on 31 July 2020).
 
@@
 \newblock Powerside, {Inc}., 2020.
-\newblock Library Catalog. Available online:  \url{Powerside.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{Powerside.com}  (accessed on 31 July 2020).
 
@@
 \newblock {ACR} {Systems}, 2020.
-\newblock Library Catalog. Available online:  \url{www.acrdatasolutions.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.acrdatasolutions.com}  (accessed on 31 July 2020).
 
@@
 \newblock {PQView} 4 {\textbar} {Electrotek} {Concepts}, 2020.
-\newblock Library Catalog. Available online:  \url{www.electrotek.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.electrotek.com}  (accessed on 31 July 2020).
 
@@
 \newblock {PQSCADA} {Sapphire}, 2016.
-\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com}  (accessed on \hl{day month year}).
+\newblock Library Catalog. Available online:  \url{www.elspec-ltd.com}  (accessed on 31 July 2020).
 
@@
 Sabin, D.
-\newblock \emph{{IEEE} 1159.3 {PQDIF} {Task} {Force}}; \hl{2020}.
+\newblock \emph{{IEEE} 1159.3 {PQDIF} {Task} {Force}}; \hl{July 31, 2020}.
+%%% Please add the link and the publisher.
 
@@
 Alliance, G.P.
-\newblock \emph{Grid {Protection} {Alliance}---{Home}}; \hl{2020}.
+\newblock \emph{Grid {Protection} {Alliance}---{Home}}; \hl{July 31, 2020}.
+%%% Please add the link and the publisher.
 
@@
 \newblock In Proceedings of the 2015 {AEIT} {International} {Annual}
-{Conference}, \hl{2015}.
+{Conference}, Naples, Italy October 14, 2015.
 
@@
 \newblock In {Proceedings of the } Ninth {International} {Conference} on {Harmonics} and {Quality} of
-{Power}, \hl{2000}.
+{Power}, Orlando, Florida October 1, 2000.
 
@@
 \newblock In {Proceedings of the } 9th {International} {Conference} on {Electrical} {Power} {Quality}
-and {Utilization}, \hl{2007}.
+and {Utilization}, Barcelona, Spain, October 9, 2007.
 
@@
 \newblock In {Proceedings of the }2012 {China} {International} {Conference} on {Electricity}
-{Distribution}, \hl{2012}.
+{Distribution}, Shanghai, China, September 10, 2012.
 
@@
 \newblock In {Proceedings of the } {IEEE} 16th {International} {Conference} on {Harmonics} and
-{Quality} of {Power}, \hl{2014}.
+{Quality} of {Power}, Orlando, Florida, October 1, 2014.
 
@@
 Sayied, O.; Spaulding, J.; Akbary, B.
-\newblock \hl{Power quality metrics calculations for the smart grid}. %%% Please add more detailed information.
+\newblock Power quality metrics calculations for the smart grid, Technical Report, ECE 401, Department of Electrical and Computer Engineering, Colorado State University, December 12, 2012.
 
@@
 \newblock In {Proceedings of the } 2016 {IEEE} {International} {Power} {Electronics} and {Motion}
-{Control} {Conference} ({PEMC}), \hl{2016}; pp. 388--393, doi:10.1109/EPEPEMC.2016.7752029.
+{Control} {Conference} ({PEMC}), Varna, Bulgaria, September 25, 2016; pp. 388--393, doi:10.1109/EPEPEMC.2016.7752029.
 
@@
 \newblock In {Proceedings of the } 2012 {IEEE} {Student} {Conference} on {Research} and {Development}
-({SCOReD}), \hl{2012}; pp. 1--6, doi:10.1109/SCOReD.2012.6518600.
+({SCOReD}), Kuala Lumpur, December 13, 2012; pp. 1--6, doi:10.1109/SCOReD.2012.6518600.
 
